---
title: "IAM-PFM: Integrated Assessment Model with Political Feasibility Module"
subtitle: "Econometric Model Results, Diagnostics and Validation Report"
author: "Renato Rodrigues"
date: "`r Sys.Date()`"
params:
  modelData: NULL
  modelDataFile: "modelData.RData"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    code_folding: hide
    fig_width: 9
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  fig.align  = "center"
)

# Create output folder for plots
plotOutputDir <- "output/plots"
if (!dir.exists(plotOutputDir)) dir.create(plotOutputDir, recursive = TRUE)

library(pfm)
library(ggplot2)
library(dplyr)
library(tidyr)
library(kableExtra)
library(patchwork)
library(scales)

# Load local helpers
source("R/kableCorrelationMatrix.R")

madrat::setConfig(
  forcecache = TRUE,
  cachefolder = "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/cache/default"
)

theme_report <- function() {
  theme_minimal(base_size = 16, base_family = "serif") +
    theme(
      plot.title = element_text(face = "bold", size = 14),
      plot.subtitle = element_text(size = 14, colour = "grey40"),
      axis.title = element_text(size = 18),
      axis.text = element_text(size = 14),
      legend.text = element_text(size = 14),
      panel.grid.minor = element_blank(),
      legend.position = "bottom",
      strip.text = element_text(face = "bold", size = 18)
    )
}

# Colour palette used throughout
col_bulk <- "#2C7BB6"
col_diffuse <- "#D7191C"
col_md <- "#1A9641"
col_md2 <- "#FDAE61"
col_md3 <- "#8e44ad"
col_md4 <- "#f39c12"

# Central Forest Plot Rendering Function
render_forest_plot <- function(coef_df, title, subtitle) {
  if (nrow(coef_df) == 0) {
    return(NULL)
  }

  p_data <- coef_df %>%
    filter(term != "(Intercept)") %>%
    mutate(
      lo = estimate - 1.96 * stdError,
      hi = estimate + 1.96 * stdError,
      sig = case_when(
        pValue < 0.001 ~ "***", pValue < 0.01 ~ "**",
        pValue < 0.05 ~ "*", pValue < 0.1 ~ ".", TRUE ~ ""
      ),
      Group = case_when(
        grepl("_x_", term) ~ "Interaction",
        grepl("Actor.Power.Index|^VRE.share|^Electrification|^Coal|^Oil.and.Gas|^Fossil.in.Industry", term) ~ "Actor Power",
        grepl("Government.Effectiveness|Voice.and.Accountability|Political.Stability", term) ~ "Institutional Quality",
        grepl("GDP.per.Capita|Land.Area|lagged_ecp|lagged_adoption", term) ~ "Controls",
        grepl("^regionFE", term) ~ "Region Fixed Effects",
        grepl("timeTrend|Year|year", term) ~ "Time Trend",
        TRUE ~ "Institutional Quality & Controls"
      )
    )

  p_data$Group <- factor(p_data$Group, levels = c(
    "Actor Power", "Institutional Quality", "Interaction",
    "Controls", "Region Fixed Effects", "Time Trend", "Institutional Quality & Controls"
  ))

  ggplot(p_data, aes(x = reorder(term, estimate), y = estimate, color = Group)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.2, linewidth = 0.8) +
    geom_text(aes(label = sig), nudge_y = 0.4, size = 4.5, color = "black", show.legend = FALSE) +
    coord_flip() +
    labs(title = title, subtitle = subtitle, x = NULL, y = "Estimate (95% CI)") +
    theme_report() +
    scale_color_manual(values = c(
      "Actor Power" = "#e74c3c", # Red
      "Institutional Quality" = "#2ecc71", # Green
      "Interaction" = "#f1c40f", # Yellow
      "Controls" = "#3498db", # Blue
      "Region Fixed Effects" = "#9b59b6", # Purple
      "Time Trend" = "#34495e" # Dark Grey
    )) +
    theme(legend.position = "right")
}
col_good <- "#2ecc71"
col_warn <- "#f39c12"
col_bad <- "#e74c3c"

# --- Static Driver Lists ---
ext_actorPowerDrivers <- c(
  "VRE share", "Electrification",
  "Coal primary energy share", "Oil/Gas primary energy share",
  "Fossil share in Industry"
)
ext_instQualityDrivers <- c(
  "Government Effectiveness", "Control of Corruption",
  "Voice and Accountability", "Political Stability", "Regulatory Quality", "Rule of Law"
)
ext_controlDrivers <- c(
  "Population", "GDP per Capita", "Land Area",
  "Urban Population Share", "Gini Income Inequality Coefficient",
  "Gender Inequality Index", "Energy Intensity"
)

# --- Load or Retrieve Model Data ---
# Decide if we load from file or use a provided object
use_file <- is.null(params$modelData) || is.character(params$modelData)
mFile <- if (is.character(params$modelData)) params$modelData else params$modelDataFile

if (use_file) {
  if (file.exists(mFile)) {
    load(mFile)
  } else {
    # --- Load panel data once ---
    panelData <- panelDataHistorical(
      aggregate = TRUE, y = 2000:2022,
      outputRegionMappingFile = "regionmapping_54.csv"
    )

    # --- Compute correlation matrices ---
    cor_df_AP <- computeCorrelationMatrix(panelData, ext_actorPowerDrivers)
    cor_df_IQ <- computeCorrelationMatrix(panelData, ext_instQualityDrivers)
    cor_df_ctrl <- computeCorrelationMatrix(panelData, ext_controlDrivers)

    # --- Selected drivers for models ---
    actorPowerDrivers <- c(
      "VRE share", "Electrification",
      "Coal primary energy share", "Oil/Gas primary energy share",
      "Fossil share in Industry"
    )
    instQualityDrivers <- c("Government Effectiveness") # "Voice and Accountability", "Political Stability"
    controlDrivers <- c("GDP per Capita", "Population") # "Land Area"

    # --- Run model selection workflows ---
    md <- modelSelectionWorkflow(
      panelData = panelData,
      outputRegionMappingFile = "regionmapping_54.csv",
      actorPowerDrivers = actorPowerDrivers,
      actorPowerIndex = "Actor Power Index",
      instQualityDrivers = instQualityDrivers,
      controlDrivers = controlDrivers,
      regionMappingFixedEffects = "regionmapping_EU_OECDp.csv"
    )

    md2 <- modelSelectionWorkflow(
      panelData = panelData,
      outputRegionMappingFile = "regionmapping_54.csv",
      actorPowerDrivers = NULL,
      actorPowerIndex = "Actor Power Index",
      instQualityDrivers = instQualityDrivers,
      controlDrivers = controlDrivers,
      regionMappingFixedEffects = "regionmapping_EU_OECDp.csv"
    )

    md3 <- modelSelectionWorkflow(
      panelData = panelData,
      outputRegionMappingFile = "regionmapping_54.csv",
      actorPowerDrivers = NULL,
      actorPowerIndex = "Actor Power Index",
      instQualityDrivers = instQualityDrivers,
      controlDrivers = controlDrivers,
      regionMappingFixedEffects = "regionmapping_54.csv" # REMIND regions instead of EU/OECD+
    )

    md4 <- modelSelectionWorkflow(
      panelData = panelData,
      outputRegionMappingFile = "regionmapping_54.csv",
      actorPowerDrivers = NULL,
      actorPowerIndex = "Actor Power Index",
      instQualityDrivers = instQualityDrivers,
      controlDrivers = controlDrivers,
      regionMappingFixedEffects = "regionmapping_EU_OECDp.csv",
      includeLaggedECP = TRUE # lagged ECP as predictor in stringency stage
    )

    # --- Load scenario data for projections ---
    gdx_path <- "C:/Users/renatoro/Desktop/Projects/Elevate/code/fulldata.gdx"
    future_mag <- if (file.exists(gdx_path)) {
      panelDataScenario(gdxFile = gdx_path, outputRegionMappingFile = "regionmapping_54.csv")
    } else {
      NULL
    }

    future_df_bulk <- NULL
    future_df_diffuse <- NULL

    if (!is.null(future_mag)) {
      future_df_bulk <- preparePanelData(
        data = future_mag, sector = "Bulk",
        actorPowerDrivers = ext_actorPowerDrivers, actorPowerIndex = "Actor Power Index",
        instQualityDrivers = ext_instQualityDrivers, controlDrivers = ext_controlDrivers,
        regionMappingFixedEffects = "regionmapping_EU_OECDp.csv", lag = 1
      )
      future_df_diffuse <- preparePanelData(
        data = future_mag, sector = "Diffuse",
        actorPowerDrivers = ext_actorPowerDrivers, actorPowerIndex = "Actor Power Index",
        instQualityDrivers = ext_instQualityDrivers, controlDrivers = ext_controlDrivers,
        regionMappingFixedEffects = "regionmapping_EU_OECDp.csv", lag = 1
      )
    }

    # Organize all data in a single object
    modelData <- list(
      panelData = panelData,
      cor_df_AP = cor_df_AP,
      cor_df_IQ = cor_df_IQ,
      cor_df_ctrl = cor_df_ctrl,
      md = md,
      md2 = md2,
      md3 = md3,
      md4 = md4,
      future_mag = future_mag,
      future_df_bulk = future_df_bulk,
      future_df_diffuse = future_df_diffuse
    )
    # Save the data for future use
    save(modelData, file = mFile)
  }
} else {
  # Use provided data object
  modelData <- params$modelData
}

# --- Unpack modelData for the rest of the report ---
panelData <- modelData$panelData
cor_df_AP <- modelData$cor_df_AP
cor_df_IQ <- modelData$cor_df_IQ
cor_df_ctrl <- modelData$cor_df_ctrl
md <- modelData$md
md2 <- modelData$md2
md3 <- modelData$md3
md4 <- modelData$md4
future_mag <- modelData$future_mag
future_df_bulk <- modelData$future_df_bulk
future_df_diffuse <- modelData$future_df_diffuse

# --- Helper: extract best model coefficients as data.frame ---
get_best_coefs <- function(wf, sector, stage) {
  comboName <- paste(sector, stage, sep = "_")
  sel <- wf$selections[[comboName]]
  if (is.null(sel) || is.null(sel$bestModel)) {
    return(data.frame(term = character(), estimate = numeric(), stdError = numeric(), statistic = numeric(), pValue = numeric()))
  }
  pfm::coeftestToDataFrame(sel$bestModel$coeftest)
}

# --- Helper: get best model diagnostics ---
get_best_diag <- function(wf, sector, stage) {
  comboName <- paste(sector, stage, sep = "_")
  sel <- wf$selections[[comboName]]
  if (is.null(sel) || is.null(sel$bestModel)) {
    return(list(pseudoR2 = NA, aic = NA, nSignificant = 0, nPredictors = 0))
  }
  sel$bestModel
}

# --- Dynamic diagnosis ---
# md  = full AP drivers + EU/OECD+ region FE
# md2 = API composite   + EU/OECD+ region FE          (preferred baseline)
# md3 = API composite   + REMIND region FE (54-region mapping, finer granularity)
# md4 = API composite   + EU/OECD+ region FE + lagged ECP in stringency
diag_results <- diagnoseModelResults(
  workflows = list(md = md, md2 = md2, md3 = md3, md4 = md4)
)
```

---

# Summary

The **IAM-PFM** (Integrated Assessment Model with Political Feasibility Module) embeds two-stage hurdle models for carbon-pricing into the REMIND integrated assessment model. For each of two electricity-sector archetypes — **Bulk** (large industrial/utility consumers) and **Diffuse** (transport, buildings, small industry) — the hurdle comprises:

* **Stage 1 — Adoption** (`logistf`): penalised logistic regression predicting whether a jurisdiction has adopted carbon pricing (binary outcome).
* **Stage 2 — Stringency** (Gamma GLM with log link, bias-reduced via `brglm2`): conditional on adoption, a GLM predicts the level of the Effective Carbon Price (ECP).

Four model variants were estimated:

```{r table-variants, echo=FALSE}
table_data <- data.frame(
  Variant = c("**md**", "**md2**", "**md3**", "**md4**"),
  `Actor Power drivers` = c(
    "Full set (VRE share, Electrification, Coal, Oil/Gas, Fossil in Industry)",
    "Actor Power Index composite",
    "Actor Power Index composite",
    "Actor Power Index composite"
  ),
  `Region Fixed effects` = c(
    "clustered regions (EU+, OECD+, other)",
    "clustered regions (EU+, OECD+, other)",
    "54-region mapping",
    "clustered regions (EU+, OECD+, other)"
  ),
  `Lagged ECP` = c(
    "NO",
    "NO",
    "NO",
    "YES"
  ),
  check.names = FALSE
)

kable(table_data, escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(1, width = "10%") %>%
  column_spec(2, width = "40%") %>%
  column_spec(3, width = "40%") %>%
  column_spec(3, width = "10%")
```

**Top-line findings:**

```{r exec-summary-bullets, echo=FALSE, results='asis'}
cat(diag_results$globalFindings)
```

---

# Model Architecture {.tabset}

## Conceptual Flow

```{r architecture-diagram, echo=FALSE, fig.height=3}
nodes <- data.frame(
  x = c(1, 3, 5, 5),
  y = c(2, 2, 3, 1),
  label = c(
    "Panel Data\n2000–2022", "Stage 1\nAdoption\n(logistf)",
    "Stage 2\nStringency\n(Gamma GLM)", "ECP = 0\n(no policy)"
  ),
  fill = c("grey70", col_md, col_bulk, "grey85")
)

ggplot(nodes, aes(x, y)) +
  geom_tile(aes(fill = fill), width = 1.4, height = 0.7, colour = "white", linewidth = 1.2) +
  geom_text(aes(label = label), size = 3.2, fontface = "bold", colour = "white") +
  scale_fill_identity() +
  annotate("segment",
    x = 1.72, xend = 2.28, y = 2, yend = 2,
    arrow = arrow(length = unit(0.25, "cm")), linewidth = 0.8, colour = "grey40"
  ) +
  annotate("segment",
    x = 3.72, xend = 4.28, y = 2.3, yend = 2.85,
    arrow = arrow(length = unit(0.25, "cm")), linewidth = 0.8, colour = col_bulk
  ) +
  annotate("segment",
    x = 3.72, xend = 4.28, y = 1.7, yend = 1.15,
    arrow = arrow(length = unit(0.25, "cm")), linewidth = 0.8, colour = "grey50"
  ) +
  annotate("text", x = 4, y = 2.7, label = "Pr(ECP > 0) = 1", size = 2.8, colour = col_bulk) +
  annotate("text", x = 4, y = 1.3, label = "Pr(ECP > 0) = 0", size = 2.8, colour = "grey50") +
  xlim(0, 6.5) +
  ylim(0.4, 3.7) +
  labs(title = "IAM-PFM Two-Stage Hurdle Architecture") +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 13))
```

## Variable Groups

```{r variable-table, echo=FALSE}
var_df <- data.frame(
  Group = c(
    rep("Actor Power (AP)", 5),
    rep("Institutional Quality (IQ)", 1),
    rep("Controls", 2),
    rep("Fixed Effects / Trend", 2)
  ),
  Variable = c(
    "VRE Share", "Electrification", "Coal Primary Energy Share",
    "Oil/Gas Primary Energy Share", "Fossil Share in Industry",
    "Government Effectiveness", # "Voice & Accountability", "Political Stability",
    "GDP per Capita", "Population",
    "Region FE (EU / OECD+ / other)", "Time Trend"
  ),
  Role = c(
    rep("Proxy for fossil-fuel incumbents' resistance", 5),
    rep("State capacity & democratic accountability", 1),
    rep("Economic & geographic controls", 2),
    "Captures unobserved regional heterogeneity", "Secular trend in adoption/stringency"
  )
)

kable(var_df, caption = "Model Variable Groups and Roles") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  collapse_rows(columns = 1, valign = "top") # %>%
# column_spec(1, bold = TRUE, color = col_md)
```

## Formal Specification

The model follows a two-stage hurdle specification to capture both the **extensive margin** (policy adoption) and **intensive margin** (policy stringency).

### Stage 1: Adoption (Logit)

Predicts the probability that a jurisdiction $i$ at time $t$ has adopted carbon pricing ($ECP > 0$) using a logit link:

$$\text{logit}(Pr(ECP_{i,t,s} > 0)) = \alpha_{0,s} + \sum_l \alpha_{1,l,s} AP_{i,t-1,s,l} + \sum_k \alpha_{2,k,s} IQ_{i,t-1,s,k} + AP\_Index_{i,t-1,s} \times \sum_k \alpha_{e,k,s} IQ_{i,t-1,s,k} + \sum_j \alpha_{4,j,s} Ctrl_{i,t-1,s,j} + CRFE + TT + \epsilon_{i,t-1,s}$$

### Stage 2: Stringency (Gamma GLM)

Conditional on adoption, predicts the level of the Effective Carbon Price (ECP) using a log-link:

$$\ln(E[ECP_{i,t,s} | ECP_{i,t,s} > 0]) = \beta_{0,s} + \sum_l \beta_{1,l,s} AP_{i,t-1,s,l} + \sum_k \beta_{2,k,s} IQ_{i,t-1,s,k} + AP\_Index_{i,t-1,s} \times \sum_k \beta_{e,k,s} IQ_{i,t-1,s,k} + \sum_j \beta_{4,j,s} Ctrl_{i,t-1,s,j} + CRFE + TT + \epsilon_{i,t-1,s}$$

**Where:**

* $AP_{i,t-1,s,l}$ denotes **Actor Power** drivers (proxies for fossil-fuel incumbents).
* $IQ_{i,t-1,s,k}$ denotes **Institutional Quality** drivers (state capacity and governance).
* $AP\_Index_{i,t-1,s}$ is the composite **Actor Power Index**.
* $Ctrl_{i,t-1,s,j}$ are geographic and economic **control variables**.
* $CRFE$ are **Carbon-pricing Region Fixed Effects** (capturing EU/OECD+ heterogeneity).
* $TT$ is a linear **Time Trend**.
* All drivers are **lagged by 1 year** ($t-1$) to avoid simultaneity bias.

---

# Correlation Analysis {.tabset}

High correlations between predictors within the same group can inflate variance and destabilise estimates.

## Actor Power Drivers

```{r cor-ap, echo=FALSE}
kableCorrelationMatrix(cor_df_AP,
  caption = "Actor Power Correlation Matrix",
  show = "lower", percentile = 0.85
)
```

```{r cor-ap-note, echo=FALSE, results='asis'}
# Find the two highest off-diagonal correlations dynamically
cor_ap_mat <- as.matrix(cor_df_AP[, -1])
rownames(cor_ap_mat) <- cor_df_AP$Driver
diag(cor_ap_mat) <- NA
high_pair <- which(abs(cor_ap_mat) == max(abs(cor_ap_mat), na.rm = TRUE), arr.ind = TRUE)[1, ]
low_pair <- which(abs(cor_ap_mat) == min(abs(cor_ap_mat), na.rm = TRUE), arr.ind = TRUE)[1, ]
cat(
  "> **Note:** ",
  rownames(cor_ap_mat)[high_pair[1]], "and", colnames(cor_ap_mat)[high_pair[2]],
  "share r =", round(cor_ap_mat[high_pair[1], high_pair[2]], 2),
  "— moderate but not disqualifying.",
  rownames(cor_ap_mat)[low_pair[1]], "and", colnames(cor_ap_mat)[low_pair[2]],
  "are essentially orthogonal (r =", round(cor_ap_mat[low_pair[1], low_pair[2]], 2), ")."
)
```

## Institutional Quality Drivers

```{r cor-iq, echo=FALSE}
kableCorrelationMatrix(cor_df_IQ,
  caption = "Institutional Quality Correlation Matrix",
  show = "lower", percentile = 0.85
)
```

```{r cor-iq-note, echo=FALSE, results='asis'}
cor_iq_mat <- as.matrix(cor_df_IQ[, -1])
diag(cor_iq_mat) <- NA
mean_r <- round(mean(abs(cor_iq_mat), na.rm = TRUE), 2)
max_r <- round(max(abs(cor_iq_mat), na.rm = TRUE), 2)
cat(
  "> **Critical finding:** The World Governance Indicators are highly inter-correlated",
  "(mean |r| =", mean_r, ", max |r| =", max_r, ").",
  "Using more than two simultaneously risks severe multicollinearity.",
  "Select a small subset (e.g. Government Effectiveness, Voice & Accountability,",
  "Political Stability) and verify VIF < 10 after model fitting."
)
```

## Control Variables

```{r cor-ctrl, echo=FALSE}
kableCorrelationMatrix(cor_df_ctrl,
  caption = "Controls Correlation Matrix",
  show = "lower", percentile = 0.85
)
```

```{r cor-ctrl-note, echo=FALSE, results='asis'}
cor_ctrl_mat <- as.matrix(cor_df_ctrl[, -1])
diag(cor_ctrl_mat) <- NA
high_ctrl <- which(abs(cor_ctrl_mat) == max(abs(cor_ctrl_mat), na.rm = TRUE), arr.ind = TRUE)[1, ]
cat(
  "> ",
  rownames(as.data.frame(cor_df_ctrl[, -1]))[high_ctrl[1]],
  "and", colnames(cor_ctrl_mat)[high_ctrl[2]],
  "are strongly correlated (r =",
  round(cor_ctrl_mat[high_ctrl[1], high_ctrl[2]], 2),
  ") and should not both appear in the same model.",
  "Use at most one from each highly correlated pair to keep VIF controlled."
)
```

---

# Model Fit Summary

```{r fit-summary, echo=FALSE}
# Build fit summary dynamically from all workflows
build_fit_row <- function(wf, model_name, sector, stage) {
  bm <- wf$bestModels
  if (is.null(bm) || nrow(bm) == 0) {
    return(NULL)
  }
  row <- bm[bm$sector == sector & bm$stage == stage, ]
  if (nrow(row) == 0) {
    return(NULL)
  }
  diag <- get_best_diag(wf, sector, stage)
  data.frame(
    Model = model_name, Sector = sector,
    Stage = ifelse(stage == "adoption", "Adoption", "Stringency"),
    PseudoR2 = row$pseudoR2, AIC = row$aic,
    nSig = row$nSignificant, nPred = row$nPredictors,
    Converged = row$converged,
    Separation = isTRUE(diag$separation),
    maxVIF = row$maxVIF,
    stringsAsFactors = FALSE
  )
}

wfs <- list(
  list(wf = md, name = "md (full AP)"),
  list(wf = md2, name = "md2 (API only)"),
  list(wf = md3, name = "md3 (REMIND FE)"),
  list(wf = md4, name = "md4 (Lagged ECP)")
)

grid <- expand.grid(
  wf_idx = 1:4,
  sector = c("Bulk", "Diffuse"),
  stage = c("adoption", "stringency"),
  stringsAsFactors = FALSE
)
grid$stage_order <- ifelse(grid$stage == "adoption", 1, 2)
grid$sector_order <- ifelse(grid$sector == "Bulk", 1, 2)
grid <- grid[order(grid$stage_order, grid$sector_order, grid$wf_idx), ]

fit_df <- do.call(rbind, lapply(1:nrow(grid), function(i) {
  wf_info <- wfs[[grid$wf_idx[i]]]
  build_fit_row(wf_info$wf, wf_info$name, grid$sector[i], grid$stage[i])
}))

kable(fit_df,
  format = "html", caption = "IAM-PFM Model Fit Summary — All Variants",
  col.names = c(
    "Model", "Sector", "Stage", "Pseudo-R²", "AIC",
    "Sig. pred.", "Total pred.", "Converged", "Sep. flag", "Max VIF"
  )
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(4,
    bold = TRUE,
    color = ifelse(fit_df$PseudoR2 >= 0.5, col_good,
      ifelse(fit_df$PseudoR2 >= 0.2, col_warn, col_bad)
    )
  )
```

```{r fit-summary-narrative, echo=FALSE, results='asis'}
cat(diag_results$fitSummaryNarrative)
```

---

```{r diagnostic-helpers, echo=FALSE}
# Central Forest Plot Rendering Function
render_forest_plot <- function(coef_df, title, subtitle) {
  if (nrow(coef_df) == 0) {
    return(NULL)
  }
  # Identify variables based on the defined driver lists
  ap_vars <- make.names(c("Actor Power Index", "Incumbent Power", ext_actorPowerDrivers))
  iq_vars <- make.names(ext_instQualityDrivers)
  ctrl_vars <- make.names(c(ext_controlDrivers, "lagged_ecp", "lagged_adoption"))

  p_data <- coef_df %>%
    filter(term != "(Intercept)") %>%
    mutate(
      lo = estimate - 1.96 * stdError,
      hi = estimate + 1.96 * stdError,
      sig = case_when(
        pValue < 0.001 ~ "***", pValue < 0.01 ~ "**",
        pValue < 0.05 ~ "*", pValue < 0.1 ~ ".", TRUE ~ ""
      ),
      Group = case_when(
        grepl("_x_|_X_", term) ~ "Interaction",
        term %in% ap_vars ~ "Actor Power",
        term %in% iq_vars ~ "Institutional Quality",
        term %in% ctrl_vars ~ "Controls",
        grepl("^regionFE", term) ~ "Region Fixed Effects",
        grepl("timeTrend|Year|year", term) ~ "Time Trend",
        TRUE ~ "Institutional Quality & Controls"
      ),
      clean_term = term %>%
        gsub("\\.", " ", .) %>%
        gsub("_x_|_X_", " x ", .)
    )

  p_data$Group <- factor(p_data$Group, levels = c(
    "Actor Power", "Institutional Quality", "Interaction",
    "Controls", "Region Fixed Effects", "Time Trend"
  ))

  ggplot(p_data, aes(x = reorder(clean_term, estimate), y = estimate, color = Group)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(size = 3) +
    geom_errorbar(aes(ymin = lo, ymax = hi), width = 0.2, linewidth = 0.8) +
    geom_text(aes(label = sig), nudge_y = 0.4, size = 4.5, color = "black", show.legend = FALSE) +
    coord_flip() +
    labs(title = title, subtitle = subtitle, x = NULL, y = "Estimate (95% CI)") +
    theme_report() +
    scale_color_manual(values = c(
      "Actor Power" = "#e74c3c", # Red
      "Institutional Quality" = "#2ecc71", # Green
      "Interaction" = "#f1c40f", # Yellow
      "Controls" = "#3498db", # Blue
      "Region Fixed Effects" = "#9b59b6", # Purple
      "Time Trend" = "#34495e" # Dark Grey
    )) +
    guides(color = guide_legend(nrow = 2, title = NULL)) +
    theme(legend.position = "bottom")
}
```

# Model Detailed Results {.tabset}

```{r detailed-results-loop, echo=FALSE, results='asis', fig.height=5}
wf_meta <- list(
  "md"  = "md: Results with Full AP Drivers",
  "md2" = "md2: Results with API Only",
  "md3" = "md3: Results with REMIND Region FE",
  "md4" = "md4: Results with Lagged ECP in Stringency"
)

for (wfid in names(wf_meta)) {
  wf_obj <- get(wfid)
  cat("\n## ", wf_meta[[wfid]], " {.tabset}\n\n")

  for (sec in c("Bulk", "Diffuse")) {
    cat("\n### ", sec, " {.tabset}\n\n")

    for (stg in c("adoption", "stringency")) {
      cat("\n#### ", tools::toTitleCase(stg), "\n\n")

      # 1. Prepare Coefficients
      coef_df <- get_best_coefs(wf_obj, sec, stg) %>%
        filter(term != "(Intercept)")

      if (nrow(coef_df) > 0) {
        # 2. Get Diagnostics
        diag_fit <- get_best_diag(wf_obj, sec, stg)

        # 3. Render Plot
        stg_lbl <- ifelse(stg == "adoption", "Log-odds", "ECP")
        p <- render_forest_plot(
          coef_df,
          title = sprintf("%s %s — Coefficient Plot", sec, tools::toTitleCase(stg)),
          subtitle = sprintf(
            "%s coefficients | Pseudo-R² = %.3f | AIC = %.1f | %d/%d significant",
            stg_lbl, diag_fit$pseudoR2, diag_fit$aic, diag_fit$nSignificant, diag_fit$nPredictors
          )
        )

        # 4. Handle Extreme z-values for Stringency
        if (!is.null(p) && stg == "stringency") {
          extreme_data <- coef_df %>% filter(abs(statistic) > 15)
          if (nrow(extreme_data) > 0) {
            p <- p + geom_text(
              data = extreme_data,
              aes(label = "⚠ Extreme z", y = estimate),
              nudge_y = -0.8, size = 3, colour = col_bad
            )
          }
        }

        if (!is.null(p)) {
          print(p)

          # Save plot
          fname <- sprintf("forest_%s_%s_%s", wfid, sec, stg)
          ggsave(file.path(plotOutputDir, paste0(fname, ".png")), p, width = 10, height = 6, dpi = 300)
          ggsave(file.path(plotOutputDir, paste0(fname, ".svg")), p, width = 10, height = 6)
        }
        cat("\n\n")

        # 5. Interpretation Text
        d <- diag_results$modelDiagnoses[[wfid]][[paste0(sec, "_", stg)]]
        if (!is.null(d)) {
          cat("**Top predictors** (by |z|):", paste0("**", d$topPredictors, "**", collapse = ", "), "\n\n")

          if (length(d$extremeZTerms) > 0) {
            cat("> ⚠ **Extreme z-values detected** on:", paste(d$extremeZTerms, collapse = ", "), "\n\n")
          }

          if (isTRUE(d$separation) && stg == "adoption") {
            cat("> ℹ Separation flag detected — logistf handles this via penalised ML; widen prediction intervals.\n\n")
          }

          # Specific comparison for md2
          if (wfid == "md2" && stg == "adoption") {
            d_md <- diag_results$modelDiagnoses$md[[paste0(sec, "_", stg)]]
            if (!is.null(d_md)) {
              delta_r2 <- round(d_md$r2 - d$r2, 3)
              delta_aic <- round(d$aic - d_md$aic, 1)
              cat(sprintf("> ℹ **Comparison with md:** ΔPseudo-R² = %.3f, ΔAIC = %.1f.\n\n", delta_r2, delta_aic))
            }
          }

          cat("> ", d$verdictIcon, "**Verdict:**", d$verdict$text, "\n")
        }
      } else {
        cat("> ❌ **Model failed to converge or select a valid specification.**\n")
      }
      cat("\n\n")
    }
  }
}
```

---

# Model Comparison

## Adoption

```{r comparison-adoption, echo=FALSE, fig.height=4}
comp_adopt <- do.call(rbind, lapply(names(diag_results$modelDiagnoses), function(wfn) {
  wf <- get(wfn)
  lbl <- switch(wfn,
    "md" = "md (full AP)",
    "md2" = "md2 (API only)",
    "md3" = "md3 (REMIND FE)",
    "md4" = "md4 (Lagged ECP)",
    wfn
  )
  if (!is.null(wf$bestModels) && nrow(wf$bestModels) > 0) {
    wf$bestModels[wf$bestModels$stage == "adoption", c("sector", "pseudoR2", "aic")] %>% mutate(Model = lbl)
  } else {
    NULL
  }
}))

p1 <- ggplot(comp_adopt, aes(x = sector, y = pseudoR2, fill = Model)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", pseudoR2)),
    position = position_dodge(0.6), vjust = -0.4, size = 3.5
  ) +
  scale_fill_manual(values = c("md (full AP)" = col_md, "md2 (API only)" = col_md2, "md3 (REMIND FE)" = col_md3, "md4 (Lagged ECP)" = col_md4)) +
  labs(title = "Adoption — Pseudo-R²", y = "McFadden Pseudo-R²", x = NULL) +
  ylim(0, 0.95) +
  theme_report()

p2 <- ggplot(comp_adopt, aes(x = sector, y = aic, fill = Model)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = sprintf("%.1f", aic)),
    position = position_dodge(0.6), vjust = -0.4, size = 3.5
  ) +
  scale_fill_manual(values = c("md (full AP)" = col_md, "md2 (API only)" = col_md2, "md3 (REMIND FE)" = col_md3, "md4 (Lagged ECP)" = col_md4)) +
  labs(title = "Adoption — AIC (lower = better)", y = "AIC", x = NULL) +
  theme_report()

p1 + p2 + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

```{r comparison-adoption-narrative, echo=FALSE, results='asis'}
cat(">", diag_results$comparisonNarrative$adoption, "\n")
```

## Stringency

```{r comparison-stringency, echo=FALSE, fig.height=4}
comp_str <- do.call(rbind, lapply(names(diag_results$modelDiagnoses), function(wfn) {
  wf <- get(wfn)
  lbl <- switch(wfn,
    "md" = "md (full AP)",
    "md2" = "md2 (API only)",
    "md3" = "md3 (REMIND FE)",
    "md4" = "md4 (Lagged ECP)",
    wfn
  )
  if (!is.null(wf$bestModels) && nrow(wf$bestModels) > 0) {
    wf$bestModels[wf$bestModels$stage == "stringency", c("sector", "pseudoR2", "aic")] %>% mutate(Model = lbl)
  } else {
    NULL
  }
}))

p3 <- ggplot(comp_str, aes(x = sector, y = pseudoR2, fill = Model)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", pseudoR2)),
    position = position_dodge(0.6), vjust = -0.4, size = 3.5
  ) +
  scale_fill_manual(values = c("md (full AP)" = col_md, "md2 (API only)" = col_md2, "md3 (REMIND FE)" = col_md3, "md4 (Lagged ECP)" = col_md4)) +
  labs(title = "Stringency — Pseudo-R²", y = "Pseudo-R²", x = NULL) +
  ylim(0, max(comp_str$pseudoR2) * 1.3) +
  theme_report()

p4 <- ggplot(comp_str, aes(x = sector, y = aic, fill = Model)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = sprintf("%.0f", aic)),
    position = position_dodge(0.6), vjust = -0.4, size = 3.5
  ) +
  scale_fill_manual(values = c("md (full AP)" = col_md, "md2 (API only)" = col_md2, "md3 (REMIND FE)" = col_md3, "md4 (Lagged ECP)" = col_md4)) +
  labs(title = "Stringency — AIC", y = "AIC", x = NULL) +
  theme_report()

p3 + p4 + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

```{r comparison-stringency-narrative, echo=FALSE, results='asis'}
cat(diag_results$comparisonNarrative$stringency, "\n")
```

---

# Diagnostic Deep-Dive

## Workflow Summary {.tabset}

### md

```{r workflow-summary-md, echo=FALSE, comment=""}
cat(md$workflowSummary)
```

### md2

```{r workflow-summary-md2, echo=FALSE, comment=""}
cat(md2$workflowSummary)
```

### md3

```{r workflow-summary-md3, echo=FALSE, comment=""}
cat(md3$workflowSummary)
```

### md4

```{r workflow-summary-md4, echo=FALSE, comment=""}
cat(md4$workflowSummary)
```


## Group Analysis {.tabset}

### md

```{r group-analysis-md, echo=FALSE}
kable(md$groupAnalysisTable,
  format = "html",
  caption = "Group-Level Variance Partitioning — md (Full AP Drivers)"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

```{r group-analysis-md-findings, echo=FALSE, results='asis'}
# groupAnalysisTable has `Pseudo-R2 contribution` as a character (e.g. "5.23%"); parse for ordering
ga_md <- md$groupAnalysisTable
ga_md$.r2num <- as.numeric(gsub("%", "", ga_md$`Pseudo-R2 contribution`))
adopt_top <- ga_md[ga_md$stage == "adoption", ]
adopt_top <- adopt_top[order(adopt_top$.r2num, decreasing = TRUE), ]
str_top <- ga_md[ga_md$stage == "stringency", ]
str_top <- str_top[order(str_top$.r2num, decreasing = TRUE), ]
cat(
  "**Findings:** The group analysis reveals how much explanatory power each predictor block",
  "contributes independently.\n\n"
)
if (nrow(adopt_top) > 0) {
  cat(
    "- **Adoption:** Top contributing group is", adopt_top$group[1],
    sprintf("(%s of variance explained).", adopt_top$`Pseudo-R2 contribution`[1]), "\n"
  )
}
if (nrow(str_top) > 0) {
  cat(
    "- **Stringency:** Top contributing group is", str_top$group[1],
    sprintf("(%s).", str_top$`Pseudo-R2 contribution`[1]),
    "If Region Fixed Effects dominate, much of the explained variance reflects",
    "between-region differences rather than within-region temporal dynamics,",
    "which limits out-of-sample projection utility.\n"
  )
}
```

### md2

```{r group-analysis-md2, echo=FALSE}
kable(md2$groupAnalysisTable,
  format = "html",
  caption = "Group-Level Variance Partitioning — md2 (API Only)"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

```{r group-analysis-md2-findings, echo=FALSE, results='asis'}
ga_md2 <- md2$groupAnalysisTable
ga_md2$.r2num <- as.numeric(gsub("%", "", ga_md2$`Pseudo-R2 contribution`))
adopt_top2 <- ga_md2[ga_md2$stage == "adoption", ]
adopt_top2 <- adopt_top2[order(adopt_top2$.r2num, decreasing = TRUE), ]
cat("**Findings:** The md2 group analysis shows a cleaner decomposition.\n\n")
if (nrow(adopt_top2) > 0) {
  cat(
    "- Without individual AP drivers competing for variance, the top contributor is",
    adopt_top2$group[1],
    sprintf("(%s of variance explained),", adopt_top2$`Pseudo-R2 contribution`[1]),
    "suggesting the composite API captures political-economy dynamics more efficiently",
    "than disaggregated fuel-specific proxies.\n"
  )
}
```

### md3

```{r group-analysis-md3, echo=FALSE}
kable(md3$groupAnalysisTable,
  format = "html",
  caption = "Group-Level Variance Partitioning — md3 (REMIND FE)"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

### md4

```{r group-analysis-md4, echo=FALSE}
kable(md4$groupAnalysisTable,
  format = "html",
  caption = "Group-Level Variance Partitioning — md4 (Lagged ECP)"
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

## Pseudo-R² Comparison

```{r r2-context, echo=FALSE, fig.height=4.5}
r2_df <- do.call(rbind, lapply(names(diag_results$modelDiagnoses), function(wfn) {
  res <- get(wfn)$bestModels
  if (!is.null(res) && nrow(res) > 0) {
    res %>% mutate(Model = paste0(sector, " ", stage, " (", wfn, ")"))
  } else {
    NULL
  }
})) %>%
  mutate(
    Quality = case_when(
      pseudoR2 >= 0.4 ~ "Good",
      pseudoR2 >= 0.2 ~ "Moderate",
      TRUE ~ "Poor"
    ),
    stage_order = ifelse(stage == "adoption", 1, 2),
    sector_order = ifelse(sector == "Bulk", 1, 2)
  ) %>%
  arrange(desc(stage_order), desc(sector_order), Model) %>%
  mutate(Model = factor(Model, levels = unique(Model)))

p <- ggplot(r2_df, aes(
  x = Model, y = pseudoR2,
  fill = factor(Quality, levels = c("Poor", "Moderate", "Good"))
)) +
  geom_col(width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", pseudoR2)), hjust = -0.15, size = 3.5) +
  scale_fill_manual(values = c(col_bad, col_warn, col_good), name = "Fit quality") +
  coord_flip() +
  ylim(0, 1) +
  labs(
    title = "McFadden Pseudo-R² Across All Models",
    subtitle = "Rules of thumb: < 0.2 = poor, 0.2–0.4 = moderate, > 0.4 = good",
    x = NULL, y = "Pseudo-R²"
  ) +
  theme_report()
print(p)

# Save plot
ggsave(file.path(plotOutputDir, "model_comparison_r2.png"), p, width = 10, height = 5, dpi = 300)
ggsave(file.path(plotOutputDir, "model_comparison_r2.svg"), p, width = 10, height = 5)
```

```{r r2-suitability-narrative, echo=FALSE, results='asis'}
r2_tiers <- lapply(names(diag_results$modelDiagnoses), function(wfn) {
  lapply(diag_results$modelDiagnoses[[wfn]], function(d) {
    list(
      model = paste0(d$sector, " ", d$stage, " (", wfn, ")"),
      r2 = d$r2, ready = d$iamReady,
      issues = length(d$extremeZTerms) > 0
    )
  })
})
r2_flat <- do.call(c, r2_tiers)
r2_good <- Filter(function(x) !is.na(x$r2) && x$r2 >= 0.4, r2_flat)
r2_mod <- Filter(function(x) !is.na(x$r2) && x$r2 >= 0.2 && x$r2 < 0.4, r2_flat)
r2_poor <- Filter(function(x) !is.na(x$r2) && x$r2 < 0.2, r2_flat)
cat("**Findings & Model Suitability Summary:**\n\n")
if (length(r2_good) > 0) {
  cat(
    "* **Top tier (Pseudo-R2 >= 0.4):**",
    paste(sapply(r2_good, `[[`, "model"), collapse = "; "),
    "-- suitable for IAM projection where no separation issues exist.\n"
  )
}
for (m in r2_mod) {
  suffix <- if (m$issues) " Despite the moderate R2, underlying separation inflates this statistic -- treat as unreliable." else ""
  cat("* **Middle tier:**", m$model, sprintf("(R2 = %.3f).", m$r2), suffix, "\n")
}
for (m in r2_poor) {
  cat(
    "* **Bottom tier:**", m$model, sprintf("(R2 = %.3f) --", m$r2),
    if (m$issues) "critical separation detected; not suitable." else "low fit; use as conservative baseline.", "\n"
  )
}
ready_models <- sapply(Filter(function(x) x$ready == "yes", r2_flat), `[[`, "model")
cat(
  "\n> **Actionable conclusion:** Use",
  if (length(ready_models) > 0) paste(ready_models, collapse = ", ") else "none",
  "for IAM deployment. Models with critical separation require structural redesign.\n"
)
```

---

# Issues, Warnings and Recommendations

## Summary Issue Table

```{r issues, echo=FALSE}
if (nrow(diag_results$issuesTable) > 0) {
  kable(diag_results$issuesTable,
    format = "html",
    caption = "Dynamically detected issues, consequences and recommended fixes",
    col.names = c("Priority", "Model", "Issue", "Consequence", "Fix")
  ) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
    column_spec(1, bold = TRUE) %>%
    collapse_rows(columns = 1)
} else {
  cat("> No issues detected across all models.")
}
```

---

# IAM Integration Notes

## What These Models Enable

The IAM-PFM module translates IAM-projected socioeconomic and energy pathways into **probabilistic carbon price trajectories** using the estimated hurdle parameters:

**1. Adoption Probability:**
$$
\Pr(\text{ECP}_t > 0) = \text{logit}^{-1}(\hat{\eta}^{(\text{adopt})}_t)
$$

**2. Conditional Stringency:**
$$
\mathbb{E}[\text{ECP}_t \mid \text{ECP}_t > 0] = \exp(\hat{\eta}^{(\text{string})}_t)
$$

**3. Expected Effective Carbon Price:**
$$
\mathbb{E}[\text{ECP}_t] = \Pr(\text{ECP}_t > 0) \times \mathbb{E}[\text{ECP}_t \mid \text{ECP}_t > 0]
$$

This decomposition allows for independent identification of policy drivers while ensuring the final ECP remains non-negative and bounded.

## Projection-Ready Models

```{r readiness, echo=FALSE}
kable(diag_results$readinessTable,
  format = "html", caption = "IAM Projection Readiness Assessment (dynamically computed)",
  col.names = c("Model", "Status", "Rationale")
) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(2,
    bold = TRUE,
    color = ifelse(grepl("\u2705", diag_results$readinessTable$Ready), col_good,
      ifelse(grepl("\u26a0", diag_results$readinessTable$Ready), col_warn, col_bad)
    )
  )
```

```{r summary-verdict, echo=FALSE, results='asis'}
rd <- diag_results$readinessTable
n_ready <- sum(grepl("\u2705", rd$Ready))
n_monitor <- sum(grepl("\u26a0", rd$Ready))
n_not <- sum(grepl("\u274c", rd$Ready))
cat("**Summary Verdict:**\n\n")
cat("Of the", nrow(rd), "model variants assessed:\n\n")
if (n_ready > 0) cat("*", n_ready, "are **fully ready** for IAM deployment.\n")
if (n_monitor > 0) cat("*", n_monitor, "**require monitoring but are usable** (widen confidence intervals or bound projections).\n")
if (n_not > 0) cat("*", n_not, "**cannot be used reliably** due to structural separation. These need fundamental redesign before deployment.\n")
cat(
  "\nFor immediate IAM integration, deploy the models marked Ready or Monitor, and replace",
  "models flagged Not Ready with externally calibrated alternatives until separation issues are resolved.\n"
)
```

## Key Projection Caveats

```{r projection-caveats, echo=FALSE, results='asis'}
cat(diag_results$projectionCaveats$asNumberedList)
```

---

# Projection Results

## Adoption Threshold {.tabset}

To determine a scientifically robust "Risk-Averse" threshold for future projections, we analyze the model's historical performance. We consider three alternatives for setting the probability at which a region is considered to have "adopted" carbon pricing:

```{r calc-thresholds, echo=FALSE}
calc_thresholds <- function(model_obj) {
  if (is.null(model_obj)) {
    return(list(Sens = 0.5, SD = 0.5, FP = 0.5))
  }

  hist_probs <- if (!is.null(model_obj$predict)) model_obj$predict else model_obj$fitted.values
  actual_outcomes <- model_obj$y

  adopters_prob <- hist_probs[actual_outcomes == 1]
  non_adopters_prob <- hist_probs[actual_outcomes == 0]

  t1 <- if (length(adopters_prob) > 0) quantile(adopters_prob, 0.10, na.rm = TRUE) else 0.5
  t2 <- if (length(adopters_prob) > 0) mean(adopters_prob, na.rm = TRUE) - sd(adopters_prob, na.rm = TRUE) else 0.5
  t3 <- if (length(non_adopters_prob) > 0) quantile(non_adopters_prob, 0.75, na.rm = TRUE) else 0.5

  return(list(
    Sens = round(max(0.25, min(0.99, t1)), 2),
    SD = round(max(0.25, min(0.99, t2)), 2),
    FP = round(max(0.25, min(0.99, t3)), 2)
  ))
}

thresh_bulk <- calc_thresholds(md2$selections$Bulk_adoption$bestModel$model)
thresh_diffuse <- calc_thresholds(md2$selections$Diffuse_adoption$bestModel$model)
```

### Bulk

- **Method 1: 90% Historical Sensitivity (`r thresh_bulk$Sens`)**: This threshold is derived from the Receiver Operating Characteristic (ROC) logic. Finds the threshold where 90% of historical adoptions would have been 'caught'. We ensure that the model successfully “identifies” 90% of actual historical policy shifts. This minimizes “False Negatives” (missing an adoption).
- **Method 2: Standard Deviation Margin (`r thresh_bulk$SD`)**: This threshold uses the variability of the “adoption signal.” By subtracting one standard deviation from the mean probability observed at the point of adoption, we include any region that falls within the “normal range” of historical early movers.
- **Method 3: False Positive Ceiling (`r thresh_bulk$FP`)**: Set at the 75th percentile of probabilities for non-adopters. Takes the risk of a false positive to ensure we don't miss the signal, while staying above the bulk of statistical noise.

### Diffuse

- **Method 1: 90% Historical Sensitivity (`r thresh_diffuse$Sens`)**: This threshold is derived from the Receiver Operating Characteristic (ROC) logic. Finds the threshold where 90% of historical adoptions would have been 'caught'. We ensure that the model successfully “identifies” 90% of actual historical policy shifts. This minimizes “False Negatives” (missing an adoption).
- **Method 2: Standard Deviation Margin (`r thresh_diffuse$SD`)**: This threshold uses the variability of the “adoption signal.” By subtracting one standard deviation from the mean probability observed at the point of adoption, we include any region that falls within the “normal range” of historical early movers.
- **Method 3: False Positive Ceiling (`r thresh_diffuse$FP`)**: Set at the 75th percentile of probabilities for non-adopters. Takes the risk of a false positive to ensure we don't miss the signal, while staying above the bulk of statistical noise.

## Projected Probability {.tabset}

```{r proj-prob, echo=FALSE, results='asis', fig.height=5, fig.width=10, warning=FALSE, message=FALSE}
if (!is.null(future_mag)) {
  get_projection_plot <- function(df, model_obj, sector_name, threshold) {
    if (is.null(model_obj)) {
      return(NULL)
    }

    needed_vars <- intersect(all.vars(model_obj$formula), colnames(df))
    df_clean <- df[complete.cases(df[, needed_vars, drop = FALSE]), ]
    if (nrow(df_clean) == 0) {
      return(NULL)
    }

    pred_link <- tryCatch(predict(model_obj, newdata = df_clean), error = function(e) NULL)
    if (is.null(pred_link)) {
      return(NULL)
    }

    df_clean$prob <- 1 / (1 + exp(-pred_link))

    top_regions <- intersect(c("USA", "CHN", "IND", "JPN", "DEU", "GBR", "FRA", "BRA", "RUS"), unique(df_clean$region))
    df_clean$highlight <- ifelse(df_clean$region %in% top_regions, as.character(df_clean$region), "Other")
    df_clean$highlight <- factor(df_clean$highlight, levels = c(top_regions, "Other"))

    my_colors <- setNames(
      c("#1f77b4", "#d62728", "#ff7f0e", "#9467bd", "#2ca02c", "#8c564b", "#e377c2", "#17becf", "#7f7f7f")[1:length(top_regions)],
      top_regions
    )
    my_colors["Other"] <- "grey80"

    p <- ggplot(df_clean, aes(x = year, y = prob, group = region, color = highlight, alpha = highlight == "Other", linewidth = highlight != "Other")) +
      geom_hline(yintercept = threshold, linetype = "dashed", color = "grey50") +
      geom_line() +
      scale_color_manual(values = my_colors) +
      scale_alpha_manual(values = c("TRUE" = 0.4, "FALSE" = 1), guide = "none") +
      scale_linewidth_manual(values = c("TRUE" = 0.5, "FALSE" = 1.2), guide = "none") +
      theme_report() +
      labs(
        title = sprintf("Projected Probability of Adoption (%s)", sector_name),
        subtitle = sprintf("The %s threshold is the minimum probability that ensures the model successfully identifies 90%% of actual historical policy shifts.", threshold),
        x = "Year", y = "Pr(Adoption)", color = "Region"
      ) +
      theme(legend.position = "right")

    # Save plot
    fname <- sprintf("prob_%s_%s", sector_name, threshold)
    ggsave(file.path(plotOutputDir, paste0(fname, ".png")), p, width = 10, height = 5, dpi = 300)
    ggsave(file.path(plotOutputDir, paste0(fname, ".svg")), p, width = 10, height = 5)

    return(p)
  }

  for (sec in c("Bulk", "Diffuse")) {
    cat("\n### ", sec, " {.tabset}\n\n")
    df <- if (sec == "Bulk") future_df_bulk else future_df_diffuse
    model_obj <- md2$selections[[paste0(sec, "_adoption")]]$bestModel$model
    t_list <- if (sec == "Bulk") thresh_bulk else thresh_diffuse

    for (t_name in c("Sens", "SD", "FP")) {
      cat("\n#### ", t_name, "\n\n")
      p <- get_projection_plot(df, model_obj, sec, t_list[[t_name]])
      if (!is.null(p)) print(p)
      cat("\n\n")
    }
  }
}
```

## Expected Adoption Timeline {.tabset}

```{r timeline-tables, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
if (!is.null(future_mag)) {
  render_adoption_timeline <- function(df, model_obj, threshold) {
    if (is.null(model_obj)) {
      return(NULL)
    }

    needed_vars <- intersect(all.vars(model_obj$formula), colnames(df))
    df_clean <- df[complete.cases(df[, needed_vars, drop = FALSE]), ]
    if (nrow(df_clean) == 0) {
      return(NULL)
    }

    pred_link <- tryCatch(predict(model_obj, newdata = df_clean), error = function(e) NULL)
    if (is.null(pred_link)) {
      return(NULL)
    }

    df_clean$prob <- 1 / (1 + exp(-pred_link))

    adopt_data <- df_clean %>%
      group_by(region) %>%
      summarise(
        AdoptionYear = ifelse(any(prob >= threshold), min(year[prob >= threshold]), Inf),
        .groups = "drop"
      ) %>%
      mutate(
        Category = case_when(
          AdoptionYear <= 2030 ~ "Forerunners",
          AdoptionYear <= 2050 ~ "Early Majority",
          AdoptionYear <= 2070 ~ "Late Majority",
          TRUE ~ "Laggards"
        ),
        YearRange = case_when(
          Category == "Forerunners" ~ "<= 2030",
          Category == "Early Majority" ~ "2031–2050",
          Category == "Late Majority" ~ "2051–2070",
          Category == "Laggards" ~ "> 2070",
        ),
        RegYear = ifelse(AdoptionYear == Inf, region, sprintf("%s (%s)", region, AdoptionYear))
      )

    final_table <- adopt_data %>%
      group_by(Category, YearRange) %>%
      summarise(Regions = paste(RegYear, collapse = ", "), .groups = "drop") %>%
      mutate(Category = factor(Category, levels = c(
        "Forerunners", "Early Majority", "Late Majority", "Laggards"
      ))) %>%
      arrange(Category)

    cat(as.character(
      kable(final_table, format = "html", col.names = c("Category", "Expected Adoption Year", "Regions")) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
    ))
    cat("\n\n")
  }

  for (sec in c("Bulk", "Diffuse")) {
    cat("\n### ", sec, " {.tabset}\n\n")
    df <- if (sec == "Bulk") future_df_bulk else future_df_diffuse
    model_obj <- md2$selections[[paste0(sec, "_adoption")]]$bestModel$model
    t_list <- if (sec == "Bulk") thresh_bulk else thresh_diffuse

    for (t_name in c("Sens", "SD", "FP")) {
      cat("\n#### ", t_name, "\n\n")
      render_adoption_timeline(df, model_obj, t_list[[t_name]])
      cat("\n\n")
    }
  }
}
```

## Spatial Distribution of Adoption {.tabset}

```{r spatial-maps, echo=FALSE, results='asis', fig.height=6, fig.width=10, warning=FALSE, message=FALSE}
if (!is.null(future_mag)) {
  mapping_54 <- tryCatch(
    {
      madrat::toolGetMapping("regionmapping_54.csv", type = "regional", where = "mappingfolder")
    },
    error = function(e) NULL
  )

  if (!is.null(mapping_54)) {
    suppressPackageStartupMessages(library(sf))
    world <- rnaturalearth::ne_countries(scale = "small", returnclass = "sf")
    world$iso_key <- as.character(world$adm0_a3)

    render_adoption_map <- function(df, model_obj, sector_name, threshold) {
      needed_vars <- intersect(all.vars(model_obj$formula), colnames(df))
      df_clean <- df[complete.cases(df[, needed_vars, drop = FALSE]), ]
      pred_link <- tryCatch(predict(model_obj, newdata = df_clean), error = function(e) NULL)

      if (is.null(pred_link)) {
        return(NULL)
      }
      df_clean$prob <- 1 / (1 + exp(-pred_link))

      adopt_data <- df_clean %>%
        group_by(region) %>%
        summarise(
          AdoptionYear = ifelse(any(prob >= threshold), min(year[prob >= threshold]), 2151),
          .groups = "drop"
        ) %>%
        mutate(
          Category = case_when(
            AdoptionYear <= 2030 ~ "Forerunners",
            AdoptionYear <= 2050 ~ "Early Majority",
            AdoptionYear <= 2070 ~ "Late Majority",
            TRUE ~ "Laggards"
          )
        )

      map_data_df <- mapping_54 %>%
        mutate(
          RegionCode = as.character(RegionCode),
          CountryCode = as.character(CountryCode)
        ) %>%
        mutate(CountryCode = ifelse(CountryCode == "SSD", "SDS", CountryCode)) %>%
        left_join(adopt_data, by = c("RegionCode" = "region")) %>%
        mutate(Category = if_else(CountryCode == "ATA", factor(NA, levels = levels(Category)), Category)) %>%
        mutate(Category = factor(Category, levels = c("Forerunners", "Early Majority", "Late Majority", "Laggards")))

      world_data <- world %>% left_join(map_data_df, by = c("iso_key" = "CountryCode"))

      p <- ggplot() +
        geom_sf(data = world_data[world_data$iso_key == "ATA", ], fill = "grey90", color = "white", size = 0.05) +
        geom_sf(data = world_data[world_data$iso_key != "ATA", ], aes(fill = Category), color = "white", size = 0.05) +
        scale_fill_manual(values = c(
          "Forerunners" = "#2ecc71",
          "Early Majority" = "#82e0aa",
          "Late Majority" = "#f7dc6f",
          "Laggards" = "#e67e22"
        ), na.value = "grey90", na.translate = FALSE) +
        theme_minimal() +
        theme(
          legend.position = "bottom",
          panel.grid = element_blank(),
          axis.text = element_blank()
        ) +
        labs(
          title = sprintf("Projected Adoption Timeline Categories (%s)", sector_name),
          subtitle = sprintf("Threshold: Pr(Adoption) > %s", threshold),
          fill = "Projected Adoption Timeline"
        )

      print(p)

      # Save plot
      fname <- sprintf("map_%s_%s", sector_name, threshold)
      ggsave(file.path(plotOutputDir, paste0(fname, ".png")), width = 10, height = 6, dpi = 300)
      ggsave(file.path(plotOutputDir, paste0(fname, ".svg")), width = 10, height = 6)
    }

    for (sec in c("Bulk", "Diffuse")) {
      cat("### ", sec, " {.tabset}\n\n")
      df <- if (sec == "Bulk") future_df_bulk else future_df_diffuse
      model_obj <- md2$selections[[paste0(sec, "_adoption")]]$bestModel$model
      t_list <- if (sec == "Bulk") thresh_bulk else thresh_diffuse

      for (t_name in c("Sens", "SD", "FP")) {
        cat("#### ", t_name, "\n\n")
        render_adoption_map(df, model_obj, sec, t_list[[t_name]])
        cat("\n\n")
      }
    }
  } else {
    cat("> ⚠️ **Could not load region mapping file.** Ensure `regionmapping_54.csv` is in the mapping folder.\n")
  }
}
```


---

# Appendix {.tabset}

## Selection Paths (md)

```{r selection-paths-md, echo=FALSE}
sp_md <- md$selectionPaths %>%
  select(sector, stage, step, description, pseudoR2, aic, nSignificant, nPredictors, converged)
kable(sp_md,
  format = "html", caption = "Model Selection Path — md (Full AP Drivers)",
  digits = 3
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
```

## Selection Paths (md2)

```{r selection-paths-md2, echo=FALSE}
sp_md2 <- md2$selectionPaths %>%
  select(sector, stage, step, description, pseudoR2, aic, nSignificant, nPredictors, converged)
kable(sp_md2,
  format = "html", caption = "Model Selection Path — md2 (API Only)",
  digits = 3
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
```

## Selection Paths (md3)

```{r selection-paths-md3, echo=FALSE}
if (!is.null(md3$selectionPaths)) {
  sp_md3 <- md3$selectionPaths %>%
    select(sector, stage, step, description, pseudoR2, aic, nSignificant, nPredictors, converged)
  kable(sp_md3,
    format = "html", caption = "Model Selection Path — md3 (REMIND Region FE)",
    digits = 3
  ) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
} else {
  cat("No data available.")
}
```

## Selection Paths (md4)

```{r selection-paths-md4, echo=FALSE}
if (!is.null(md4$selectionPaths)) {
  sp_md4 <- md4$selectionPaths %>%
    select(sector, stage, step, description, pseudoR2, aic, nSignificant, nPredictors, converged)
  kable(sp_md4,
    format = "html", caption = "Model Selection Path — md4 (Lagged ECP)",
    digits = 3
  ) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
} else {
  cat("No data available.")
}
```

## Best Coefficients (md)

```{r best-coefficients-md, echo=FALSE}
kable(md$bestCoefficients,
  format = "html", caption = "Best Model Coefficients — md",
  digits = 4
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
```

## Best Coefficients (md2)

```{r best-coefficients-md2, echo=FALSE}
kable(md2$bestCoefficients,
  format = "html", caption = "Best Model Coefficients — md2",
  digits = 4
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
```

## Best Coefficients (md3)

```{r best-coefficients-md3, echo=FALSE}
if (!is.null(md3$bestCoefficients)) {
  kable(md3$bestCoefficients,
    format = "html", caption = "Best Model Coefficients — md3",
    digits = 4
  ) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
} else {
  cat("No data available.")
}
```

## Best Coefficients (md4)

```{r best-coefficients-md4, echo=FALSE}
if (!is.null(md4$bestCoefficients)) {
  kable(md4$bestCoefficients,
    format = "html", caption = "Best Model Coefficients — md4",
    digits = 4
  ) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
} else {
  cat("No data available.")
}
```

## Session Information

```{r session, echo=FALSE}
tryCatch(
  {
    print(sessionInfo())
  },
  error = function(e) {
    cat("Note: sessionInfo() could not be fully rendered due to a package metadata error.\n")
    cat("R Version:", R.version.string, "\n")
    cat("Platform:", R.version$platform, "\n")
  }
)
```
