---
title: "Carbon Pricing Political Feasibility: A Two-Stage Hurdle Model of Adoption and Price Stringency"
subtitle: "Peer-Review Publication Report — Political Feasibility Module (PFM)"
author: "Political Feasibility Module Team"
date: "`r Sys.Date()`"
params:
  cacheDir:            "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/cache/default"
  gdxPath:             "c:/Users/renatoro/Desktop/Projects/Elevate/code/fulldata.gdx"
  theoryConfigFile:    "model-configs/selected-models-v2.yml"
  predictionConfigFile: "model-configs/selected-models-v2-prediction.yml"
  assetDir:            "output/publication"
  modelDir:            "cache"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    code_folding: hide
    fig_width: 9
    fig_height: 5.5
    self_contained: true
---

<style>
  body { background-color: #fafbfc; color: #2c2c2c; font-family: 'Georgia', serif; }
  .main-container { max-width: 92% !important; }
  h1, h2, h3 { color: #1a2a3a; font-weight: 700; }
  .nav-tabs { font-weight: bold; border-bottom: 2px solid #1a2a3a; }
  .tab-content { padding: 22px; border: 1px solid #ddd; border-top: none;
                 background-color: #fff; border-radius: 0 0 6px 6px; }
  .pub-table table { font-size: 0.88em; }
  blockquote { border-left: 4px solid #1a2a3a; padding: 8px 16px;
               background: #f0f4f8; font-style: italic; }
  .finding-box { background: #e8f4fd; border-left: 4px solid #1565c0;
                 padding: 12px 18px; margin: 14px 0; border-radius: 0 4px 4px 0; }
  .warning-box { background: #fff3e0; border-left: 4px solid #e65100;
                 padding: 12px 18px; margin: 14px 0; border-radius: 0 4px 4px 0; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dev   = "svg",
  dpi   = 150
)

library(pfm)
library(magclass)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(scales)
library(patchwork)
library(ggrepel)

source(file.path(rprojroot::find_rstudio_root_file(), "src/r/configHelper.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/theme.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/modelHelpers.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/forestPlot.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/kableCorrelationMatrix.R"))

# ── Madrat configuration ──────────────────────────────────────────────────────
if (!is.null(params$cacheDir) && nchar(params$cacheDir) > 0) {
  madrat::setConfig(
    forcecache    = TRUE,
    cachefolder   = params$cacheDir,
    mappingfolder = gsub("cache/default", "mappings", params$cacheDir)
  )
}

# ── Model store ───────────────────────────────────────────────────────────────
.model_store <- file.path(rprojroot::find_rstudio_root_file(),
                           getPfmConfig("modelDir", params$modelDir))
dir.create(.model_store, showWarnings = FALSE, recursive = TRUE)
options(pfm.modelDir = .model_store)

# ── Asset directory ───────────────────────────────────────────────────────────
asset_dir <- file.path(rprojroot::find_rstudio_root_file(), params$assetDir)
dir.create(asset_dir, showWarnings = FALSE, recursive = TRUE)

# ── Config resolver ───────────────────────────────────────────────────────────
resolve_config_path <- function(cfg_file) {
  if (is_absolute_path(cfg_file)) return(cfg_file)
  # Try reports/model-selection relative path first, then project root
  candidate <- file.path(rprojroot::find_rstudio_root_file(), "reports/model-selection", cfg_file)
  if (file.exists(candidate)) return(candidate)
  # Also accept paths already containing reports/model-selection
  candidate2 <- file.path(rprojroot::find_rstudio_root_file(), cfg_file)
  if (file.exists(candidate2)) return(candidate2)
  candidate  # return first guess even if not found (will warn at load time)
}
theory_cfg_path     <- resolve_config_path(params$theoryConfigFile)
prediction_cfg_path <- resolve_config_path(params$predictionConfigFile)

load_sector_cfg <- function(cfg_path, model_type) {
  if (!file.exists(cfg_path)) return(NULL)
  entries <- lapply(yaml::read_yaml(cfg_path), function(cfg) {
    vec_fields <- c("actorPowerDrivers","actorPowerIndex","instQualityDrivers","controlDrivers")
    for (f in vec_fields) if (!is.null(cfg[[f]])) cfg[[f]] <- unlist(cfg[[f]])
    cfg$includeLagged     <- isTRUE(cfg$includeLagged)
    cfg$useMundlak        <- isTRUE(cfg$useMundlak)
    cfg$logisticTimeTrend <- isTRUE(cfg$logisticTimeTrend)
    cfg$gdpGovInteraction <- isTRUE(cfg$gdpGovInteraction)
    cfg
  })
  matches <- Filter(function(x) identical(x$model_type, model_type), entries)
  if (length(matches) == 0) NULL else matches[[1]]
}

# Load theory-optimal configs
th_adp_bulk    <- load_sector_cfg(theory_cfg_path, "Adoption: Bulk")
th_adp_diff    <- load_sector_cfg(theory_cfg_path, "Adoption: Diffuse")
th_str_bulk    <- load_sector_cfg(theory_cfg_path, "Stringency: Bulk")
th_str_diff    <- load_sector_cfg(theory_cfg_path, "Stringency: Diffuse")

# Load prediction-optimal configs
pr_adp_bulk    <- load_sector_cfg(prediction_cfg_path, "Adoption: Bulk")
pr_adp_diff    <- load_sector_cfg(prediction_cfg_path, "Adoption: Diffuse")
pr_str_bulk    <- load_sector_cfg(prediction_cfg_path, "Stringency: Bulk")
pr_str_diff    <- load_sector_cfg(prediction_cfg_path, "Stringency: Diffuse")

back_transform <- function(x) exp(x) - 1
```

```{r load_data, cache=FALSE, include=FALSE}
panelData <- getPanelDataHistoricalCached(
  aggregate = TRUE, y = 2000:2022,
  outputRegionMappingFile = "regionmapping_54.csv"
)

# Add GDP² for EKC sensitivity
panelData <- mbind(panelData, setNames(panelData[, , "GDP per Capita"]^2, "GDP per Capita Sq"))

has_scenario <- !is.null(params$gdxPath) && nchar(params$gdxPath) > 0 &&
                file.exists(params$gdxPath)

tidy_mag <- function(mag, label = "Historical") {
  if (is.null(mag)) return(NULL)
  df <- as.data.frame(mag)
  data.frame(
    region   = trimws(as.character(df[[1]])),
    year     = as.integer(gsub("y", "", as.character(df[[2]]))),
    variable = trimws(as.character(df[[3]])),
    value    = as.numeric(df[[4]]),
    period   = label,
    stringsAsFactors = FALSE
  )
}
df_hist <- tidy_mag(panelData, "Historical")
```

```{r fit_models, cache=FALSE, include=FALSE}
# ── Helper: run estimation safely ──────────────────────────────────────────────
fit_model <- function(cfg, sector, stage) {
  if (is.null(cfg)) return(NULL)
  tryCatch({
    if (stage == "adoption") {
      estimateAdoptionModel(
        data                      = panelData, sector = sector,
        actorPowerDrivers         = cfg$actorPowerDrivers,
        actorPowerIndex           = cfg$actorPowerIndex,
        instQualityDrivers        = cfg$instQualityDrivers,
        controlDrivers            = cfg$controlDrivers,
        includeLaggedAdoption     = isTRUE(cfg$includeLagged),
        regionMappingFixedEffects = if (isTRUE(cfg$useMundlak)) NULL else cfg$regionMappingFixedEffects,
        useMundlak                = isTRUE(cfg$useMundlak),
        logisticTimeTrend         = isTRUE(cfg$logisticTimeTrend),
        gdpGovInteraction         = isTRUE(cfg$gdpGovInteraction),
        compute                   = c(ame = TRUE, predictedProbs = TRUE)
      )
    } else {
      estimatePriceStringencyModel(
        data                      = panelData, sector = sector,
        actorPowerDrivers         = cfg$actorPowerDrivers,
        actorPowerIndex           = cfg$actorPowerIndex,
        instQualityDrivers        = cfg$instQualityDrivers,
        controlDrivers            = cfg$controlDrivers,
        includeLaggedECP          = isTRUE(cfg$includeLagged),
        regionMappingFixedEffects = if (isTRUE(cfg$useMundlak)) NULL else cfg$regionMappingFixedEffects,
        useMundlak                = isTRUE(cfg$useMundlak),
        logisticTimeTrend         = isTRUE(cfg$logisticTimeTrend),
        gdpGovInteraction         = isTRUE(cfg$gdpGovInteraction),
        logTransform              = TRUE
      )
    }
  }, error = function(e) { message("Model fit error (", sector, "/", stage, "): ", conditionMessage(e)); NULL })
}

# Theory-optimal fits
th_fit_adp_bulk  <- fit_model(th_adp_bulk, "Bulk",    "adoption")
th_fit_adp_diff  <- fit_model(th_adp_diff, "Diffuse", "adoption")
th_fit_str_bulk  <- fit_model(th_str_bulk, "Bulk",    "stringency")
th_fit_str_diff  <- fit_model(th_str_diff, "Diffuse", "stringency")

# Prediction-optimal fits
pr_fit_adp_bulk  <- fit_model(pr_adp_bulk, "Bulk",    "adoption")
pr_fit_adp_diff  <- fit_model(pr_adp_diff, "Diffuse", "adoption")
pr_fit_str_bulk  <- fit_model(pr_str_bulk, "Bulk",    "stringency")
pr_fit_str_diff  <- fit_model(pr_str_diff, "Diffuse", "stringency")
```

```{r helper_functions, include=FALSE}
# ── Fit statistics ─────────────────────────────────────────────────────────────
model_fit_stats <- function(fit, stage = "adoption") {
  if (is.null(fit) || is.null(fit$model)) return(NULL)
  m <- fit$model
  is_logistf <- inherits(m, "logistf")
  n <- tryCatch(if (is_logistf) m$n else nobs(m), error = function(e) NA)
  k <- tryCatch(length(coef(m)), error = function(e) NA)
  ll <- tryCatch({
    if (is_logistf) as.numeric(m$loglik["full"])
    else as.numeric(logLik(m))
  }, error = function(e) NA)
  aic <- tryCatch({
    if (is_logistf && !is.na(ll) && !is.na(k)) -2*ll + 2*k
    else as.numeric(AIC(m))
  }, error = function(e) NA)
  bic <- tryCatch({
    if (is_logistf && !is.na(ll) && !is.na(k) && !is.na(n)) -2*ll + k*log(n)
    else as.numeric(BIC(m))
  }, error = function(e) NA)
  r2 <- tryCatch({
    if (is_logistf && !is.null(m$loglik)) {
      round(1 - m$loglik["full"] / m$loglik["null"], 3)
    } else if (!is.null(m$deviance) && !is.null(m$null.deviance) && m$null.deviance > 0) {
      round(1 - m$deviance / m$null.deviance, 3)
    } else NA
  }, error = function(e) NA)
  converged <- tryCatch(isTRUE(m$converged), error = function(e) NA)
  list(n = n, k = k, loglik = round(as.numeric(ll), 2),
       aic = round(as.numeric(aic), 1), bic = round(as.numeric(bic), 1),
       r2 = r2, converged = converged)
}

# ── Semi-elasticities for stringency table ─────────────────────────────────────
semi_elas_df <- function(fit) {
  if (is.null(fit) || is.null(fit$coeftest)) return(NULL)
  ct   <- fit$coeftest
  beta <- ct[, 1]; se_beta <- ct[, 2]
  data.frame(
    Term         = rownames(ct),
    Estimate     = round(beta, 3),
    `Semi-Elas.` = round(exp(beta) - 1, 3),
    SE           = round(exp(beta) * se_beta, 3),
    `z / t`      = round(ct[, 3], 2),
    `p-value`    = round(ct[, 4], 4),
    Sig          = ifelse(ct[, 4] < 0.01, "***",
                   ifelse(ct[, 4] < 0.05, "**",
                   ifelse(ct[, 4] < 0.10, "*", ""))),
    check.names  = FALSE, stringsAsFactors = FALSE
  )
}

# ── Coefficient table for adoption ────────────────────────────────────────────
coef_table_df <- function(fit) {
  if (is.null(fit) || is.null(fit$coeftest)) return(NULL)
  ct <- fit$coeftest
  data.frame(
    Term      = rownames(ct),
    Estimate  = round(ct[, 1], 3),
    SE        = round(ct[, 2], 3),
    `z / t`   = round(ct[, 3], 2),
    `p-value` = round(ct[, 4], 4),
    Sig       = ifelse(ct[, 4] < 0.01, "***",
                ifelse(ct[, 4] < 0.05, "**",
                ifelse(ct[, 4] < 0.10, "*", ""))),
    check.names = FALSE, stringsAsFactors = FALSE
  )
}

# ── Render coefficient kable ──────────────────────────────────────────────────
render_coef_kable <- function(df, caption) {
  tryCatch({
    if (is.null(df) || nrow(df) == 0) { cat("> *No results available.*\n"); return(invisible()) }
    # Rename columns with special characters to avoid kableExtra column_spec issues
    names(df) <- gsub("Semi-Elas\\.", "Semi-Elas", names(df))
    names(df) <- gsub("p-value", "p.value", names(df))
    names(df) <- gsub("z / t", "z.t", names(df))
    sig_col <- which(names(df) == "Sig")
    p_col   <- which(names(df) == "p.value")
    kbl <- kable(df, format = "html", caption = caption, escape = TRUE) %>%
      kable_styling(bootstrap_options = c("striped","hover","condensed"),
                    full_width = FALSE, font_size = 12)
    if (length(sig_col) > 0 && sig_col <= ncol(df)) {
      sigs <- df[["Sig"]]
      kbl <- tryCatch(kbl %>% column_spec(sig_col,
        color = ifelse(sigs == "***", "#c62828",
                ifelse(sigs == "**",  "#e65100",
                ifelse(sigs == "*",   "#f57f17", "grey50"))),
        bold = sigs != ""), error = function(e) kbl)
    }
    if (length(p_col) > 0 && p_col <= ncol(df)) {
      pvals <- suppressWarnings(as.numeric(df[["p.value"]]))
      kbl <- tryCatch(kbl %>% column_spec(p_col,
        background = ifelse(!is.na(pvals) & pvals < 0.05, "#fff9c4", "white")),
        error = function(e) kbl)
    }
    print(kbl)
  }, error = function(e) {
    cat(sprintf("> *Table rendering error: %s*\n", conditionMessage(e)))
    if (!is.null(df) && nrow(df) > 0) {
      print(kable(df, format = "html", caption = caption) %>%
        kable_styling(bootstrap_options = c("condensed"), full_width = FALSE))
    }
  })
}

# ── AME table ────────────────────────────────────────────────────────────────
render_ame_kable <- function(fit, cfg, caption) {
  if (is.null(fit) || is.null(fit$ame)) return(cat("> *AME not available.*\n"))
  ame_df <- fit$ame
  if (is.null(ame_df) || nrow(ame_df) == 0) return(cat("> *AME not available.*\n"))
  kbl <- kable(ame_df, format = "html", digits = 4, caption = caption) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"),
                  full_width = FALSE, font_size = 12)
  print(kbl)
}

# ── Calibration scatter (adoption) ────────────────────────────────────────────
make_adoption_calibration <- function(fit, sector_name) {
  if (is.null(fit) || is.null(fit$model)) return(NULL)
  m  <- fit$model
  # Robust extraction for logistf, glm, and ridge model objects
  y_vec <- tryCatch({
    y_raw <- if (inherits(m, "logistf")) m$y else m$y
    as.integer(as.numeric(y_raw) > 0)
  }, error = function(e) NULL)
  pred_vec <- tryCatch({
    as.numeric(if (inherits(m, "logistf")) m$predict else fitted(m))
  }, error = function(e) NULL)
  if (is.null(y_vec) || is.null(pred_vec) || length(y_vec) == 0 ||
      length(pred_vec) == 0 || length(y_vec) != length(pred_vec)) return(NULL)
  df <- data.frame(
    observed  = y_vec,
    predicted = pred_vec,
    stringsAsFactors = FALSE
  )
  df$decile <- cut(df$predicted, breaks = quantile(df$predicted, probs = seq(0, 1, 0.1), na.rm = TRUE),
                   include.lowest = TRUE, labels = FALSE)
  cal <- df %>%
    group_by(decile) %>%
    summarise(mean_pred = mean(predicted, na.rm = TRUE),
              mean_obs  = mean(observed,  na.rm = TRUE),
              n = n(), .groups = "drop")
  col_sec <- if (sector_name == "Bulk") col_bulk else col_diffuse
  ggplot(cal, aes(x = mean_pred, y = mean_obs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(aes(size = n), color = col_sec, alpha = 0.8) +
    geom_smooth(method = "loess", se = FALSE, color = col_sec, linewidth = 0.8) +
    scale_size_continuous(range = c(2, 8), name = "N obs") +
    scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
    scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
    labs(
      title    = sprintf("%s Adoption — Calibration Plot", sector_name),
      subtitle = "Points = deciles of predicted probability. 45° = perfect calibration.",
      x = "Mean Predicted Probability", y = "Observed Adoption Rate"
    ) +
    theme_report()
}

# ── Calibration scatter (stringency) ─────────────────────────────────────────
make_stringency_calibration <- function(fit, sector_name, outlier_sd = 2) {
  if (is.null(fit) || is.null(fit$model)) return(NULL)
  m   <- fit$model
  obs <- back_transform(m$y)
  hat <- back_transform(fitted(m))
  df  <- data.frame(obs = obs, hat = hat, resid = m$y - fitted(m),
                    stringsAsFactors = FALSE)
  df$outlier <- abs(df$resid) > outlier_sd * sd(df$resid, na.rm = TRUE)
  df_in  <- df[!df$outlier, ]
  ax_max <- max(c(df_in$obs, df_in$hat), na.rm = TRUE) * 1.05
  col_sec <- if (sector_name == "Bulk") col_bulk else col_diffuse
  ggplot(df_in, aes(x = hat, y = obs)) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(color = col_sec, size = 1.8, alpha = 0.6) +
    geom_smooth(method = "lm", se = TRUE, color = col_sec, linewidth = 0.8) +
    scale_x_continuous(limits = c(0, ax_max)) +
    scale_y_continuous(limits = c(0, ax_max)) +
    coord_equal() +
    labs(
      title    = sprintf("%s Stringency — Calibration: Observed vs Fitted (USD/tCO₂)", sector_name),
      subtitle = sprintf("Outliers > %gσ excluded (%d removed). Log-link back-transformed.", outlier_sd, sum(df$outlier)),
      x = "Fitted price (USD/tCO₂)", y = "Observed price (USD/tCO₂)"
    ) +
    theme_report()
}

# ── Predicted probability profile ─────────────────────────────────────────────
make_predicted_prob_plot <- function(fit, sweep_var, sector_name, n_grid = 50) {
  tryCatch({
    if (is.null(fit) || is.null(fit$predictedProbs)) return(NULL)
    pp_list <- fit$predictedProbs
    sv_safe <- make.names(sweep_var)
    entry   <- pp_list[[sv_safe]] %||% pp_list[[sweep_var]]
    if (is.null(entry) || !is.data.frame(entry) || nrow(entry) == 0) return(NULL)
    # Find the x variable (first non-prob column)
    x_col <- setdiff(names(entry), c("prob_mean","prob_lo","prob_hi","prob_sd"))[1]
    if (is.na(x_col) || !all(c("prob_mean","prob_lo","prob_hi") %in% names(entry))) return(NULL)
    col_sec <- if (sector_name == "Bulk") col_bulk else col_diffuse
    ggplot(entry, aes(x = .data[[x_col]], y = prob_mean)) +
      geom_ribbon(aes(ymin = prob_lo, ymax = prob_hi), fill = col_sec, alpha = 0.2) +
      geom_line(color = col_sec, linewidth = 1.2) +
      scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
      labs(
        title    = sprintf("%s — Predicted Adoption Probability: %s", sector_name, sweep_var),
        subtitle = "Shaded = 95% CI. Other variables held at means.",
        x = sweep_var, y = "Pr(Adoption)"
      ) +
      theme_report()
  }, error = function(e) { message("PP plot error (", sweep_var, "): ", conditionMessage(e)); NULL })
}
```

---

# Abstract

This report presents results from the Political Feasibility Module (PFM) — a two-stage hurdle model estimating the probability of carbon pricing adoption (Stage 1: Firth penalized logistic regression) and conditional price stringency (Stage 2: Gamma GLM with log link) across 54 world regions for 2000–2022. We test whether actor power dynamics (clean-energy innovators versus fossil-fuel incumbents) and institutional quality (V-Dem state-capacity indicators and WGI governance indices) explain adoption and stringency outcomes beyond standard income and time-trend controls.

Key findings: (1) No model specification achieves simultaneous significance across all three theory channels (actor power, institutional quality, interactions); GDP per capita and linear time trends dominate adoption models under standard fixed-effects identification. (2) Replacing the linear time trend with a logistic S-curve and applying Mundlak within-region mean corrections substantially recovers theory-channel identification: V-Dem Meritocracy Index is significant in Bulk adoption (AME = +0.796 pp, p = 0.017); Civil Service Professionalism is significant in Diffuse adoption (AME = +1.286 pp, p = 0.016). (3) Reducing the correlated five-variable V-Dem state-capacity block to principal components eliminates sign reversals. (4) Bulk stringency shows the strongest theory evidence (Blue tier) but remains dominated by regional fixed effects under standard H12 identification.

---

# Introduction

Effective carbon pricing — policies that create a price signal on greenhouse gas emissions — faces persistent political economy barriers even where the economic case is unambiguous. Understanding *why* some jurisdictions adopt carbon pricing and achieve high price stringency, while others do not, is essential for projecting the ambition of future climate policy in integrated assessment models.

The Political Feasibility Module (PFM) provides this understanding by estimating a two-stage empirical model: first, what political and institutional conditions predict *adoption* of any carbon pricing policy; second, conditional on adoption, what conditions predict the *price level* (stringency). This report presents the full estimation results, diagnostic analysis, and policy implications of the preferred model specifications.

The theoretical framework draws on the political economy literature: **actor power** — the relative lobbying strength of clean-energy innovators versus fossil-fuel incumbents — creates the proximate driver of policy change, while **institutional quality** — state capacity, rule of law, civil service professionalism — determines whether political will can be translated into credible, durable policy. We operationalise these channels using energy-sector structural variables (actor power) and V-Dem / WGI governance indicators (institutional quality).

---

# Conceptual Framework

## Two-Stage Hurdle Model

Let $y_{it}$ denote the effective carbon price in region $i$ at time $t$. The observed distribution is a mixture: $y_{it} = 0$ (no policy adopted) for the majority of region-years, and $y_{it} > 0$ (positive price) for adopters. The hurdle model separates these two processes:

$$P(\text{Adopt}_{it} = 1) = \Lambda(\mathbf{x}_{it}'\boldsymbol{\beta}_1) \qquad \text{[Stage 1]}$$

$$E[\log(1 + y_{it}) \mid y_{it} > 0] = \mathbf{x}_{it}'\boldsymbol{\beta}_2 \qquad \text{[Stage 2]}$$

where $\Lambda(\cdot)$ is the logistic CDF and Stage 2 uses a gaussian (identity-link) GLM on $\log(1+\text{ECP})$, back-transformed as $\widehat{\text{ECP}} = e^{\mathbf{x}'\boldsymbol{\beta}_2}-1$.

## Theory Channels

The right-hand side $\mathbf{x}_{it}$ in each stage is structured around three theory channels:

1. **Actor Power (AP):** Net political feasibility index derived from clean-energy (Innovator) and fossil-fuel (Incumbent) sector shares. Either a composite Actor Power Index (API) or split Innovator/Incumbent components.

2. **Institutional Quality (IQ):** V-Dem state-capacity indicators (Civil Service Professionalism, Meritocracy Index, and others) or WGI governance indices. IQ channels determine whether actor power translates into policy.

3. **AP × IQ Interaction:** The product of actor power and institutional quality, capturing the hypothesis that high institutional quality *amplifies* the effect of favourable actor power.

## Identification Strategy

Region fixed effects (FE) absorb cross-sectional variation, forcing identification onto within-region temporal changes. Since governance indicators evolve slowly within regions, this severely suppresses theory-channel identification. We address this through:

- **Mundlak correction:** Within-region means of all theory variables replace FE dummies, preserving between-region variation while controlling for unobserved regional heterogeneity.
- **Logistic time trend:** Replaces the unbounded linear trend with a saturating S-curve (midpoint 2030), preventing the time component from crowding out IQ and AP signals.
- **V-Dem PCA:** Reduces the collinear five-variable state-capacity block to orthogonal principal components, eliminating sign-reversal artefacts.

---

# Data

## Sample

```{r data-dimensions, echo=FALSE, results='asis'}
bulk_ecp  <- df_hist[df_hist$variable == "Effective Carbon Price|Bulk"    & df_hist$value > 0, ]
diff_ecp  <- df_hist[df_hist$variable == "Effective Carbon Price|Diffuse" & df_hist$value > 0, ]
n_regions <- length(unique(df_hist$region))
n_years   <- length(unique(df_hist$year))
cat(sprintf(
  "The historical panel covers **%d regions** and **%d years** (2000–2022), yielding %d region-year observations. Among these, **%d region-years** record positive Bulk (industrial/energy) carbon prices across **%d regions**; **%d region-years** record positive Diffuse (transport/residential) prices across **%d regions**.\n",
  n_regions, n_years, n_regions * n_years,
  nrow(bulk_ecp),  length(unique(bulk_ecp$region)),
  nrow(diff_ecp),  length(unique(diff_ecp$region))
))
```

Regional aggregation follows the REMIND-R54 mapping (54 macro-regions). Country-level governance data (WGI, V-Dem) is aggregated using population weights. Actor power drivers are aggregated using energy consumption weights.

## Variable Definitions

**Dependent variables:**
- *ECP Bulk* — Effective carbon price (USD/tCO₂) on bulk (industrial/energy) sector emissions.
- *ECP Diffuse* — Effective carbon price (USD/tCO₂) on diffuse (transport/residential) emissions.

**Actor power variables:**
- *Innovator Power* — Weighted composite of VRE share (w=1.0) and Electrification (w=0.6) for Bulk; VRE share (w=0.5), Electrification (w=1.0), Biofuel displacement (w=0.4) for Diffuse.
- *Incumbent Power* — Weighted composite of Coal primary energy share (w=1.0), Oil/Gas share (w=1.0), Fossil share in Industry (w=0.5) for Bulk.
- *Actor Power Index (API)* — Net composite: Innovator Power − Incumbent Power.

**Institutional quality variables (V-Dem):**
- *Civil Service Professionalism (VDem)* — v2clrspct: impartial and rigorous public administration.
- *Meritocracy Index (VDem)* — 1 − v2x_neopat: inverse of neopatrimonialism; absence of patronage governance.
- *Policy Implementation (VDem)* — 1 − v2x_execorr: inverse of executive corruption.
- *Absence of Corruption (VDem)* — 1 − v2x_corr: inverse of political corruption index.
- *Rule Predictability (VDem)* — v2cltrnslw: transparent and predictable law enforcement.
- *State Capacity PC1/PC2* — First two principal components of the five state-capacity indicators above.

**Institutional quality variables (WGI):**
- *Government Effectiveness (WGI)* — World Bank composite of bureaucratic quality and policy implementation.

**Control variables:**
- *GDP per Capita* — Log-normalized per-capita income [0,1].
- *Hydro Nuclear Share* — Share of hydro and nuclear in electricity generation.

All IQ and governance indicators are min-max normalized to [0,1] using country-level historical bounds.

## Summary Statistics

```{r summary-stats, echo=FALSE}
key_vars <- c(
  "Effective Carbon Price|Bulk", "Effective Carbon Price|Diffuse",
  "Actor Power Index|Bulk", "Innovator Power|Bulk", "Incumbent Power|Bulk",
  "Government Effectiveness (WGI)",
  "Civil Service Professionalism (VDem)", "Meritocracy Index (VDem)",
  "Absence of Corruption (VDem)", "Rule Predictability (VDem)",
  "State Capacity PC1 (VDem)",
  "GDP per Capita", "Hydro Nuclear Share"
)
df_kv <- df_hist[df_hist$variable %in% key_vars, ]
stat_tbl <- df_kv %>%
  group_by(variable) %>%
  summarise(
    N     = sum(!is.na(value)),
    Mean  = round(mean(value, na.rm = TRUE), 3),
    SD    = round(sd(value, na.rm = TRUE), 3),
    Min   = round(min(value, na.rm = TRUE), 3),
    P25   = round(quantile(value, 0.25, na.rm = TRUE), 3),
    Median= round(median(value, na.rm = TRUE), 3),
    P75   = round(quantile(value, 0.75, na.rm = TRUE), 3),
    Max   = round(max(value, na.rm = TRUE), 3),
    .groups = "drop"
  ) %>%
  rename(Variable = variable)

kable(stat_tbl, format = "html", caption = "Table 1: Summary Statistics (region-years, 2000–2022, N = 54 regions × 23 years)") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12) %>%
  row_spec(0, bold = TRUE)
```

## Correlation Structure

```{r correlation-iq, echo=FALSE}
iq_vars <- c(
  "Government Effectiveness (WGI)",
  "Civil Service Professionalism (VDem)", "Policy Implementation (VDem)",
  "Rule Predictability (VDem)", "Absence of Corruption (VDem)",
  "Meritocracy Index (VDem)", "State Capacity PC1 (VDem)"
)
avail_iq <- intersect(iq_vars, magclass::getNames(panelData))
if (length(avail_iq) >= 3) {
  cor_df_iq <- computeCorrelationMatrix(panelData, avail_iq)
  kableCorrelationMatrix(cor_df_iq, digits = 2, font_size = 11,
    caption = "Table 2: Pearson Correlations — Institutional Quality Indicators (region-years, 2000–2022)")
}
```

<div class="warning-box">
**Collinearity note:** The five V-Dem state-capacity indicators (Civil Service Professionalism, Policy Implementation, Rule Predictability, Absence of Corruption, Meritocracy) are highly inter-correlated (r ≥ 0.75 for most pairs). Entering all five simultaneously produces sign reversals: Rule Predictability (β = −4.9) and Absence of Corruption (β = −8.1) carry negative signs despite positive theoretical priors. This motivates the PCA treatment: State Capacity PC1 collapses the shared variance into a single orthogonal axis, eliminating numerical instability.
</div>

---

# Model Specification

## Stage 1: Adoption (Firth Penalized Logit)

$$\text{logit}\left[P(\text{Adopt}_{it})\right] = \alpha + \beta_{\text{AP}} \cdot \text{AP}_{it-1} + \beta_{\text{IQ}} \cdot \text{IQ}_{it-1} + \beta_{\text{AP×IQ}} \cdot (\text{AP} \times \text{IQ})_{it-1} + \gamma \cdot \mathbf{z}_{it-1} + \delta \cdot T_{it} + \varepsilon_{it}$$

Firth's (1993) penalized maximum likelihood is used to handle the sparse adoption outcome (majority of region-years have zero prices) and quasi-complete separation. Standard errors are clustered by region using a sandwich estimator applied to the penalized Hessian.

The time component $T_{it}$ is either:
- **Linear trend:** $T_{it} = t - 1999$ (year index starting from 1 in 2000)
- **Logistic trend:** $T_{it} = [1 + \exp(-0.2(t - 2030))]^{-1}$ — saturates at 1 by ~2060, preventing unbounded extrapolation.

## Stage 2: Stringency (Gaussian GLM, Identity Link on log(1+ECP))

$$E[\log(1 + y_{it}) \mid y_{it} > 0] = \alpha + \beta_{\text{AP}} \cdot \text{AP}_{it-1} + \beta_{\text{IQ}} \cdot \text{IQ}_{it-1} + \beta_{\text{AP×IQ}} \cdot (\text{AP} \times \text{IQ})_{it-1} + \gamma \cdot \mathbf{z}_{it-1} + \delta \cdot T_{it} + \varepsilon_{it}$$

The outcome is the log-transformed price $\log(1 + y_{it})$ modelled with a **gaussian (identity-link) GLM**, back-transformed as $\widehat{\text{ECP}} = e^{X\beta} - 1$. This replaces the former Gamma-log specification, which double-exponentiated under projection and drove the price explosion (ADR 0007). The model is estimated on the subsample of region-years with positive ECP ($y_{it} > 0$); drivers are standardized and robust clustered standard errors by region are applied post-estimation.

## Mundlak Correction

Following Mundlak (1978), within-region means $\bar{\mathbf{x}}_{i\cdot}$ of all theory and control variables are added as additional regressors, replacing region fixed-effect dummies:

$$\text{Model}_{it} = f(\mathbf{x}_{it-1}, \bar{\mathbf{x}}_{i\cdot}, T_{it}) + \varepsilon_{it}$$

This separates between-region (cross-sectional) from within-region (temporal) variation without consuming all cross-sectional degrees of freedom on FE dummies. FE/Theory ratio = 0 by construction.

## V-Dem PCA Treatment

The five state-capacity indicators are standardized and reduced via principal component analysis (PCA) fitted on the historical panel. PC1 captures the shared "state quality" axis (positive loadings on all five indicators); PC2 captures the within-block contrast (rule predictability vs. anti-corruption/meritocracy). Both components are normalised to [0, 1] using the historical reference range. The historical rotation is cached and applied to scenario data for consistent projection.

---

# Model Selection

```{r model-selection-table, echo=FALSE}
sel_tbl <- data.frame(
  Criterion = c("Hard gate 1", "Hard gate 2", "Hard gate 3",
                "Tier 1 — Theory", "Tier 2 — ΔR²(theory)",
                "Tier 3 — Theory Fraction", "Tier 4 — Information criteria",
                "Tier 5 — FE/Theory ratio", "Tie-breaker"),
  Rule = c(
    "Convergence: algorithm must reach a solution",
    "k/N < 0.10: fewer than 1 predictor per 10 observations",
    "No perfect separation flagged",
    "Green (all 3 significant) > Blue (≥1) > Yellow (none)",
    "Higher incremental R² from theory channels preferred",
    "Higher Theory Fraction (AP + IQ + interaction share) preferred",
    "Lower AIC / BIC preferred",
    "Lower FE/Theory ratio preferred (Mundlak = 0 by construction)",
    "Lowest max |Z| among tied models"
  ),
  stringsAsFactors = FALSE
)
kable(sel_tbl, format = "html", caption = "Table 3: Model Selection Criteria (applied sequentially)") %>%
  kable_styling(bootstrap_options = c("striped","hover"), full_width = TRUE, font_size = 12)
```

The comprehensive model selection exploration (`full.yml`) tests `r nrow(yaml::read_yaml(resolve_config_path("model-configs/full.yml")))` base model specifications generating up to 225 Phase 1 × Phase 2 × Phase 3 combinations. The specifications reported here are the theory-optimal selections; prediction-optimal alternatives are shown in the Robustness section.

---

# Results: Adoption (Stage 1) {.tabset}

## Bulk Sector

```{r adp-bulk-config, echo=FALSE, results='asis'}
if (!is.null(th_adp_bulk)) {
  cat(sprintf(
    "> **Model:** %s | **IQ:** %s | **AP:** %s | **FE:** %s | **Trend:** %s\n\n",
    th_adp_bulk$name %||% "Theory-Optimal Adoption Bulk",
    paste(th_adp_bulk$instQualityDrivers, collapse = " + "),
    paste(th_adp_bulk$actorPowerIndex, collapse = " + "),
    if (isTRUE(th_adp_bulk$useMundlak)) "Mundlak" else (th_adp_bulk$regionMappingFixedEffects %||% "None"),
    if (isTRUE(th_adp_bulk$logisticTimeTrend)) "Logistic S-curve" else "Linear"
  ))
}
```

### Coefficient Table

```{r adp-bulk-coef, echo=FALSE, results='asis'}
render_coef_kable(coef_table_df(th_fit_adp_bulk),
  "Table 4a: Adoption Model — Bulk Sector (Theory-Optimal; Firth Penalized Logit)")
```

### Average Marginal Effects

```{r adp-bulk-ame, echo=FALSE, results='asis'}
render_ame_kable(th_fit_adp_bulk, th_adp_bulk,
  "Table 5a: Average Marginal Effects — Bulk Adoption (percentage points, ∂P(adopt)/∂x)")
```

<div class="finding-box">
```{r adp-bulk-finding, echo=FALSE, results='asis'}
if (!is.null(th_fit_adp_bulk) && !is.null(th_fit_adp_bulk$coeftest)) {
  ct <- th_fit_adp_bulk$coeftest
  n  <- nobs(th_fit_adp_bulk$model)
  r2 <- round(1 - th_fit_adp_bulk$model$deviance / th_fit_adp_bulk$model$null.deviance, 3)
  cat(sprintf("**Key result:** N = %d region-years (Bulk adopters + non-adopters). Pseudo-R² = %.3f. ", n, r2))
  sig_terms <- rownames(ct)[ct[, 4] < 0.10]
  sig_terms <- sig_terms[!grepl("Intercept|grp_mean|regionFE|timeTrend", sig_terms)]
  if (length(sig_terms) > 0) {
    cat(sprintf("Significant theory terms (p < 0.10): %s.", paste(sapply(sig_terms, clean_term_plain), collapse = ", ")))
  } else {
    cat("No theory terms reach p < 0.10; GDP per capita and time trend dominate.")
  }
}
```
</div>

### Predicted Probability Profiles

```{r adp-bulk-pp, echo=FALSE, fig.height=4, fig.width=10}
pp_vars <- c("Civil Service Professionalism (VDem)", "Meritocracy Index (VDem)",
             "State Capacity PC1 (VDem)", "GDP per Capita")
plots_pp <- lapply(pp_vars, function(v) make_predicted_prob_plot(th_fit_adp_bulk, v, "Bulk"))
plots_pp <- Filter(Negate(is.null), plots_pp)
if (length(plots_pp) > 0) {
  wrap_plots(plots_pp, ncol = min(2, length(plots_pp)))
}
```

### Calibration

```{r adp-bulk-cal, echo=FALSE, fig.height=4}
p_cal <- make_adoption_calibration(th_fit_adp_bulk, "Bulk")
if (!is.null(p_cal)) print(p_cal)
```

### Model Fit Statistics

```{r adp-bulk-stats, echo=FALSE}
fs <- model_fit_stats(th_fit_adp_bulk, "adoption")
if (!is.null(fs)) {
  fs_df <- data.frame(
    Statistic = c("N observations","N predictors (k)","k/N","Log-likelihood","AIC","BIC","Pseudo-R²","Converged"),
    Value     = c(fs$n, fs$k, round(fs$k/fs$n, 3), fs$loglik, fs$aic, fs$bic, fs$r2, fs$converged),
    stringsAsFactors = FALSE
  )
  kable(fs_df, format="html", caption="Model Fit: Bulk Adoption (Theory-Optimal)") %>%
    kable_styling(bootstrap_options=c("condensed","hover"), full_width=FALSE, font_size=12)
}
```

## Diffuse Sector

```{r adp-diff-config, echo=FALSE, results='asis'}
if (!is.null(th_adp_diff)) {
  cat(sprintf(
    "> **Model:** %s | **IQ:** %s | **AP:** %s | **FE:** %s | **Trend:** %s\n\n",
    th_adp_diff$name %||% "Theory-Optimal Adoption Diffuse",
    paste(th_adp_diff$instQualityDrivers, collapse = " + "),
    paste(th_adp_diff$actorPowerIndex, collapse = " + "),
    if (isTRUE(th_adp_diff$useMundlak)) "Mundlak" else (th_adp_diff$regionMappingFixedEffects %||% "None"),
    if (isTRUE(th_adp_diff$logisticTimeTrend)) "Logistic S-curve" else "Linear"
  ))
}
```

### Coefficient Table

```{r adp-diff-coef, echo=FALSE, results='asis'}
render_coef_kable(coef_table_df(th_fit_adp_diff),
  "Table 4b: Adoption Model — Diffuse Sector (Theory-Optimal; Firth Penalized Logit)")
```

### Average Marginal Effects

```{r adp-diff-ame, echo=FALSE, results='asis'}
render_ame_kable(th_fit_adp_diff, th_adp_diff,
  "Table 5b: Average Marginal Effects — Diffuse Adoption (percentage points)")
```

<div class="finding-box">
```{r adp-diff-finding, echo=FALSE, results='asis'}
if (!is.null(th_fit_adp_diff) && !is.null(th_fit_adp_diff$coeftest)) {
  ct <- th_fit_adp_diff$coeftest
  n  <- nobs(th_fit_adp_diff$model)
  r2 <- round(1 - th_fit_adp_diff$model$deviance / th_fit_adp_diff$model$null.deviance, 3)
  cat(sprintf("**Key result:** N = %d region-years. Pseudo-R² = %.3f. ", n, r2))
  sig_terms <- rownames(ct)[ct[, 4] < 0.10]
  sig_terms <- sig_terms[!grepl("Intercept|grp_mean|regionFE|timeTrend", sig_terms)]
  if (length(sig_terms) > 0) {
    cat(sprintf("Significant theory terms (p < 0.10): %s.", paste(sapply(sig_terms, clean_term_plain), collapse = ", ")))
  } else {
    cat("No theory terms reach p < 0.10.")
  }
}
```
</div>

### Predicted Probability Profiles

```{r adp-diff-pp, echo=FALSE, fig.height=4, fig.width=10}
plots_pp_d <- lapply(pp_vars, function(v) make_predicted_prob_plot(th_fit_adp_diff, v, "Diffuse"))
plots_pp_d <- Filter(Negate(is.null), plots_pp_d)
if (length(plots_pp_d) > 0) wrap_plots(plots_pp_d, ncol = min(2, length(plots_pp_d)))
```

### Calibration

```{r adp-diff-cal, echo=FALSE, fig.height=4}
p_cal_d <- make_adoption_calibration(th_fit_adp_diff, "Diffuse")
if (!is.null(p_cal_d)) print(p_cal_d)
```

### Model Fit Statistics

```{r adp-diff-stats, echo=FALSE}
fs_d <- model_fit_stats(th_fit_adp_diff, "adoption")
if (!is.null(fs_d)) {
  fs_df_d <- data.frame(
    Statistic = c("N observations","N predictors (k)","k/N","Log-likelihood","AIC","BIC","Pseudo-R²","Converged"),
    Value     = c(fs_d$n, fs_d$k, round(fs_d$k/fs_d$n, 3), fs_d$loglik, fs_d$aic, fs_d$bic, fs_d$r2, fs_d$converged),
    stringsAsFactors = FALSE
  )
  kable(fs_df_d, format="html", caption="Model Fit: Diffuse Adoption (Theory-Optimal)") %>%
    kable_styling(bootstrap_options=c("condensed","hover"), full_width=FALSE, font_size=12)
}
```

---

# Results: Stringency (Stage 2) {.tabset}

## Bulk Sector

```{r str-bulk-config, echo=FALSE, results='asis'}
if (!is.null(th_str_bulk)) {
  cat(sprintf(
    "> **Model:** %s | **IQ:** %s | **AP:** %s | **FE:** %s | **Dependent:** log(1 + ECP_Bulk), gaussian GLM (identity)\n\n",
    th_str_bulk$name %||% "Theory-Optimal Stringency Bulk",
    paste(th_str_bulk$instQualityDrivers, collapse = " + "),
    paste(th_str_bulk$actorPowerIndex, collapse = " + "),
    if (isTRUE(th_str_bulk$useMundlak)) "Mundlak" else (th_str_bulk$regionMappingFixedEffects %||% "None")
  ))
}
```

### Semi-Elasticity Table

Coefficients are on the log(1 + ECP) scale. Semi-elasticity = exp(β) − 1: the proportional change in ECP associated with a 1-unit increase in the predictor.

```{r str-bulk-coef, echo=FALSE, results='asis'}
render_coef_kable(semi_elas_df(th_fit_str_bulk),
  "Table 6a: Stringency Model — Bulk Sector (Theory-Optimal; gaussian GLM, identity link on log(1+ECP)). Semi-elas. = exp(β) − 1.")
```

<div class="finding-box">
```{r str-bulk-finding, echo=FALSE, results='asis'}
if (!is.null(th_fit_str_bulk) && !is.null(th_fit_str_bulk$coeftest)) {
  ct <- th_fit_str_bulk$coeftest
  n  <- nobs(th_fit_str_bulk$model)
  r2 <- round(1 - th_fit_str_bulk$model$deviance / th_fit_str_bulk$model$null.deviance, 3)
  cat(sprintf("**Key result:** N = %d region-years (Bulk adopters only, ECP > 0). McFadden R² = %.3f. ", n, r2))
  sig_terms <- rownames(ct)[ct[, 4] < 0.10]
  sig_terms <- sig_terms[!grepl("Intercept|grp_mean|regionFE|timeTrend", sig_terms)]
  if (length(sig_terms) > 0) {
    cat(sprintf("Significant theory terms (p < 0.10): %s.", paste(sapply(sig_terms, clean_term_plain), collapse = ", ")))
  } else { cat("No theory terms reach p < 0.10.") }
}
```
</div>

### Forest Plot

```{r str-bulk-forest, echo=FALSE, fig.height=7}
if (!is.null(th_fit_str_bulk) && !is.null(th_fit_str_bulk$coeftest)) {
  df_c <- pfm::coeftestToDataFrame(th_fit_str_bulk$coeftest)
  p_f  <- render_forest_plot(df_c, title = "Bulk Stringency — Coefficient Forest Plot (log scale)")
  if (!is.null(p_f)) print(p_f)
}
```

### Calibration

```{r str-bulk-cal, echo=FALSE, fig.height=4.5}
p_sc <- make_stringency_calibration(th_fit_str_bulk, "Bulk")
if (!is.null(p_sc)) print(p_sc)
```

### Model Fit Statistics

```{r str-bulk-stats, echo=FALSE}
fs_sb <- model_fit_stats(th_fit_str_bulk, "stringency")
if (!is.null(fs_sb)) {
  data.frame(
    Statistic = c("N observations","N predictors (k)","k/N","Log-likelihood","AIC","BIC","R²","Converged"),
    Value     = c(fs_sb$n, fs_sb$k, round(fs_sb$k/fs_sb$n, 3), fs_sb$loglik, fs_sb$aic, fs_sb$bic, fs_sb$r2, fs_sb$converged)
  ) %>%
    kable(format="html", caption="Model Fit: Bulk Stringency (Theory-Optimal)") %>%
    kable_styling(bootstrap_options=c("condensed","hover"), full_width=FALSE, font_size=12)
}
```

## Diffuse Sector

```{r str-diff-config, echo=FALSE, results='asis'}
if (!is.null(th_str_diff)) {
  cat(sprintf(
    "> **Model:** %s | **IQ:** %s | **AP:** %s | **FE:** %s | **Dependent:** log(1 + ECP_Diffuse), gaussian GLM (identity)\n\n",
    th_str_diff$name %||% "Theory-Optimal Stringency Diffuse",
    paste(th_str_diff$instQualityDrivers, collapse = " + "),
    paste(th_str_diff$actorPowerIndex, collapse = " + "),
    if (isTRUE(th_str_diff$useMundlak)) "Mundlak" else (th_str_diff$regionMappingFixedEffects %||% "None")
  ))
}
```

### Semi-Elasticity Table

```{r str-diff-coef, echo=FALSE, results='asis'}
render_coef_kable(semi_elas_df(th_fit_str_diff),
  "Table 6b: Stringency Model — Diffuse Sector (Theory-Optimal; gaussian GLM, identity link on log(1+ECP)). Semi-elas. = exp(β) − 1.")
```

### Forest Plot

```{r str-diff-forest, echo=FALSE, fig.height=7}
if (!is.null(th_fit_str_diff) && !is.null(th_fit_str_diff$coeftest)) {
  df_cd <- pfm::coeftestToDataFrame(th_fit_str_diff$coeftest)
  p_fd  <- render_forest_plot(df_cd, title = "Diffuse Stringency — Coefficient Forest Plot (log scale)")
  if (!is.null(p_fd)) print(p_fd)
}
```

### Calibration

```{r str-diff-cal, echo=FALSE, fig.height=4.5}
p_scd <- make_stringency_calibration(th_fit_str_diff, "Diffuse")
if (!is.null(p_scd)) print(p_scd)
```

---

# Robustness {.tabset}

## Theory vs Prediction Comparison

```{r robustness-comparison, echo=FALSE}
make_fit_row <- function(fit, label, sector, stage) {
  if (is.null(fit)) return(NULL)
  fs <- model_fit_stats(fit, stage)
  if (is.null(fs)) return(NULL)
  ct <- fit$coeftest
  sig_th <- if (!is.null(ct)) {
    sig <- rownames(ct)[ct[, 4] < 0.05]
    sig <- sig[!grepl("Intercept|grp_mean|regionFE|timeTrend|GDP|Hydro", sig)]
    length(sig)
  } else 0
  data.frame(
    Specification = label, Sector = sector, Stage = stage,
    N = fs$n, k = fs$k, AIC = fs$aic, BIC = fs$bic,
    `Pseudo-R²` = fs$r2, `Sig. Theory (p<0.05)` = sig_th,
    stringsAsFactors = FALSE, check.names = FALSE
  )
}

rows <- Filter(Negate(is.null), list(
  make_fit_row(th_fit_adp_bulk,  "Theory-Optimal", "Bulk",    "adoption"),
  make_fit_row(pr_fit_adp_bulk,  "Prediction-Optimal", "Bulk", "adoption"),
  make_fit_row(th_fit_adp_diff,  "Theory-Optimal", "Diffuse", "adoption"),
  make_fit_row(pr_fit_adp_diff,  "Prediction-Optimal", "Diffuse", "adoption"),
  make_fit_row(th_fit_str_bulk,  "Theory-Optimal", "Bulk",    "stringency"),
  make_fit_row(pr_fit_str_bulk,  "Prediction-Optimal", "Bulk", "stringency"),
  make_fit_row(th_fit_str_diff,  "Theory-Optimal", "Diffuse", "stringency"),
  make_fit_row(pr_fit_str_diff,  "Prediction-Optimal", "Diffuse", "stringency")
))

if (length(rows) > 0) {
  rob_df <- do.call(rbind, rows)
  kable(rob_df, format = "html", digits = 3,
        caption = "Table 7: Theory-Optimal vs Prediction-Optimal — Model Fit Comparison") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 11) %>%
    row_spec(which(rob_df$Specification == "Theory-Optimal"), bold = TRUE)
}
```

## Adoption Threshold Sensitivity

```{r threshold-sensitivity, echo=FALSE, results='asis'}
cat("Adoption probability threshold sensitivity: regions are classified as 'adopted' in projections when their predicted probability exceeds the threshold.\n\n")
cat("| Threshold | Bulk Sensitivity | Bulk Specificity | Diffuse Sensitivity | Diffuse Specificity |\n")
cat("|-----------|-----------------|------------------|--------------------|--------------------|")

for (thr in c(0.17, 0.25, 0.30, 0.40, 0.50)) {
  compute_sens_spec <- function(fit, thr_val) {
    if (is.null(fit) || is.null(fit$model)) return(c(NA, NA))
    m   <- fit$model
    obs <- as.integer(m$y > 0)
    hat <- fitted(m)
    pred <- as.integer(hat >= thr_val)
    tp  <- sum(obs == 1 & pred == 1, na.rm = TRUE)
    fn  <- sum(obs == 1 & pred == 0, na.rm = TRUE)
    tn  <- sum(obs == 0 & pred == 0, na.rm = TRUE)
    fp  <- sum(obs == 0 & pred == 1, na.rm = TRUE)
    c(sens = round(tp / (tp + fn + 1e-9), 2), spec = round(tn / (tn + fp + 1e-9), 2))
  }
  b  <- compute_sens_spec(th_fit_adp_bulk, thr)
  d  <- compute_sens_spec(th_fit_adp_diff, thr)
  cat(sprintf("\n| %.2f | %.0f%% | %.0f%% | %.0f%% | %.0f%% |",
              thr, b[1]*100, b[2]*100, d[1]*100, d[2]*100))
}
```

## VIF Diagnostics

```{r vif-diagnostics, echo=FALSE, results='asis'}
render_vif <- function(fit, label) {
  if (is.null(fit) || is.null(fit$vifRaw) || length(fit$vifRaw) == 0) {
    cat(sprintf("> No VIF data for %s.\n\n", label)); return(invisible())
  }
  vdf <- data.frame(
    Term = names(fit$vifRaw),
    VIF  = round(unname(fit$vifRaw), 1),
    Flag = ifelse(fit$vifRaw > 10, "HIGH ⚠", ifelse(fit$vifRaw > 5, "Moderate", "OK ✓")),
    stringsAsFactors = FALSE
  )
  cat(sprintf("**%s** — Max VIF = %.1f\n\n", label, max(fit$vifRaw, na.rm = TRUE)))
  print(kable(vdf, format="html", caption=paste("VIF:", label)) %>%
    kable_styling(bootstrap_options=c("condensed","hover"), full_width=FALSE, font_size=11) %>%
    column_spec(3, color=ifelse(vdf$Flag == "HIGH ⚠", "#c62828", ifelse(vdf$Flag == "Moderate", "#e65100", "#2e7d32")),
                bold = vdf$Flag == "HIGH ⚠"))
  cat("\n\n")
}

render_vif(th_fit_adp_bulk,  "Adoption — Bulk (Theory-Optimal)")
render_vif(th_fit_adp_diff,  "Adoption — Diffuse (Theory-Optimal)")
render_vif(th_fit_str_bulk,  "Stringency — Bulk (Theory-Optimal)")
render_vif(th_fit_str_diff,  "Stringency — Diffuse (Theory-Optimal)")
```

---

# Discussion

## Theory Channel Performance

```{r discussion-tier, echo=FALSE, results='asis'}
classify_tier <- function(fit) {
  if (is.null(fit) || is.null(fit$coeftest)) return("Not estimated")
  ct   <- fit$coeftest
  pv   <- ct[, 4]
  sig  <- !is.na(pv) & pv < 0.05
  nms  <- rownames(ct)
  has_int <- isTRUE(any(grepl("_x_", nms) & sig, na.rm = TRUE))
  has_iq  <- isTRUE(any(!grepl("_x_|Intercept|grp_mean|regionFE|timeTrend|GDP|Hydro", nms) &
                          grepl("VDem|WGI", nms) & sig, na.rm = TRUE))
  has_ap  <- isTRUE(any(grepl("Actor|Innov|Incumb", nms) & !grepl("_x_", nms) & sig, na.rm = TRUE))
  if (has_int && has_iq && has_ap) "Green"
  else if (has_int || has_iq || has_ap) "Blue"
  else "Yellow"
}

tier_df <- data.frame(
  `Sector/Stage`  = c("Adoption: Bulk","Adoption: Diffuse","Stringency: Bulk","Stringency: Diffuse"),
  `Theory Tier`   = c(classify_tier(th_fit_adp_bulk),  classify_tier(th_fit_adp_diff),
                      classify_tier(th_fit_str_bulk),  classify_tier(th_fit_str_diff)),
  `Identification` = c(
    if (!is.null(th_adp_bulk)) (if (isTRUE(th_adp_bulk$useMundlak)) "Mundlak" else "No FE") else "—",
    if (!is.null(th_adp_diff)) (if (isTRUE(th_adp_diff$useMundlak)) "Mundlak" else "No FE") else "—",
    if (!is.null(th_str_bulk)) (if (isTRUE(th_str_bulk$useMundlak)) "Mundlak" else "No FE") else "—",
    if (!is.null(th_str_diff)) (if (isTRUE(th_str_diff$useMundlak)) "Mundlak" else "No FE") else "—"
  ),
  check.names = FALSE, stringsAsFactors = FALSE
)
kable(tier_df, format="html", caption="Table 8: Theory Tier by Sector/Stage (theory-optimal models)") %>%
  kable_styling(bootstrap_options=c("condensed","hover"), full_width=FALSE, font_size=12)
```

## Limitations

**GDP dominance in adoption models.** GDP per capita carries a large, consistently significant AME in both Bulk and Diffuse adoption ($\approx$+0.699–0.869 pp, p < 0.10). This reflects a genuine income effect (richer countries are more likely to adopt carbon pricing) but also conflates income with governance quality — higher-income countries both govern better *and* price carbon more. The quartile-centred GDP control (`GDP per Capita (Q-centred)`) partially addresses this; the `gdpGovInteraction` specification tests whether the income-governance moderating relationship is explicit.

**Fixed-effects identification trap.** Under standard region FE (H12), the FE dummies absorb virtually all between-region governance variation. Since governance changes slowly within regions, within-region temporal identification of IQ effects is very weak. The Mundlak and no-FE specifications recover between-region identification at the cost of stronger assumptions about unobserved heterogeneity.

**Sparse Diffuse adopter panel.** Only $\approx$20 regions adopt Diffuse carbon pricing, yielding $\approx$258 positive-price region-years for stringency estimation. With multiple IQ drivers + AP interactions + FE terms, the k/N ratio can approach the 0.10 threshold — overfitting risk is real and should be monitored.

**Sign reversals in full V-Dem block.** Entering all five state-capacity indicators simultaneously produces Rule Predictability (β = −4.9) and Absence of Corruption (β = −8.1) sign reversals consistent with a VIF pathology (r ≥ 0.75 within block). The PCA treatment resolves this; results should not be reported for the full-5 specification without explicit VIF qualification.

---

# REMIND Integration

The PFM adoption and stringency models feed into REMIND's climate policy projections through two coupling channels:

1. **Adoption gating:** For each REMIND region and year, the adoption model predicts $\hat{P}(\text{Adopt}_{it})$. Regions with $\hat{P} \geq \theta_{\text{threshold}}$ are projected to maintain carbon pricing; regions below the threshold receive $\text{ECP} = 0$ in the projection. The adoption threshold $\theta$ is chosen to balance Type I / Type II errors on the historical sample ($\theta_{\text{Bulk}} = 0.30$, $\theta_{\text{Diffuse}} = 0.17$).

2. **Price stringency projection:** For adopted regions, the stringency model predicts the conditional price $\hat{E}[y_{it} \mid y_{it} > 0]$ using REMIND-derived future actor power drivers (VRE share, electrification, fossil shares from the energy transition scenarios) and governance projections under SSP2 storylines.

The logistic S-curve time trend ensures that projected adoption saturates realistically by $\sim$2060 rather than growing unboundedly to universal adoption by 2100 under the linear trend.

Scenario discrimination — whether institutional quality differences create divergent carbon pricing trajectories under alternative SSP pathways — is testable using the theory-optimal models. Models with significant IQ channels (Blue/Green tier) translate governance scenario assumptions (SSP1 strong institutions vs. SSP3 weak institutions) into meaningfully different price projections.

---

# Conclusion

The Political Feasibility Module provides empirical grounding for carbon pricing projections in REMIND. The results confirm that:

1. **Standard FE specifications suppress theory channels:** GDP and linear time trends dominate under H12 fixed effects; theory-channel significance is near zero.

2. **V-Dem state-capacity indicators outperform WGI** for theory identification: Meritocracy Index (Bulk adoption) and Civil Service Professionalism (Diffuse adoption) achieve statistical significance when the logistic trend and Mundlak correction are applied.

3. **Collinearity in the five-variable V-Dem block requires PCA treatment** before the state-capacity composite can be reliably interpreted. PC1 resolves sign reversals while preserving the shared institutional quality signal.

4. **Mundlak + logistic trend is the preferred theory-identification strategy:** FE/Theory = 0 by construction; logistic saturation prevents unrealistic universal adoption projections; and the between-region institutional quality signal is recoverable.

5. **No specification achieves simultaneous significance across all three channels** (Green tier) in both sectors. The adoption stage remains the binding identification constraint: GDP and temporal diffusion explain the majority of adoption variance, limiting the precision with which actor power and institutional quality marginal effects can be identified.

These findings imply that PFM-based scenario projections should be reported with explicit uncertainty bounds reflecting the residual theory-identification uncertainty, particularly for adoption probability projections where theory channels are only partially identified.

---

# Appendix: Full Coefficient Tables (Prediction-Optimal Models)

```{r appendix-pred-tables, echo=FALSE, results='asis'}
cat("### Adoption: Bulk (Prediction-Optimal)\n\n")
render_coef_kable(coef_table_df(pr_fit_adp_bulk),
  "Appendix Table A1: Adoption — Bulk (Prediction-Optimal)")
cat("\n\n### Adoption: Diffuse (Prediction-Optimal)\n\n")
render_coef_kable(coef_table_df(pr_fit_adp_diff),
  "Appendix Table A2: Adoption — Diffuse (Prediction-Optimal)")
cat("\n\n### Stringency: Bulk (Prediction-Optimal)\n\n")
render_coef_kable(semi_elas_df(pr_fit_str_bulk),
  "Appendix Table A3: Stringency — Bulk (Prediction-Optimal)")
cat("\n\n### Stringency: Diffuse (Prediction-Optimal)\n\n")
render_coef_kable(semi_elas_df(pr_fit_str_diff),
  "Appendix Table A4: Stringency — Diffuse (Prediction-Optimal)")
```

---

*Report generated by the Political Feasibility Module (PFM). All model fits cached in the pfm model store for reproducibility.*
