---
title: "`r params$modelName` — PFM Adoption Model Report"
subtitle: "Input Data, Estimation, and Future Projections for Bulk & Diffuse Adoption"
author: "Political Feasibility Module"
date: "`r Sys.Date()`"
params:
  cacheDir: "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/cache/default"
  gdxPath: "c:/Users/renatoro/Desktop/Projects/Elevate/code/fulldata.gdx"
  assetDir: "output/adoption_model"
  modelName: "Baseline + Rule of Law"
  instQualityDrivers: !r c("Government Effectiveness (WGI)", "Rule of Law (VDem)")
  controlDrivers: !r c("GDP per Capita", "Hydro Nuclear Share")
  regionMappingFE: "regionmappingH12.csv"
  includeLagged: false
  adoptionThreshold: 0.5
  snapshotYears: !r c(2030, 2040, 2050, 2070, 2100)
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    theme: flatly
    highlight: tango
    number_sections: true
    code_folding: hide
    fig_width: 10
    fig_height: 6
---

<style>
  body { background-color: #f5f7fa; color: #333; }
  .main-container { max-width: 95% !important; }
  .toc-content { padding-left: 0px !important; }
  h1, h2, h3 { color: #2C3E50; font-weight: 700; }
  .nav-tabs {
    font-weight: bold;
    margin-bottom: 20px;
    border-bottom: 2px solid #2C3E50;
  }
  .nav-tabs > li.active > a,
  .nav-tabs > li.active > a:focus,
  .nav-tabs > li.active > a:hover {
    background-color: #2C3E50; color: #fff;
    border: 1px solid #2C3E50; border-bottom-color: transparent;
  }
  .tab-content {
    padding: 25px; border: 1px solid #ddd; border-top: none;
    border-radius: 0 0 6px 6px; background-color: #fff;
    box-shadow: 0 4px 6px rgba(0,0,0,0.05);
  }
  .card {
    background-color: #fff; border-radius: 8px;
    border: 1px solid #eef2f5; box-shadow: 0 4px 6px rgba(0,0,0,0.02);
    margin-bottom: 25px;
  }
  .card-header {
    background-color: #f8fafc; border-bottom: 1px solid #eef2f5;
    padding: 12px 20px; font-weight: bold; font-size: 1.1em; color: #2C3E50;
  }
  .card-body { padding: 20px; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE,
  fig.align = "center", dev = "svg"
)

library(pfm)
library(magclass)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(scales)

source(file.path(rprojroot::find_rstudio_root_file(), "src/r/configHelper.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/theme.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/forestPlot.R"))

if (!is.null(params$cacheDir) && nchar(params$cacheDir) > 0) {
  madrat::setConfig(
    forcecache    = TRUE,
    cachefolder   = params$cacheDir,
    mappingfolder = gsub("cache/default", "mappings", params$cacheDir)
  )
}

plotOutputDir <- file.path(params$assetDir, "plots")
if (!dir.exists(plotOutputDir)) dir.create(plotOutputDir, recursive = TRUE)

model_instQualityDrivers <- params$instQualityDrivers
model_controlDrivers     <- params$controlDrivers
model_regionMappingFE    <- if (is.null(params$regionMappingFE) ||
                                  nchar(as.character(params$regionMappingFE)) == 0)
                              NULL else params$regionMappingFE
model_includeLagged      <- isTRUE(params$includeLagged)
adoption_threshold       <- as.numeric(params$adoptionThreshold)
snapshot_years           <- as.integer(params$snapshotYears)

# Variables shown in the input data section (only those used by this model)
model_vars_input <- c(
  "Actor Power Index|Bulk",
  "Actor Power Index|Diffuse",
  model_instQualityDrivers,
  model_controlDrivers
)

# Resolve gdxPath as an absolute path relative to the project root if it is relative
gdx_path <- params$gdxPath
if (!is.null(gdx_path) && nchar(gdx_path) > 0 && !is_absolute_path(gdx_path)) {
  gdx_path <- file.path(rprojroot::find_rstudio_root_file(), gdx_path)
}
has_scenario <- !is.null(gdx_path) && nchar(gdx_path) > 0 && file.exists(gdx_path)

tidy_magpie <- function(mag_obj, period_label = "Historical") {
  if (is.null(mag_obj)) return(NULL)
  df      <- as.data.frame(mag_obj)
  reg_col <- if ("Region" %in% colnames(df)) "Region" else colnames(df)[1]
  yr_col  <- if ("Year"   %in% colnames(df)) "Year"   else colnames(df)[2]
  var_col <- if ("Data1"  %in% colnames(df)) "Data1"  else colnames(df)[3]
  val_col <- if ("Value"  %in% colnames(df)) "Value"  else colnames(df)[4]
  data.frame(
    region   = trimws(as.character(df[[reg_col]])),
    year     = as.integer(gsub("y", "", as.character(df[[yr_col]]))),
    variable = trimws(as.character(df[[var_col]])),
    value    = as.numeric(df[[val_col]]),
    period   = period_label,
    stringsAsFactors = FALSE
  )
}

key_regions <- c("USA", "CHA", "DEU", "IND", "JPN", "BRA", "RUS", "ZAF")
```

```{r load_data, cache=FALSE, include=FALSE}
panelData <- getPanelDataHistoricalCached(
  aggregate = TRUE, y = 2000:2022,
  outputRegionMappingFile = "regionmapping_54.csv"
)

df_hist <- tidy_magpie(panelData, "Historical")

future_mag <- NULL
df_scen    <- NULL
if (has_scenario) {
  future_mag <- getPanelDataScenarioCached(
    gdxFile = gdx_path,
    aggregate = TRUE,
    outputRegionMappingFile = "regionmapping_54.csv"
  )
  df_scen <- tidy_magpie(future_mag, "Scenario")
}

df_all <- do.call(rbind, Filter(Negate(is.null), list(df_hist, df_scen)))
```

```{r fit_models, cache=FALSE, include=FALSE}
fit_bulk <- estimateAdoptionModel(
  data                      = panelData,
  sector                    = "Bulk",
  actorPowerDrivers         = NULL,
  actorPowerIndex           = "Actor Power Index",
  instQualityDrivers        = model_instQualityDrivers,
  controlDrivers            = model_controlDrivers,
  includeLaggedAdoption     = model_includeLagged,
  interactRegionFE          = FALSE,
  regionMappingFixedEffects = model_regionMappingFE
)

fit_diffuse <- estimateAdoptionModel(
  data                      = panelData,
  sector                    = "Diffuse",
  actorPowerDrivers         = NULL,
  actorPowerIndex           = "Actor Power Index",
  instQualityDrivers        = model_instQualityDrivers,
  controlDrivers            = model_controlDrivers,
  includeLaggedAdoption     = model_includeLagged,
  interactRegionFE          = FALSE,
  regionMappingFixedEffects = model_regionMappingFE
)

future_df_bulk    <- NULL
future_df_diffuse <- NULL

if (!is.null(future_mag)) {
  future_df_bulk <- preparePanelData(
    data                      = future_mag,
    sector                    = "Bulk",
    actorPowerDrivers         = NULL,
    actorPowerIndex           = "Actor Power Index",
    instQualityDrivers        = model_instQualityDrivers,
    controlDrivers            = model_controlDrivers,
    regionMappingFixedEffects = model_regionMappingFE,
    lag                       = 1
  )
  future_df_diffuse <- preparePanelData(
    data                      = future_mag,
    sector                    = "Diffuse",
    actorPowerDrivers         = NULL,
    actorPowerIndex           = "Actor Power Index",
    instQualityDrivers        = model_instQualityDrivers,
    controlDrivers            = model_controlDrivers,
    regionMappingFixedEffects = model_regionMappingFE,
    lag                       = 1
  )
}

compute_fit_stats <- function(fit) {
  m          <- fit$model
  is_logistf <- inherits(m, "logistf")
  k          <- length(coef(m))
  nObs       <- if (is_logistf) m$n else nobs(m)
  loglik     <- if (is_logistf) as.numeric(m$loglik["full"]) else as.numeric(logLik(m))
  aic        <- if (is_logistf) -2 * loglik + 2 * k else AIC(m)
  bic        <- if (is_logistf) -2 * loglik + k * log(nObs) else BIC(m)
  pseudoR2   <- if (is_logistf) {
    round(1 - (m$loglik["full"] / m$loglik["null"]), 3)
  } else if (!is.null(m$deviance) && !is.null(m$null.deviance)) {
    round(1 - (m$deviance / m$null.deviance), 3)
  } else NA
  pvals <- if (!is.null(fit$coeftest)) fit$coeftest[, 4] else NULL
  nSig  <- if (!is.null(pvals)) sum(pvals < 0.05, na.rm = TRUE) else NA
  list(
    k = k, nObs = nObs,
    loglik   = round(loglik, 2),
    aic      = round(aic, 2),
    bic      = round(bic, 2),
    pseudoR2 = pseudoR2,
    nSig     = nSig,
    converged = if (is_logistf) !any(is.na(m$coefficients)) else isTRUE(m$converged)
  )
}

stats_bulk    <- compute_fit_stats(fit_bulk)
stats_diffuse <- compute_fit_stats(fit_diffuse)

get_probs <- function(fit, future_df) {
  if (is.null(fit) || is.null(future_df)) return(NULL)
  m           <- fit$model
  needed_vars <- intersect(all.vars(m$formula), colnames(future_df))
  df_clean    <- future_df[complete.cases(future_df[, needed_vars, drop = FALSE]), ]
  if (nrow(df_clean) == 0) return(NULL)
  pred_link   <- tryCatch(predict(m, newdata = df_clean), error = function(e) NULL)
  if (is.null(pred_link)) return(NULL)
  df_clean$prob <- 1 / (1 + exp(-pred_link))
  df_clean
}

proj_bulk    <- get_probs(fit_bulk,    future_df_bulk)
proj_diffuse <- get_probs(fit_diffuse, future_df_diffuse)
```

---

# Executive Summary

This report documents a single-model deep dive into the **`r params$modelName`** specification of the PFM adoption hurdle model. It covers both the **Bulk** (industrial) and **Diffuse** (transport/buildings) sectors and includes three sections: input data inspection (historical & scenario), model estimation diagnostics, and future adoption projections.

<div class="row">
<div class="col-md-6">
<div class="card">
<div class="card-header">Model Specification</div>
<div class="card-body">
<ul>
  <li><strong>Model name:</strong> `r params$modelName`</li>
  <li><strong>Actor Power:</strong> Actor Power Index (composite)</li>
  <li><strong>Institutional Quality:</strong> `r paste(model_instQualityDrivers, collapse = ", ")`</li>
  <li><strong>Controls:</strong> `r paste(model_controlDrivers, collapse = ", ")`</li>
  <li><strong>Region FE mapping:</strong> `r if (is.null(model_regionMappingFE)) "None" else model_regionMappingFE`</li>
  <li><strong>Lagged adoption:</strong> `r model_includeLagged`</li>
</ul>
</div>
</div>
</div>
<div class="col-md-6">
<div class="card">
<div class="card-header">Projection Settings</div>
<div class="card-body">
<ul>
  <li><strong>Adoption threshold:</strong> `r adoption_threshold` (Pr(Adoption) &ge; threshold &rarr; adopted)</li>
  <li><strong>Snapshot years:</strong> `r paste(snapshot_years, collapse = ", ")`</li>
  <li><strong>Scenario data:</strong> `r if (has_scenario) paste0("Loaded from ", basename(params$gdxPath)) else "Not provided — projection sections disabled"`</li>
  <li><strong>Historical range:</strong> 2000–2022</li>
</ul>
</div>
</div>
</div>
</div>

```{r summary-fit-table, echo=FALSE}
summary_df <- data.frame(
  Sector     = c("Bulk", "Diffuse"),
  `Pseudo-R²`= c(stats_bulk$pseudoR2, stats_diffuse$pseudoR2),
  AIC        = c(stats_bulk$aic,      stats_diffuse$aic),
  BIC        = c(stats_bulk$bic,      stats_diffuse$bic),
  `Log-lik`  = c(stats_bulk$loglik,   stats_diffuse$loglik),
  `N obs`    = c(stats_bulk$nObs,     stats_diffuse$nObs),
  `N pred`   = c(stats_bulk$k,        stats_diffuse$k),
  `N sig`    = c(stats_bulk$nSig,     stats_diffuse$nSig),
  Converged  = c(stats_bulk$converged, stats_diffuse$converged),
  check.names = FALSE
)

kable(summary_df, format = "html", caption = "Adoption Model Fit Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(2, bold = TRUE,
    color = ifelse(summary_df$`Pseudo-R²` >= 0.4, col_good,
      ifelse(summary_df$`Pseudo-R²` >= 0.2, col_warn, col_bad)))
```

---

# Data Quality Check

```{r data-qa, echo=FALSE, results='asis', cache=FALSE}
check_panel_nas <- function(mag_obj, label) {
  if (is.null(mag_obj)) {
    cat(sprintf(
      '<div style="background:#fef9e7;border-left:4px solid %s;padding:10px 16px;margin-bottom:12px;">',
      col_warn
    ))
    cat(sprintf('<strong style="color:%s;">⊘ %s</strong>: not loaded.</div>\n', col_warn, label))
    return(invisible(NULL))
  }

  # Use as.array() — magclass's as.data.frame() uses sparse representation and
  # silently drops NaN/NA cells, giving false-clean results. The full array
  # includes every region × year × variable cell, including NaN.
  arr      <- as.array(mag_obj)
  n_total  <- length(arr)
  bad_mask <- !is.finite(arr)
  n_bad    <- sum(bad_mask)

  if (n_bad == 0) {
    cat(sprintf(
      '<div style="background:#eafaf1;border-left:4px solid %s;padding:10px 16px;margin-bottom:12px;">',
      col_good
    ))
    cat(sprintf(
      '<strong style="color:%s;">&#10003; %s</strong>: no missing or non-finite values across %s variable–region–year combinations.</div>\n',
      col_good, label, format(n_total, big.mark = ",")
    ))
    return(invisible(NULL))
  }

  pct     <- round(100 * n_bad / n_total, 2)
  cat(sprintf(
    '<div style="background:#fef9e7;border-left:4px solid %s;padding:10px 16px;margin-bottom:4px;">',
    col_warn
  ))
  cat(sprintf(
    '<strong style="color:%s;">&#9888; %s</strong>: %s non-finite values (NA / NaN / Inf) — %.2f%% of %s total.</div>\n',
    col_warn, label, format(n_bad, big.mark = ","), pct, format(n_total, big.mark = ",")
  ))

  dn       <- dimnames(arr)
  bad_idx  <- which(bad_mask, arr.ind = TRUE)
  bad_vals <- arr[bad_mask]
  bad_df   <- data.frame(
    region   = dn[[1]][bad_idx[, 1]],
    year     = as.integer(gsub("y", "", dn[[2]][bad_idx[, 2]])),
    variable = dn[[3]][bad_idx[, 3]],
    kind     = ifelse(is.nan(bad_vals), "NaN",
                 ifelse(is.na(bad_vals), "NA",
                   ifelse(bad_vals == Inf, "Inf", "-Inf"))),
    stringsAsFactors = FALSE
  )

  by_var <- bad_df %>%
    dplyr::group_by(variable, kind) %>%
    dplyr::summarise(
      `Count`            = dplyr::n(),
      `Affected regions` = paste(sort(unique(region)), collapse = ", "),
      `Years`            = if (min(year) == max(year)) as.character(min(year))
                           else paste0(min(year), "–", max(year)),
      .groups = "drop"
    ) %>%
    dplyr::arrange(dplyr::desc(`Count`))

  print(
    knitr::kable(by_var, format = "html",
                 caption = paste0(label, " — non-finite values by variable")) %>%
      kableExtra::kable_styling(
        bootstrap_options = c("striped", "hover", "condensed"),
        full_width = TRUE
      )
  )
  invisible(NULL)
}

check_panel_nas(panelData,  "Historical panel")
check_panel_nas(future_mag, "Scenario panel")

# --- Region coverage: which regions appear in each stage of the pipeline ---
banner <- function(ok, msg) {
  bg  <- if (ok) "#eafaf1" else "#fef9e7"
  col <- if (ok) col_good  else col_warn
  ico <- if (ok) "&#10003;" else "&#9888;"
  cat(sprintf(
    '<div style="background:%s;border-left:4px solid %s;padding:10px 16px;margin-bottom:8px;"><strong style="color:%s;">%s %s</strong></div>\n',
    bg, col, col, ico, msg
  ))
}

hist_regions <- sort(magclass::getItems(panelData, dim = 1))

if (!is.null(future_mag)) {
  scen_regions <- sort(magclass::getItems(future_mag, dim = 1))
  missing_in_scen <- setdiff(hist_regions, scen_regions)
  extra_in_scen   <- setdiff(scen_regions, hist_regions)

  cat("\n\n**Region coverage — Historical vs Scenario panel:**\n\n")
  if (length(missing_in_scen) == 0) {
    banner(TRUE, sprintf("Scenario panel covers all %d historical regions.", length(hist_regions)))
  } else {
    banner(FALSE, sprintf(
      "%d region(s) present in historical panel but ABSENT from scenario panel: %s",
      length(missing_in_scen), paste(missing_in_scen, collapse = ", ")
    ))
  }
  if (length(extra_in_scen) > 0) {
    banner(FALSE, sprintf(
      "%d region(s) in scenario panel but NOT in historical panel: %s",
      length(extra_in_scen), paste(extra_in_scen, collapse = ", ")
    ))
  }
}

# Coverage and year-count after complete.cases() inside get_probs()
check_proj_coverage <- function(proj_df, hist_regs, label) {
  if (is.null(proj_df)) return(invisible(NULL))

  expected_years <- length(unique(proj_df$year))
  yr_counts <- proj_df %>%
    dplyr::count(region, name = "n_years")

  proj_regs <- sort(yr_counts$region)
  missing   <- setdiff(hist_regs, proj_regs)
  sparse    <- yr_counts %>%
    dplyr::filter(n_years < expected_years) %>%
    dplyr::arrange(n_years)

  cat(sprintf("\n\n**Region coverage after model prediction — %s:**\n\n", label))

  if (length(missing) == 0) {
    banner(TRUE, sprintf("All %d historical regions are present in %s projections.", length(hist_regs), label))
  } else {
    banner(FALSE, sprintf(
      "%d region(s) dropped before/after prediction in %s: %s",
      length(missing), label, paste(missing, collapse = ", ")
    ))
  }

  if (nrow(sparse) == 0) {
    banner(TRUE, sprintf("All regions have the full %d projected years.", expected_years))
  } else {
    banner(FALSE, sprintf(
      "%d region(s) have fewer than the expected %d projected years (sparse rows may cause near-zero-only trajectories):",
      nrow(sparse), expected_years
    ))
    print(
      knitr::kable(sparse, col.names = c("Region", "Years with predictions"),
                   format = "html",
                   caption = paste0(label, " — regions with incomplete year coverage")) %>%
        kableExtra::kable_styling(
          bootstrap_options = c("striped", "hover", "condensed"),
          full_width = FALSE
        )
    )
  }
}

check_proj_coverage(proj_bulk,    hist_regions, "Bulk")
check_proj_coverage(proj_diffuse, hist_regions, "Diffuse")
```

---

# Input Data {.tabset}

Variables used by this model: **Actor Power Index** (Bulk & Diffuse), **`r paste(model_instQualityDrivers, collapse = "**, **")`**, **`r paste(model_controlDrivers, collapse = "**, **")`**.

## Historical Distributions

```{r hist-distributions, echo=FALSE, fig.width=12, fig.height=7}
df_model_hist <- df_hist %>%
  filter(variable %in% model_vars_input)

df_model_hist$var_short <- df_model_hist$variable %>%
  gsub("\\|Bulk",   " (Bulk)",   .) %>%
  gsub("\\|Diffuse", " (Diffuse)", .) %>%
  gsub(" \\(WGI\\)", " [WGI]", .) %>%
  gsub(" \\(VDem\\)", " [VDem]", .)

ggplot(df_model_hist, aes(x = value)) +
  geom_density(fill = "#2C7BB6", alpha = 0.6, color = "white") +
  geom_rug(alpha = 0.2, color = "#2C3E50") +
  facet_wrap(~var_short, scales = "free", ncol = 3) +
  labs(
    title    = paste0(params$modelName, " — Historical Variable Distributions (2000–2022)"),
    subtitle = "Aggregated 54-region panel; density estimated over region-year observations",
    x = "Variable value", y = "Density"
  ) +
  theme_report()
```

### Summary Statistics

```{r hist-stats, echo=FALSE}
hist_stats <- df_model_hist %>%
  group_by(variable) %>%
  summarize(
    N      = sum(!is.na(value)),
    Mean   = mean(value,   na.rm = TRUE),
    Median = median(value, na.rm = TRUE),
    SD     = sd(value,     na.rm = TRUE),
    Min    = min(value,    na.rm = TRUE),
    Max    = max(value,    na.rm = TRUE),
    .groups = "drop"
  )

kable(hist_stats, digits = 3, caption = "Historical Variable Statistics (Aggregated, 2000–2022)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

## Historical Trends

```{r hist-trends, echo=FALSE, fig.width=12, fig.height=8}
df_trends <- df_model_hist %>%
  filter(region %in% key_regions)

df_trends$var_short <- df_trends$variable %>%
  gsub("\\|Bulk",   " (Bulk)",   .) %>%
  gsub("\\|Diffuse", " (Diffuse)", .) %>%
  gsub(" \\(WGI\\)", " [WGI]", .) %>%
  gsub(" \\(VDem\\)", " [VDem]", .)

ggplot(df_trends, aes(x = year, y = value, color = region)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 1.5, alpha = 0.7) +
  facet_wrap(~var_short, scales = "free_y", ncol = 3) +
  labs(
    title    = paste0(params$modelName, " — Historical Trends by Region (2000–2022)"),
    subtitle = "Selected major economies; aggregated 54-region mapping",
    x = "Year", y = "Variable value", color = "Region"
  ) +
  theme_report() +
  theme(legend.position = "bottom")
```

## Scenario Projections

```{r scenario-projections, echo=FALSE, fig.width=12, fig.height=8}
if (has_scenario) {
  df_model_all <- df_all %>%
    filter(variable %in% model_vars_input, region %in% key_regions)

  df_model_all$var_short <- df_model_all$variable %>%
    gsub("\\|Bulk",    " (Bulk)",    .) %>%
    gsub("\\|Diffuse", " (Diffuse)", .) %>%
    gsub(" \\(WGI\\)", " [WGI]", .) %>%
    gsub(" \\(VDem\\)", " [VDem]", .)

  ggplot(df_model_all, aes(x = year, y = value, color = region, linetype = period)) +
    geom_line(linewidth = 1.1) +
    geom_vline(xintercept = 2022, linetype = "dotted", color = "grey40") +
    facet_wrap(~var_short, scales = "free_y", ncol = 3) +
    labs(
      title    = paste0(params$modelName, " — Historical & Scenario Variable Trajectories"),
      subtitle = "Vertical dotted line = 2022 (historical / scenario boundary)",
      x = "Year", y = "Variable value", color = "Region", linetype = "Period"
    ) +
    theme_report() +
    theme(legend.position = "bottom")
} else {
  cat("> [!NOTE] Scenario GDX file not provided — long-term projections are disabled.")
}
```

---

# Model Estimation {.tabset}

## Bulk — Adoption

```{r forest-bulk, echo=FALSE, fig.height=6, fig.width=10}
coef_df_bulk <- pfm::coeftestToDataFrame(fit_bulk$coeftest) %>%
  filter(term != "(Intercept)")

p_bulk <- render_forest_plot(
  coef_df_bulk,
  title    = "Bulk Adoption — Coefficient Forest Plot",
  subtitle = sprintf(
    "Log-odds coefficients | Pseudo-R² = %.3f | AIC = %.1f | %d/%d significant",
    stats_bulk$pseudoR2, stats_bulk$aic, stats_bulk$nSig, stats_bulk$k
  )
)

if (!is.null(p_bulk)) {
  print(p_bulk)
  ggsave(file.path(plotOutputDir, "forest_bulk_adoption.png"), p_bulk, width = 10, height = 6, dpi = 300)
  ggsave(file.path(plotOutputDir, "forest_bulk_adoption.svg"), p_bulk, width = 10, height = 6)
}
```

### Coefficient Table

```{r coef-table-bulk, echo=FALSE}
coef_display_bulk <- pfm::coeftestToDataFrame(fit_bulk$coeftest) %>%
  mutate(
    Significance = case_when(
      pValue < 0.001 ~ "***", pValue < 0.01 ~ "**",
      pValue < 0.05  ~ "*",   pValue < 0.1  ~ ".", TRUE ~ ""
    )
  ) %>%
  rename(
    Term = term, Estimate = estimate,
    `Std. Error` = stdError, `z / t` = statistic, `p-value` = pValue
  )

kable(coef_display_bulk, digits = 4, format = "html",
  caption = "Bulk Adoption — Coefficient Estimates") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(5, bold = TRUE,
    color = ifelse(coef_display_bulk$`p-value` < 0.05, col_good, "grey50"))
```

### Fit Diagnostics

```{r diag-bulk, echo=FALSE}
diag_bulk_df <- data.frame(
  Metric = c("Pseudo-R²", "AIC", "BIC", "Log-likelihood",
             "N observations", "N predictors", "N significant (p<0.05)", "Converged"),
  Value = c(
    stats_bulk$pseudoR2, stats_bulk$aic, stats_bulk$bic, stats_bulk$loglik,
    stats_bulk$nObs, stats_bulk$k, stats_bulk$nSig,
    as.character(stats_bulk$converged)
  )
)

kable(diag_bulk_df, format = "html", caption = "Bulk Adoption — Model Diagnostics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE)
```

## Diffuse — Adoption

```{r forest-diffuse, echo=FALSE, fig.height=6, fig.width=10}
coef_df_diffuse <- pfm::coeftestToDataFrame(fit_diffuse$coeftest) %>%
  filter(term != "(Intercept)")

p_diffuse <- render_forest_plot(
  coef_df_diffuse,
  title    = "Diffuse Adoption — Coefficient Forest Plot",
  subtitle = sprintf(
    "Log-odds coefficients | Pseudo-R² = %.3f | AIC = %.1f | %d/%d significant",
    stats_diffuse$pseudoR2, stats_diffuse$aic, stats_diffuse$nSig, stats_diffuse$k
  )
)

if (!is.null(p_diffuse)) {
  print(p_diffuse)
  ggsave(file.path(plotOutputDir, "forest_diffuse_adoption.png"), p_diffuse, width = 10, height = 6, dpi = 300)
  ggsave(file.path(plotOutputDir, "forest_diffuse_adoption.svg"), p_diffuse, width = 10, height = 6)
}
```

### Coefficient Table

```{r coef-table-diffuse, echo=FALSE}
coef_display_diffuse <- pfm::coeftestToDataFrame(fit_diffuse$coeftest) %>%
  mutate(
    Significance = case_when(
      pValue < 0.001 ~ "***", pValue < 0.01 ~ "**",
      pValue < 0.05  ~ "*",   pValue < 0.1  ~ ".", TRUE ~ ""
    )
  ) %>%
  rename(
    Term = term, Estimate = estimate,
    `Std. Error` = stdError, `z / t` = statistic, `p-value` = pValue
  )

kable(coef_display_diffuse, digits = 4, format = "html",
  caption = "Diffuse Adoption — Coefficient Estimates") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(5, bold = TRUE,
    color = ifelse(coef_display_diffuse$`p-value` < 0.05, col_good, "grey50"))
```

### Fit Diagnostics

```{r diag-diffuse, echo=FALSE}
diag_diffuse_df <- data.frame(
  Metric = c("Pseudo-R²", "AIC", "BIC", "Log-likelihood",
             "N observations", "N predictors", "N significant (p<0.05)", "Converged"),
  Value = c(
    stats_diffuse$pseudoR2, stats_diffuse$aic, stats_diffuse$bic, stats_diffuse$loglik,
    stats_diffuse$nObs, stats_diffuse$k, stats_diffuse$nSig,
    as.character(stats_diffuse$converged)
  )
)

kable(diag_diffuse_df, format = "html", caption = "Diffuse Adoption — Model Diagnostics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE)
```

## Sector Comparison

```{r sector-comparison, echo=FALSE, fig.width=10, fig.height=4}
comp_df <- data.frame(
  Sector   = c("Bulk", "Diffuse"),
  PseudoR2 = c(stats_bulk$pseudoR2, stats_diffuse$pseudoR2),
  AIC      = c(stats_bulk$aic, stats_diffuse$aic),
  Quality  = c(
    ifelse(stats_bulk$pseudoR2 >= 0.4,    "Good",
      ifelse(stats_bulk$pseudoR2 >= 0.2,  "Moderate", "Poor")),
    ifelse(stats_diffuse$pseudoR2 >= 0.4, "Good",
      ifelse(stats_diffuse$pseudoR2 >= 0.2, "Moderate", "Poor"))
  )
)

p_r2 <- ggplot(comp_df, aes(x = Sector, y = PseudoR2,
    fill = factor(Quality, levels = c("Poor", "Moderate", "Good")))) +
  geom_col(width = 0.5) +
  geom_text(aes(label = sprintf("%.3f", PseudoR2)), vjust = -0.5, size = 4, fontface = "bold") +
  scale_fill_manual(values = c("Poor" = col_bad, "Moderate" = col_warn, "Good" = col_good),
    name = "Fit quality") +
  labs(title = "Pseudo-R² by Sector", y = "McFadden Pseudo-R²", x = NULL) +
  ylim(0, 1) +
  theme_report()

p_aic <- ggplot(comp_df, aes(x = Sector, y = AIC, fill = Sector)) +
  geom_col(width = 0.5) +
  geom_text(aes(label = sprintf("%.1f", AIC)), vjust = -0.5, size = 4, fontface = "bold") +
  scale_fill_manual(values = c("Bulk" = col_bulk, "Diffuse" = col_diffuse)) +
  labs(title = "AIC by Sector (lower = better)", y = "AIC", x = NULL) +
  theme_report() +
  theme(legend.position = "none")

library(patchwork)
p_r2 + p_aic + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

---

# Historical Model Fit {.tabset}

```{r hist-fit-helper, include=FALSE}
build_hist_fit <- function(fit, sector, data) {
  df <- pfm::preparePanelData(
    data                      = data,
    sector                    = sector,
    actorPowerDrivers         = NULL,
    actorPowerIndex           = "Actor Power Index",
    instQualityDrivers        = model_instQualityDrivers,
    controlDrivers            = model_controlDrivers,
    regionMappingFixedEffects = model_regionMappingFE,
    lag                       = 1
  )
  df$adoption_raw <- as.integer(df$ecp > 0)

  needed_vars <- intersect(all.vars(fit$model$formula), colnames(df))
  ok          <- complete.cases(df[, needed_vars, drop = FALSE])
  df$fitted_prob <- NA_real_
  pred <- tryCatch(
    predict(fit$model, newdata = df[ok, ], type = "link"),
    error = function(e) NULL
  )
  if (!is.null(pred)) df$fitted_prob[ok] <- 1 / (1 + exp(-as.numeric(pred)))
  df
}

hist_fit_bulk    <- build_hist_fit(fit_bulk,    "Bulk",    panelData)
hist_fit_diffuse <- build_hist_fit(fit_diffuse, "Diffuse", panelData)

# --- Trajectory plot (adopted or non-adopted regions) ---
make_traj_plot <- function(df, sector_name, adopted_group = TRUE) {
  col_sector  <- if (sector_name == "Bulk") col_bulk else col_diffuse
  group_label <- if (adopted_group) "Adopted" else "Non-Adopted"

  ever_adopted <- df %>%
    group_by(region) %>%
    summarise(adopted = any(adoption_raw == 1, na.rm = TRUE), .groups = "drop")

  regs <- ever_adopted %>% filter(adopted == adopted_group) %>% pull(region)
  df_f <- df %>% filter(region %in% regs)
  if (nrow(df_f) == 0) return(NULL)

  n_regs <- length(unique(df_f$region))
  ncols  <- min(6L, n_regs)

  ggplot(df_f, aes(x = year)) +
    geom_step(aes(y = adoption_raw, color = "Actual adoption (0/1)"),
              linewidth = 0.7, alpha = 0.6) +
    geom_line(aes(y = fitted_prob, color = "Model fitted probability"),
              linewidth = 1.0, na.rm = TRUE) +
    facet_wrap(~region, ncol = ncols) +
    scale_color_manual(
      values = c("Actual adoption (0/1)" = "grey55",
                 "Model fitted probability" = col_sector),
      name = NULL
    ) +
    scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
    labs(
      title    = paste0(sector_name, " — ", group_label, " Regions: In-Sample Fit"),
      subtitle = "Grey step = actual adoption (0/1); solid = model fitted probability",
      x = "Year", y = "Probability / Adoption"
    ) +
    theme_report() +
    theme(legend.position = "bottom")
}

# --- Density plot of predicted probabilities by actual outcome ---
make_density_plot <- function(df, sector_name) {
  df_clean <- df %>%
    filter(!is.na(fitted_prob)) %>%
    mutate(outcome = factor(
      ifelse(adoption_raw == 1, "Adopted", "Not Adopted"),
      levels = c("Adopted", "Not Adopted")
    ))

  ggplot(df_clean, aes(x = fitted_prob, fill = outcome, color = outcome)) +
    geom_density(alpha = 0.4, linewidth = 0.8) +
    scale_fill_manual(values  = c("Adopted" = "#E74C3C", "Not Adopted" = "#3498DB"),
                      name = "Actual Outcome") +
    scale_color_manual(values = c("Adopted" = "#E74C3C", "Not Adopted" = "#3498DB"),
                       name = "Actual Outcome") +
    scale_x_continuous(labels = scales::percent, limits = c(0, 1)) +
    labs(
      title    = paste0("Model Fit: Adoption Probability (", sector_name, ")"),
      subtitle = "Distribution of predicted probabilities for actual outcomes",
      x        = "Predicted Probability of Adopting a Carbon Price",
      y        = "Density",
      caption  = paste0(
        "Figure: Fit of the Logit (adoption) model component.\n",
        "Shows predicted probability distributions for actual adoption outcomes.\n",
        "Colors indicate the observed outcome (adopted vs not adopted)."
      )
    ) +
    theme_report() +
    theme(legend.position = "right",
          plot.caption = element_text(hjust = 0, size = 8, color = "grey40"))
}

# --- Adoption timeline (Gantt chart) ---
make_timeline_plot <- function(df, sector_name, threshold) {
  col_sector <- if (sector_name == "Bulk") col_bulk else col_diffuse

  adopted_regs <- df %>%
    group_by(region) %>%
    filter(any(adoption_raw == 1, na.rm = TRUE)) %>%
    ungroup()

  actual <- adopted_regs %>%
    filter(adoption_raw == 1) %>%
    group_by(region) %>%
    summarise(start_year = min(year), end_year = max(year), .groups = "drop")

  predicted <- adopted_regs %>%
    filter(!is.na(fitted_prob), fitted_prob >= threshold) %>%
    group_by(region) %>%
    summarise(pred_year = min(year), .groups = "drop")

  timeline <- actual %>%
    left_join(predicted, by = "region") %>%
    arrange(start_year, region)

  timeline$region <- factor(timeline$region, levels = rev(timeline$region))

  ggplot(timeline, aes(y = region)) +
    geom_segment(aes(x = start_year, xend = end_year, yend = region,
                     color = "Real Adoption"),
                 linewidth = 5, alpha = 0.45, lineend = "round") +
    geom_point(aes(x = pred_year, color = "Predicted Adoption"),
               size = 3, na.rm = TRUE) +
    scale_color_manual(
      values = c("Real Adoption" = "grey55", "Predicted Adoption" = col_sector),
      name = NULL
    ) +
    labs(
      title    = paste0("Carbon Price Adoption: Real vs Predicted (",
                        sector_name, ", threshold = ", threshold, ")"),
      subtitle = "Grey bar = actual adoption period; dot = first year model predicts adoption",
      x = "Year", y = NULL,
      caption  = paste0(
        "Note: Predicted adoption based on probability > ", threshold, ".\n",
        "Logit model only predicts adoption probability, not price levels."
      )
    ) +
    theme_report() +
    theme(legend.position = "bottom",
          plot.caption = element_text(hjust = 0, size = 8, color = "grey40"))
}
```

## Bulk {.tabset}

### Adopted Regions

```{r hist-traj-bulk-adopted, echo=FALSE, fig.width=14, fig.height=10}
p <- make_traj_plot(hist_fit_bulk, "Bulk", adopted_group = TRUE)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "hist_fit_bulk_adopted.png"), p, width = 14, height = 10, dpi = 300)
}
```

### Non-Adopted Regions

```{r hist-traj-bulk-nonadopted, echo=FALSE, fig.width=14, fig.height=10}
p <- make_traj_plot(hist_fit_bulk, "Bulk", adopted_group = FALSE)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "hist_fit_bulk_nonadopted.png"), p, width = 14, height = 10, dpi = 300)
}
```

### Probability Distribution

```{r hist-density-bulk, echo=FALSE, fig.width=10, fig.height=6}
p <- make_density_plot(hist_fit_bulk, "Bulk")
print(p)
ggsave(file.path(plotOutputDir, "hist_density_bulk.png"), p, width = 10, height = 6, dpi = 300)
```

### Adoption Timeline

```{r hist-timeline-bulk, echo=FALSE, fig.width=11, fig.height=9}
p <- make_timeline_plot(hist_fit_bulk, "Bulk", adoption_threshold)
print(p)
ggsave(file.path(plotOutputDir, "hist_timeline_bulk.png"), p, width = 11, height = 9, dpi = 300)
```

## Diffuse {.tabset}

### Adopted Regions

```{r hist-traj-diffuse-adopted, echo=FALSE, fig.width=14, fig.height=10}
p <- make_traj_plot(hist_fit_diffuse, "Diffuse", adopted_group = TRUE)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "hist_fit_diffuse_adopted.png"), p, width = 14, height = 10, dpi = 300)
}
```

### Non-Adopted Regions

```{r hist-traj-diffuse-nonadopted, echo=FALSE, fig.width=14, fig.height=10}
p <- make_traj_plot(hist_fit_diffuse, "Diffuse", adopted_group = FALSE)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "hist_fit_diffuse_nonadopted.png"), p, width = 14, height = 10, dpi = 300)
}
```

### Probability Distribution

```{r hist-density-diffuse, echo=FALSE, fig.width=10, fig.height=6}
p <- make_density_plot(hist_fit_diffuse, "Diffuse")
print(p)
ggsave(file.path(plotOutputDir, "hist_density_diffuse.png"), p, width = 10, height = 6, dpi = 300)
```

### Adoption Timeline

```{r hist-timeline-diffuse, echo=FALSE, fig.width=11, fig.height=9}
p <- make_timeline_plot(hist_fit_diffuse, "Diffuse", adoption_threshold)
print(p)
ggsave(file.path(plotOutputDir, "hist_timeline_diffuse.png"), p, width = 11, height = 9, dpi = 300)
```

---

# Future Projections {.tabset}

```{r proj-check, echo=FALSE, results='asis'}
if (!has_scenario) {
  cat("> [!WARNING]\n> **Scenario GDX file not provided.** All projection sections below are disabled.\n> Re-render with `gdxPath` pointing to a valid `fulldata.gdx` to enable projections.\n\n")
}
```

## Probability Trajectories {.tabset}

```{r prob-trajectory-helper, include=FALSE}
make_trajectory_plot <- function(proj_df, sector_name, key_regs, threshold, full_facet = FALSE) {
  if (is.null(proj_df) || nrow(proj_df) == 0) return(NULL)

  col_sector <- if (sector_name == "Bulk") col_bulk else col_diffuse
  all_regs   <- unique(proj_df$region)
  highlight  <- intersect(key_regs, all_regs)

  proj_df$grp <- ifelse(proj_df$region %in% highlight,
                        as.character(proj_df$region), "Other")
  proj_df$grp <- factor(proj_df$grp, levels = c(highlight, "Other"))

  my_colors <- setNames(
    c(scales::hue_pal()(length(highlight)), "grey80"),
    c(highlight, "Other")
  )

  if (full_facet) {
    ggplot(proj_df, aes(x = year, y = prob, group = region)) +
      geom_hline(yintercept = threshold, linetype = "dashed", color = "grey50", linewidth = 0.5) +
      geom_line(color = col_sector, linewidth = 0.7, alpha = 0.8) +
      facet_wrap(~region, ncol = 9) +
      scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
      labs(
        title    = sprintf("%s Adoption — All Regions Probability Trajectories", sector_name),
        subtitle = sprintf("Dashed line = threshold %.2f", threshold),
        x = "Year", y = "Pr(Adoption)"
      ) +
      theme_report() +
      theme(axis.text = element_text(size = 8), strip.text = element_text(size = 8))
  } else {
    ggplot(proj_df, aes(x = year, y = prob, group = region, color = grp,
        alpha = I(ifelse(grp == "Other", 0.35, 1)),
        linewidth = I(ifelse(grp == "Other", 0.5, 1.2)))) +
      geom_hline(yintercept = threshold, linetype = "dashed", color = "grey50") +
      geom_line() +
      scale_color_manual(values = my_colors) +
      scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
      labs(
        title    = sprintf("%s Adoption — Projected Probability (Key Regions)", sector_name),
        subtitle = sprintf("Dashed line = adoption threshold %.2f | Grey = all other regions", threshold),
        x = "Year", y = "Pr(Adoption)", color = "Region"
      ) +
      theme_report() +
      theme(legend.position = "right")
  }
}
```

### Bulk

```{r traj-bulk-key, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario && !is.null(proj_bulk)) {
  p <- make_trajectory_plot(proj_bulk, "Bulk", key_regions, adoption_threshold, full_facet = FALSE)
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "prob_trajectory_bulk.png"), p, width = 12, height = 6, dpi = 300)
    ggsave(file.path(plotOutputDir, "prob_trajectory_bulk.svg"), p, width = 12, height = 6)
  }
} else {
  cat("> Scenario data not available.")
}
```

```{r traj-bulk-full, echo=FALSE, fig.width=16, fig.height=20, class.output="fold-hide"}
if (has_scenario && !is.null(proj_bulk)) {
  p_full <- make_trajectory_plot(proj_bulk, "Bulk", key_regions, adoption_threshold, full_facet = TRUE)
  if (!is.null(p_full)) print(p_full)
}
```

### Diffuse

```{r traj-diffuse-key, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario && !is.null(proj_diffuse)) {
  p <- make_trajectory_plot(proj_diffuse, "Diffuse", key_regions, adoption_threshold, full_facet = FALSE)
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "prob_trajectory_diffuse.png"), p, width = 12, height = 6, dpi = 300)
    ggsave(file.path(plotOutputDir, "prob_trajectory_diffuse.svg"), p, width = 12, height = 6)
  }
} else {
  cat("> Scenario data not available.")
}
```

```{r traj-diffuse-full, echo=FALSE, fig.width=16, fig.height=20, class.output="fold-hide"}
if (has_scenario && !is.null(proj_diffuse)) {
  p_full <- make_trajectory_plot(proj_diffuse, "Diffuse", key_regions, adoption_threshold, full_facet = TRUE)
  if (!is.null(p_full)) print(p_full)
}
```

## Adoption Timeline {.tabset}

```{r timeline-helper, include=FALSE}
compute_timeline <- function(proj_df, threshold) {
  if (is.null(proj_df) || nrow(proj_df) == 0) return(NULL)
  proj_df %>%
    group_by(region) %>%
    summarise(
      AdoptionYear = ifelse(any(prob >= threshold, na.rm = TRUE),
                            min(year[prob >= threshold], na.rm = TRUE), Inf),
      .groups = "drop"
    ) %>%
    mutate(
      Category  = case_when(
        AdoptionYear <= 2030 ~ "Forerunners",
        AdoptionYear <= 2050 ~ "Early Majority",
        AdoptionYear <= 2070 ~ "Late Majority",
        TRUE                 ~ "Laggards"
      ),
      YearRange = case_when(
        Category == "Forerunners"    ~ "≤ 2030",
        Category == "Early Majority" ~ "2031–2050",
        Category == "Late Majority"  ~ "2051–2070",
        Category == "Laggards"       ~ "> 2070"
      ),
      RegYear = ifelse(is.finite(AdoptionYear),
                       paste0(region, " (", as.integer(AdoptionYear), ")"), region)
    )
}

render_timeline_table <- function(tl_df, sector_name) {
  if (is.null(tl_df)) {
    cat("> Scenario data not available.")
    return(invisible(NULL))
  }
  final_table <- tl_df %>%
    mutate(Category = factor(Category,
      levels = c("Forerunners", "Early Majority", "Late Majority", "Laggards"))) %>%
    group_by(Category, YearRange) %>%
    summarise(Regions = paste(RegYear, collapse = ", "), .groups = "drop") %>%
    arrange(Category)

  cat(as.character(
    kable(final_table, format = "html",
      col.names = c("Category", "Expected Year", "Regions")) %>%
      kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
      column_spec(1, bold = TRUE, color = c("#2ecc71", "#82e0aa", "#f7dc6f", "#e67e22"))
  ))
  cat("\n\n")
}

make_full_gantt_plot <- function(hist_df, proj_df, sector_name, threshold) {
  col_sector <- if (sector_name == "Bulk") col_bulk else col_diffuse

  hist_actual <- hist_df %>%
    filter(adoption_raw == 1) %>%
    group_by(region) %>%
    summarise(hist_start = min(year), hist_end = max(year), .groups = "drop")

  proj_adoption <- if (!is.null(proj_df) && nrow(proj_df) > 0) {
    proj_df %>%
      group_by(region) %>%
      summarise(
        proj_year = ifelse(any(prob >= threshold, na.rm = TRUE),
                           min(year[prob >= threshold], na.rm = TRUE), NA_real_),
        .groups = "drop"
      )
  } else {
    data.frame(region = character(0), proj_year = numeric(0))
  }

  all_regions <- unique(c(as.character(hist_df$region),
                           if (!is.null(proj_df)) as.character(proj_df$region)))
  gantt <- data.frame(region = all_regions, stringsAsFactors = FALSE) %>%
    left_join(hist_actual,    by = "region") %>%
    left_join(proj_adoption,  by = "region") %>%
    filter(!is.na(hist_start) | !is.na(proj_year)) %>%
    mutate(
      sort_key = case_when(
        !is.na(hist_start) ~ as.numeric(hist_start),
        !is.na(proj_year)  ~ as.numeric(proj_year) + 1000,
        TRUE               ~ Inf
      )
    ) %>%
    arrange(sort_key) %>%
    mutate(region = factor(region, levels = rev(region)))

  max_hist_year <- max(hist_df$year, na.rm = TRUE)

  ggplot(gantt, aes(y = region)) +
    geom_segment(
      data = gantt %>% filter(!is.na(hist_start)),
      aes(x = hist_start, xend = pmax(hist_end, max_hist_year), yend = region,
          color = "Historical Adoption"),
      linewidth = 4, alpha = 0.55, lineend = "round"
    ) +
    geom_segment(
      data = gantt %>% filter(!is.na(proj_year)),
      aes(x = proj_year, xend = 2150, yend = region,
          color = "Projected Adoption"),
      linewidth = 3, alpha = 0.35, lineend = "round"
    ) +
    scale_color_manual(
      values = c("Historical Adoption" = "grey55", "Projected Adoption" = col_sector),
      name = NULL
    ) +
    labs(
      title    = paste0(sector_name, " Carbon Price Adoption: Historical + Projected"),
      subtitle = paste0("Grey = observed adoption period; Colour = projected onset (prob ≥ ", threshold, ")"),
      x = "Year", y = NULL
    ) +
    theme_report() +
    theme(
      legend.position  = "bottom",
      axis.text.y      = element_text(size = 7)
    )
}
```

### Bulk

```{r timeline-bulk, echo=FALSE, results='asis'}
if (has_scenario && !is.null(proj_bulk)) {
  tl_bulk <- compute_timeline(proj_bulk, adoption_threshold)
  render_timeline_table(tl_bulk, "Bulk")
} else {
  cat("> Scenario data not available.")
}
```

### Bulk Gantt

```{r timeline-gantt-bulk, echo=FALSE, fig.width=12, fig.height=16}
if (has_scenario && !is.null(proj_bulk)) {
  p <- make_full_gantt_plot(hist_fit_bulk, proj_bulk, "Bulk", adoption_threshold)
  print(p)
  ggsave(file.path(plotOutputDir, "gantt_full_bulk.png"), p, width = 12, height = 16, dpi = 300)
} else {
  cat("> Scenario data not available.")
}
```

### Diffuse

```{r timeline-diffuse, echo=FALSE, results='asis'}
if (has_scenario && !is.null(proj_diffuse)) {
  tl_diffuse <- compute_timeline(proj_diffuse, adoption_threshold)
  render_timeline_table(tl_diffuse, "Diffuse")
} else {
  cat("> Scenario data not available.")
}
```

### Diffuse Gantt

```{r timeline-gantt-diffuse, echo=FALSE, fig.width=12, fig.height=16}
if (has_scenario && !is.null(proj_diffuse)) {
  p <- make_full_gantt_plot(hist_fit_diffuse, proj_diffuse, "Diffuse", adoption_threshold)
  print(p)
  ggsave(file.path(plotOutputDir, "gantt_full_diffuse.png"), p, width = 12, height = 16, dpi = 300)
} else {
  cat("> Scenario data not available.")
}
```

## Adoption Year Maps {.tabset}

```{r adoption-map-setup, include=FALSE}
mapping_54 <- NULL
world      <- NULL

if (has_scenario) {
  mapping_54 <- tryCatch(
    madrat::toolGetMapping("regionmapping_54.csv", type = "regional", where = "mappingfolder"),
    error = function(e) NULL
  )
  if (!is.null(mapping_54)) {
    suppressPackageStartupMessages(library(sf))
    world           <- rnaturalearth::ne_countries(scale = "small", returnclass = "sf")
    world$iso_key   <- as.character(world$adm0_a3)
  }
}

render_adoption_year_map <- function(tl_df, sector_name, threshold) {
  if (is.null(tl_df) || is.null(mapping_54) || is.null(world)) {
    cat("> Map rendering requires scenario data and the regionmapping_54.csv file.\n")
    return(invisible(NULL))
  }

  adopt_data <- tl_df %>%
    mutate(Category = factor(Category,
      levels = c("Forerunners", "Early Majority", "Late Majority", "Laggards")))

  map_data_df <- mapping_54 %>%
    mutate(
      RegionCode  = as.character(RegionCode),
      CountryCode = as.character(CountryCode),
      CountryCode = ifelse(CountryCode == "SSD", "SDS", CountryCode)
    ) %>%
    left_join(adopt_data, by = c("RegionCode" = "region"))

  world_data <- world %>% left_join(map_data_df, by = c("iso_key" = "CountryCode"))

  p <- ggplot() +
    geom_sf(data = world_data[world_data$iso_key == "ATA", ],
            fill = "grey90", color = "white", linewidth = 0.05) +
    geom_sf(data = world_data[world_data$iso_key != "ATA", ],
            aes(fill = Category), color = "white", linewidth = 0.05) +
    scale_fill_manual(
      values = c(
        "Forerunners"   = "#2ecc71",
        "Early Majority"= "#82e0aa",
        "Late Majority" = "#f7dc6f",
        "Laggards"      = "#e67e22"
      ),
      na.value = "grey90", na.translate = FALSE
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      panel.grid = element_blank(),
      axis.text  = element_blank()
    ) +
    labs(
      title    = sprintf("Projected Adoption Timeline — %s (%s)", sector_name, params$modelName),
      subtitle = sprintf("Threshold: Pr(Adoption) ≥ %.2f", threshold),
      fill     = "Projected Adoption Timeline"
    )

  print(p)
  fname <- sprintf("map_adoption_year_%s", tolower(sector_name))
  ggsave(file.path(plotOutputDir, paste0(fname, ".png")), p, width = 12, height = 6, dpi = 300)
  ggsave(file.path(plotOutputDir, paste0(fname, ".svg")), p, width = 12, height = 6)
}
```

### Bulk

```{r map-adoption-bulk, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario && !is.null(proj_bulk)) {
  tl_bulk <- compute_timeline(proj_bulk, adoption_threshold)
  render_adoption_year_map(tl_bulk, "Bulk", adoption_threshold)
} else {
  cat("> Scenario data not available.")
}
```

### Diffuse

```{r map-adoption-diffuse, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario && !is.null(proj_diffuse)) {
  tl_diffuse <- compute_timeline(proj_diffuse, adoption_threshold)
  render_adoption_year_map(tl_diffuse, "Diffuse", adoption_threshold)
} else {
  cat("> Scenario data not available.")
}
```

## Probability Snapshot Maps {.tabset}

Maps show the predicted adoption probability per region at each snapshot year, using a continuous colour scale (red = low, green = high). The adoption threshold is overlaid as a reference.

```{r snapshot-map-helper, include=FALSE}
render_snapshot_map <- function(proj_df, snap_year, sector_name, threshold) {
  if (is.null(proj_df) || is.null(mapping_54) || is.null(world)) return(NULL)

  available_years <- sort(unique(proj_df$year))
  used_year <- snap_year
  if (!snap_year %in% available_years) {
    used_year <- available_years[which.min(abs(available_years - snap_year))]
  }

  df_year <- proj_df %>% filter(year == used_year) %>% select(region, prob)
  if (nrow(df_year) == 0) return(NULL)

  map_data_df <- mapping_54 %>%
    mutate(
      RegionCode  = as.character(RegionCode),
      CountryCode = as.character(CountryCode),
      CountryCode = ifelse(CountryCode == "SSD", "SDS", CountryCode)
    ) %>%
    left_join(df_year, by = c("RegionCode" = "region"))

  world_data <- world %>% left_join(map_data_df, by = c("iso_key" = "CountryCode"))

  title_suffix <- if (used_year != snap_year)
    sprintf(" (closest year: %d)", used_year) else ""

  p <- ggplot() +
    geom_sf(data = world_data[world_data$iso_key == "ATA", ],
            fill = "grey90", color = "white", linewidth = 0.05) +
    geom_sf(data = world_data[world_data$iso_key != "ATA", ],
            aes(fill = prob), color = "white", linewidth = 0.05) +
    scale_fill_gradientn(
      colors   = c("#e74c3c", "#f39c12", "#f7dc6f", "#2ecc71"),
      limits   = c(0, 1),
      labels   = scales::percent,
      na.value = "grey90",
      name     = "Pr(Adoption)"
    ) +
    geom_sf(
      data = world_data[!is.na(world_data$prob) & world_data$prob >= threshold &
                          world_data$iso_key != "ATA", ],
      fill = NA, color = "#2C3E50", linewidth = 0.4
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      panel.grid = element_blank(),
      axis.text  = element_blank()
    ) +
    labs(
      title    = sprintf("%s Adoption Probability — %d%s", sector_name, snap_year, title_suffix),
      subtitle = sprintf("Dark outline = regions at or above threshold %.2f | Model: %s",
                         threshold, params$modelName)
    )

  print(p)
  fname <- sprintf("map_prob_%s_%d", tolower(sector_name), snap_year)
  ggsave(file.path(plotOutputDir, paste0(fname, ".png")), p, width = 12, height = 6, dpi = 300)
  ggsave(file.path(plotOutputDir, paste0(fname, ".svg")), p, width = 12, height = 6)
}
```

### Bulk {.tabset}

```{r snapshot-maps-bulk, echo=FALSE, results='asis', fig.width=12, fig.height=6}
if (has_scenario && !is.null(proj_bulk) && !is.null(mapping_54)) {
  for (yr in snapshot_years) {
    cat(sprintf("\n#### %d\n\n", yr))
    render_snapshot_map(proj_bulk, yr, "Bulk", adoption_threshold)
    cat("\n\n")
  }
} else {
  cat("> Scenario data or region mapping not available.")
}
```

### Diffuse {.tabset}

```{r snapshot-maps-diffuse, echo=FALSE, results='asis', fig.width=12, fig.height=6}
if (has_scenario && !is.null(proj_diffuse) && !is.null(mapping_54)) {
  for (yr in snapshot_years) {
    cat(sprintf("\n#### %d\n\n", yr))
    render_snapshot_map(proj_diffuse, yr, "Diffuse", adoption_threshold)
    cat("\n\n")
  }
} else {
  cat("> Scenario data or region mapping not available.")
}
```

---

# Appendix {.tabset}

## Model Formula

**Bulk Adoption Model Equation:**

```{r model-formula-bulk, echo=FALSE, comment=""}
cat(pfm::formatModelEquation(
  coefficients = pfm::coeftestToDataFrame(fit_bulk$coeftest),
  actorPowerIndex = "Actor Power Index",
  actorPowerDrivers = NULL,
  instQualityDrivers = model_instQualityDrivers,
  controlDrivers = model_controlDrivers,
  includeLegend = FALSE,
  prefix = "  Pr(Bulk Adoption) ~ "
), "\n")
```

**Diffuse Adoption Model Equation:**

```{r model-formula-diffuse, echo=FALSE, comment=""}
cat(pfm::formatModelEquation(
  coefficients = pfm::coeftestToDataFrame(fit_diffuse$coeftest),
  actorPowerIndex = "Actor Power Index",
  actorPowerDrivers = NULL,
  instQualityDrivers = model_instQualityDrivers,
  controlDrivers = model_controlDrivers,
  includeLegend = TRUE,
  prefix = "  Pr(Diffuse Adoption) ~ "
), "\n")
```

## Adoption Threshold Guide

```{r calc-alternative-thresholds, echo=FALSE}
calc_thresholds_local <- function(fit_obj) {
  if (is.null(fit_obj) || is.null(fit_obj$model)) return(list(Sens = 0.5, SD = 0.5, FP = 0.5))
  m <- fit_obj$model
  hist_probs <- if (inherits(m, "logistf")) m$predict else m$fitted.values
  actual_outcomes <- m$y
  adopters_prob <- hist_probs[actual_outcomes == 1]
  non_adopters_prob <- hist_probs[actual_outcomes == 0]
  
  t_sens <- if (length(adopters_prob) > 0) quantile(adopters_prob, 0.10, na.rm = TRUE) else 0.5
  t_sd   <- if (length(adopters_prob) > 0) mean(adopters_prob, na.rm = TRUE) - sd(adopters_prob, na.rm = TRUE) else 0.5
  t_fp   <- if (length(non_adopters_prob) > 0) quantile(non_adopters_prob, 0.75, na.rm = TRUE) else 0.5
  
  list(
    Sens = round(max(0.01, min(0.99, t_sens)), 3),
    SD   = round(max(0.01, min(0.99, t_sd)), 3),
    FP   = round(max(0.01, min(0.99, t_fp)), 3)
  )
}

thresh_bulk_local <- calc_thresholds_local(fit_bulk)
thresh_diffuse_local <- calc_thresholds_local(fit_diffuse)
```

The default adoption threshold (`r adoption_threshold`) classifies a region as having adopted carbon pricing when the model's predicted probability first exceeds that value.

### Standard Guidance

| Threshold | Interpretation |
|-----------|----------------|
| **0.5** | Standard decision boundary (logit model default) — balanced false positive / false negative rate |
| **< 0.5** | More sensitive — catches early movers but accepts more false positives |
| **> 0.5** | More conservative — higher confidence required before flagging adoption |

### Data-Driven Alternative Thresholds (from Section 10.1 of IAM-PFM Report)

Rather than using an arbitrary threshold like `0.5`, we can calibrate the threshold based on the model's historical predictive performance. Three alternative methods can be used:

1. **Method 1: 90% Historical Sensitivity (Sens)**
   - **Definition:** Finds the threshold where 90% of actual historical adoptions are correctly caught (10th percentile of predicted probabilities among historical adopters).
   - **Rationale:** Minimizes False Negatives (failures to predict adoption when it actually occurred), making it highly sensitive to early signals.
   
2. **Method 2: Standard Deviation Margin (SD)**
   - **Definition:** Mean predicted probability among historical adopters minus one standard deviation.
   - **Rationale:** Includes regions within the normal range of historical early movers, creating a representative threshold for "standard" policy readiness.

3. **Method 3: False Positive Ceiling (FP)**
   - **Definition:** 75th percentile of predicted probabilities among non-adopters.
   - **Rationale:** Accepts some false positives to ensure we do not miss the policy adoption signal, serving as an upper ceiling for classification.

#### Calibrated Threshold Values for Current Model

The calculated values for this specific model configuration (**`r params$modelName`**) are:

| Sector | Method 1: 90% Sensitivity (Sens) | Method 2: Standard Deviation Margin (SD) | Method 3: False Positive Ceiling (FP) |
|--------|----------------------------------|-----------------------------------------|--------------------------------------|
| **Bulk** | **`r sprintf("%.1f%%", thresh_bulk_local$Sens * 100)`** | **`r sprintf("%.1f%%", thresh_bulk_local$SD * 100)`** | **`r sprintf("%.1f%%", thresh_bulk_local$FP * 100)`** |
| **Diffuse** | **`r sprintf("%.1f%%", thresh_diffuse_local$Sens * 100)`** | **`r sprintf("%.1f%%", thresh_diffuse_local$SD * 100)`** | **`r sprintf("%.1f%%", thresh_diffuse_local$FP * 100)`** |

*Note: You can re-render this report and update projections by setting the `adoptionThreshold` parameter (e.g. `adoptionThreshold = 0.35` or `0.28`) in your render call.*

## Session Information

```{r session, echo=FALSE}
tryCatch(
  print(sessionInfo()),
  error = function(e) {
    cat("R Version:", R.version.string, "\n")
    cat("Platform:", R.version$platform, "\n")
  }
)
```
