---
title: "IAM-PFM: Integrated Assessment Model with Political Feasibility Module"
subtitle: "Econometric Model Results, Diagnostics and Validation Report"
author: "Renato Rodrigues"
date: "`r Sys.Date()`"
params:
  modelData:     NULL
  modelDataFile: "data/modelData.RData"
  modelDir:      "cache"
  cachefolder:      ""
  gdxPath:       ""
  projectionsDir: ""
  assetDir:      "output/IAM_PFM_report"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    code_folding: hide
    fig_width: 9
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.align = "center")

library(dplyr)
library(tidyr)
library(kableExtra)
library(patchwork)
library(scales)

# Shared helpers (theme, colours, driver lists)
library(pfmreports)
library(pfmreports)
library(pfmreports)

# Report-specific helpers
library(pfmreports)
library(pfmreports)

# Asset output folder (plots, etc.)
plotOutputDir <- file.path(params$assetDir, "plots")
if (!dir.exists(plotOutputDir)) dir.create(plotOutputDir, recursive = TRUE)

# Load or compute model data
modelData <- if (!is.null(params$modelData) && is.list(params$modelData)) {
  params$modelData
} else {
  loadModelData(params)
}

# Unpack
panelData         <- modelData$panelData
cor_df_AP         <- modelData$cor_df_AP
cor_df_IQ         <- modelData$cor_df_IQ
cor_df_ctrl       <- modelData$cor_df_ctrl
cor_df_vdem       <- modelData$cor_df_vdem
md                <- modelData$md
md2               <- modelData$md2
md3               <- modelData$md3
md4               <- modelData$md4
md5               <- modelData$md5
md6               <- modelData$md6
future_mag        <- modelData$future_mag
future_df_bulk    <- modelData$future_df_bulk
future_df_diffuse <- modelData$future_df_diffuse

diag_results <- diagnoseModelResults(
  workflows = list(md = md, md2 = md2, md3 = md3, md4 = md4, md5 = md5, md6 = md6)
)
```

---

# Summary

The **IAM-PFM** (Integrated Assessment Model with Political Feasibility Module) embeds two-stage hurdle models for carbon-pricing into the REMIND integrated assessment model. For each of two electricity-sector archetypes — **Bulk** (large industrial/utility consumers) and **Diffuse** (transport, buildings, small industry) — the hurdle comprises:

* **Stage 1 — Adoption** (`logistf`): penalised logistic regression predicting whether a jurisdiction has adopted carbon pricing (binary outcome).
* **Stage 2 — Stringency** (gaussian GLM, identity link on log(1+ECP); ADR 0007): conditional on adoption, a GLM predicts log(1+ECP), back-transformed to the Effective Carbon Price (ECP) as `exp(Xβ)−1`.

Four model variants were estimated:

```{r table-variants, echo=FALSE}
table_data <- data.frame(
  Variant = c("**md**", "**md2**", "**md3**", "**md4**", "**md5**", "**md6**"),
  `Actor Power drivers` = c(
    "Full set (VRE share, Electrification, Coal, Oil/Gas, Fossil in Industry)",
    "Actor Power Index composite",
    "Actor Power Index composite",
    "Actor Power Index composite",
    "Actor Power Index composite",
    "Actor Power Index composite"
  ),
  `IQ source` = c(
    "WGI: Government Effectiveness",
    "WGI: Government Effectiveness",
    "WGI: Government Effectiveness",
    "WGI: Government Effectiveness",
    "V-Dem: Rule of Law + Vertical Accountability",
    "WGI + V-Dem: Gov. Effectiveness + Rule of Law (VDem)"
  ),
  `Region Fixed effects` = c(
    "clustered regions (EU+, OECD+, other)",
    "clustered regions (EU+, OECD+, other)",
    "54-region mapping",
    "clustered regions (EU+, OECD+, other)",
    "clustered regions (EU+, OECD+, other)",
    "clustered regions (EU+, OECD+, other)"
  ),
  `Lagged ECP` = c("NO", "NO", "NO", "YES", "NO", "NO"),
  check.names = FALSE
)

kable(table_data, escape = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(1, width = "8%") %>%
  column_spec(2, width = "32%") %>%
  column_spec(3, width = "32%") %>%
  column_spec(4, width = "20%") %>%
  column_spec(5, width = "8%") %>%
  row_spec(5:6, background = "#f0fff0")
```

**Top-line findings:**

```{r exec-summary-bullets, echo=FALSE, results='asis'}
cat(diag_results$globalFindings)
```

---

# Model Architecture {.tabset}

## Conceptual Flow

```{r architecture-diagram, echo=FALSE, fig.height=3}
nodes <- data.frame(
  x     = c(1, 3, 5, 5),
  y     = c(2, 2, 3, 1),
  label = c(
    "Panel Data\n2000–2022", "Stage 1\nAdoption\n(logistf)",
    "Stage 2\nStringency\n(gaussian GLM)", "ECP = 0\n(no policy)"
  ),
  fill  = c("grey70", col_md, col_bulk, "grey85")
)

ggplot(nodes, aes(x, y)) +
  geom_tile(aes(fill = fill), width = 1.4, height = 0.7, colour = "white", linewidth = 1.2) +
  geom_text(aes(label = label), size = 3.2, fontface = "bold", colour = "white") +
  scale_fill_identity() +
  annotate("segment", x = 1.72, xend = 2.28, y = 2, yend = 2,
    arrow = arrow(length = unit(0.25, "cm")), linewidth = 0.8, colour = "grey40") +
  annotate("segment", x = 3.72, xend = 4.28, y = 2.3, yend = 2.85,
    arrow = arrow(length = unit(0.25, "cm")), linewidth = 0.8, colour = col_bulk) +
  annotate("segment", x = 3.72, xend = 4.28, y = 1.7, yend = 1.15,
    arrow = arrow(length = unit(0.25, "cm")), linewidth = 0.8, colour = "grey50") +
  annotate("text", x = 4, y = 2.7, label = "Pr(ECP > 0) = 1", size = 2.8, colour = col_bulk) +
  annotate("text", x = 4, y = 1.3, label = "Pr(ECP > 0) = 0", size = 2.8, colour = "grey50") +
  xlim(0, 6.5) + ylim(0.4, 3.7) +
  labs(title = "IAM-PFM Two-Stage Hurdle Architecture") +
  theme_void() +
  theme(plot.title = element_text(face = "bold", hjust = 0.5, size = 13))
```

## Variable Groups

```{r variable-table, echo=FALSE}
var_df <- data.frame(
  Group = c(
    rep("Actor Power (AP)", 5),
    rep("IQ — WGI (md1–md4)", 1),
    rep("IQ — V-Dem (md5)", 2),
    rep("IQ — Combined (md6)", 2),
    rep("Controls", 2),
    rep("Fixed Effects / Trend", 2)
  ),
  Variable = c(
    "VRE Share", "Electrification", "Coal Primary Energy Share",
    "Oil/Gas Primary Energy Share", "Fossil Share in Industry",
    "Government Effectiveness (WGI)",
    "Rule of Law (VDem)", "Vertical Accountability (VDem)",
    "Government Effectiveness (WGI)", "Rule of Law (VDem)",
    "GDP per Capita", "Population",
    "Region FE (EU / OECD+ / other)", "Time Trend"
  ),
  Role = c(
    rep("Proxy for fossil-fuel incumbents' resistance", 5),
    "State capacity — WGI source",
    "Rule of law quality — V-Dem source", "Electoral accountability — V-Dem source",
    "State capacity — WGI", "Rule of law quality — V-Dem",
    rep("Economic & geographic controls", 2),
    "Captures unobserved regional heterogeneity", "Secular trend in adoption/stringency"
  )
)

kable(var_df, caption = "Model Variable Groups and Roles") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  collapse_rows(columns = 1, valign = "top")
```

## Formal Specification

The model follows a two-stage hurdle specification to capture both the **extensive margin** (policy adoption) and **intensive margin** (policy stringency).

### Stage 1: Adoption (Logit)

Predicts the probability that a jurisdiction $i$ at time $t$ has adopted carbon pricing ($ECP > 0$) using a logit link:

$$\text{logit}(Pr(ECP_{i,t,s} > 0)) = \alpha_{0,s} + \sum_l \alpha_{1,l,s} AP_{i,t-1,s,l} + \sum_k \alpha_{2,k,s} IQ_{i,t-1,s,k} + AP\_Index_{i,t-1,s} \times \sum_k \alpha_{e,k,s} IQ_{i,t-1,s,k} + \sum_j \alpha_{4,j,s} Ctrl_{i,t-1,s,j} + CRFE + TT + \epsilon_{i,t-1,s}$$

### Stage 2: Stringency (Gaussian GLM, identity link on log(1+ECP))

Conditional on adoption, predicts the level of the Effective Carbon Price (ECP) using a log-link:

$$\ln(E[ECP_{i,t,s} | ECP_{i,t,s} > 0]) = \beta_{0,s} + \sum_l \beta_{1,l,s} AP_{i,t-1,s,l} + \sum_k \beta_{2,k,s} IQ_{i,t-1,s,k} + AP\_Index_{i,t-1,s} \times \sum_k \beta_{e,k,s} IQ_{i,t-1,s,k} + \sum_j \beta_{4,j,s} Ctrl_{i,t-1,s,j} + CRFE + TT + \epsilon_{i,t-1,s}$$

**Where:**

* $AP_{i,t-1,s,l}$ denotes **Actor Power** drivers (proxies for fossil-fuel incumbents).
* $IQ_{i,t-1,s,k}$ denotes **Institutional Quality** drivers (state capacity and governance).
* $AP\_Index_{i,t-1,s}$ is the composite **Actor Power Index**.
* $Ctrl_{i,t-1,s,j}$ are geographic and economic **control variables**.
* $CRFE$ are **Carbon-pricing Region Fixed Effects** (capturing EU/OECD+ heterogeneity).
* $TT$ is a linear **Time Trend**.
* All drivers are **lagged by 1 year** ($t-1$) to avoid simultaneity bias.

---

# Correlation Analysis {.tabset}

High correlations between predictors within the same group can inflate variance and destabilise estimates.

## Actor Power Drivers

```{r cor-ap, echo=FALSE}
kableCorrelationMatrix(cor_df_AP, caption = "Actor Power Correlation Matrix",
  show = "lower", percentile = 0.85)
```

```{r cor-ap-note, echo=FALSE, results='asis'}
cor_ap_mat <- as.matrix(cor_df_AP[, -1])
rownames(cor_ap_mat) <- cor_df_AP$Driver
diag(cor_ap_mat) <- NA
high_pair <- which(abs(cor_ap_mat) == max(abs(cor_ap_mat), na.rm = TRUE), arr.ind = TRUE)[1, ]
low_pair  <- which(abs(cor_ap_mat) == min(abs(cor_ap_mat), na.rm = TRUE), arr.ind = TRUE)[1, ]
cat(
  "> **Note:** ",
  rownames(cor_ap_mat)[high_pair[1]], "and", colnames(cor_ap_mat)[high_pair[2]],
  "share r =", round(cor_ap_mat[high_pair[1], high_pair[2]], 2),
  "— moderate but not disqualifying.",
  rownames(cor_ap_mat)[low_pair[1]], "and", colnames(cor_ap_mat)[low_pair[2]],
  "are essentially orthogonal (r =", round(cor_ap_mat[low_pair[1], low_pair[2]], 2), ")."
)
```

## Institutional Quality Drivers (WGI)

```{r cor-iq, echo=FALSE}
kableCorrelationMatrix(cor_df_IQ, caption = "Institutional Quality Correlation Matrix (WGI)",
  show = "lower", percentile = 0.85)
```

```{r cor-iq-note, echo=FALSE, results='asis'}
cor_iq_mat <- as.matrix(cor_df_IQ[, -1])
diag(cor_iq_mat) <- NA
mean_r <- round(mean(abs(cor_iq_mat), na.rm = TRUE), 2)
max_r  <- round(max(abs(cor_iq_mat), na.rm = TRUE), 2)
cat(
  "> **Critical finding:** The World Governance Indicators are highly inter-correlated",
  "(mean |r| =", mean_r, ", max |r| =", max_r, ").",
  "Using more than two simultaneously risks severe multicollinearity.",
  "Select a small subset (e.g. Government Effectiveness (WGI), Voice & Accountability (WGI),",
  "Political Stability (WGI)) and verify VIF < 10 after model fitting."
)
```

## Institutional Quality Drivers (V-Dem)

```{r cor-vdem, echo=FALSE}
kableCorrelationMatrix(cor_df_vdem,
  caption = "Institutional Quality Correlation Matrix — WGI vs V-Dem",
  show = "lower", percentile = 0.85)
```

```{r cor-vdem-note, echo=FALSE, results='asis'}
vdem_cols <- c("Rule of Law (VDem)", "Vertical Accountability (VDem)",
               "Horizontal Accountability (VDem)", "Diagonal Accountability (VDem)")
vdem_present <- intersect(vdem_cols, colnames(cor_df_vdem))
if (length(vdem_present) >= 2) {
  cor_vdem_mat <- as.matrix(cor_df_vdem[cor_df_vdem$Driver %in% vdem_cols, vdem_present])
  diag(cor_vdem_mat) <- NA
  mean_vdem <- round(mean(abs(cor_vdem_mat), na.rm = TRUE), 2)
  max_vdem  <- round(max(abs(cor_vdem_mat), na.rm = TRUE), 2)
  cat(
    "> **V-Dem internal correlation:** The four V-Dem indicators are correlated",
    "(mean |r| =", mean_vdem, ", max |r| =", max_vdem, ").",
    "**md5** uses Rule of Law (VDem) + Vertical Accountability (VDem) as IQ drivers.",
    "**md6** cross-tests Government Effectiveness (WGI) alongside Rule of Law (VDem).",
    "Cross-source pairs near the diagonal of this matrix reveal whether WGI and V-Dem",
    "measure the same construct or capture complementary dimensions."
  )
} else {
  cat("> V-Dem indicators not found in panel data. Ensure V-Dem source data is installed.")
}
```

## Control Variables

```{r cor-ctrl, echo=FALSE}
kableCorrelationMatrix(cor_df_ctrl, caption = "Controls Correlation Matrix",
  show = "lower", percentile = 0.85)
```

```{r cor-ctrl-note, echo=FALSE, results='asis'}
cor_ctrl_mat <- as.matrix(cor_df_ctrl[, -1])
diag(cor_ctrl_mat) <- NA
high_ctrl <- which(abs(cor_ctrl_mat) == max(abs(cor_ctrl_mat), na.rm = TRUE), arr.ind = TRUE)[1, ]
cat(
  "> ",
  rownames(as.data.frame(cor_df_ctrl[, -1]))[high_ctrl[1]],
  "and", colnames(cor_ctrl_mat)[high_ctrl[2]],
  "are strongly correlated (r =",
  round(cor_ctrl_mat[high_ctrl[1], high_ctrl[2]], 2),
  ") and should not both appear in the same model.",
  "Use at most one from each highly correlated pair to keep VIF controlled."
)
```

---

# Model Fit Summary

```{r fit-summary, echo=FALSE}
build_fit_row <- function(wf, model_name, sector, stage) {
  bm <- wf$bestModels
  if (is.null(bm) || nrow(bm) == 0) return(NULL)
  row <- bm[bm$sector == sector & bm$stage == stage, ]
  if (nrow(row) == 0) return(NULL)
  diag <- get_best_diag(wf, sector, stage)
  data.frame(
    Model     = model_name, Sector = sector,
    Stage     = ifelse(stage == "adoption", "Adoption", "Stringency"),
    PseudoR2  = row$pseudoR2, AIC = row$aic,
    nSig      = row$nSignificant, nPred = row$nPredictors,
    Converged = row$converged,
    Separation = isTRUE(diag$separation),
    maxVIF    = row$maxVIF,
    stringsAsFactors = FALSE
  )
}

wfs <- list(
  list(wf = md,  name = "md (full AP)"),
  list(wf = md2, name = "md2 (API only)"),
  list(wf = md3, name = "md3 (REMIND FE)"),
  list(wf = md4, name = "md4 (Lagged ECP)"),
  list(wf = md5, name = "md5 (V-Dem IQ)"),
  list(wf = md6, name = "md6 (WGI+VDem IQ)")
)

grid <- expand.grid(wf_idx = 1:6, sector = c("Bulk", "Diffuse"),
  stage = c("adoption", "stringency"), stringsAsFactors = FALSE)
grid$stage_order  <- ifelse(grid$stage == "adoption", 1, 2)
grid$sector_order <- ifelse(grid$sector == "Bulk", 1, 2)
grid <- grid[order(grid$stage_order, grid$sector_order, grid$wf_idx), ]

fit_df <- do.call(rbind, lapply(1:nrow(grid), function(i) {
  wf_info <- wfs[[grid$wf_idx[i]]]
  build_fit_row(wf_info$wf, wf_info$name, grid$sector[i], grid$stage[i])
}))

kable(fit_df, format = "html", caption = "IAM-PFM Model Fit Summary — All Variants",
  col.names = c("Model", "Sector", "Stage", "Pseudo-R²", "AIC",
    "Sig. pred.", "Total pred.", "Converged", "Sep. flag", "Max VIF")
) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(4, bold = TRUE,
    color = ifelse(fit_df$PseudoR2 >= 0.5, col_good,
      ifelse(fit_df$PseudoR2 >= 0.2, col_warn, col_bad)))
```

```{r fit-summary-narrative, echo=FALSE, results='asis'}
cat(diag_results$fitSummaryNarrative)
```

---

# Model Detailed Results {.tabset}

```{r detailed-results-loop, echo=FALSE, results='asis', fig.height=5}
wf_meta <- list(
  "md"  = "md: Results with Full AP Drivers",
  "md2" = "md2: Results with API Only",
  "md3" = "md3: Results with REMIND Region FE",
  "md4" = "md4: Results with Lagged ECP in Stringency",
  "md5" = "md5: V-Dem IQ (Rule of Law + Vertical Accountability)",
  "md6" = "md6: Combined WGI + V-Dem IQ"
)

for (wfid in names(wf_meta)) {
  wf_obj <- get(wfid)
  cat("\n## ", wf_meta[[wfid]], " {.tabset}\n\n")

  for (sec in c("Bulk", "Diffuse")) {
    cat("\n### ", sec, " {.tabset}\n\n")

    for (stg in c("adoption", "stringency")) {
      cat("\n#### ", tools::toTitleCase(stg), "\n\n")

      coef_df <- get_best_coefs(wf_obj, sec, stg) %>% filter(term != "(Intercept)")

      if (nrow(coef_df) > 0) {
        diag_fit <- get_best_diag(wf_obj, sec, stg)
        stg_lbl  <- ifelse(stg == "adoption", "Log-odds", "ECP")

        p <- render_forest_plot(
          coef_df,
          title    = sprintf("%s %s — Coefficient Plot", sec, tools::toTitleCase(stg)),
          subtitle = sprintf(
            "%s coefficients | Pseudo-R² = %.3f | AIC = %.1f | %d/%d significant",
            stg_lbl, diag_fit$pseudoR2, diag_fit$aic,
            diag_fit$nSignificant, diag_fit$nPredictors
          )
        )

        if (!is.null(p) && stg == "stringency") {
          extreme_data <- coef_df %>% filter(abs(statistic) > 15)
          if (nrow(extreme_data) > 0) {
            p <- p + geom_text(data = extreme_data,
              aes(label = "⚠ Extreme z", y = estimate),
              nudge_y = -0.8, size = 3, colour = col_bad)
          }
        }

        if (!is.null(p)) {
          print(p)
          fname <- sprintf("forest_%s_%s_%s", wfid, sec, stg)
          ggsave(file.path(plotOutputDir, paste0(fname, ".png")), p, width = 10, height = 6, dpi = 300)
          ggsave(file.path(plotOutputDir, paste0(fname, ".svg")), p, width = 10, height = 6)
        }
        cat("\n\n")

        d <- diag_results$modelDiagnoses[[wfid]][[paste0(sec, "_", stg)]]
        if (!is.null(d)) {
          cat("**Top predictors** (by |z|):", paste0("**", vapply(d$topPredictors, clean_term_plain, character(1)), "**", collapse = ", "), "\n\n")
          if (length(d$extremeZTerms) > 0)
            cat("> ⚠ **Extreme z-values detected** on:", paste(d$extremeZTerms, collapse = ", "), "\n\n")
          if (isTRUE(d$separation) && stg == "adoption")
            cat("> ℹ Separation flag detected — logistf handles this via penalised ML; widen prediction intervals.\n\n")
          if (wfid == "md2" && stg == "adoption") {
            d_md <- diag_results$modelDiagnoses$md[[paste0(sec, "_", stg)]]
            if (!is.null(d_md)) {
              delta_r2  <- round(d_md$r2 - d$r2, 3)
              delta_aic <- round(d$aic - d_md$aic, 1)
              cat(sprintf("> ℹ **Comparison with md:** ΔPseudo-R² = %.3f, ΔAIC = %.1f.\n\n", delta_r2, delta_aic))
            }
          }
          cat("> ", d$verdictIcon, "**Verdict:**", d$verdict$text, "\n")
        }
      } else {
        cat("> ❌ **Model failed to converge or select a valid specification.**\n")
      }
      cat("\n\n")
    }
  }
}
```

---

# Model Comparison

## Adoption

```{r comparison-adoption, echo=FALSE, fig.height=4}
comp_adopt <- do.call(rbind, lapply(names(diag_results$modelDiagnoses), function(wfn) {
  wf  <- get(wfn)
  lbl <- switch(wfn, "md" = "md (full AP)", "md2" = "md2 (API only)",
    "md3" = "md3 (REMIND FE)", "md4" = "md4 (Lagged ECP)", wfn)
  if (!is.null(wf$bestModels) && nrow(wf$bestModels) > 0)
    wf$bestModels[wf$bestModels$stage == "adoption", c("sector", "pseudoR2", "aic")] %>%
      mutate(Model = lbl)
  else NULL
}))

variant_colours <- c("md (full AP)" = col_md, "md2 (API only)" = col_md2,
  "md3 (REMIND FE)" = col_md3, "md4 (Lagged ECP)" = col_md4)

p1 <- ggplot(comp_adopt, aes(x = sector, y = pseudoR2, fill = Model)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", pseudoR2)),
    position = position_dodge(0.6), vjust = -0.4, size = 3.5) +
  scale_fill_manual(values = variant_colours) +
  labs(title = "Adoption — Pseudo-R²", y = "McFadden Pseudo-R²", x = NULL) +
  ylim(0, 0.95) + theme_report()

p2 <- ggplot(comp_adopt, aes(x = sector, y = aic, fill = Model)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = sprintf("%.1f", aic)),
    position = position_dodge(0.6), vjust = -0.4, size = 3.5) +
  scale_fill_manual(values = variant_colours) +
  labs(title = "Adoption — AIC (lower = better)", y = "AIC", x = NULL) +
  theme_report()

p1 + p2 + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

```{r comparison-adoption-narrative, echo=FALSE, results='asis'}
cat(">", diag_results$comparisonNarrative$adoption, "\n")
```

## Stringency

```{r comparison-stringency, echo=FALSE, fig.height=4}
comp_str <- do.call(rbind, lapply(names(diag_results$modelDiagnoses), function(wfn) {
  wf  <- get(wfn)
  lbl <- switch(wfn, "md" = "md (full AP)", "md2" = "md2 (API only)",
    "md3" = "md3 (REMIND FE)", "md4" = "md4 (Lagged ECP)", wfn)
  if (!is.null(wf$bestModels) && nrow(wf$bestModels) > 0)
    wf$bestModels[wf$bestModels$stage == "stringency", c("sector", "pseudoR2", "aic")] %>%
      mutate(Model = lbl)
  else NULL
}))

p3 <- ggplot(comp_str, aes(x = sector, y = pseudoR2, fill = Model)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", pseudoR2)),
    position = position_dodge(0.6), vjust = -0.4, size = 3.5) +
  scale_fill_manual(values = variant_colours) +
  labs(title = "Stringency — Pseudo-R²", y = "Pseudo-R²", x = NULL) +
  ylim(0, max(comp_str$pseudoR2) * 1.3) + theme_report()

p4 <- ggplot(comp_str, aes(x = sector, y = aic, fill = Model)) +
  geom_col(position = "dodge", width = 0.6) +
  geom_text(aes(label = sprintf("%.0f", aic)),
    position = position_dodge(0.6), vjust = -0.4, size = 3.5) +
  scale_fill_manual(values = variant_colours) +
  labs(title = "Stringency — AIC", y = "AIC", x = NULL) + theme_report()

p3 + p4 + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

```{r comparison-stringency-narrative, echo=FALSE, results='asis'}
cat(diag_results$comparisonNarrative$stringency, "\n")
```

---

# Diagnostic Deep-Dive

## Workflow Summary {.tabset}

### md

```{r workflow-summary-md, echo=FALSE, comment=""}
cat(md$workflowSummary)
```

### md2

```{r workflow-summary-md2, echo=FALSE, comment=""}
cat(md2$workflowSummary)
```

### md3

```{r workflow-summary-md3, echo=FALSE, comment=""}
cat(md3$workflowSummary)
```

### md4

```{r workflow-summary-md4, echo=FALSE, comment=""}
cat(md4$workflowSummary)
```

## Group Analysis {.tabset}

### md

```{r group-analysis-md, echo=FALSE}
kable(md$groupAnalysisTable, format = "html",
  caption = "Group-Level Variance Partitioning — md (Full AP Drivers)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

```{r group-analysis-md-findings, echo=FALSE, results='asis'}
ga_md <- md$groupAnalysisTable
ga_md$.r2num <- as.numeric(gsub("%", "", ga_md$`Pseudo-R2 contribution`))
adopt_top <- ga_md[ga_md$stage == "adoption", ]
adopt_top <- adopt_top[order(adopt_top$.r2num, decreasing = TRUE), ]
str_top   <- ga_md[ga_md$stage == "stringency", ]
str_top   <- str_top[order(str_top$.r2num, decreasing = TRUE), ]
cat("**Findings:** The group analysis reveals how much explanatory power each predictor block contributes independently.\n\n")
if (nrow(adopt_top) > 0)
  cat("- **Adoption:** Top contributing group is", adopt_top$group[1],
    sprintf("(%s of variance explained).\n", adopt_top$`Pseudo-R2 contribution`[1]))
if (nrow(str_top) > 0)
  cat("- **Stringency:** Top contributing group is", str_top$group[1],
    sprintf("(%s).", str_top$`Pseudo-R2 contribution`[1]),
    "If Region Fixed Effects dominate, much of the explained variance reflects",
    "between-region differences rather than within-region temporal dynamics,",
    "which limits out-of-sample projection utility.\n")
```

### md2

```{r group-analysis-md2, echo=FALSE}
kable(md2$groupAnalysisTable, format = "html",
  caption = "Group-Level Variance Partitioning — md2 (API Only)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

```{r group-analysis-md2-findings, echo=FALSE, results='asis'}
ga_md2 <- md2$groupAnalysisTable
ga_md2$.r2num <- as.numeric(gsub("%", "", ga_md2$`Pseudo-R2 contribution`))
adopt_top2 <- ga_md2[ga_md2$stage == "adoption", ]
adopt_top2 <- adopt_top2[order(adopt_top2$.r2num, decreasing = TRUE), ]
cat("**Findings:** The md2 group analysis shows a cleaner decomposition.\n\n")
if (nrow(adopt_top2) > 0)
  cat("- Without individual AP drivers competing for variance, the top contributor is",
    adopt_top2$group[1],
    sprintf("(%s of variance explained),", adopt_top2$`Pseudo-R2 contribution`[1]),
    "suggesting the composite API captures political-economy dynamics more efficiently",
    "than disaggregated fuel-specific proxies.\n")
```

### md3

```{r group-analysis-md3, echo=FALSE}
kable(md3$groupAnalysisTable, format = "html",
  caption = "Group-Level Variance Partitioning — md3 (REMIND FE)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

### md4

```{r group-analysis-md4, echo=FALSE}
kable(md4$groupAnalysisTable, format = "html",
  caption = "Group-Level Variance Partitioning — md4 (Lagged ECP)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
```

## Pseudo-R² Comparison

```{r r2-context, echo=FALSE, fig.height=4.5}
r2_df <- do.call(rbind, lapply(names(diag_results$modelDiagnoses), function(wfn) {
  res <- get(wfn)$bestModels
  if (!is.null(res) && nrow(res) > 0)
    res %>% mutate(Model = paste0(sector, " ", stage, " (", wfn, ")"))
  else NULL
})) %>%
  mutate(
    Quality      = case_when(pseudoR2 >= 0.4 ~ "Good", pseudoR2 >= 0.2 ~ "Moderate", TRUE ~ "Poor"),
    stage_order  = ifelse(stage == "adoption", 1, 2),
    sector_order = ifelse(sector == "Bulk", 1, 2)
  ) %>%
  arrange(desc(stage_order), desc(sector_order), Model) %>%
  mutate(Model = factor(Model, levels = unique(Model)))

p <- ggplot(r2_df, aes(x = Model, y = pseudoR2,
    fill = factor(Quality, levels = c("Poor", "Moderate", "Good")))) +
  geom_col(width = 0.6) +
  geom_text(aes(label = sprintf("%.3f", pseudoR2)), hjust = -0.15, size = 3.5) +
  scale_fill_manual(values = c(col_bad, col_warn, col_good), name = "Fit quality") +
  coord_flip() + ylim(0, 1) +
  labs(title = "McFadden Pseudo-R² Across All Models",
    subtitle = "Rules of thumb: < 0.2 = poor, 0.2–0.4 = moderate, > 0.4 = good",
    x = NULL, y = "Pseudo-R²") +
  theme_report()
print(p)
ggsave(file.path(plotOutputDir, "model_comparison_r2.png"), p, width = 10, height = 5, dpi = 300)
ggsave(file.path(plotOutputDir, "model_comparison_r2.svg"), p, width = 10, height = 5)
```

```{r r2-suitability-narrative, echo=FALSE, results='asis'}
r2_tiers <- lapply(names(diag_results$modelDiagnoses), function(wfn) {
  lapply(diag_results$modelDiagnoses[[wfn]], function(d) {
    list(model = paste0(d$sector, " ", d$stage, " (", wfn, ")"),
      r2 = d$r2, ready = d$iamReady, issues = length(d$extremeZTerms) > 0)
  })
})
r2_flat <- do.call(c, r2_tiers)
r2_good <- Filter(function(x) !is.na(x$r2) && x$r2 >= 0.4, r2_flat)
r2_mod  <- Filter(function(x) !is.na(x$r2) && x$r2 >= 0.2 && x$r2 < 0.4, r2_flat)
r2_poor <- Filter(function(x) !is.na(x$r2) && x$r2 < 0.2, r2_flat)
cat("**Findings & Model Suitability Summary:**\n\n")
if (length(r2_good) > 0)
  cat("* **Top tier (Pseudo-R2 >= 0.4):**",
    paste(sapply(r2_good, `[[`, "model"), collapse = "; "),
    "-- suitable for IAM projection where no separation issues exist.\n")
for (m in r2_mod) {
  suffix <- if (m$issues) " Despite the moderate R2, underlying separation inflates this statistic -- treat as unreliable." else ""
  cat("* **Middle tier:**", m$model, sprintf("(R2 = %.3f).", m$r2), suffix, "\n")
}
for (m in r2_poor)
  cat("* **Bottom tier:**", m$model, sprintf("(R2 = %.3f) --", m$r2),
    if (m$issues) "critical separation detected; not suitable." else "low fit; use as conservative baseline.", "\n")
ready_models <- sapply(Filter(function(x) x$ready == "yes", r2_flat), `[[`, "model")
cat("\n> **Actionable conclusion:** Use",
  if (length(ready_models) > 0) paste(ready_models, collapse = ", ") else "none",
  "for IAM deployment. Models with critical separation require structural redesign.\n")
```

---

# Issues, Warnings and Recommendations

## Summary Issue Table

```{r issues, echo=FALSE}
if (nrow(diag_results$issuesTable) > 0) {
  kable(diag_results$issuesTable, format = "html",
    caption = "Dynamically detected issues, consequences and recommended fixes",
    col.names = c("Priority", "Model", "Issue", "Consequence", "Fix")) %>%
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
    column_spec(1, bold = TRUE) %>%
    collapse_rows(columns = 1)
} else {
  cat("> No issues detected across all models.")
}
```

---

# IAM Integration Notes

## What These Models Enable

The IAM-PFM module translates IAM-projected socioeconomic and energy pathways into **probabilistic carbon price trajectories** using the estimated hurdle parameters:

**1. Adoption Probability:**
$$\Pr(\text{ECP}_t > 0) = \text{logit}^{-1}(\hat{\eta}^{(\text{adopt})}_t)$$

**2. Conditional Stringency:**
$$\mathbb{E}[\text{ECP}_t \mid \text{ECP}_t > 0] = \exp(\hat{\eta}^{(\text{string})}_t)$$

**3. Expected Effective Carbon Price:**
$$\mathbb{E}[\text{ECP}_t] = \Pr(\text{ECP}_t > 0) \times \mathbb{E}[\text{ECP}_t \mid \text{ECP}_t > 0]$$

## Projection-Ready Models

```{r readiness, echo=FALSE}
kable(diag_results$readinessTable, format = "html",
  caption = "IAM Projection Readiness Assessment (dynamically computed)",
  col.names = c("Model", "Status", "Rationale")) %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) %>%
  column_spec(2, bold = TRUE,
    color = ifelse(grepl("✅", diag_results$readinessTable$Ready), col_good,
      ifelse(grepl("⚠", diag_results$readinessTable$Ready), col_warn, col_bad)))
```

```{r summary-verdict, echo=FALSE, results='asis'}
rd       <- diag_results$readinessTable
n_ready  <- sum(grepl("✅", rd$Ready))
n_monitor <- sum(grepl("⚠", rd$Ready))
n_not    <- sum(grepl("❌", rd$Ready))
cat("**Summary Verdict:**\n\nOf the", nrow(rd), "model variants assessed:\n\n")
if (n_ready > 0)   cat("*", n_ready,   "are **fully ready** for IAM deployment.\n")
if (n_monitor > 0) cat("*", n_monitor, "**require monitoring but are usable** (widen confidence intervals or bound projections).\n")
if (n_not > 0)     cat("*", n_not,     "**cannot be used reliably** due to structural separation. These need fundamental redesign before deployment.\n")
cat("\nFor immediate IAM integration, deploy the models marked Ready or Monitor, and replace",
  "models flagged Not Ready with externally calibrated alternatives until separation issues are resolved.\n")
```

## Key Projection Caveats

```{r projection-caveats, echo=FALSE, results='asis'}
cat(diag_results$projectionCaveats$asNumberedList)
```

## Policy-Scenario Contrast

```{r scenario-contrast, echo=FALSE, results='asis', fig.height=4}
# Locate the projection fan-out (ADR 0035): an explicit projectionsDir, else auto-discover a
# `projections/` directory (with *.rds) under the modelData / modelDir / asset roots.
.findProjDir <- function() {
  cand <- character(0)
  if (!is.null(params$projectionsDir) && nzchar(params$projectionsDir)) cand <- params$projectionsDir
  roots <- unique(c(dirname(params$modelDataFile), params$modelDir, dirname(params$assetDir), getwd()))
  roots <- roots[!is.na(roots) & nzchar(roots)]
  for (r in roots) cand <- c(cand, file.path(r, "projections"), Sys.glob(file.path(r, "*", "projections")))
  cand <- unique(cand)
  hit <- cand[vapply(cand, function(d) dir.exists(d) && length(list.files(d, pattern = "\\.rds$")) > 0, logical(1))]
  if (length(hit)) hit[1] else NULL
}
projAll <- tryCatch(pfmreports::loadGroupProjections(.findProjDir()), error = function(e) NULL)
if (!is.null(projAll) && length(unique(projAll$scenario)) >= 2) {
  ord <- pfmreports::scenarioOrder(projAll, 2050)                       # least -> most ambitious
  nm  <- stats::setNames(projAll$scenarioName[match(ord, projAll$scenario)], ord)
  pa  <- projAll; pa$scenarioName <- factor(pa$scenarioName, levels = nm)
  hi  <- ord[length(ord)]; lo <- ord[1]
  adoptM <- stats::aggregate(prob ~ scenarioName + year, data = pa, FUN = mean)
  g1 <- ggplot(adoptM, aes(year, prob, colour = scenarioName)) + geom_line(linewidth = 0.9) +
    coord_cartesian(ylim = c(0, 1)) + labs(title = "Mean P(adoption)", x = NULL, y = NULL, colour = NULL) +
    theme_minimal() + theme(legend.position = "top")
  priceM <- stats::aggregate(expectedPrice ~ scenarioName + year, data = pa, FUN = stats::median)
  pcap <- as.numeric(stats::quantile(pa$expectedPrice, 0.99, na.rm = TRUE)); priceM$expectedPrice <- pmin(priceM$expectedPrice, pcap)
  g2 <- ggplot(priceM, aes(year, expectedPrice, colour = scenarioName)) + geom_line(linewidth = 0.9) +
    labs(title = "Median expected price (USD/tCO2)", x = NULL, y = NULL, colour = NULL) +
    theme_minimal() + theme(legend.position = "top")
  print(g1 + g2 + patchwork::plot_layout(guides = "collect") & theme(legend.position = "top"))
  cat("\n\n")
  m <- merge(pa[pa$scenario == hi, c("region", "year", "prob")],
             pa[pa$scenario == lo, c("region", "year", "prob")],
             by = c("region", "year"), suffixes = c(".hi", ".lo"))
  ag <- stats::aggregate(abs(m$prob.hi - m$prob.lo), list(year = m$year), mean); py <- ag$year[which.max(ag$x)]
  cat(sprintf(paste0("Deployed model projected onto **%s** vs **%s** (both SSP2, one shared fit, ADR 0035): ",
    "near-identical near-term and diverging by mid-century — the adoption-probability gap peaks ~**%d** ",
    "(mean %.0f percentage points), with higher ambition also lifting the feasible price path. These are ",
    "relative feasibility signals (above-historical-range extrapolation), not forecasts.\n\n"),
    nm[length(nm)], nm[1], py, 100 * max(ag$x)))
} else {
  cat("*Policy-scenario contrast unavailable: no multi-scenario `projections/*.rds` found. Pass ",
      "`projectionsDir=<group>/projections` or run the `projection` fan-out over a multi-scenario ",
      "registry (ADR 0035).*\n\n")
}
```

---

# Projection Results

## Adoption Threshold {.tabset}

```{r calc-thresholds, echo=FALSE}
calc_thresholds <- function(model_obj) {
  if (is.null(model_obj)) return(list(Sens = 0.5, SD = 0.5, FP = 0.5))
  hist_probs       <- if (!is.null(model_obj$predict)) model_obj$predict else model_obj$fitted.values
  actual_outcomes  <- model_obj$y
  adopters_prob    <- hist_probs[actual_outcomes == 1]
  non_adopters_prob <- hist_probs[actual_outcomes == 0]
  t1 <- if (length(adopters_prob) > 0)     quantile(adopters_prob, 0.10, na.rm = TRUE) else 0.5
  t2 <- if (length(adopters_prob) > 0)     mean(adopters_prob, na.rm = TRUE) - sd(adopters_prob, na.rm = TRUE) else 0.5
  t3 <- if (length(non_adopters_prob) > 0) quantile(non_adopters_prob, 0.75, na.rm = TRUE) else 0.5
  list(
    Sens = round(max(0.25, min(0.99, t1)), 2),
    SD   = round(max(0.25, min(0.99, t2)), 2),
    FP   = round(max(0.25, min(0.99, t3)), 2)
  )
}

thresh_bulk    <- calc_thresholds(md2$selections$Bulk_adoption$bestModel$model)
thresh_diffuse <- calc_thresholds(md2$selections$Diffuse_adoption$bestModel$model)
```

### Bulk

- **Method 1: 90% Historical Sensitivity (`r thresh_bulk$Sens`)**: Finds the threshold where 90% of historical adoptions are caught. Minimises False Negatives.
- **Method 2: Standard Deviation Margin (`r thresh_bulk$SD`)**: Mean adoption probability minus one SD — includes regions within the normal range of historical early movers.
- **Method 3: False Positive Ceiling (`r thresh_bulk$FP`)**: 75th percentile of probabilities for non-adopters. Accepts some false positives to avoid missing the signal.

### Diffuse

- **Method 1: 90% Historical Sensitivity (`r thresh_diffuse$Sens`)**: Finds the threshold where 90% of historical adoptions are caught. Minimises False Negatives.
- **Method 2: Standard Deviation Margin (`r thresh_diffuse$SD`)**: Mean adoption probability minus one SD — includes regions within the normal range of historical early movers.
- **Method 3: False Positive Ceiling (`r thresh_diffuse$FP`)**: 75th percentile of probabilities for non-adopters. Accepts some false positives to avoid missing the signal.

## Projected Probability {.tabset}

```{r proj-prob, echo=FALSE, results='asis', fig.height=5, fig.width=10, warning=FALSE, message=FALSE}
if (!is.null(future_mag)) {
  get_projection_plot <- function(df, model_obj, sector_name, threshold) {
    if (is.null(model_obj)) return(NULL)
    needed_vars <- intersect(all.vars(model_obj$formula), colnames(df))
    df_clean    <- df[complete.cases(df[, needed_vars, drop = FALSE]), ]
    if (nrow(df_clean) == 0) return(NULL)
    pred_link <- tryCatch(predict(model_obj, newdata = df_clean), error = function(e) NULL)
    if (is.null(pred_link)) return(NULL)
    df_clean$prob <- 1 / (1 + exp(-pred_link))

    top_regions      <- intersect(c("USA", "CHN", "IND", "JPN", "DEU", "GBR", "FRA", "BRA", "RUS"), unique(df_clean$region))
    df_clean$highlight <- ifelse(df_clean$region %in% top_regions, as.character(df_clean$region), "Other")
    df_clean$highlight <- factor(df_clean$highlight, levels = c(top_regions, "Other"))
    my_colors <- setNames(
      c("#1f77b4", "#d62728", "#ff7f0e", "#9467bd", "#2ca02c", "#8c564b", "#e377c2", "#17becf", "#7f7f7f")[1:length(top_regions)],
      top_regions
    )
    my_colors["Other"] <- "grey80"

    p <- ggplot(df_clean, aes(x = year, y = prob, group = region, color = highlight,
        alpha = highlight == "Other", linewidth = highlight != "Other")) +
      geom_hline(yintercept = threshold, linetype = "dashed", color = "grey50") +
      geom_line() +
      scale_color_manual(values = my_colors) +
      scale_alpha_manual(values = c("TRUE" = 0.4, "FALSE" = 1), guide = "none") +
      scale_linewidth_manual(values = c("TRUE" = 0.5, "FALSE" = 1.2), guide = "none") +
      theme_report() +
      labs(title = sprintf("Projected Probability of Adoption (%s)", sector_name),
        subtitle = sprintf("Dashed line = threshold %.2f (90%% historical sensitivity)", threshold),
        x = "Year", y = "Pr(Adoption)", color = "Region") +
      theme(legend.position = "right")

    fname <- sprintf("prob_%s_%.2f", sector_name, threshold)
    ggsave(file.path(plotOutputDir, paste0(fname, ".png")), p, width = 10, height = 5, dpi = 300)
    ggsave(file.path(plotOutputDir, paste0(fname, ".svg")), p, width = 10, height = 5)
    p
  }

  for (sec in c("Bulk", "Diffuse")) {
    cat("\n### ", sec, " {.tabset}\n\n")
    df      <- if (sec == "Bulk") future_df_bulk else future_df_diffuse
    model_obj <- md2$selections[[paste0(sec, "_adoption")]]$bestModel$model
    t_list  <- if (sec == "Bulk") thresh_bulk else thresh_diffuse
    for (t_name in c("Sens", "SD", "FP")) {
      cat("\n#### ", t_name, "\n\n")
      p <- get_projection_plot(df, model_obj, sec, t_list[[t_name]])
      if (!is.null(p)) print(p)
      cat("\n\n")
    }
  }
}
```

## Expected Adoption Timeline {.tabset}

```{r timeline-tables, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
if (!is.null(future_mag)) {
  render_adoption_timeline <- function(df, model_obj, threshold) {
    if (is.null(model_obj)) return(NULL)
    needed_vars <- intersect(all.vars(model_obj$formula), colnames(df))
    df_clean    <- df[complete.cases(df[, needed_vars, drop = FALSE]), ]
    if (nrow(df_clean) == 0) return(NULL)
    pred_link <- tryCatch(predict(model_obj, newdata = df_clean), error = function(e) NULL)
    if (is.null(pred_link)) return(NULL)
    df_clean$prob <- 1 / (1 + exp(-pred_link))

    adopt_data <- df_clean %>%
      group_by(region) %>%
      summarise(AdoptionYear = ifelse(any(prob >= threshold), min(year[prob >= threshold]), Inf), .groups = "drop") %>%
      mutate(
        Category = case_when(
          AdoptionYear <= 2030 ~ "Forerunners", AdoptionYear <= 2050 ~ "Early Majority",
          AdoptionYear <= 2070 ~ "Late Majority", TRUE ~ "Laggards"
        ),
        YearRange = case_when(
          Category == "Forerunners"    ~ "<= 2030",   Category == "Early Majority" ~ "2031–2050",
          Category == "Late Majority"  ~ "2051–2070", Category == "Laggards"       ~ "> 2070"
        ),
        RegYear = ifelse(AdoptionYear == Inf, region, sprintf("%s (%s)", region, AdoptionYear))
      )

    final_table <- adopt_data %>%
      group_by(Category, YearRange) %>%
      summarise(Regions = paste(RegYear, collapse = ", "), .groups = "drop") %>%
      mutate(Category = factor(Category, levels = c("Forerunners", "Early Majority", "Late Majority", "Laggards"))) %>%
      arrange(Category)

    cat(as.character(
      kable(final_table, format = "html", col.names = c("Category", "Expected Adoption Year", "Regions")) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
    ))
    cat("\n\n")
  }

  for (sec in c("Bulk", "Diffuse")) {
    cat("\n### ", sec, " {.tabset}\n\n")
    df        <- if (sec == "Bulk") future_df_bulk else future_df_diffuse
    model_obj <- md2$selections[[paste0(sec, "_adoption")]]$bestModel$model
    t_list    <- if (sec == "Bulk") thresh_bulk else thresh_diffuse
    for (t_name in c("Sens", "SD", "FP")) {
      cat("\n#### ", t_name, "\n\n")
      render_adoption_timeline(df, model_obj, t_list[[t_name]])
      cat("\n\n")
    }
  }
}
```

## Spatial Distribution of Adoption {.tabset}

```{r spatial-maps, echo=FALSE, results='asis', fig.height=6, fig.width=10, warning=FALSE, message=FALSE}
if (!is.null(future_mag)) {
  mapping_54 <- tryCatch(
    madrat::toolGetMapping("regionmapping_54.csv", type = "regional", where = "mappingfolder"),
    error = function(e) NULL
  )

  if (!is.null(mapping_54)) {
    suppressPackageStartupMessages(library(sf))
    world     <- rnaturalearth::ne_countries(scale = "small", returnclass = "sf")
    world$iso_key <- as.character(world$adm0_a3)

    render_adoption_map <- function(df, model_obj, sector_name, threshold) {
      needed_vars <- intersect(all.vars(model_obj$formula), colnames(df))
      df_clean    <- df[complete.cases(df[, needed_vars, drop = FALSE]), ]
      pred_link   <- tryCatch(predict(model_obj, newdata = df_clean), error = function(e) NULL)
      if (is.null(pred_link)) return(NULL)
      df_clean$prob <- 1 / (1 + exp(-pred_link))

      adopt_data <- df_clean %>%
        group_by(region) %>%
        summarise(AdoptionYear = ifelse(any(prob >= threshold), min(year[prob >= threshold]), 2151), .groups = "drop") %>%
        mutate(Category = case_when(
          AdoptionYear <= 2030 ~ "Forerunners", AdoptionYear <= 2050 ~ "Early Majority",
          AdoptionYear <= 2070 ~ "Late Majority", TRUE ~ "Laggards"
        ))

      map_data_df <- mapping_54 %>%
        mutate(RegionCode = as.character(RegionCode), CountryCode = as.character(CountryCode)) %>%
        mutate(CountryCode = ifelse(CountryCode == "SSD", "SDS", CountryCode)) %>%
        left_join(adopt_data, by = c("RegionCode" = "region")) %>%
        mutate(
          Category = if_else(CountryCode == "ATA", factor(NA, levels = levels(Category)), Category),
          Category = factor(Category, levels = c("Forerunners", "Early Majority", "Late Majority", "Laggards"))
        )

      world_data <- world %>% left_join(map_data_df, by = c("iso_key" = "CountryCode"))

      p <- ggplot() +
        geom_sf(data = world_data[world_data$iso_key == "ATA", ], fill = "grey90", color = "white", size = 0.05) +
        geom_sf(data = world_data[world_data$iso_key != "ATA", ], aes(fill = Category), color = "white", size = 0.05) +
        scale_fill_manual(values = c(
          "Forerunners"   = "#2ecc71", "Early Majority" = "#82e0aa",
          "Late Majority" = "#f7dc6f", "Laggards"       = "#e67e22"
        ), na.value = "grey90", na.translate = FALSE) +
        theme_minimal() +
        theme(legend.position = "bottom", panel.grid = element_blank(), axis.text = element_blank()) +
        labs(title = sprintf("Projected Adoption Timeline Categories (%s)", sector_name),
          subtitle = sprintf("Threshold: Pr(Adoption) > %s", threshold),
          fill = "Projected Adoption Timeline")

      print(p)
      fname <- sprintf("map_%s_%.2f", sector_name, threshold)
      ggsave(file.path(plotOutputDir, paste0(fname, ".png")), width = 10, height = 6, dpi = 300)
      ggsave(file.path(plotOutputDir, paste0(fname, ".svg")), width = 10, height = 6)
    }

    for (sec in c("Bulk", "Diffuse")) {
      cat("### ", sec, " {.tabset}\n\n")
      df        <- if (sec == "Bulk") future_df_bulk else future_df_diffuse
      model_obj <- md2$selections[[paste0(sec, "_adoption")]]$bestModel$model
      t_list    <- if (sec == "Bulk") thresh_bulk else thresh_diffuse
      for (t_name in c("Sens", "SD", "FP")) {
        cat("#### ", t_name, "\n\n")
        render_adoption_map(df, model_obj, sec, t_list[[t_name]])
        cat("\n\n")
      }
    }
  } else {
    cat("> ⚠️ **Could not load region mapping file.** Ensure `regionmapping_54.csv` is in the mapping folder.\n")
  }
}
```

---

# Appendix {.tabset}

## Selection Paths (md)

```{r selection-paths-md, echo=FALSE}
sp_md <- md$selectionPaths %>%
  select(sector, stage, step, description, pseudoR2, aic, nSignificant, nPredictors, converged)
kable(sp_md, format = "html", caption = "Model Selection Path — md (Full AP Drivers)", digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
```

## Selection Paths (md2)

```{r selection-paths-md2, echo=FALSE}
sp_md2 <- md2$selectionPaths %>%
  select(sector, stage, step, description, pseudoR2, aic, nSignificant, nPredictors, converged)
kable(sp_md2, format = "html", caption = "Model Selection Path — md2 (API Only)", digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
```

## Selection Paths (md3)

```{r selection-paths-md3, echo=FALSE}
if (!is.null(md3$selectionPaths)) {
  sp_md3 <- md3$selectionPaths %>%
    select(sector, stage, step, description, pseudoR2, aic, nSignificant, nPredictors, converged)
  kable(sp_md3, format = "html", caption = "Model Selection Path — md3 (REMIND Region FE)", digits = 3) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
} else {
  cat("No data available.")
}
```

## Selection Paths (md4)

```{r selection-paths-md4, echo=FALSE}
if (!is.null(md4$selectionPaths)) {
  sp_md4 <- md4$selectionPaths %>%
    select(sector, stage, step, description, pseudoR2, aic, nSignificant, nPredictors, converged)
  kable(sp_md4, format = "html", caption = "Model Selection Path — md4 (Lagged ECP)", digits = 3) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
} else {
  cat("No data available.")
}
```

## Best Coefficients (md)

```{r best-coefficients-md, echo=FALSE}
kable(md$bestCoefficients, format = "html", caption = "Best Model Coefficients — md", digits = 4) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
```

## Best Coefficients (md2)

```{r best-coefficients-md2, echo=FALSE}
kable(md2$bestCoefficients, format = "html", caption = "Best Model Coefficients — md2", digits = 4) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
```

## Best Coefficients (md3)

```{r best-coefficients-md3, echo=FALSE}
if (!is.null(md3$bestCoefficients)) {
  kable(md3$bestCoefficients, format = "html", caption = "Best Model Coefficients — md3", digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
} else {
  cat("No data available.")
}
```

## Best Coefficients (md4)

```{r best-coefficients-md4, echo=FALSE}
if (!is.null(md4$bestCoefficients)) {
  kable(md4$bestCoefficients, format = "html", caption = "Best Model Coefficients — md4", digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE, font_size = 11)
} else {
  cat("No data available.")
}
```

## Session Information

```{r session, echo=FALSE}
tryCatch(
  print(sessionInfo()),
  error = function(e) {
    cat("Note: sessionInfo() could not be fully rendered.\n")
    cat("R Version:", R.version.string, "\n")
    cat("Platform:", R.version$platform, "\n")
  }
)
```
