---
title: "`r paste0('PFM Model Selection — sweep analysis (', params$reportName, ')')`"
author: "PFM Modeling Output"
date: "`r Sys.Date()`"
params:
  workflowRds: "results/exhaustive/sweep.rds"
  reportName: "default"
  topN: 40
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    code_folding: hide
---

<style>
  .main-container { max-width: 95% !important; }
  .toc-content { padding-left: 0px !important; }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      dev = "svglite", fig.ext = ".svg",
                      fig.width = 11, fig.height = 6, dpi = 100)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)

root <- getwd()
library(pfmreports)
library(pfmreports)
library(pfmreports)

# ── Pure consumer of the Run-Group sweep (ADR 0018) — NO model fitting ────────
# Everything below is derived from the precomputed sweep artifact written by
# pfm::runSweep(): per-fit metrics (`results`), per-term estimates (`coefficients`),
# the maximin rankings, the selected deliverable, and the Stage-0 screens. The report
# never estimates a model (the old in-Rmd fitting + AME recomputation were removed when
# compute moved into pfm; see ADR 0018). The cross-model AME comparison, which needed a
# fit, now lives in the dedicated adoption results report.
rds_path <- params$workflowRds
if (!is_absolute_path(rds_path)) rds_path <- file.path(root, rds_path)
if (!file.exists(rds_path)) {
  stop("Sweep RDS not found: ", rds_path,
       " — run pfm::runSweep()/runModelGroup() for this group first.")
}
wf <- readRDS(rds_path)
results <- wf$results
if (is.null(results$tier)) {
  results$tier <- pfm::computeTheoryTier(results$sigActorPower, results$sigInstQual,
                                         results$sigInteractions)
}
coefs <- wf$coefficients
topN  <- as.integer(params$topN)

tier_bg <- c(Green = "#c8e6c9", Blue = "#bbdefb", Yellow = "#fff9c4")

# k (number of estimated terms) per model × sector × stage, from the coefficients table.
kByFit <- if (!is.null(coefs)) {
  coefs %>% group_by(model, sector, stage) %>% summarise(k = dplyr::n(), .groups = "drop")
} else {
  results %>% transmute(model, sector, stage, k = NA_integer_)
}
results <- results %>% left_join(kByFit, by = c("model", "sector", "stage")) %>%
  mutate(kN = ifelse(is.finite(k / nObs), k / nObs, NA_real_))

sectors <- intersect(c("Diffuse", "Bulk"), unique(results$sector))
stages  <- intersect(c("Adoption", "Stringency"), unique(results$stage))
```

# Overview

```{r overview, results='asis'}
nSpec <- length(unique(results$model))
cat(sprintf(paste0(
  "Pure-consumer analysis of the **%s** sweep: %d specifications × %d sector/stages = ",
  "**%d fits** (generated %s). This report visualises the *full sweep* — every fitted ",
  "specification ranked and compared. The single selected deliverable per stage, the ",
  "Projection-Sanity trace and the Stage-0 channel screen are detailed in the companion ",
  "**Selection** report; here the focus is the multi-criterion ranking and the ",
  "distribution of results across the whole specification space.\n\n"),
  wf$mode %||% "?", nSpec, length(stages), nrow(results),
  format(wf$generated %||% Sys.time(), "%Y-%m-%d %H:%M")))
for (stg in stages) {
  sel <- wf$selected[[stg]]
  cat(sprintf("- **%s — selected deliverable:** `%s`\n", stg, sel %||% "(none)"))
}
```

# Tier distribution across the sweep

```{r tier-dist, results='asis'}
td <- results %>%
  mutate(tier = factor(tier, levels = c("Green", "Blue", "Yellow"))) %>%
  count(stage, sector, tier, .drop = FALSE)
p <- ggplot(td, aes(x = sector, y = n, fill = tier)) +
  geom_col(position = "stack") +
  facet_wrap(~stage) +
  scale_fill_manual(values = tier_bg) +
  theme_report() +
  labs(title = "Theory-tier composition of the sweep", x = NULL, y = "specifications",
       fill = "Tier")
print(p)
howToRead(
  "How many fitted specifications fall in each theory tier, per sector and stage. ",
  "Green = a significant term in all three theory groups (Actor Power, Inst. Quality, ",
  "Interaction); Blue = at least one; Yellow = none. A sweep dominated by Yellow in one ",
  "sector/stage means theory rarely survives there — a structural finding, not a single ",
  "model's failure.")
```

# ΔR²(theory) distribution

```{r dr2-dist, results='asis'}
dd <- results %>% filter(converged %in% TRUE, is.finite(deltaR2Theory))
if (nrow(dd) > 0) {
  p <- ggplot(dd, aes(x = sector, y = deltaR2Theory, fill = sector)) +
    geom_violin(alpha = 0.5, colour = NA) +
    geom_boxplot(width = 0.15, outlier.size = 0.6) +
    facet_wrap(~stage, scales = "free_y") +
    theme_report() + theme(legend.position = "none") +
    labs(title = "Distribution of theory's unique explanatory contribution",
         x = NULL, y = "ΔR²(theory)")
  print(p)
}
howToRead(
  "ΔR²(theory) is the pseudo-R² lost when all theory terms are stripped from a fit — the ",
  "unique signal carried by Actor Power + Institutional Quality + their interaction. Each ",
  "violin is the spread over all converged specifications in that sector/stage; a mass ",
  "near zero means most specifications carry little theory signal there.")
```

# ΔR²(theory) distribution — gate-passing specifications only

```{r dr2-dist-gated, results='asis'}
# Redrawn distribution restricted to specifications that PASS the hard reliability gates
# (per-sector VIF < 10, converged, no lagged dependent-variable terms, non-inflated ΔR² ≤ 1,
# and — stringency only — the time trend not dominating, trendShare ≤ 0.5). This matches the
# hard gates in pfm::computeMaximinScore, so the reject pile (which is ~half of all converged
# specs and contaminates the raw violin's upper tail with collinear/degenerate fits) is removed.
# The deployed model per stage is marked so the reader sees WHERE the pick sits in the valid set.
gp <- results %>%
  filter(converged %in% TRUE, is.finite(deltaR2Theory)) %>%
  mutate(gatePass = (is.na(maxVIF) | maxVIF < 10) &
                    !(usesLagged %in% TRUE) &
                    deltaR2Theory <= 1 &
                    (is.na(trendShare) | trendShare <= 0.5)) %>%
  filter(gatePass)
nAll  <- results %>% filter(converged %in% TRUE, is.finite(deltaR2Theory)) %>% nrow()
nKept <- nrow(gp)
dep <- do.call(rbind, lapply(stages, function(stg) {
  sel <- wf$selected[[stg]]
  if (is.null(sel)) return(NULL)
  results %>% filter(model == sel, stage == stg, is.finite(deltaR2Theory)) %>%
    transmute(stage, sector, deltaR2Theory)
}))
depLayer <- if (!is.null(dep) && nrow(dep) > 0)
  geom_point(data = dep, aes(x = sector, y = deltaR2Theory), inherit.aes = FALSE,
             shape = 23, size = 3.2, fill = "#D55E00", colour = "white", stroke = 0.8) else NULL
if (nKept > 0) {
  p <- ggplot(gp, aes(x = sector, y = deltaR2Theory, fill = sector)) +
    geom_violin(alpha = 0.5, colour = NA) +
    geom_boxplot(width = 0.15, outlier.size = 0.6) +
    depLayer +
    facet_wrap(~stage, scales = "free_y") +
    theme_report() + theme(legend.position = "none") +
    labs(title = "ΔR²(theory) among gate-passing specifications, with the deployed model marked (◆)",
         subtitle = sprintf("Reliability-gate survivors only: %d of %d converged fits retained (%.0f%% dropped for VIF ≥ 10, inflation, lagged terms, or trend dominance)",
                            nKept, nAll, 100 * (1 - nKept / nAll)),
         x = NULL, y = "ΔR²(theory)")
  print(p)
}
howToRead(
  "The same theory-signal distribution as above, but restricted to specifications that clear ",
  "the HARD reliability gates (VIF < 10, converged, no lagged terms, ΔR²(theory) ≤ 1, trend not ",
  "dominant). Roughly half of all converged fits are removed — including most of the raw chart's ",
  "high-ΔR² tail, which is inflated by collinear/degenerate specs the selection rule rejects. The ",
  "orange diamond (◆) marks the deployed model: it sits HIGH within the valid distribution without ",
  "chasing the (rejected) inflated tail — the honest read of 'we did not cherry-pick the max-ΔR² spec.'")
```

# Model Rankings {.tabset}

```{r model-rankings, results='asis'}
rank_asc  <- function(x) rank(x,  ties.method = "min", na.last = TRUE)
rank_desc <- function(x) rank(-x, ties.method = "min", na.last = TRUE)
tierRank  <- c(Green = 1L, Blue = 2L, Yellow = 3L)

for (stg in stages) {
  for (sec in sectors) {
    sec_stg <- paste0(stg, ": ", sec)
    cat(sprintf("\n## %s\n\n", sec_stg))
    sub <- results %>% filter(stage == stg, sector == sec)
    if (nrow(sub) == 0) { cat("> No fits for this sector/stage.\n\n"); next }

    is_flagged <- !(sub$converged %in% TRUE) | (!is.na(sub$maxVIF) & sub$maxVIF >= 10) |
                  (sub$usesLagged %in% TRUE)
    ttier <- tierRank[sub$tier]
    ord <- order(as.integer(is_flagged), ttier, -sub$deltaR2Theory, sub$aic,
                 sub$maxVIF, na.last = TRUE)
    sub <- sub[ord, ]; is_flagged <- is_flagged[ord]; ttier <- ttier[ord]
    sub <- utils::head(sub, topN); is_flagged <- is_flagged[seq_len(nrow(sub))]
    ttier <- ttier[seq_len(nrow(sub))]
    n <- nrow(sub)

    rk_flag <- rank(as.integer(is_flagged), ties.method = "min")
    rk_tier <- rank(ttier, ties.method = "min")
    rk_dr2  <- rank_desc(sub$deltaR2Theory)
    rk_aic  <- rank_asc(sub$aic)
    rk_bic  <- rank_asc(sub$bic)
    rk_r2   <- rank_desc(sub$pseudoR2)
    rk_tf   <- rank_desc(sub$theoryFrac)
    rk_vif  <- rank_asc(sub$maxVIF)
    rk_kn   <- rank_asc(sub$kN)

    vr <- function(vals, ranks, digits = 2) mapply(function(v, r) {
      vstr <- if (is.na(v) || !is.finite(v)) "—" else formatC(round(v, digits), digits = digits, format = "f")
      paste0(r, " (", vstr, ")")
    }, vals, ranks)
    tier_names <- c("1" = "Green", "2" = "Blue", "3" = "Yellow")

    out <- data.frame(
      `#`   = seq_len(n),
      Model = sub$model,
      Transform = sub$panelTransform,
      Flag  = paste0(rk_flag, " (", ifelse(is_flagged, "Yes", "No"), ")"),
      Tier  = paste0(rk_tier, " (", tier_names[as.character(ttier)], ")"),
      `ΔR²`  = vr(sub$deltaR2Theory, rk_dr2, 3),
      AIC   = vr(sub$aic, rk_aic, 1),
      BIC   = vr(sub$bic, rk_bic, 1),
      `R²`   = vr(sub$pseudoR2, rk_r2, 3),
      TFrac = vr(sub$theoryFrac, rk_tf, 3),
      MaxVIF = vr(sub$maxVIF, rk_vif, 1),
      `k/N` = vr(sub$kN, rk_kn, 3),
      check.names = FALSE, stringsAsFactors = FALSE
    )

    ramp_green <- grDevices::colorRampPalette(c("#a5d6a7", "white"))(max(n, 1))
    make_bg <- function(ranks) ifelse(is.na(ranks), "white",
                                      ramp_green[pmin(pmax(as.integer(ranks), 1L), n)])
    k <- kable(out, format = "html", escape = FALSE,
               caption = paste0("Model Rankings — ", sec_stg,
                                " (top ", n, " by tier → ΔR²(theory) → AIC)")) %>%
      kable_styling(bootstrap_options = c("hover", "condensed", "responsive"),
                    font_size = 12, full_width = TRUE) %>%
      column_spec(1, bold = TRUE, width = "2em") %>%
      column_spec(2, width = "24em") %>%
      column_spec(4, background = ifelse(is_flagged, "#ffcdd2", make_bg(rk_flag)),
                  color = ifelse(is_flagged, "#c62828", "black")) %>%
      column_spec(5, background = tier_bg[sub("^\\d+\\s*\\((.+)\\)$", "\\1", out$Tier)]) %>%
      column_spec(6, background = make_bg(rk_dr2)) %>%
      column_spec(7, background = make_bg(rk_aic)) %>%
      column_spec(8, background = make_bg(rk_bic)) %>%
      column_spec(9, background = make_bg(rk_r2)) %>%
      column_spec(10, background = make_bg(rk_tf)) %>%
      column_spec(11, background = ifelse(!is.na(sub$maxVIF) & sub$maxVIF > 10, "#ffcdd2",
                                   ifelse(!is.na(sub$maxVIF) & sub$maxVIF > 5, "#fff9c4",
                                          make_bg(rk_vif))),
                  color = ifelse(!is.na(sub$maxVIF) & sub$maxVIF > 10, "#c62828", "black"),
                  bold = !is.na(sub$maxVIF) & sub$maxVIF > 10) %>%
      column_spec(12, background = make_bg(rk_kn)) %>%
      scroll_box(width = "100%")
    print(k)
    cat("\n<br>\n")
  }
}
```

```{r model-rankings-cap, results='asis'}
howToRead(
  "Each cell shows **rank (value)**: the rank is that model's standing on that single ",
  "criterion (1 = best), the value in parentheses is the raw measure; greener = better ",
  "rank. Rows are pre-sorted by the deliverable logic (flag → theory tier → ΔR²(theory) ",
  "→ AIC). 'Flag' (red) marks a fit that failed a hard gate — non-convergence, maxVIF ≥ 10, ",
  "or a lagged dependent variable — and is sorted last. Capped at the top ",
  paste0("`", topN, "`"), " specifications per sector/stage (raise with the `topN` param).",
  type = "table")
```

# Best models {.tabset}

## By theory fraction

```{r best-theory, results='asis'}
bt <- results %>%
  filter(converged %in% TRUE) %>%
  arrange(as.integer(!is.na(maxVIF) & maxVIF >= 10), tierRank[tier], desc(theoryFrac)) %>%
  transmute(`Sector/Stage` = paste0(stage, ": ", sector), Model = model,
            Transform = panelTransform, Tier = tier, `Theory Frac.` = theoryFrac,
            `ΔR²(theory)` = deltaR2Theory, AIC = aic, `Pseudo-R²` = pseudoR2,
            `Max VIF` = maxVIF, `N obs` = nObs) %>%
  head(30)
kbl <- kable(bt, format = "html", digits = 3,
             caption = "Top 30 fits by Theory Fraction (all sector/stages)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                font_size = 11, full_width = TRUE) %>%
  column_spec(4, background = tier_bg[bt$Tier], bold = bt$Tier == "Green")
print(kbl)
howToRead(
  "Theory Fraction = |theory contribution| / Σ|all non-intercept contributions| of the ",
  "linear predictor — the share of the model's signal carried by theory vs controls, FE, ",
  "and trend. High Theory Fraction with a Green tier is the theoretically richest corner ",
  "of the sweep.", type = "table")
```

## By AIC (prediction)

```{r best-pred, results='asis'}
bp <- results %>%
  filter(converged %in% TRUE, !(!is.na(maxVIF) & maxVIF >= 10)) %>%
  arrange(stage, sector, aic) %>%
  group_by(stage, sector) %>% slice_head(n = 10) %>% ungroup() %>%
  transmute(`Sector/Stage` = paste0(stage, ": ", sector), Model = model,
            Transform = panelTransform, Tier = tier, AIC = aic, BIC = bic,
            `Pseudo-R²` = pseudoR2, `ΔR²(theory)` = deltaR2Theory, `Max VIF` = maxVIF)
kbl <- kable(bp, format = "html", digits = 3,
             caption = "Top 10 fits by AIC within each sector/stage") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                font_size = 11, full_width = TRUE) %>%
  column_spec(4, background = tier_bg[bp$Tier], bold = bp$Tier == "Green")
print(kbl)
howToRead(
  "Lower AIC = better fit-complexity trade-off, comparable only within a sector/stage and ",
  "panelTransform. This is the prediction-leaning ranking; contrast the top AIC models' ",
  "tiers with the theory ranking above to see where predictive and theoretical strength ",
  "diverge.", type = "table")
```

## Maximin shortlist (deliverable rule)

```{r maximin-short, results='asis'}
for (stg in stages) {
  mm <- wf$maximin[[stg]]
  if (is.null(mm)) next
  top <- mm %>% filter(gatePass) %>% head(10)
  cat(sprintf("### %s\n\n", stg))
  print(interactiveTable(top, caption = paste0(stg, " — top 10 gate-passing specs (maximin)")))
  cat("\n\n")
}
howToRead(
  "The deliverable shortlist per stage under the Maximin Selection Rule (worse-sector tier ",
  "first, then mean ΔR²(theory); ADR 0012). These shared specifications apply to both ",
  "sectors — the candidates worth a confirmation pass. The single chosen model and its ",
  "Projection-Sanity verdict are in the Selection report.", type = "table")
```

# Selection uncertainty (ADR 0025)

How stable is the chosen deliverable to the specification search? Two in-report layers (the
region-block bootstrap is a separate opt-in cluster step).

> **Note — these layers reflect the *maximin ranking* (before the Projection-Sanity gate).** The
> deployed deliverable is the maximin band member that *also* passes Projection-Sanity (clamp /
> explosion checks), applied only in the full pipeline. So the band leader / knob winner shown here
> can differ from the deployed spec — e.g. the stringency maximin-top may be a Vertical-Accountability
> spec while the deployed one is the Diagonal-Accountability spec that passed sanity. Read these for
> **channel-set stability**, and see the Selection report for the single deployed model.

## Near-tie confidence set {.tabset}

```{r conf-set, results='asis'}
eps <- 0.025  # deliverable near-tie tolerance (ADR 0022)
chOf <- function(m) vapply(strsplit(as.character(m), " ", fixed = TRUE),
                           function(x) if (length(x) >= 2) x[[2]] else "", character(1))  # IQ-channel token from the spec name
for (stg in stages) {
  mm <- wf$maximin[[stg]]; if (is.null(mm)) next
  p <- mm[mm$gatePass %in% TRUE, , drop = FALSE]; if (!nrow(p)) next
  lead <- p$meanDeltaR2[1]
  band <- p[p$minTier == p$minTier[1] & !is.na(p$meanDeltaR2) & p$meanDeltaR2 >= lead - eps, ,
            drop = FALSE]
  band$channels <- chOf(band$model)
  cat(sprintf("### %s\n\n", stg))
  cat(sprintf("**%d spec(s)** are within ε = %.3f of the leader (statistically tied on theory). ",
              nrow(band), eps))
  uch <- unique(band$channels)
  cat(sprintf("They span **%d distinct channel set(s)**%s.\n\n", length(uch),
              if (length(uch) == 1) " — i.e. the channel choice is unambiguous in the tie" else ""))
  print(interactiveTable(
    band[, intersect(c("model", "channels", "minTier", "meanDeltaR2", "sumBIC", "idleControl",
                       "fragility"), names(band))],
    caption = paste0(stg, " — near-tie confidence set (within ε of the leader)")))
  cat("\n\n")
}
howToRead(
  "The specifications statistically tied with the deliverable (same worse-sector tier, mean ",
  "ΔR²(theory) within ε). A small set sharing one channel set means the pick is robust; a large ",
  "set spanning many channel sets means the exact winner is partly arbitrary — read the *channel* ",
  "agreement, not the single row.", type = "table")
```

## Knob-sensitivity {.tabset}

```{r knob-sens, results='asis'}
realFE <- "fe:(H12|OECDp|Mundlak)"
mmCols <- intersect(c("model", "sector", "sigActorPower", "sigInstQual", "sigInteractions",
                      "deltaR2Theory", "pseudoR2", "bic", "maxVIF", "converged", "usesLagged",
                      "nFE", "nObs", "sigControl", "nControl", "trendShare"), names(results))
base_args <- list(nearTieEps = 0.025, feParsimonyWeight = 0, dropIdleControls = TRUE,
                  softVifGate = 6)
knobs <- list(
  "deliverable (default)"             = list(),
  "classic BIC (feParsimonyWeight=1)" = list(feParsimonyWeight = 1),
  "keep idle controls"                = list(dropIdleControls = FALSE),
  "no VIF preference"                 = list(softVifGate = Inf),
  "wider band (ε=0.05)"               = list(nearTieEps = 0.05),
  "tighter band (ε=0.01)"             = list(nearTieEps = 0.01))
chOf <- function(m) vapply(strsplit(as.character(m), " ", fixed = TRUE),
                           function(x) if (length(x) >= 2) x[[2]] else "", character(1))
for (stg in stages) {
  sub_s <- results[results$stage == stg & grepl(realFE, results$model), , drop = FALSE]
  if (!nrow(sub_s)) next
  rows <- lapply(names(knobs), function(nm) {
    args <- utils::modifyList(c(list(df = sub_s[, mmCols, drop = FALSE]), base_args), knobs[[nm]])
    mm <- tryCatch(do.call(pfm::computeMaximinScore, args), error = function(e) NULL)
    if (is.null(mm)) return(NULL)
    top <- mm[mm$gatePass %in% TRUE, , drop = FALSE]
    if (!nrow(top)) return(data.frame(knob = nm, winner = "(none gate-passing)", channels = "",
                                      tier = "", meanDeltaR2 = NA_real_))
    data.frame(knob = nm, winner = top$model[1], channels = chOf(top$model[1]),
               tier = top$minTier[1], meanDeltaR2 = round(top$meanDeltaR2[1], 3))
  })
  tab <- do.call(rbind, rows)
  nWin <- length(unique(tab$winner)); nCh <- length(unique(tab$channels[nzchar(tab$channels)]))
  cat(sprintf("### %s\n\n", stg))
  cat(sprintf(paste0("Across %d knob settings: **%d distinct winning spec(s)**, **%d distinct ",
              "channel set(s)**.%s\n\n"), nrow(tab), nWin, nCh,
              if (nCh == 1) " The channel set is invariant — the result is a knob-robust channel claim."
              else " The exact winner moves with the knobs; report channel agreement, not the row."))
  print(interactiveTable(tab, caption = paste0(stg, " — winner under each selection-knob setting")))
  cat("\n\n")
}
howToRead(
  "Re-runs the Maximin Selection Rule on the sweep under each knob flipped one at a time. The ",
  "scientific claim is the **channel set**, which is far more stable than the exact winning row; ",
  "a single distinct channel set across knobs is the robustness headline. Sensitivity to ε and ",
  "to the FE-discount is expected — see the Parsimony Frontier.", type = "table")
```

## Region-block bootstrap

The gold-standard selection-uncertainty check — a region-block bootstrap of the entire selection
(ADR 0025) — now lives in its **own report, "Selection Stability"** (`selection_stability_*.html`),
where it is presented in full (channel-set + per-channel frequency charts, exact-spec frequency, the
90% selection confidence set, a deliverable check, and a paper-ready summary). It is a separate,
heavy cluster step, so isolating it keeps this report fast to render.

# All fits

```{r all-fits, results='asis'}
interactiveTable(
  results %>%
    select(model, sector, stage, panelTransform, tier, deltaR2Theory, theoryFrac,
           aic, bic, pseudoR2, nObs, maxVIF, converged, sigActorPower, sigInstQual,
           sigInteractions),
  caption = "Every fit in the sweep — selection measures", pageLength = 25, filter = "top")
howToRead(
  "One row per model × sector × stage. Use the column filters to slice the sweep — e.g. ",
  "tier = Green + stage = Stringency for the Green stringency shortlist, or sort by ",
  "deltaR2Theory for the theory-content ranking. Predictive diagnostics (AUC/Brier/RMSE) ",
  "and the full measure glossary are in the Selection report.", type = "table")
```
