---
title: "PSM Results — Policy Stringency Model — `r params$reportName`"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 2
    toc_float:
      collapsed: true
    number_sections: true
    code_folding: hide
params:
  sweepRds: "results/psm-exhaustive/sweep.rds"
  selectedConfig: "results/psm-exhaustive/selected-models-psm.yml"
  agreementRds: "results/psm-exhaustive/estimator-agreement.rds"
  projectionsDir: "results/psm-exhaustive/projections"
  manifestJson: "results/psm-exhaustive/manifest.json"
  temporalRds: "results/psm-exhaustive/temporal-validation.rds"
  frontierRds: "results/psm-exhaustive/frontier.rds"
  influenceRds: "results/psm-exhaustive/influence.rds"
  reportName: "default"
  indexMax: 10
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 5)
library(ggplot2)
library(dplyr)
library(pfmreports)

root <- getwd()
absify <- function(p) {
  if (is.null(p) || !nzchar(p)) return(p)
  if (grepl("^([A-Za-z]:|/|\\\\)", p)) p else file.path(root, p)
}
readIf <- function(p, reader) {
  p <- absify(p)
  if (!is.null(p) && nzchar(p) && file.exists(p)) tryCatch(reader(p), error = function(e) NULL) else NULL
}
sweep <- readIf(params$sweepRds, readRDS)
sel <- readIf(params$selectedConfig, yaml::read_yaml)
agree <- readIf(params$agreementRds, readRDS)
manifest <- readIf(params$manifestJson, function(p) jsonlite::fromJSON(p, simplifyVector = FALSE))
temporal <- readIf(params$temporalRds, readRDS)
frontierArt <- readIf(params$frontierRds, readRDS)
influence <- readIf(params$influenceRds, readRDS)
projDir <- absify(params$projectionsDir)
projFiles <- if (!is.null(projDir) && nzchar(projDir) && dir.exists(projDir)) {
  list.files(projDir, pattern = "\\.rds$", full.names = TRUE)
} else character(0)
projections <- if (length(projFiles)) {
  dplyr::bind_rows(lapply(projFiles, function(f) tryCatch(readRDS(f), error = function(e) NULL)))
} else NULL
indexMax <- as.numeric(params$indexMax)
deployedSpec <- if (!is.null(sweep$selected$PolicyStringency)) sweep$selected$PolicyStringency else
  if (!is.null(sel) && length(sel)) sel[[1]]$name else NA_character_
sectors <- c("Bulk", "Diffuse")
fmt <- function(x, d = 3) ifelse(is.na(x), "", formatC(x, digits = d, format = "f"))
```

# Overview

*What this report is for.* The **Policy Stringency Model (PSM, ADR 0036)** results for one
Run-Group: a single-stage bounded-response model of the CAPMF policy-stringency index
(0–`r indexMax`) on Actor Power, Institutional Quality and their interactions — the PFM's
policy-*portfolio* implementability signal, sibling to the carbon-price hurdle model. This report
is a **pure consumer** of the Run-Group artifacts (`sweep.rds`, `selected-models-psm.yml`,
`estimator-agreement.rds`, `projections/`): it estimates nothing itself. Two properties frame
everything below: the outcome is **bounded by construction** (`index = indexMax·logit⁻¹(η)`), so
**no extrapolation clamp exists anywhere** in this model; and estimators are **never compared on
AIC/BIC** — the satP engine drives selection, the fractional logit headlines the coefficients,
and the agreement between them is the robustness claim.

```{r provenance, results='asis'}
cat("**Artifact checklist:** ",
    sprintf("sweep.rds %s | selected-models-psm.yml %s | estimator-agreement.rds %s | projections/ %s\n\n",
            if (!is.null(sweep)) "✓" else "✗ missing",
            if (!is.null(sel)) "✓" else "✗ missing",
            if (!is.null(agree)) "✓" else "✗ missing",
            if (!is.null(projections)) sprintf("✓ (%d scenario file(s))", length(projFiles)) else "✗ missing"))
if (!is.null(manifest)) {
  cat(sprintf("- **Run-Group:** `%s` (mode `%s`), pfm %s / mrpfm %s\n",
              manifest$group %||% "?", manifest$mode %||% "?",
              manifest$pfm_version %||% "?", manifest$mrpfm_version %||% "?"))
  if (!is.null(manifest$training_years)) {
    cat(sprintf("- **Training years:** %s–%s | panel hash `%s`\n",
                manifest$training_years[[1]], manifest$training_years[[2]],
                manifest$panel_hash %||% "?"))
  }
  if (!is.null(manifest$steps)) {
    st <- vapply(names(manifest$steps), function(s) {
      e <- manifest$steps[[s]]
      if (is.list(e)) paste0(s, ": ", e$status %||% "?") else paste0(s, ": done")
    }, character(1))
    cat("- **Steps:** ", paste(st, collapse = " | "), "\n")
  }
}
cat(sprintf("- **Deployed spec:** `%s`\n", deployedSpec))
if (is.null(sweep)) {
  cat("\n> **sweep.rds not found — run `pfm::runPSMSweep()` first.** Sections below that depend on it are skipped.\n")
}
```

# Selection (sweep + maximin + bounded-index sanity)

*What this section is for.* How the deployed specification was chosen: every swept spec's fit
metrics, the worse-sector Maximin ranking (identical rule to the carbon-price model — the satP
engine's full likelihood keeps AIC/BIC/ΔR² valid), and the **bounded-index Projection Sanity
gate**. Because the index saturates structurally, the price model's explosion/clamp-reliance
rules have no analogue here; what is screened is *implausibility on the index scale* (dead
region blocks, saturation timing, seams, spikes).

```{r sweep-results, results='asis'}
if (!is.null(sweep$results) && nrow(sweep$results)) {
  res <- sweep$results
  keep <- intersect(c("model", "sector", "tier", "deltaR2Theory", "pseudoR2", "aic", "bic",
                      "maxVIF", "trendShare", "nObs", "converged"), colnames(res))
  tab <- res[, keep, drop = FALSE]
  for (cc in intersect(c("deltaR2Theory", "pseudoR2", "trendShare"), colnames(tab))) tab[[cc]] <- round(tab[[cc]], 3)
  for (cc in intersect(c("aic", "bic", "maxVIF"), colnames(tab))) tab[[cc]] <- round(tab[[cc]], 1)
  tab$deployed <- ifelse(tab$model == deployedSpec, "◄ deployed", "")
  interactiveTable(tab, caption = "Per-fit metrics — every swept PSM spec × sector", pageLength = 12)
  howToRead("One row per fitted (spec, sector). `tier` is the Theory Tier (Green > Blue > Yellow), ",
            "`deltaR2Theory` the incremental pseudo-R² of the theory terms over FE + controls + trend ",
            "(the primary identification metric), `maxVIF` the collinearity gate input. The deployed ",
            "spec is flagged. A spec must convergence- and VIF-pass in BOTH sectors to compete.",
            type = "table")
} else cat("> *No sweep results available.*\n")
```

```{r maximin, results='asis'}
mm <- sweep$maximin$PolicyStringency
if (!is.null(mm) && nrow(mm)) {
  mt <- mm
  numc <- vapply(mt, is.numeric, logical(1))
  mt[numc] <- lapply(mt[numc], function(v) round(v, 3))
  mt$deployed <- ifelse(mt$model == deployedSpec, "◄ deployed", "")
  interactiveTable(utils::head(mt, 40), caption = "Maximin ranking (worse-sector tier → mean ΔR²(theory) → tie-breaks)",
                   pageLength = 12)
  howToRead("The cross-sector ranking that picks the shared specification: rank by the WORSE sector's ",
            "Theory Tier, tie-break by mean ΔR²(theory), then the BIC-parsimony band rules. `gatePass` ",
            "combines the hard gates (VIF, convergence, no unintended lag). The deployed spec is the ",
            "best gate-passing row that also cleared the bounded-index sanity gate below.", type = "table")
} else cat("> *No maximin table available.*\n")
```

```{r sanity, results='asis'}
sn <- sweep$sanity$PolicyStringency
if (!is.null(sn) && !is.null(sn$trace) && nrow(sn$trace)) {
  cat(sprintf("**Sanity walk:** %d candidate(s) evaluated; chosen `%s`%s.\n\n",
              nrow(sn$trace), sn$chosen,
              if (isTRUE(sn$forced)) " — **LEAST-FLAGGED fallback (no candidate passed all severe rules)**" else ""))
  print(knitr::kable(sn$trace, row.names = FALSE,
                     caption = "Bounded-index Projection-Sanity trace (expanding batches, maximin order)"))
  cat("\n\n")
  howToRead("Candidates are projected onto the gating scenario in maximin order and screened by the ",
            "bounded-index rules — severe: non-finite index, dead region block (a whole block below ",
            sprintf("%.1f", 0.05 * indexMax), " through the horizon); warnings: all-region saturation above ",
            sprintf("%.1f", 0.95 * indexMax), ", seam jumps, single-year spikes, NA coverage. The first ",
            "severe-free candidate wins. There is NO price-explosion or clamp-reliance rule: nothing in ",
            "this model is unbounded.", type = "table")
  fl <- sn$flags[[sn$chosen]]
  if (!is.null(fl) && nrow(fl)) {
    print(knitr::kable(fl, row.names = FALSE, caption = paste0("Flags for the deployed spec `", sn$chosen, "`")))
    cat("\n\n")
    howToRead("Remaining (non-severe) flags on the deployed spec's gating-scenario projection — ",
              "disclosed, not disqualifying.", type = "table")
  } else cat("*The deployed spec carries no sanity flags on the gating scenario.*\n\n")
} else cat("> *No sanity trace (sweep ran without a scenario panel — selection was pure maximin).*\n")
```

## Selection variants and the sharing cost (Tournament v2, ADR 0039)

*What this subsection is for.* Under Tournament v2 the ranking is the **worse sector's
ΔR²(theory)** and tier is a **gate** (deployment requires Green in both sectors). Three
things are documented rather than hidden: the **Blue-gate (tier-relaxed) winner**, each
sector's **own best** spec, and the **Specification-Sharing Cost** — the ΔR² each sector
gives up because one shared spec serves both. Per-sector winners are never deployed
(the Shared Specification rule stands); this exhibit quantifies what the rule costs.

```{r selection-variants, results='asis'}
sv <- sweep$selectionVariants
if (is.null(sv)) sv <- readIf(file.path(dirname(absify(params$sweepRds)), "selection-variants.rds"), readRDS)
if (!is.null(sv)) {
  w <- data.frame(tierGate = names(sv$winners),
                  winner = unlist(lapply(sv$winners, function(x) if (is.na(x)) "(gate passes nothing)" else x)),
                  deployed = names(sv$winners) == sv$deployedGate,
                  row.names = NULL)
  print(knitr::kable(w, row.names = FALSE, caption = "Tier-gate winners (worse-sector ΔR² ranking)"))
  cat("\n\n")
  ps <- sv$perSector
  numc <- vapply(ps, is.numeric, logical(1)); ps[numc] <- lapply(ps[numc], function(v) round(v, 3))
  print(knitr::kable(ps, row.names = FALSE, caption = "Specification-Sharing Cost exhibit"))
  cat("\n\n")
  howToRead("`sharingCost` = the sector's own best worse-case ΔR²(theory) minus what it gets ",
            "under the shared deployed spec; `sharingCostShare` expresses it as a fraction of ",
            "the achievable. A large share is grounds to re-open the dual-deployment question ",
            "in its own decision session — never a silent switch.", type = "table")
} else cat("> *No selection-variants exhibit (pre-v2 sweep, or documentTierGates = NULL).*\n")
```

# Deployed model — coefficients

*What this section is for.* The deployed specification's estimated coefficients per sector, from
the **satP selection engine** (gaussian GLM on `logit(y/indexMax)`, cluster-robust SEs by
region). Coefficients are in per-standard-deviation units of each driver on the logit scale;
the natural-scale magnitudes are in the AME comparison of the next section.

```{r coefficients, results='asis'}
cf <- sweep$coefficients
if (!is.null(cf) && nrow(cf) && !is.na(deployedSpec)) {
  d <- cf[cf$model == deployedSpec & cf$stage == "PolicyStringency", , drop = FALSE]
  d <- d[!grepl("^regionFE", d$term), , drop = FALSE]
  if (nrow(d)) {
    d$sig <- cut(d$pValue, c(-Inf, 0.001, 0.01, 0.05, 0.1, Inf),
                 labels = c("***", "**", "*", ".", ""))
    tab <- d %>%
      mutate(estimate = round(estimate, 3), stdError = round(stdError, 3),
             zValue = round(zValue, 2), pValue = signif(pValue, 3)) %>%
      select(sector, term, estimate, stdError, zValue, pValue, sig)
    interactiveTable(tab, caption = paste0("Deployed spec `", deployedSpec, "` — cluster-robust coefficients (satP engine)"),
                     pageLength = 15)
    howToRead("Coefficients on the logit(index/indexMax) scale, per sector, with clustered (region) ",
              "standard errors. Signs are directly interpretable (positive = more policy stringency); ",
              "magnitudes are per-SD of the standardized driver. Interaction terms carry the `_x_` infix.",
              type = "table")
    thy <- d[!grepl("Intercept|timeTrend|logisticTimeTrend|lagged_ecp", d$term), , drop = FALSE]
    if (nrow(thy)) {
      p <- ggplot(thy, aes(x = estimate, y = term, colour = sector)) +
        geom_vline(xintercept = 0, linetype = 2, colour = "grey50") +
        geom_pointrange(aes(xmin = estimate - 1.96 * stdError, xmax = estimate + 1.96 * stdError),
                        position = position_dodge(width = 0.5)) +
        theme_report() +
        labs(title = paste0("Deployed spec `", deployedSpec, "` — theory coefficients (95% CI)"),
             x = "coefficient (logit scale, per SD)", y = NULL)
      print(p); cat("\n\n")
      howToRead("Point estimate ± 1.96 clustered SE per sector. Intervals clear of the zero line = ",
                "significant at 5%; sector agreement in sign is the Shared-Specification expectation.")
    }
  } else cat("> *No coefficient rows for the deployed spec.*\n")
} else cat("> *No coefficient table available.*\n")
```

# Estimator agreement (the estimator-invariance claim)

*What this section is for.* The deployed spec re-estimated under the whole **PSM Estimator
Suite** — satP engine, **fractional logit (the paper's headline)**, beta regression, and the
levels-OLS benchmark. The published claim is that the political-economy channels are
**estimator-invariant**: same signs, comparable natural-scale AMEs. Estimators are compared only
on scale-free natural-scale metrics; the AIC column is transparency, never a ranking.

```{r agreement, results='asis', fig.height=5.5}
if (!is.null(agree) && length(agree$bySector)) {
  for (sec in intersect(sectors, names(agree$bySector))) {
    a <- agree$bySector[[sec]]
    if (is.null(a)) next
    cat(sprintf("## %s\n\n", sec))
    fs <- a$fitStats
    if (!is.null(fs)) {
      numc <- vapply(fs, is.numeric, logical(1))
      fs[numc] <- lapply(fs[numc], function(v) round(v, 3))
      print(knitr::kable(fs, row.names = FALSE, caption = paste0(sec, " — per-estimator fit statistics")))
      cat("\n\n")
      howToRead("`rmseNatural`/`corNatural` are computed on the natural 0–", indexMax, " scale and are the ",
                "ONLY cross-estimator comparable fit numbers. `aic` is NA for the fractional logit by ",
                "construction (quasi-likelihood) — that is why it can never drive selection. ",
                "`shareAtZero`/`shareAtMax` are the boundary masses: the empirical gate for the ",
                "conditional Tobit/two-part rung.", type = "table")
    }
    if (!is.null(a$agreement) && nrow(a$agreement)) {
      print(knitr::kable(a$agreement, row.names = FALSE,
                         caption = paste0(sec, " — per-term sign agreement across the fitted suite")))
      cat("\n\n")
      howToRead("`signsAgree` = every fitted estimator gives this term the same sign; `nSignificant05` ",
                "counts in how many estimators it clears 5%. Theory terms with full agreement are the ",
                "estimator-invariance result; a disagreeing term is fragile and must be reported as such.",
                type = "table")
    }
    tb <- a$table
    if (!is.null(tb) && nrow(tb)) {
      thy <- tb[!grepl("Intercept|timeTrend|logisticTimeTrend|lagged_ecp", tb$term), , drop = FALSE]
      thy <- thy[is.finite(thy$ameIndex), , drop = FALSE]
      if (nrow(thy)) {
        p <- ggplot(thy, aes(x = ameIndex, y = term, colour = estimator)) +
          geom_vline(xintercept = 0, linetype = 2, colour = "grey50") +
          geom_pointrange(aes(xmin = ameIndex - 1.96 * ameSE, xmax = ameIndex + 1.96 * ameSE),
                          position = position_dodge(width = 0.6)) +
          theme_report() +
          labs(title = paste0(sec, " — natural-scale AMEs across the estimator suite"),
               x = paste0("average marginal effect (index points, 0–", indexMax, ")"), y = NULL)
        print(p); cat("\n\n")
        howToRead("Average marginal effects on the natural index scale (delta-method 95% intervals, ",
                  "clustered vcov) — the cross-estimator comparable effect size. Overlapping intervals ",
                  "with the same sign across estimators = the robustness claim in one picture. The ",
                  "fractional-logit AMEs are the numbers the paper leads with.")
      }
    }
  }
  if (length(agree$skipped)) {
    cat(sprintf("> Skipped estimator(s): %s.\n\n",
                paste(sprintf("`%s` (%s)", names(agree$skipped), agree$skipped), collapse = "; ")))
  }
} else cat("> *estimator-agreement.rds not found — run `pfm::runPSMEstimatorAgreement()`.*\n")
```

## Marginal effects over observed support

*What this section is for.* The interaction claims live or die here: the marginal effect of
each Actor-Power term on the 0–`r params$indexMax` index, traced over the **observed**
(standardized) range of its institutional-quality moderator. The solid segment is where the
moderation is identified; the faded extension shows how the fitted line continues just outside
support — which is exactly what a scenario projection does when drivers leave their history
(the driver-support guard now winsorizes there).

```{r me-support, results='asis', fig.height=4.5}
mes <- if (!is.null(agree) && length(agree$bySector)) {
  dplyr::bind_rows(lapply(names(agree$bySector), function(sec) {
    m <- agree$bySector[[sec]]$meSupport
    if (is.null(m)) return(NULL)
    m$sector <- sec
    m
  }))
} else NULL
if (!is.null(mes) && nrow(mes)) {
  mes$pair <- paste(gsub("\\.", " ", mes$apTerm), "×", gsub("\\.", " ", mes$moderator))
  p <- ggplot(mes, aes(x = m, y = me)) +
    geom_hline(yintercept = 0, linetype = "dashed", colour = "grey40") +
    geom_ribbon(aes(ymin = lo, ymax = hi, alpha = inSupport), fill = "steelblue") +
    geom_line(aes(linetype = inSupport), colour = "steelblue4") +
    scale_alpha_manual(values = c(`TRUE` = 0.35, `FALSE` = 0.12), guide = "none") +
    scale_linetype_manual(values = c(`TRUE` = "solid", `FALSE` = "dotted"), guide = "none") +
    facet_grid(sector ~ pair, scales = "free_x") +
    theme_minimal() +
    labs(x = "moderator value (standardized; solid = observed support)",
         y = "marginal effect of AP (index points)")
  print(p); cat("\n\n")
  howToRead("Marginal effect of the actor-power term at each moderator value (satP engine, ",
            "delta-method 95% band, clustered vcov). Where the band excludes zero INSIDE the ",
            "solid segment, the moderation is an identified claim; dotted segments are ",
            "extrapolation and carry no evidence.")
} else cat("> *No marginal-effect-support exhibit (older estimator-agreement.rds — re-run ",
           "`pfm::runPSMEstimatorAgreement()`).*\n")
```

# Influence diagnostics (pivotal countries; ADR 0037)

*What this section is for.* Explains **which clusters (regions/countries) make a theory term
marginally (non-)significant** instead of pressuring selection with p-values. Every cluster is
dropped in turn and the deployed spec refit (analytic CR1 clustered SEs, the suite's convention);
a **pivotal** cluster is one whose removal moves a term across the 5% line in either direction
(`gain` = significant only without it; `loss` = significance depends on it). The wild-cluster
bootstrap is re-run at pivotal folds only. **Report-only** — this never changes the deliverable;
a term that owes its star to one country must be claimed accordingly.

```{r influence, results='asis', fig.height=4.5}
if (!is.null(influence) && length(influence$bySector)) {
  for (sec in intersect(sectors, names(influence$bySector))) {
    inf <- influence$bySector[[sec]]
    if (is.null(inf$byTerm) || !nrow(inf$byTerm)) next
    cat(sprintf("## %s\n\n", sec))
    bt <- inf$byTerm
    numc <- vapply(bt, is.numeric, logical(1))
    bt[numc] <- lapply(bt[numc], function(v) round(v, 3))
    print(knitr::kable(bt, row.names = FALSE,
                       caption = paste0(sec, " — per-term leave-one-cluster-out summary")))
    cat("\n\n")
    howToRead("`p` is the full-sample cluster-robust p; `pMin`/`pMax` bound the leave-one-",
              "cluster-out p path. `pivotalClusters` lists the folds crossing 5%; ",
              "`topInfluencer` moves the coefficient most; `tChangeDriver` says whether that ",
              "fold changes the t-statistic through the coefficient or through the SE ",
              "(coefficient shrinkage vs SE inflation).", type = "table")
    pp <- inf$path[inf$path$crossing != "fold-failed" & is.finite(inf$path$p), , drop = FALSE]
    focus <- inf$byTerm$term[inf$byTerm$nPivotal > 0 |
                               (inf$byTerm$p > 0.01 & inf$byTerm$p < 0.10)]
    pp <- pp[pp$term %in% focus, , drop = FALSE]
    if (nrow(pp)) {
      p <- ggplot(pp, aes(x = p, y = cluster, colour = crossing)) +
        geom_vline(xintercept = inf$alpha, linetype = 2, colour = "grey40") +
        geom_point(size = 1.8) +
        scale_colour_manual(values = c(gain = "seagreen", loss = "firebrick",
                                       none = "grey60")) +
        facet_wrap(~term, scales = "free_x") +
        theme_report() +
        labs(title = paste0(sec, " — leave-one-cluster-out p paths (marginal / pivotal terms)"),
             x = "cluster-robust p when this cluster is dropped", y = NULL)
      print(p); cat("\n\n")
      howToRead("Each dot is the term's p-value with that cluster removed; the dashed line is ",
                "alpha = 5%. Green (`gain`) folds show the cluster suppressing an otherwise-",
                "significant term; red (`loss`) folds show the significance depends on that ",
                "one cluster — either way the claim must not rest on a single country.")
    }
  }
} else cat("> *influence.rds not found — run `pfm::runPSMInfluence()`.*\n")
```

# Scenario projections (fan-out)

*What this section is for.* The deployed spec projected onto every registered **Policy
Scenario** (ADR 0035) — the exogenous scenario calculation that is the paper's v1 results
scope. Values are **bounded by construction**; the display needs no cap and there is no clamp
whose reliance would have to be disclosed. Rows for regions outside the CAPMF estimation sample
(USA, Brazil, …) are driver-based extrapolations and flagged `outOfCoverage`. The
**Implementability Factor** columns (`index/indexMax` ∈ [0,1]) are the ready-to-apply IAM
multiplier; the raw index remains the canonical output.

```{r projections, results='asis', fig.height=5.5}
if (!is.null(projections) && nrow(projections)) {
  pr <- projections
  pr$coverage <- ifelse(pr$outOfCoverage, "out-of-coverage (extrapolated)", "in-coverage")
  summ <- pr %>%
    filter(is.finite(index)) %>%
    group_by(scenarioName, sector, year) %>%
    summarise(mid = median(index), lo = quantile(index, 0.25), hi = quantile(index, 0.75),
              .groups = "drop")
  p <- ggplot(summ, aes(x = year, y = mid, colour = scenarioName, fill = scenarioName)) +
    geom_ribbon(aes(ymin = lo, ymax = hi), alpha = 0.15, colour = NA) +
    geom_line(linewidth = 0.9) +
    facet_wrap(~sector) +
    coord_cartesian(ylim = c(0, indexMax)) +
    theme_report() +
    labs(title = "Projected policy-stringency index by scenario (median, IQR across regions)",
         x = NULL, y = paste0("CAPMF index (0–", indexMax, ")"), colour = NULL, fill = NULL)
  print(p); cat("\n\n")
  howToRead("Median projected index across regions per scenario and sector, ribbon = interquartile ",
            "range. The y-axis is the FULL structural range — nothing can leave it, so scenario ",
            "separation here is driver-driven (actor power / institutions), not a runaway functional ",
            "form. A widening gap between scenarios is the political-economy signal delivered to the IAM.")

  horizonYear <- max(pr$year[is.finite(pr$index)], na.rm = TRUE)
  horizon <- pr %>%
    filter(year == horizonYear, is.finite(index)) %>%
    group_by(Scenario = scenarioName, Sector = sector) %>%
    summarise(`median index` = round(median(index), 2),
              `median implementability` = round(median(implementability), 3),
              `p10–p90 implementability` = sprintf("%.2f–%.2f", quantile(implementability, 0.1),
                                                   quantile(implementability, 0.9)),
              `out-of-coverage regions` = length(unique(region[outOfCoverage])),
              .groups = "drop")
  print(knitr::kable(horizon, row.names = FALSE,
                     caption = paste0("Horizon summary (", horizonYear, ") — index and Implementability Factor")))
  cat("\n\n")
  howToRead("End-of-horizon medians per scenario × sector. The Implementability Factor is the 0–1 ",
            "rescale REMIND can apply to non-price policy assumptions (raw index stays canonical). ",
            "Out-of-coverage counts state how many regions are extrapolations rather than in-sample ",
            "predictions.", type = "table")

  ooc <- pr %>% filter(outOfCoverage) %>% distinct(region)
  if (nrow(ooc)) {
    cat(sprintf("**Out-of-coverage regions (driver-based extrapolation):** %s.\n\n",
                paste(sort(ooc$region), collapse = ", ")))
  }
} else cat("> *No projection artifacts — run `pfm::runPSMProjection()` with the Scenario Registry.*\n")
```

# Feasibility frontier and political slack

*What this section is for.* The frontier reading of feasibility: instead of the conditional
*mean* ("how much policy do comparable polities typically have"), a stochastic frontier
estimates the *maximum attainable* stringency given political-economy fundamentals, and
attributes the shortfall to one-sided **political slack** — the ambition gap. Two statistics
qualify everything else: `gamma` (share of error variance from slack; near 0 = the data show
no frontier structure) and the LR test against the no-frontier null. The ratio
observed/frontier is the principled Implementability Factor.

```{r frontier, results='asis', fig.height=4.5}
if (!is.null(frontierArt) && length(frontierArt$bySector)) {
  cat(sprintf("Deployed spec: `%s`.\n\n", if (is.null(frontierArt$spec)) "?" else frontierArt$spec))
  gt <- dplyr::bind_rows(lapply(names(frontierArt$bySector), function(sec) {
    e <- frontierArt$bySector[[sec]]
    data.frame(sector = sec, gamma = e$gamma, LRvsNoFrontier = e$lr, n = e$n,
               medianSlack = stats::median(e$scores$slackIndex, na.rm = TRUE))
  }))
  print(knitr::kable(gt, digits = 3, caption = "Frontier structure per sector"))
  cat("\n\n")
  sc <- dplyr::bind_rows(lapply(names(frontierArt$bySector), function(sec) {
    s <- frontierArt$bySector[[sec]]$scores
    if (is.null(s)) return(NULL)
    s$sector <- sec
    s
  }))
  if (!is.null(sc) && nrow(sc)) {
    scLast <- sc |> dplyr::group_by(sector, region) |> dplyr::slice_max(year, n = 1) |> dplyr::ungroup()
    p <- ggplot(scLast, aes(x = observedIndex, y = frontierIndex)) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
      geom_point(alpha = 0.7, colour = "steelblue4") +
      facet_wrap(~sector) +
      coord_equal(xlim = c(0, indexMax), ylim = c(0, indexMax)) +
      theme_minimal() +
      labs(x = "observed index (last estimation year)", y = "feasibility frontier")
    print(p); cat("\n\n")
    howToRead("Each point is a region at its last estimation year; vertical distance to the ",
              "dashed diagonal is the political ambition gap (frontier minus observed). If ",
              "gamma is near zero, the frontier collapses onto the regression mean and the ",
              "gaps carry no information — check the table before reading the plot.")
  }
} else cat("> *frontier.rds not found — run `pfm::runPSMFrontier()` (step `psm-frontier`).*\n")
```

# Pseudo-out-of-sample validation

*What this section is for.* The decisive test of the projection claim: the deployed spec is
re-fit on the panel truncated at `trainEnd`, projected onto the **observed** drivers of the
held-out years under full deployment discipline (frozen scaling, frozen trend, driver-support
guard), and scored against the **observed** CAPMF index. `skillVsPersistence` compares against
carrying each region's last observed value forward — the naive baseline any projection must
beat (positive = beats persistence).

```{r temporal-validation, results='asis', fig.height=4.5}
if (!is.null(temporal) && length(temporal$bySector)) {
  cat(sprintf("Deployed spec: `%s` — trained ≤ %s, scored on later observed years.\n\n",
              if (is.null(temporal$spec)) "?" else temporal$spec, temporal$trainEnd))
  mt <- dplyr::bind_rows(lapply(names(temporal$bySector), function(sec) {
    data.frame(sector = sec, as.data.frame(temporal$bySector[[sec]]$metrics))
  }))
  print(knitr::kable(mt, digits = 3, caption = "Held-out fit metrics per sector"))
  cat("\n\n")
  rows <- dplyr::bind_rows(lapply(temporal$bySector, `[[`, "rows"))
  if (nrow(rows)) {
    p <- ggplot(rows, aes(x = observed, y = predicted, colour = factor(year))) +
      geom_abline(slope = 1, intercept = 0, linetype = "dashed", colour = "grey40") +
      geom_point(alpha = 0.7) +
      facet_wrap(~sector) +
      coord_equal(xlim = c(0, indexMax), ylim = c(0, indexMax)) +
      theme_minimal() +
      labs(x = "observed CAPMF index", y = "predicted (held-out)", colour = "year")
    print(p); cat("\n\n")
    howToRead("Each point is a held-out region-year. Points on the dashed diagonal are perfect ",
              "predictions; systematic displacement above/below reveals level bias, and a ",
              "vertical spread wider than the historical index movement means the drivers carry ",
              "little forecasting signal. Read together with skillVsPersistence in the table.")
  }
} else cat("> *temporal-validation.rds not found — run `pfm::runPSMTemporalValidation()` ",
           "(step `psm-temporal`).*\n")
```

# Definitions

```{r defs}
defs <- data.frame(
  Term = c("Policy Stringency Model (PSM)", "satP engine", "Fractional logit (headline)",
           "Estimator agreement", "ΔR²(theory)", "Bounded-index Projection Sanity",
           "Implementability Factor", "Out-of-coverage"),
  Definition = c(
    "Single-stage bounded-response model of the CAPMF index (0–10) on AP + IQ + AP×IQ; no adoption hurdle (ADR 0036).",
    "Gaussian GLM on logit(index/indexMax) — full likelihood, so it drives selection (maximin, BIC tie-breaks).",
    "Papke–Wooldridge quasi-binomial logit — canonical for bounded outcomes, exact boundary handling; the paper's coefficient/AME tables. No likelihood ⇒ never used for selection.",
    "The deployed spec re-estimated under the whole suite; signs/AMEs compared, never information criteria.",
    "Pseudo-R² of the full model minus a baseline with all theory terms removed — the primary identification metric.",
    "Severe: non-finite index, dead region block. Warnings: saturation timing, seam jumps, spikes, NA coverage. No explosion/clamp rules — nothing is unbounded.",
    "index/indexMax ∈ [0,1]: the post-processing multiplier for IAM coupling; the raw index remains canonical.",
    "Region absent from the CAPMF estimation sample; its projection is a driver-based extrapolation, flagged per row."),
  stringsAsFactors = FALSE
)
interactiveTable(defs, caption = "PSM report glossary", pageLength = 10)
```
