---
title: "PFM Results — Adoption Model (`r params$reportName`)"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 1
    toc_float:
      collapsed: true
    number_sections: true
    code_folding: hide
params:
  modelConfig: "results/exhaustive/selected-models.yml"
  reportName: "default"
  gdxPath: "data/fulldata.gdx"
  cachefolder: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 5.5)
library(ggplot2)
library(dplyr)
library(magclass)
library(pfm)

root <- getwd()
library(pfmreports)

if (nzchar(params$cachefolder)) madrat::setConfig(cachefolder = params$cachefolder)
madrat::setConfig(forcecache = TRUE)
model_dir <- getPfmConfig("modelDir", "../output")
if (!is_absolute_path(model_dir)) model_dir <- file.path(root, model_dir)
options(pfm.modelDir = model_dir)

# ── Config: one entry per sector (model_type "Adoption: <Sector>") ─────────────
cfg_path <- params$modelConfig
if (!is_absolute_path(cfg_path)) cfg_path <- file.path(root, cfg_path)
all_cfg <- yaml::read_yaml(cfg_path)
norm_cfg <- function(cfg) {
  for (f in c("actorPowerDrivers", "actorPowerIndex", "instQualityDrivers", "controlDrivers"))
    if (!is.null(cfg[[f]])) cfg[[f]] <- unlist(cfg[[f]])
  cfg$panelTransform <- cfg$panelTransform %||% "levels"
  cfg
}
sectors <- c("Bulk", "Diffuse")
cfg <- list()
for (sec in sectors) {
  hit <- Filter(function(x) identical(x$model_type, paste0("Adoption: ", sec)), all_cfg)
  cfg[[sec]] <- if (length(hit) > 0) norm_cfg(hit[[1]]) else NULL
}
if (all(vapply(cfg, is.null, logical(1)))) stop("No 'Adoption: <Sector>' entries in ", cfg_path)
for (sec in sectors) if (is.null(cfg[[sec]])) cfg[[sec]] <- cfg[[setdiff(sectors, sec)]]

# ── Data + fits (cached via model store) ───────────────────────────────────────
hist_mag <- getPanelDataHistoricalCached(
  aggregate = TRUE, y = 2000:2022, outputRegionMappingFile = "regionmapping_54.csv"
)
fits <- list()
for (sec in sectors) {
  c0 <- cfg[[sec]]
  fits[[sec]] <- estimateAdoptionModel(
    data = hist_mag, sector = sec,
    actorPowerDrivers = c0$actorPowerDrivers, actorPowerIndex = c0$actorPowerIndex,
    instQualityDrivers = c0$instQualityDrivers, controlDrivers = c0$controlDrivers,
    includeLaggedAdoption = isTRUE(c0$includeLagged),
    interactRegionFE = isTRUE(c0$interactRegionFE),
    regionMappingFixedEffects = c0$regionMappingFixedEffects,
    useMundlak = isTRUE(c0$useMundlak), gdpGovInteraction = isTRUE(c0$gdpGovInteraction),
    logisticTimeTrend = isTRUE(c0$logisticTimeTrend),
    panelTransform = c0$panelTransform,
    modelDir = model_dir, verbose = FALSE,
    compute = c(ame = TRUE, predictedProbs = TRUE)
  )
}

# Fitted probabilities aligned to estimation rows
fittedProbs <- function(fit) {
  m <- fit$model
  if (inherits(m, "logistf")) as.numeric(m$predict) else as.numeric(fitted(m))
}

# ── Scenario projections (selected spec; skipped for FD specs) ─────────────────
gdx_path <- params$gdxPath
if (nzchar(gdx_path) && !is_absolute_path(gdx_path)) gdx_path <- file.path(root, gdx_path)
scen_mag <- NULL
if (nzchar(gdx_path) && file.exists(gdx_path)) {
  scen_mag <- tryCatch(
    getPanelDataScenarioCached(gdxFile = gdx_path, aggregate = TRUE,
                               outputRegionMappingFile = "regionmapping_54.csv"),
    error = function(e) NULL
  )
}
proj <- list()
if (!is.null(scen_mag)) {
  for (sec in sectors) {
    proj[[sec]] <- tryCatch(
      projectSpecScenario(cfg[[sec]], sec, histData = hist_mag, scenarioData = scen_mag,
                          modelDir = model_dir),
      error = function(e) NULL
    )
  }
}
has_proj <- any(!vapply(proj, is.null, logical(1)))

# ── Historical region × year fitted probabilities (feeds thresholds, classification, maps) ──────
histProb <- bind_rows(lapply(sectors, function(sec) {
  f <- fits[[sec]]
  p <- fittedProbs(f); idx <- usedRows(f)
  d <- f$data[idx, , drop = FALSE]
  n <- min(length(p), nrow(d))
  data.frame(sector = sec, region = as.character(d$region[seq_len(n)]),
             year = as.integer(d$year[seq_len(n)]), y = as.numeric(f$model$y)[seq_len(n)],
             prob = p[seq_len(n)], stringsAsFactors = FALSE)
}))

# Combined historical and projected probabilities per region, year and sector (for future categories).
# NOTE: keep the `sector` column — downstream code filters `fullProb[fullProb$sector == sec, ]`, which
# silently returns 0 rows (empty classification/maps) if sector is dropped.
fullProb <- bind_rows(lapply(sectors, function(sec) {
  histD <- histProb[histProb$sector == sec, c("sector", "region", "year", "prob")]
  if (has_proj && !is.null(proj[[sec]])) {
    projD <- proj[[sec]][, c("region", "year", "prob")]
    projD$sector <- sec
    projD <- projD[!(projD$year %in% histD$year), ]  # avoid duplicate years at the seam
    bind_rows(histD, projD[, c("sector", "region", "year", "prob")])
  } else {
    histD
  }
}))

# Reference adoption thresholds per sector (ADR 0029): Sens-90 (default), Youden, base-rate.
thr <- lapply(sectors, function(sec) {
  d <- histProb[histProb$sector == sec, ]
  adoptionReferenceThresholds(d$y, d$prob)
})
names(thr) <- sectors

# Region → H12 fixed-effect group (for the regional-groups map, C1). Majority H12 per 54-region.
regionH12 <- tryCatch({
  m54 <- madrat::toolGetMapping("regionmapping_54.csv", type = "regional", where = "mappingfolder")
  m12 <- madrat::toolGetMapping("regionmappingH12.csv", type = "regional", where = "mappingfolder")
  c2h <- stats::setNames(as.character(m12$RegionCode), as.character(m12$CountryCode))
  tapply(as.character(m54$CountryCode), as.character(m54$RegionCode),
         function(cc) { tb <- sort(table(c2h[cc]), decreasing = TRUE); if (length(tb)) names(tb)[1] else NA_character_ })
}, error = function(e) NULL)
```

# Executive Summary

*What this section is for.* A high-level summary of the selected carbon pricing adoption model specifications and key diagnostics (including convergence, max VIF, theory fraction, and predictive metrics) for the Bulk and Diffuse sectors.

```{r exec, results='asis'}
for (sec in sectors) {
  f <- fits[[sec]]
  pd <- f$predictiveDiagnostics
  cat(sprintf(paste0(
    "- **%s** — `%s` | panelTransform: %s | N = %d | converged: %s | maxVIF: %.1f | ",
    "Theory Frac.: %.2f | AUC: %s | Brier: %s | calibration slope: %s\n"
  ), sec, cfg[[sec]]$name, cfg[[sec]]$panelTransform,
     if (inherits(f$model, "logistf")) f$model$n else stats::nobs(f$model),
     isTRUE(f$model$converged) || (inherits(f$model, "logistf") && !any(is.na(f$model$coefficients))),
     f$maxVIF %||% NA, f$theoryFrac %||% NA,
     if (!is.null(pd)) sprintf("%.3f", pd$auc) else "—",
     if (!is.null(pd)) sprintf("%.3f", pd$brier) else "—",
     if (!is.null(pd)) sprintf("%.2f", pd$calibrationSlope) else "—"))
}
cat(sprintf("\nScenario projections: %s\n",
            if (has_proj) "available (SSP2)" else "not available (no gdx or FD spec)"))
```

# Model Equations & Coefficients {.tabset}

*What this section is for.* The mathematical equations of the estimated logistic regression models and the standardized coefficient estimates (with robust 95% confidence intervals) for the Bulk and Diffuse sectors.

```{r coefs, results='asis', fig.height=6}
for (sec in sectors) {
  cat(sprintf("## %s\n\n", sec))
  f <- fits[[sec]]
  ct <- as.data.frame(unclass(f$coeftest))
  names(ct) <- c("estimate", "stdError", "zValue", "pValue")
  ct$term <- rownames(f$coeftest)
  ct$signif <- sigStars(ct$pValue)
  main <- ct %>% filter(!grepl("^regionFE", term))
  eq_terms <- sprintf("%+.3f·%s%s", main$estimate,
                      ifelse(main$term == "(Intercept)", "1", main$term), main$signif)
  cat(sprintf("```\nlogit P(adoption) = %s + [region FE]\n```\n\n",
              paste(eq_terms, collapse = " ")))
  cc <- ct %>% filter(!grepl("^regionFE|^\\(Intercept", term)) %>%
    mutate(lo = estimate - 1.96 * stdError, hi = estimate + 1.96 * stdError,
           sig = pValue < 0.05)
  p <- ggplot(cc, aes(estimate, reorder(term, estimate), colour = sig)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_pointrange(aes(xmin = lo, xmax = hi)) +
    scale_colour_manual(values = c(`TRUE` = "#d62728", `FALSE` = "grey60")) +
    theme_report() + theme(legend.position = "none") +
    labs(title = paste0(sec, " — coefficients (robust 95% CI)"),
         x = "Log-odds per 1 SD of the driver", y = NULL)
  print(p)
  cat("\n\n")
  print(knitr::kable(ct %>% select(term, estimate, stdError, zValue, pValue, signif) %>%
                       mutate(across(where(is.numeric), ~round(.x, 3))),
                     caption = paste0(sec, " — full coefficient table (per-SD drivers)")))
  cat("\n\n")
}
howToRead("The equation is on the log-odds scale with significance stars from ",
          "cluster-robust standard errors; the forest plot omits intercept and region ",
          "FE. Drivers are **standardized**, so a coefficient is the log-odds change per ",
          "**one-standard-deviation** increase in that driver — magnitudes are directly ",
          "comparable across drivers. Divide by a driver's SD for its natural-unit effect.")
```

# Real-World Effects — Average Marginal Effects {.tabset}

*What this section is for.* The coefficients above are on the log-odds scale, which is hard to
interpret directly. **Average Marginal Effects (AMEs)** translate each driver into the metric that
matters — **percentage points of adoption probability** — so you can read off "a one-SD increase in
this driver raises the adoption probability by X pp, on average". This is the closest single number
to a real-world effect size for the adoption stage, and the basis for the channel story told to
REMIND. One tab per sector.

```{r ame, results='asis'}
for (sec in sectors) {
  cat(sprintf("## %s\n\n", sec))
  am <- fits[[sec]]$ame
  if (!is.null(am)) {
    # Visual forest plot
    am_plot <- am %>%
      filter(term != "(Intercept)") %>%
      mutate(
        ame_pp = ame * 100,
        lower_pp = lower * 100,
        upper_pp = upper * 100,
        sig = p < 0.05
      )
    p <- ggplot(am_plot, aes(ame_pp, reorder(term, ame_pp), colour = sig)) +
      geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
      geom_pointrange(aes(xmin = lower_pp, xmax = upper_pp)) +
      scale_colour_manual(values = c(`TRUE` = "#d62728", `FALSE` = "grey60")) +
      theme_report() + theme(legend.position = "none") +
      labs(title = paste0(sec, " — Average Marginal Effects (percentage points, robust 95% CI)"),
           x = "Change in adoption probability (percentage points) per 1 SD", y = NULL)
    print(p)
    cat("\n\n")

    # Full data table
    am2 <- am
    num <- vapply(am2, is.numeric, logical(1))
    am2[num] <- lapply(am2[num], function(x) round(x * 100, 2))
    print(knitr::kable(am2, caption = paste0(sec, " — AME table in percentage points of P(adoption), per 1 SD")))
    cat("\n\n")
  } else cat("*AMEs not available.*\n\n")
}
howToRead("The chart (forest plot) and table display Average Marginal Effects (AMEs) in percentage points. ",
          "An AME is the average change in adoption probability (in percentage points) when a driver ",
          "increases by **one standard deviation** (drivers are standardized, making magnitudes directly ",
          "comparable), holding other covariates fixed. Red points and lines denote statistical significance (p < 0.05).")
```

# Historical Fit & Outliers {.tabset}

*What this section is for.* Diagnostics of how well the models fit the historical data (2000–2022), featuring fitted probability density plots by actual outcome and a detailed list of region-years with large Pearson residuals (outliers).

## Predicted probability by outcome

```{r fit-density, fig.height=4.5}
dd <- bind_rows(lapply(sectors, function(sec) {
  f <- fits[[sec]]
  data.frame(sector = sec, prob = fittedProbs(f),
             outcome = factor(f$model$y, labels = c("Not adopted", "Adopted")))
}))
ggplot(dd, aes(prob, fill = outcome)) +
  geom_density(alpha = 0.45, colour = NA) +
  facet_wrap(~sector) +
  scale_fill_manual(values = c("Not adopted" = "#1f77b4", "Adopted" = "#d62728")) +
  theme_report() +
  labs(title = "Fitted adoption probability by actual outcome",
       x = "Fitted P(adoption)", y = "Density", fill = NULL)
```

```{r fit-density-cap, results='asis'}
howToRead("Distributions of in-sample fitted probabilities, split by what actually ",
          "happened. Good discrimination = red mass near 1, blue mass near 0; heavy ",
          "overlap means the model cannot separate adopters from non-adopters.")
```

## Mispredicted region-years (|Pearson residual| > 3)

```{r outliers, results='asis'}
out_rows <- bind_rows(lapply(sectors, function(sec) {
  f <- fits[[sec]]
  p <- fittedProbs(f)
  y <- as.numeric(f$model$y)
  idx <- usedRows(f) # align fitted values to complete-case rows
  df <- f$data[idx, , drop = FALSE]
  n <- min(length(p), nrow(df))
  pr <- (y - p) / sqrt(pmax(p * (1 - p), 1e-12))
  data.frame(sector = sec, region = df$region[seq_len(n)], year = df$year[seq_len(n)],
             outcome = y[seq_len(n)], fittedProb = round(p[seq_len(n)], 3),
             pearsonResid = round(pr[seq_len(n)], 2))
})) %>% filter(abs(pearsonResid) > 3) %>% arrange(desc(abs(pearsonResid)))
if (nrow(out_rows) > 0) {
  interactiveTable(out_rows, caption = "Region-years the model badly mispredicts")
} else cat("*No region-years with |Pearson residual| > 3 — no adoption-side outliers.*\n")
```

```{r outliers-cap, results='asis'}
howToRead("Adoption-side outlier rule: observations where the model was confidently ",
          "wrong (adopted while fitted P was tiny, or vice versa). These region-years ",
          "deserve a story — a policy adopted against the model's odds usually marks a ",
          "driver the specification misses.", type = "table")
```

# Predicted Probability Profiles {.tabset}

*What this section is for.* Visualizing how the predicted adoption probability changes as each theory driver sweeps from its minimum to maximum value, holding other variables constant at their sample means.

```{r margins, results='asis', fig.height=4}
for (sec in sectors) {
  pp <- fits[[sec]]$predictedProbs
  if (is.null(pp) || length(pp) == 0) next
  cat(sprintf("## %s\n\n", sec))
  for (v in names(pp)) {
    prof <- pp[[v]]
    if (is.null(prof) || nrow(prof) == 0) next
    # The profile df is (variable, x, prob, lower, upper): x = the SWEPT driver value. Use it
    # explicitly — picking the first non-prob column grabbed the constant `variable` label and
    # drew every point at one x (a vertical line). 2026-06-26.
    xcol <- if ("x" %in% colnames(prof)) "x" else
      setdiff(colnames(prof), c("variable", "prob", "lower", "upper"))[1]
    p <- ggplot(prof, aes(x = .data[[xcol]], y = prob)) +
      geom_ribbon(aes(ymin = lower, ymax = upper), fill = "#1f77b4", alpha = 0.2) +
      geom_line(colour = "#1f77b4", linewidth = 1) +
      coord_cartesian(ylim = c(0, 1)) +
      theme_report() +
      labs(title = paste0(sec, " — P(adoption) vs ", v),
           x = paste0(v, " (standardized, per SD)"), y = "P(adoption)")
    print(p)
    cat("\n\n")
  }
}
howToRead("Each curve sweeps one theory variable from its observed minimum to maximum ",
          "while all other variables sit at sample means; the ribbon is the delta-method ",
          "95% interval. A flat curve with a wide ribbon = the channel is not doing ",
          "real work despite a significant coefficient.")
```

# Adoption Thresholds {.tabset}

*What this section is for.* The model outputs a continuous P(adoption); to talk about *which*
regions "adopt" you need a probability cutoff. This section shows, per sector, how the
classification quality changes across cutoffs and pins three reference thresholds (ADR 0029). The
**Sens-90** threshold is the default used by the classification and maps below.

```{r thresholds, results='asis'}
rogersPal <- c("Innovators" = "#1a9850", "Early Adopters" = "#66bd63", "Early Majority" = "#fee08b",
               "Late Majority" = "#fdae61", "Laggards" = "#d73027", "Non-Adopters" = "#bdbdbd")
for (sec in sectors) {
  cat(sprintf("## %s\n\n", sec))
  d <- histProb[histProb$sector == sec, ]
  tbl <- adoptionThresholdTable(d$y, d$prob)
  refs <- thr[[sec]]
  print(knitr::kable(tbl %>% mutate(across(where(is.numeric), ~round(.x, 3))),
                     caption = sprintf("%s — classification quality across thresholds", sec)))
  cat(sprintf(paste0("\n\n**Reference thresholds — %s:** Sens-90 = **%.2f** (default; catches ≥90%% ",
    "of historical adopters) · Youden-optimal = **%.2f** · base-rate = **%.2f**.\n\n"),
    sec, refs$Sens90, refs$Youden, refs$BaseRate))
  howToRead(
    "Each row is a candidate probability cutoff. **sensitivity** = share of historical adopter ",
    "region-years correctly flagged; **specificity** = share of non-adopters correctly left out; ",
    "**Youden J** = sensitivity + specificity − 1 (peak = best single cutoff); **predAdoptRate** = ",
    "share of region-years predicted to adopt. Sens-90 deliberately favours catching adopters ",
    "(few misses) over avoiding false positives.", type = "table")
}
```

# Adopter Classification (Rogers) {.tabset}

*What this section is for.* Ranking regions by **when the model says they adopt** (first year fitted
P crosses the threshold) and binning them into the classic Rogers diffusion groups — Innovators,
Early Adopters, Early/Late Majority, Laggards, Non-Adopters (ADR 0029). It answers "who leads and who
trails on carbon-pricing feasibility". The master table uses the **Sens-90** threshold and also shows
each region's group under the Youden and base-rate thresholds, so you can see which regions are
threshold-sensitive.

## Classification table

```{r classif-table}
clsAll <- bind_rows(lapply(sectors, function(sec) {
  d <- fullProb[fullProb$sector == sec, c("region", "year", "prob")]
  if (nrow(d) == 0) {
    return(data.frame(sector = character(0), region = character(0), adoptionYear = numeric(0),
                      Sens90_group = factor(character(0)), Youden_group = factor(character(0)),
                      BaseRate_group = factor(character(0)), changesWithThreshold = logical(0),
                      stringsAsFactors = FALSE))
  }
  s  <- rogersClassify(d, thr[[sec]]$Sens90)[, c("region", "adoptionYear", "group")]
  yo <- rogersClassify(d, thr[[sec]]$Youden)[, c("region", "group")]
  br <- rogersClassify(d, thr[[sec]]$BaseRate)[, c("region", "group")]
  m <- Reduce(function(a, b) merge(a, b, by = "region"),
              list(setNames(s, c("region", "adoptionYear", "Sens90_group")),
                   setNames(yo, c("region", "Youden_group")),
                   setNames(br, c("region", "BaseRate_group"))))
  m$sector <- sec
  m$changesWithThreshold <- !(as.character(m$Sens90_group) == as.character(m$Youden_group) &
                              as.character(m$Sens90_group) == as.character(m$BaseRate_group))
  m[, c("sector", "region", "adoptionYear", "Sens90_group", "Youden_group",
        "BaseRate_group", "changesWithThreshold")]
}))
interactiveTable(clsAll, caption = "Rogers adopter classification (Sens-90 default; group under each threshold)",
                 pageLength = 25, filter = "top")
```

```{r classif-table-cap, results='asis'}
howToRead(
  "One row per region/sector. `adoptionYear` is the first year the fitted probability crosses the ",
  "Sens-90 threshold (blank = never → Non-Adopter). `*_group` is the Rogers band under each ",
  "reference threshold; **changesWithThreshold = TRUE** flags regions whose band depends on the ",
  "threshold choice — read their classification cautiously.", type = "table")
```

## Band counts by threshold {.tabset}

```{r classif-counts, results='asis'}
for (tn in c("Sens90", "Youden", "BaseRate")) {
  cat(sprintf("### %s\n\n", tn))
  g_all <- bind_rows(lapply(sectors, function(sec) {
    d <- fullProb[fullProb$sector == sec, c("region", "year", "prob")]
    if (nrow(d) == 0) {
      return(data.frame(sector = character(0), group = factor(character(0)), stringsAsFactors = FALSE))
    }
    data.frame(sector = sec, group = rogersClassify(d, thr[[sec]][[tn]])$group, stringsAsFactors = FALSE)
  }))
  if (nrow(g_all) == 0) {
    cat("*No data available to construct the table.*\n\n")
    next
  }
  print(knitr::kable(as.data.frame.matrix(table(g_all$group, g_all$sector)),
                     caption = sprintf("Region counts per Rogers band — %s threshold", tn)))
  cat("\n\n")
}
howToRead(
  "How the band membership shifts with the threshold: a stricter (higher) cutoff pushes regions ",
  "later and swells Non-Adopters; a looser cutoff pulls regions into the early bands. Compare the ",
  "three tabs to gauge how robust the classification is.", type = "table")
```

# Spatial Distribution {.tabset}

*What this section is for.* The same information on a world map, so regional patterns are visible at
a glance. Regions are mapped to countries via `regionmapping_54`. (Maps need the `sf` and
`rnaturalearth` packages; without them this section shows a note instead.)

```{r maps, results='asis', fig.height=5.2}
lastY <- max(histProb$year)
mapsAvailable <- requireNamespace("sf", quietly = TRUE) && requireNamespace("rnaturalearth", quietly = TRUE)
if (!mapsAvailable) {
  cat("> Maps unavailable — install `sf` and `rnaturalearth` to render the spatial section.\n")
} else {
  ## A. Adoption probability (historical, last year)
  cat("## P(adoption) — historical {.tabset}\n\n")
  for (sec in sectors) {
    cat(sprintf("### %s\n\n", sec))
    d <- histProb[histProb$sector == sec & histProb$year == lastY, ]
    p <- plotRegionWorldMap(setNames(d$prob, d$region), discrete = FALSE, limits = c(0, 1),
      title = sprintf("Fitted P(adoption) — %s, %d", sec, lastY), legend = "P(adoption)")
    if (!is.null(p)) { print(p); cat("\n\n") } else cat("*Map could not be drawn.*\n\n")
  }
  howToRead("Each region shaded by its fitted adoption probability in the last historical year — ",
            "darker = more likely to have carbon pricing. The between-region gradient is the ",
            "cross-sectional feasibility picture.")

  ## A'. Scenario probability (if available)
  if (has_proj) {
    cat("## P(adoption) — scenario {.tabset}\n\n")
    projY <- 2050
    for (sec in names(proj)) {
      if (is.null(proj[[sec]])) next
      cat(sprintf("### %s\n\n", sec))
      dd <- proj[[sec]]; yy <- dd$year[which.min(abs(dd$year - projY))][1]
      d <- dd[dd$year == yy, ]
      p <- plotRegionWorldMap(setNames(d$prob, d$region), discrete = FALSE, limits = c(0, 1),
        title = sprintf("Projected P(adoption) — %s, %d (SSP2)", sec, yy), legend = "P(adoption)")
      if (!is.null(p)) { print(p); cat("\n\n") } else cat("*Map could not be drawn.*\n\n")
    }
    howToRead("Projected adoption probability around 2050 under SSP2 — compare with the historical ",
              "map to see where the scenario expects pricing to spread.")
  }

  ## B. Adopter category (Sens-90)
  cat("## Adopter category {.tabset}\n\n")
  for (sec in sectors) {
    cat(sprintf("### %s\n\n", sec))
    d <- fullProb[fullProb$sector == sec, c("region", "year", "prob")]
    cls <- rogersClassify(d, thr[[sec]]$Sens90)
    p <- plotRegionWorldMap(cls[, c("region", "group")], discrete = TRUE, palette = rogersPal,
      title = sprintf("Rogers adopter category — %s (Sens-90)", sec), legend = "Adopter group")
    if (!is.null(p)) { print(p); cat("\n\n") } else cat("*Map could not be drawn.*\n\n")
  }
  howToRead("Regions coloured by Rogers band (green = early, red = laggard, grey = never crosses ",
            "the threshold in sample) — the spatial view of the classification table above.")

  ## C1. H12 fixed-effect grouping
  cat("## Regional groups (H12 FE)\n\n")
  if (!is.null(regionH12)) {
    p <- plotRegionWorldMap(regionH12, discrete = TRUE,
      title = "H12 fixed-effect region grouping", legend = "H12 group")
    if (!is.null(p)) { print(p); cat("\n\n") } else cat("*Map could not be drawn.*\n\n")
    howToRead("The 12 REMIND macro-regions that the model's region fixed effects are grouped into — ",
              "regions sharing a colour share an estimated intercept (the FE strategy, ADR 0011).")
  } else cat("*Region→H12 mapping unavailable.*\n\n")
}
```

# Adoption Timeline — Early Warning {.tabset}

*What this section is for.* Where does the model **lead or lag** reality? The bar spans each region's
*observed* adoption period; the dot is the year the model's fitted probability first crosses the
Sens-90 threshold. A dot left of the bar = the model saw conditions ripen early; right of the bar =
the model was late.

```{r timeline, results='asis', fig.height=7}
for (sec in sectors) {
  cat(sprintf("## %s\n\n", sec))
  d <- histProb[histProb$sector == sec, ]
  obs <- d %>% filter(y == 1) %>% group_by(region) %>%
    summarise(obsStart = min(year), obsEnd = max(year), .groups = "drop")
  cls <- rogersClassify(d[, c("region", "year", "prob")], thr[[sec]]$Sens90)
  tl <- merge(obs, cls[, c("region", "adoptionYear")], by = "region", all = TRUE)
  tl <- tl %>% filter(is.finite(obsStart) | is.finite(adoptionYear)) %>%
    arrange(adoptionYear, obsStart)
  if (!nrow(tl)) { cat("*No adopters in sample.*\n\n"); next }
  tl$region <- factor(tl$region, levels = rev(unique(tl$region)))
  p <- ggplot(tl) +
    geom_segment(aes(x = obsStart, xend = obsEnd, y = region, yend = region),
                 colour = "grey70", linewidth = 2, na.rm = TRUE) +
    geom_point(aes(x = adoptionYear, y = region), colour = "#d62728", size = 2, na.rm = TRUE) +
    theme_report() +
    labs(title = sprintf("%s — observed adoption period (bar) vs model crossing (dot)", sec),
         x = NULL, y = NULL)
  print(p); cat("\n\n")
}
howToRead("Grey bar = observed adoption period; red dot = first year fitted P crosses Sens-90. ",
          "Dot before the bar = model anticipates (early warning); dot after = model lags; ",
          "bar with no dot = the model never reaches the threshold for an observed adopter.")
```

# Per-Region Fit {.tabset}

*What this section is for.* The granular diagnostic: each region's fitted probability over its
observed 0/1 adoption path. Collapsed by default (54 panels per sector).

```{r per-region, results='asis', fig.height=9}
for (sec in sectors) {
  cat(sprintf("## %s\n\n", sec))
  d <- histProb[histProb$sector == sec, ]
  p <- ggplot(d, aes(year)) +
    geom_step(aes(y = y), colour = "grey60") +
    geom_line(aes(y = prob), colour = "#1f77b4", linewidth = 0.7) +
    facet_wrap(~region, ncol = 6) +
    coord_cartesian(ylim = c(0, 1)) +
    theme_report() +
    theme(strip.text = element_text(size = 6), axis.text = element_text(size = 5)) +
    labs(title = sprintf("%s — fitted P(adoption) (blue) vs observed adoption (grey step)", sec),
         x = NULL, y = NULL)
  print(p); cat("\n\n")
}
howToRead("Per region: grey step = observed 0/1 adoption, blue = fitted probability. A good fit has ",
          "blue rising into the grey step. Blue high before the step = early signal; blue flat-low ",
          "under a step = the spec misses that region's adoption.")
```

# Future Projections (SSP2) {.tabset}

*What this section is for.* Projections of carbon pricing adoption probabilities under the SSP2 scenario up to 2100. It shows probability trajectories over time, region snapshots at specific horizons, and a seam check between the last historical and first scenario year.

```{r proj, results='asis', fig.height=6}
if (has_proj) {
  pall <- bind_rows(proj)
  cat("## Probability trajectories\n\n")
  p <- ggplot(pall, aes(year, prob, group = region)) +
    geom_line(alpha = 0.35, linewidth = 0.4, colour = "#1f77b4") +
    stat_summary(aes(group = 1), fun = median, geom = "line",
                 colour = "#d62728", linewidth = 1.2) +
    facet_wrap(~sector) +
    coord_cartesian(ylim = c(0, 1)) +
    theme_report() +
    labs(title = "Projected adoption probabilities (SSP2)",
         x = NULL, y = "P(adoption)")
  print(p)
  cat("\n\n")
  howToRead("One line per region; red = cross-region median. Watch for: everyone ",
            "saturating at 1 (the trend assumption doing all the work — see the ",
            "logistic-trend chart in the panel-data report) and regions stuck at 0 ",
            "(dead blocks, a severe Projection Sanity rule).")

  cat("## Snapshot table\n\n")
  # Base-R pivot (no tidyr dependency): one column per snapshot year present in the projection.
  yrs <- intersect(c(2030, 2050, 2070, 2100), sort(unique(pall$year)))
  sl <- pall[pall$year %in% yrs, c("sector", "region", "year", "prob")]
  if (length(yrs) && nrow(sl)) {
    snap <- Reduce(function(a, b) merge(a, b, by = c("sector", "region"), all = TRUE),
      lapply(yrs, function(yy) {
        d <- sl[sl$year == yy, c("sector", "region", "prob")]
        names(d)[3] <- paste0("P", yy); d
      }))
    snap <- snap %>% mutate(across(where(is.numeric), ~round(.x, 3))) %>% arrange(sector, region)
    print(knitr::kable(snap, caption = "P(adoption) at snapshot years"))
  } else cat("*No projection rows at the snapshot years.*\n")
  cat("\n\n")
  howToRead("Projected adoption probability per region at four snapshot years — sort a ",
            "column to rank regions by projected feasibility at that horizon.",
            type = "table")

  cat("## Seam check\n\n")
  seam <- bind_rows(lapply(sectors, function(sec) {
    if (is.null(proj[[sec]]) || !nrow(proj[[sec]])) return(NULL)   # skip non-projectable sector
    f <- fits[[sec]]
    p <- fittedProbs(f)
    idx <- usedRows(f)
    df <- f$data[idx, , drop = FALSE]
    n <- min(length(p), nrow(df))
    df <- df[seq_len(n), , drop = FALSE]
    p <- p[seq_len(n)]
    lastY <- max(df$year)
    hist_last <- data.frame(region = df$region[df$year == lastY],
                            histProb = p[df$year == lastY])
    firstY <- min(proj[[sec]]$year)
    scen_first <- proj[[sec]] %>% filter(year == firstY) %>% select(region, prob)
    merge(hist_last, scen_first, by = "region") %>%
      mutate(sector = sec, jump = prob - histProb) %>%
      arrange(desc(abs(jump)))
  }))
  if (!is.null(seam) && nrow(seam)) {
    print(knitr::kable(seam %>% mutate(across(where(is.numeric), ~round(.x, 3))),
                       caption = "Fitted (last historical) vs projected (first scenario) probability"))
    cat("\n\n")
    howToRead("The model-level seam test: fitted probability at the last historical year ",
              "vs the projection at the first scenario year, per region. |jump| > 0.3 is ",
              "the Projection Sanity warning threshold — large jumps mean scenario drivers ",
              "are discontinuous with history for that region.", type = "table")
  } else cat("*No projectable sector for the seam check.*\n")
} else {
  cat("> Projections unavailable: no scenario panel, or the selected spec is an FD ",
      "transform (not projectable — ADR 0005).\n")
}
```

# Measure Definitions

*What this section is for.* A glossary defining all statistical measures and diagnostics used throughout this adoption report.

```{r defs}
interactiveTable(
  measureDefinitions(groups = c("Predictive (report-only)", "Model form", "Theory content",
                                "Tier", "Projection")),
  caption = "Measure / statistic glossary (adoption-relevant; full card in the selection report)",
  pageLength = 30, filter = "top")
```

```{r defs-cap, results='asis'}
howToRead("Definitions for the statistics shown in this report, drawn from the shared canonical ",
          "glossary (`pfmreports::measureDefinitions`) used across all PFM reports. 'Report-only' ",
          "measures never influence which model is chosen.", type = "table")
```
