---
title: "PFM Results — Price Stringency Model (`r params$reportName`)"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 1
    toc_float:
      collapsed: true
    number_sections: true
    code_folding: hide
params:
  modelConfig: "results/exhaustive/selected-models.yml"
  reportName: "default"
  gdxPath: "data/fulldata.gdx"
  cachefolder: ""
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 5.5)
library(ggplot2)
library(dplyr)
library(magclass)
library(pfm)

root <- getwd()
library(pfmreports)

if (nzchar(params$cachefolder)) madrat::setConfig(cachefolder = params$cachefolder)
madrat::setConfig(forcecache = TRUE)
model_dir <- getPfmConfig("modelDir", "../output")
if (!is_absolute_path(model_dir)) model_dir <- file.path(root, model_dir)
options(pfm.modelDir = model_dir)

cfg_path <- params$modelConfig
if (!is_absolute_path(cfg_path)) cfg_path <- file.path(root, cfg_path)
all_cfg <- yaml::read_yaml(cfg_path)
norm_cfg <- function(cfg) {
  for (f in c("actorPowerDrivers", "actorPowerIndex", "instQualityDrivers", "controlDrivers"))
    if (!is.null(cfg[[f]])) cfg[[f]] <- unlist(cfg[[f]])
  cfg$panelTransform <- cfg$panelTransform %||% "levels"
  cfg
}
sectors <- c("Bulk", "Diffuse")
get_cfg <- function(stage) {
  out <- list()
  for (sec in sectors) {
    hit <- Filter(function(x) identical(x$model_type, paste0(stage, ": ", sec)), all_cfg)
    out[[sec]] <- if (length(hit) > 0) norm_cfg(hit[[1]]) else NULL
  }
  for (sec in sectors) if (is.null(out[[sec]])) out[[sec]] <- out[[setdiff(sectors, sec)]]
  out
}
cfgS <- get_cfg("Stringency")
cfgA <- get_cfg("Adoption") # used only for the hurdle expectation
if (all(vapply(cfgS, is.null, logical(1)))) stop("No 'Stringency: <Sector>' entries in ", cfg_path)

hist_mag <- getPanelDataHistoricalCached(
  aggregate = TRUE, y = 2000:2022, outputRegionMappingFile = "regionmapping_54.csv"
)
fits <- list()
for (sec in sectors) {
  c0 <- cfgS[[sec]]
  fits[[sec]] <- estimatePriceStringencyModel(
    data = hist_mag, sector = sec, family = "gaussian",
    actorPowerDrivers = c0$actorPowerDrivers, actorPowerIndex = c0$actorPowerIndex,
    instQualityDrivers = c0$instQualityDrivers, controlDrivers = c0$controlDrivers,
    includeLaggedECP = isTRUE(c0$includeLaggedECP) || isTRUE(c0$includeLagged),
    interactRegionFE = isTRUE(c0$interactRegionFE),
    regionMappingFixedEffects = c0$regionMappingFixedEffects,
    useMundlak = isTRUE(c0$useMundlak), gdpGovInteraction = isTRUE(c0$gdpGovInteraction),
    logisticTimeTrend = isTRUE(c0$logisticTimeTrend),
    panelTransform = c0$panelTransform,
    nickellCorrection = isTRUE(c0$nickellCorrection),
    priceLink = c0$priceLink %||% "log1p",
    priceCeilingMax = c0$priceCeilingMax %||% 1000,
    modelDir = model_dir, verbose = FALSE
  )
}

# ── Scenario projections: stringency price x adoption probability ─────────────
gdx_path <- params$gdxPath
if (nzchar(gdx_path) && !is_absolute_path(gdx_path)) gdx_path <- file.path(root, gdx_path)
scen_mag <- NULL
if (nzchar(gdx_path) && file.exists(gdx_path)) {
  scen_mag <- tryCatch(
    getPanelDataScenarioCached(gdxFile = gdx_path, aggregate = TRUE,
                               outputRegionMappingFile = "regionmapping_54.csv"),
    error = function(e) NULL
  )
}
proj <- list()
if (!is.null(scen_mag)) {
  for (sec in sectors) {
    pS <- tryCatch(projectSpecScenario(cfgS[[sec]], sec, hist_mag, scen_mag,
                                       modelDir = model_dir), error = function(e) NULL)
    pA <- if (!is.null(cfgA[[sec]])) {
      tryCatch(projectSpecScenario(cfgA[[sec]], sec, hist_mag, scen_mag,
                                   modelDir = model_dir), error = function(e) NULL)
    } else NULL
    if (!is.null(pS)) {
      # Carry the UNCLAMPED model price + the clamp-pinned flag (ADR 0032): the report displays the
      # raw projection and uses the clamp only as a "beyond the extrapolation guard" indicator.
      keep <- intersect(c("region", "year", "sector", "price", "priceUnclamped",
                          "clampPinned", "pmaxPinned"), colnames(pS))
      out <- pS[, keep, drop = FALSE]
      if (!"priceUnclamped" %in% names(out)) out$priceUnclamped <- out$price          # older fits
      if (!"clampPinned" %in% names(out)) out$clampPinned <- FALSE
      if (!"pmaxPinned" %in% names(out)) out$pmaxPinned <- FALSE
      if (!is.null(pA)) {
        out <- merge(out, pA %>% select(region, year, prob), by = c("region", "year"))
        out$expectedPrice <- out$prob * out$priceUnclamped                            # unclamped
      } else {
        out$prob <- NA_real_
        out$expectedPrice <- out$priceUnclamped
      }
      proj[[sec]] <- out
    }
  }
}
has_proj <- length(proj) > 0

# Shared display cap for ALL price charts (ADR 0032): the p99/p99.5/p99.9 of the pooled prices that is
# closest to 1000 USD/tCO2 without exceeding it. A disclosed display cap (not the model clamp).
histPx <- unlist(lapply(sectors, function(sec) {
  f <- fits[[sec]]; expm1(c(as.numeric(f$model$y), as.numeric(stats::fitted(f$model))))
}))
projPx <- if (has_proj) { p0 <- bind_rows(proj); c(p0$priceUnclamped, p0$expectedPrice) } else numeric(0)
priceCap <- priceDisplayCap(c(histPx, projPx), target = 1000)
```

# Executive Summary

*What this section is for.* A high-level overview of the selected Price Stringency model specifications and key performance diagnostics (RMSE, obs-fit correlation) for the Bulk and Diffuse sectors.

```{r exec, results='asis'}
for (sec in sectors) {
  f <- fits[[sec]]
  pd <- f$predictiveDiagnostics
  cat(sprintf(paste0(
    "- **%s** — `%s` | panelTransform: %s | family: %s | N = %d | maxVIF: %.1f | ",
    "in-sample RMSE (estimation scale): %s | obs-fit correlation: %s\n"
  ), sec, cfgS[[sec]]$name, cfgS[[sec]]$panelTransform, f$family,
     length(f$model$y), f$maxVIF %||% NA,
     if (!is.null(pd)) sprintf("%.3f", pd$rmse) else "—",
     if (!is.null(pd)) sprintf("%.3f", pd$corObsPred) else "—"))
}
cat(sprintf("\nHurdle projections (adoption-gated expected price): %s\n",
            if (has_proj) "available (SSP2)" else "not available"))
```

# Model Equations & Coefficients {.tabset}

*What this section is for.* The mathematical price equations and standardized coefficient estimates (with robust 95% confidence intervals) for each sector, allowing direct comparison of driver influence on price levels.

```{r coefs, results='asis', fig.height=6}
for (sec in sectors) {
  cat(sprintf("## %s\n\n", sec))
  f <- fits[[sec]]
  ct <- as.data.frame(unclass(f$coeftest))
  names(ct) <- c("estimate", "stdError", "zValue", "pValue")
  ct$term <- rownames(f$coeftest)
  ct$signif <- sigStars(ct$pValue)
  main <- ct %>% filter(!grepl("^regionFE", term))
  eq_terms <- sprintf("%+.3f·%s%s", main$estimate,
                      ifelse(main$term == "(Intercept)", "1", main$term), main$signif)
  cat(sprintf("```\nE[log(1+ECP)] = %s + [region FE]   (gaussian identity; price = expm1(.))\n```\n\n",
              paste(eq_terms, collapse = " ")))
  cc <- ct %>% filter(!grepl("^regionFE|^\\(Intercept", term)) %>%
    mutate(lo = estimate - 1.96 * stdError, hi = estimate + 1.96 * stdError,
           sig = pValue < 0.05)
  p <- ggplot(cc, aes(estimate, reorder(term, estimate), colour = sig)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_pointrange(aes(xmin = lo, xmax = hi)) +
    scale_colour_manual(values = c(`TRUE` = "#d62728", `FALSE` = "grey60")) +
    theme_report() + theme(legend.position = "none") +
    labs(title = paste0(sec, " — coefficients (robust 95% CI)"),
         x = "Estimate (per 1 SD of the driver)", y = NULL)
  print(p)
  cat("\n\n")
  print(knitr::kable(ct %>% select(term, estimate, stdError, zValue, pValue, signif) %>%
                       mutate(across(where(is.numeric), ~round(.x, 3))),
                     caption = paste0(sec, " — full coefficient table (per-SD drivers)")))
  cat("\n\n")
}
howToRead("Stringency coefficients are on the log(1+ECP) scale (a log-linear model: ",
          "gaussian identity link on the log-transformed price, so E[log(1+ECP)] = Xβ ",
          "and price = expm1(Xβ)). Drivers are **standardized**, so each coefficient is ",
          "the effect of a **one-standard-deviation** increase in that driver — magnitudes ",
          "are directly comparable across drivers (the largest is the most influential). ",
          "To express a coefficient in natural units, divide it by that driver's SD. Read ",
          "direction and significance here; the next section gives real-world magnitudes.")
```

# Real-World Effects — Semi-Elasticities & USD Effects {.tabset}

*What this section is for.* Translating log-linear coefficients into real-world price changes. It reports the percentage change in price (semi-elasticity) and the absolute dollar effect per one standard deviation increase in each driver, calculated at the mean fitted historical price.

```{r semi, results='asis', fig.height=4.5}
for (sec in sectors) {
  cat(sprintf("## %s\n\n", sec))
  f <- fits[[sec]]
  ct <- as.data.frame(unclass(f$coeftest))
  names(ct) <- c("estimate", "stdError", "zValue", "pValue")
  ct$term <- rownames(f$coeftest)
  meanFitted <- mean(expm1(as.numeric(fitted(f$model))), na.rm = TRUE)
  eff <- ct %>%
    filter(!grepl("^regionFE|^\\(Intercept", term)) %>%
    mutate(
      semiElasticityPct = round((exp(estimate) - 1) * 100, 1),
      effectUSDtCO2 = round((exp(estimate) - 1) * meanFitted, 2),
      sig = pValue < 0.05,
      signif = sigStars(pValue)
    )
  print(knitr::kable(eff %>% select(term, estimate, semiElasticityPct, effectUSDtCO2, pValue, signif) %>%
                       mutate(across(where(is.numeric), ~round(.x, 3))), caption = sprintf(
    "%s — effect per 1 SD of driver (mean fitted price %.1f USD/tCO2)", sec, meanFitted)))
  cat("\n\n")
  # Lollipop chart of the dollar effect per driver (sorted, signed, significance-coloured).
  pe <- ggplot(eff, aes(x = effectUSDtCO2, y = reorder(term, effectUSDtCO2), colour = sig)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_segment(aes(x = 0, xend = effectUSDtCO2, yend = term), linewidth = 0.6) +
    geom_point(size = 2.6) +
    scale_colour_manual(values = c(`TRUE` = "#d62728", `FALSE` = "grey60"),
                        labels = c(`TRUE` = "p < 0.05", `FALSE` = "n.s."), name = NULL) +
    theme_report() +
    labs(title = sprintf("%s — price effect per 1 SD of driver", sec),
         x = "Δ price (USD/tCO2) at the mean fitted price", y = NULL)
  print(pe)
  cat("\n\n")
}
howToRead("The lollipop chart shows the dollar price effect (effectUSDtCO2) per one-SD increase in ",
          "each driver, evaluated at the mean fitted price — bars right of zero raise the price, ",
          "left lower it; red = significant at 5%, length = magnitude (the longest bar is the most ",
          "influential driver). The table adds the percentage (semi-elasticity) view. ",
          "semiElasticityPct = (e^β − 1)·100: the % change in the (log1p) price index ",
          "per **one standard deviation** increase in the driver (drivers are standardized, ",
          "so effects are comparable across drivers). effectUSDtCO2 translates that to ",
          "dollars at the mean fitted price — a rough but intuitive 'so what' number. ",
          "For a natural-unit effect, rescale by the driver's SD.",
          type = "table")
```

# Historical Fit & Outliers {.tabset}

*What this section is for.* Diagnostic tools to evaluate model performance on the historical dataset (2000–2022). It includes calibration curves, historical price trajectory plots, list of price outliers, and residual diagnostics.

## Calibration (45°)

```{r calib, fig.height=5}
cal <- bind_rows(lapply(sectors, function(sec) {
  f <- fits[[sec]]
  data.frame(sector = sec,
             observed = as.numeric(f$model$y),
             fitted = as.numeric(fitted(f$model)))
}))
ggplot(cal, aes(fitted, observed)) +
  geom_abline(slope = 1, intercept = 0, colour = "grey50", linetype = "dashed") +
  geom_point(alpha = 0.4, colour = "#1f77b4", size = 1.2) +
  geom_smooth(method = "lm", se = FALSE, colour = "#d62728", linewidth = 0.8) +
  facet_wrap(~sector) +
  theme_report() +
  labs(title = "Observed vs fitted (estimation scale: log(1+ECP))",
       x = "Fitted", y = "Observed")
```

```{r calib-cap, results='asis'}
howToRead("Each dot is an adopting region-year on the estimation scale (log(1+ECP) for ",
          "levels specs, Δlog(1+ECP) for FD). Perfect calibration = dots on the dashed ",
          "45° line; the red regression line bending away from it shows where the model ",
          "systematically over/under-predicts price levels.")
```

## Price trajectories (USD/tCO2)

```{r traj, fig.height=5}
tr <- bind_rows(lapply(sectors, function(sec) {
  f <- fits[[sec]]
  idx <- usedRows(f) # fitted values align to complete-case rows only
  df <- f$data[idx, , drop = FALSE]
  data.frame(sector = sec, region = df$region, year = df$year,
             observed = expm1(as.numeric(f$model$y)),
             fitted = expm1(as.numeric(fitted(f$model))))
}))
nAbove <- sum(c(tr$observed, tr$fitted) > priceCap, na.rm = TRUE)
trL <- tr %>%
  tidyr::pivot_longer(c(observed, fitted), names_to = "kind", values_to = "price") %>%
  mutate(price = pmin(price, priceCap))
ggplot(trL, aes(year, price, colour = kind, group = interaction(region, kind))) +
  geom_line(alpha = 0.35, linewidth = 0.35) +
  facet_wrap(~sector) +
  scale_colour_manual(values = c(observed = "#1f77b4", fitted = "#d62728")) +
  coord_cartesian(ylim = c(0, priceCap)) +
  theme_report() +
  labs(title = "Observed vs fitted carbon prices (adopting region-years)",
       x = NULL, y = "USD/tCO2", colour = NULL,
       subtitle = sprintf("Shared display cap %.0f USD/tCO2 (%d value(s) above)", priceCap, nAbove))
```

```{r traj-cap, results='asis'}
howToRead("Back-transformed to USD/tCO2 (display-capped at p99 — Price Outlier rule). ",
          "Blue = observed, red = model fit, one line per adopting region. Persistent ",
          "red-above-blue (or below) for a region indicates a level bias the region FE ",
          "did not absorb.")
```

## Price outliers

```{r outliers, results='asis'}
obs_pos <- tr$observed[is.finite(tr$observed) & tr$observed > 0]
p99 <- as.numeric(quantile(obs_pos, 0.99, na.rm = TRUE))
out_rows <- tr %>%
  filter(observed > p99 | fitted > p99) %>%
  mutate(across(c(observed, fitted), ~round(.x, 1))) %>%
  arrange(desc(pmax(observed, fitted)))
cat(sprintf("**Price Outlier threshold (p99 of positive historical ECP): %.1f USD/tCO2.**\n\n", p99))
if (nrow(out_rows) > 0) {
  interactiveTable(out_rows, caption = "Region-years above the p99 price threshold (observed or fitted)")
} else cat("*No outliers above p99.*\n")
```

```{r outliers-cap, results='asis'}
howToRead("Every region-year whose observed or fitted price exceeds the p99 threshold. ",
          "These are kept in estimation (never dropped) but capped in charts; a fitted ",
          "value here without a matching observed one means the model invents extreme ",
          "prices for that region — check its drivers.", type = "table")
```

## Residual diagnostics

```{r resid, fig.height=4.5}
rd <- bind_rows(lapply(sectors, function(sec) {
  f <- fits[[sec]]
  data.frame(sector = sec,
             fitted = as.numeric(fitted(f$model)),
             residual = as.numeric(f$model$y) - as.numeric(fitted(f$model)))
}))
ggplot(rd, aes(fitted, residual)) +
  geom_hline(yintercept = 0, colour = "grey50", linetype = "dashed") +
  geom_point(alpha = 0.4, colour = "#1f77b4", size = 1.1) +
  geom_smooth(se = FALSE, colour = "#d62728", linewidth = 0.8) +
  facet_wrap(~sector) +
  theme_report() +
  labs(title = "Residuals vs fitted (estimation scale)", x = "Fitted", y = "Residual")
```

```{r resid-cap, results='asis'}
howToRead("Raw residuals against fitted values. A flat red smoother around zero is ",
          "healthy; funnels (heteroskedasticity) or curvature (missed nonlinearity) ",
          "show where the log-linear functional form struggles.")
```

# Future Projections (SSP2) {.tabset}

*What this section is for.* Projections of hurdle-expected carbon prices under the SSP2 scenario up to 2100. It combines the estimated adoption probability and the price stringency model to compute the expected price (probability of adoption multiplied by the conditional price) for each region, and displays trajectories, snapshot rankings, and price distributions.

```{r proj, results='asis', fig.height=6}
if (has_proj) {
  pall <- bind_rows(proj)

  cat("## Hurdle-expected price trajectories\n\n")
  nAboveP <- sum(pall$expectedPrice > priceCap, na.rm = TRUE)
  pp <- pall %>% mutate(expCapped = pmin(expectedPrice, priceCap))
  p <- ggplot(pp, aes(year, expCapped, group = region)) +
    geom_line(alpha = 0.35, linewidth = 0.4, colour = "#1f77b4") +
    stat_summary(aes(group = 1), fun = median, geom = "line",
                 colour = "#d62728", linewidth = 1.2) +
    facet_wrap(~sector) +
    theme_report() +
    coord_cartesian(ylim = c(0, priceCap)) +
    labs(title = "Projected expected carbon price: P(adoption) × price | adoption (UNCLAMPED)",
         x = NULL, y = "USD/tCO2",
         subtitle = sprintf("Shared display cap %.0f USD/tCO2 (%d region-year(s) above)", priceCap, nAboveP))
  print(p)
  cat("\n\n")

  # Sanity indicator (ADR 0032): region-years where the UNCLAMPED model price exceeds the
  # extrapolation guard (clampPinned) — i.e. the model is projecting beyond its data there. The clamp
  # is NOT applied to the displayed value; it is shown here as a "way off" flag.
  pinFrac <- if ("clampPinned" %in% names(pall)) mean(pall$clampPinned, na.rm = TRUE) else NA_real_
  cat(sprintf("**Extrapolation indicator:** %s of projected region-years are beyond the extrapolation guard%s.\n\n",
              if (is.na(pinFrac)) "n/a" else sprintf("%.0f%%", 100 * pinFrac),
              if (!is.na(pinFrac) && pinFrac > 0.25) " — **> 25%, a severe clamp-reliance flag (this spec would be rejected in selection)**" else ""))
  # Saturating form: bounded by Pmax, so it bypasses the clamp/clamp-reliance and instead reports a
  # NON-severe Pmax-ceiling indicator (ADR 0032).
  pmaxFrac <- if ("pmaxPinned" %in% names(pall)) mean(pall$pmaxPinned, na.rm = TRUE) else NA_real_
  if (!is.na(pmaxFrac) && pmaxFrac > 0) {
    cat(sprintf(paste0("**Pmax-ceiling indicator (saturating form):** %.0f%% of region-years project ",
                "at the Pmax ceiling — bounded by construction, **non-severe** (the saturating form ",
                "bypasses the extrapolation clamp/gate, ADR 0032).\n\n"), 100 * pmaxFrac))
  }
  guardrows <- if ("clampPinned" %in% names(pall)) {
    pall %>% filter(clampPinned) %>%
      mutate(priceUnclamped = round(priceUnclamped, 1), expectedPrice = round(expectedPrice, 1)) %>%
      select(sector, region, year, priceUnclamped, expectedPrice) %>%
      arrange(desc(priceUnclamped)) %>% head(50)
  } else pall[0, ]
  if (nrow(guardrows) > 0) {
    print(knitr::kable(guardrows, caption = "Region-years beyond the extrapolation guard (model extrapolating past data)"))
  } else cat("*No region-years exceed the extrapolation guard — the projection stays within data range.*\n")
  cat("\n\n")

  howToRead("The hurdle expectation per region: adoption probability times the **unclamped** ",
            "conditional price (the genuine model projection, ADR 0032 — the extrapolation clamp is ",
            "NOT applied to the displayed value), back-transformed to USD/tCO2. Red = median region. ",
            "The y-axis is a shared *display* cap (the p99/p99.5/p99.9 nearest 1000 USD/tCO2) common ",
            "to all price charts; the few region-years above it are listed in the extrapolation table ",
            "above. Levels are modest because this is the *expected* price = P(adoption) × conditional ",
            "price (P < 1), not because of a clamp. Specs that lean on the guard for > 25% of ",
            "region-years are rejected from selection — so the deliverable projects within its data.")

  cat("## Price ranking at snapshots\n\n")
  snap <- pall %>%
    filter(year %in% c(2030, 2050, 2070, 2100)) %>%
    mutate(expectedPrice = round(expectedPrice, 1)) %>%
    select(sector, region, year, expectedPrice) %>%
    tidyr::pivot_wider(names_from = year, values_from = expectedPrice, names_prefix = "y")
  print(knitr::kable(snap %>% mutate(across(where(is.numeric), ~round(.x, 1))),
                     caption = "Hurdle-expected price (USD/tCO2) at snapshot years"))
  cat("\n\n")
  howToRead("Sortable per-region expected prices at four horizons — the ranking view ",
            "for 'which regions can sustain high carbon prices, when'. Values above the ",
            "2000 USD/tCO2 sanity threshold would be severe Projection Sanity flags.",
            type = "table")

  cat("## Projected price distribution\n\n")
  pv <- pall %>% filter(year %in% c(2030, 2050, 2070, 2100))
  pv$expectedPrice <- pmin(pv$expectedPrice, priceCap)
  p2 <- ggplot(pv, aes(factor(year), expectedPrice)) +
    geom_violin(fill = "#1f77b4", alpha = 0.3, colour = "#1f77b4", scale = "width") +
    geom_boxplot(width = 0.15, outlier.size = 1, outlier.colour = "grey40", fill = "white", alpha = 0.6) +
    facet_wrap(~sector) +
    theme_report() +
    coord_cartesian(ylim = c(0, priceCap)) +
    labs(title = "Distribution of expected prices across regions",
         x = NULL, y = "USD/tCO2", subtitle = sprintf("Shared display cap %.0f USD/tCO2", priceCap))
  print(p2)
  cat("\n\n")
  howToRead("Cross-region distribution of the hurdle-expected price at each snapshot. ",
            "A widening violin = growing regional divergence in feasible carbon pricing; ",
            "the median line of the box is the 'typical region'.")
} else {
  cat("> Projections unavailable: no scenario panel, or the selected spec is an FD ",
      "transform (not projectable — ADR 0005).\n")
}
```
