---
title: "PFM — Robustness & Alternative Specifications"
subtitle: "Overfitting audit, parsimony frontier, and control specification-curve for the selected models"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    df_print: kable
params:
  robustnessRds: "results/exhaustive/robustness.rds"
  workflowRds: "results/exhaustive/sweep.rds"
  difRds: "results/exhaustive/difference-first.rds"
  temporalRds: "results/exhaustive/temporal-split.rds"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, fig.width = 10, fig.height = 5.5)
library(ggplot2); library(dplyr); library(knitr); library(kableExtra); library(tidyr)
rb  <- readRDS(params$robustnessRds)
wf  <- tryCatch(readRDS(params$workflowRds), error = function(e) NULL)
dif <- tryCatch(readRDS(params$difRds), error = function(e) NULL)
temporal <- tryCatch(readRDS(params$temporalRds)$temporal, error = function(e) NULL)
ladder <- rb$ladder; frontier <- rb$frontier; cc <- rb$ctrlCurve
deliv <- rb$specs
tierColor <- c(Green = "#1a9850", Blue = "#4575b4", Yellow = "#d6a000")
fmt <- function(x, d = 3) ifelse(is.na(x), "—", formatC(round(x, d), format = "f", digits = d))
specName <- function(stage) sub(" .*", "", deliv[[stage]]$name)  # X-id
```

# Executive summary

This report audits the two **deliverable** specifications against alternative models, to bound the
over-fitting / specification-search / small-N limitations identified in the paper draft. It does **not**
change the deliverable; it characterises how much of the headline fit is structural and which simpler
alternatives a stricter analyst might prefer.

- **Adoption deliverable:** `r deliv$Adoption$name`
- **Stringency deliverable:** `r deliv$Stringency$name`

```{r headline, results='asis'}
fd  <- ladder %>% filter(rung == "4. First-difference")
lev <- ladder %>% filter(rung == "1. Deliverable")
ds_lev <- lev$pseudoR2[lev$stage=="Stringency" & lev$sector=="Diffuse"]
ds_fd  <- fd$pseudoR2 [fd$stage =="Stringency" & fd$sector =="Diffuse"]
cat(sprintf(
"> **Headline.** The Diffuse-stringency pseudo-R² falls from **%.2f** in the levels deliverable to **%.2f** under first-differences — most of the levels fit is cross-sectional persistence (long-standing OECD transport taxes), not year-on-year responsiveness of prices to changes in actor power or governance. The theory channels' *signs* survive, but the **level of fit should be read as a relative feasibility signal, not a forecast**.\n", ds_lev, ds_fd))
```

# 1. Robustness Ladder

Anchored, one-knob-at-a-time: each rung changes exactly one modelling choice from the deliverable, so its
effect is attributable. `ΔR²(theory)` = incremental McFadden pseudo-R² of the theory channels; `Tier` =
Green (all 3 channels significant) / Blue (≥1) / Yellow (none).

```{r ladder-tables, results='asis'}
for (stg in c("Adoption","Stringency")) {
  cat("\n## ", stg, "\n\n")
  d <- ladder %>% filter(stage == stg) %>%
    transmute(Rung = rung, Sector = sector, Params = nParams,
              `ΔR²(theory)` = round(deltaR2Theory,3), `Pseudo-R²` = round(pseudoR2,3),
              BIC = round(bic,1), `AP×IQ coef` = round(apIQcoef,3), Tier = tier)
  k <- kable(d, format = "html", align = "llrrrrrl") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12)
  for (i in seq_len(nrow(d))) k <- row_spec(k, i, color = tierColor[d$Tier[i]])
  k <- row_spec(k, which(d$Rung == "1. Deliverable"), bold = TRUE)
  cat(k); cat("\n\n")
}
```

# 2. The first-difference test (where the R² goes)

First-differencing strips time-invariant regional levels, isolating whether *changes* in the drivers
track *changes* in policy. A large pseudo-R² drop from levels → FD is the signature of a fit driven by
cross-sectional persistence rather than dynamic responsiveness.

```{r fd-collapse}
fdcmp <- ladder %>% filter(rung %in% c("1. Deliverable","4. First-difference")) %>%
  mutate(spec = ifelse(rung=="1. Deliverable","Levels (deliverable)","First-difference"),
         facet = paste(stage, sector, sep=" — "))
ggplot(fdcmp, aes(spec, pseudoR2, fill = spec)) +
  geom_col(width = 0.6) + facet_wrap(~facet, nrow = 1) +
  geom_text(aes(label = sprintf("%.2f", pseudoR2)), vjust = -0.3, size = 3.2) +
  scale_fill_manual(values = c("Levels (deliverable)"="#4575b4","First-difference"="#d73027")) +
  labs(x = NULL, y = "Pseudo-R²", fill = NULL,
       title = "Pseudo-R²: levels vs first-difference (by stage × sector)") +
  theme_minimal(base_size = 12) + theme(axis.text.x = element_text(angle = 20, hjust = 1), legend.position = "none") +
  expand_limits(y = max(fdcmp$pseudoR2)*1.15)
```

# 3. Parsimony frontier (ε-sensitivity)

The deliverable uses the BIC parsimony tie-break at ε = 0.05 (ADR 0012). As the near-tie tolerance ε
grows, more specifications count as theory-equivalent and the lowest-BIC (most parsimonious) one wins —
tracing how much simpler the choice becomes as theory strength is traded for parsimony.

```{r frontier}
if (!is.null(frontier) && nrow(frontier)) {
  ft <- frontier %>% transmute(Stage = stage, `ε` = eps, `Selected (X-id)` = sub(" .*","",model),
                               `Worse-sector tier` = minTier, `mean ΔR²` = round(meanDeltaR2,3),
                               `summed BIC` = round(sumBIC,1))
  kable(ft, format="html") %>% kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=TRUE, font_size=12) %>%
    collapse_rows(columns = 1, valign = "top")
} else cat("_Frontier unavailable (BIC not present in the sweep RDS)._")
```

# 3b. Temporal-stability frontier

Temporal sign-stability (whether a spec's *theory-term* signs hold when the model is trained on the
early window only) does **not** drive the deliverable selection (decision 2026-06-25, ADR 0027) — it
is reported here instead. As we raise the *minimum required* worse-sector sign-stability τ from 0
(no constraint = the deliverable) toward 1.0 (fully stable), the best maximin spec at each level is
shown — collapsed to the τ-points where the **winner changes**. A deliverable that holds across the
whole frontier is temporally robust; where the winner shifts, the table shows what a
stability-first analyst would have picked instead, and at what stability/fit cost.

```{r temporal-frontier}
tf <- rb$temporalFrontier
if (!is.null(tf) && nrow(tf)) {
  tft <- tf %>% transmute(Stage = stage, `min stability τ` = round(minStability, 2),
    `Selected (X-id)` = sub(" .*", "", model), Channels = channels,
    `Worse tier` = minTier, `mean ΔR²` = round(meanDeltaR2, 3), `maxVIF` = round(maxVIF, 1),
    `summed BIC` = round(sumBIC, 1), `stability` = round(stability, 2),
    `AUC` = ifelse(is.na(auc), "", sprintf("%.3f", auc)),
    `RMSE` = ifelse(is.na(rmse), "", sprintf("%.3f", rmse)))
  kable(tft, format = "html") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 12) %>%
    collapse_rows(columns = 1, valign = "top")
} else cat("_Temporal-stability frontier unavailable (run the robustness step on a sweep with the ",
           "`trendShare`/temporal columns)._")
```

# 4. Control specification-curve

Each ladder rung is re-fit against the full enumerated set of control combinations — income-form ∈
{none, GDP-Q, GDP-Q+², logGDP, logGDP+²} × {±log Population} × {±Hydro/Nuclear} (income forms mutually
exclusive). A theory channel that survives *any* control choice is control-robust (cf. Extreme-Bounds /
specification-curve analysis).

```{r ctrl-curve, results='asis'}
if (is.null(cc) || !nrow(cc)) {
  cat("_Control specification-curve not present — run `pfm::runRobustness(group, quick = FALSE)` (or `start.R --steps=robustness`) to populate it._\n")
} else {
  delivTier <- sapply(c("Adoption","Stringency"), function(s)
    ladder$tier[ladder$rung=="1. Deliverable" & ladder$stage==s & ladder$sector=="Diffuse"])
  summ <- cc %>% group_by(stage, sector, rung) %>%
    summarise(nCombos = n(),
              dR2_min = min(deltaR2Theory, na.rm=TRUE), dR2_med = median(deltaR2Theory, na.rm=TRUE),
              dR2_max = max(deltaR2Theory, na.rm=TRUE),
              apIQ_min = min(apIQcoef, na.rm=TRUE), apIQ_max = max(apIQcoef, na.rm=TRUE),
              pctGreenBlue = round(100*mean(tier %in% c("Green","Blue")),0), .groups="drop")
  for (stg in c("Adoption","Stringency")) {
    cat("\n## ", stg, "\n\n")
    d <- summ %>% filter(stage==stg) %>%
      transmute(Rung=rung, Sector=sector, `#combos`=nCombos,
                `ΔR² min`=round(dR2_min,3), `ΔR² median`=round(dR2_med,3), `ΔR² max`=round(dR2_max,3),
                `AP×IQ coef range`=sprintf("[%.2f, %.2f]", apIQ_min, apIQ_max),
                `% Green/Blue`=pctGreenBlue)
    print(kable(d, format="html") %>% kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=TRUE, font_size=11))
    cat("\n\n")
  }
  cat("\n### Specification-curve — deliverable core, Diffuse sector\n\n")
  sc <- cc %>% filter(rung=="1. Deliverable", sector=="Diffuse") %>% arrange(stage, deltaR2Theory) %>%
    group_by(stage) %>% mutate(idx = row_number()) %>% ungroup()
  print(
    ggplot(sc, aes(idx, deltaR2Theory, color = tier)) + geom_point(size=2) + facet_wrap(~stage, scales="free") +
      scale_color_manual(values = tierColor) +
      labs(x = "Control combination (sorted)", y = "ΔR²(theory)", color = "Tier",
           title = "Theory ΔR² across all control specifications (deliverable core, Diffuse)") +
      theme_minimal(base_size = 12)
  )
}
```

# 5. AP × IQ interaction stability

The institutional-conditionality interaction (H1) is the theoretical lynchpin. Here is its coefficient
across ladder rungs — a stable sign/magnitude is evidence the interaction is not a fragile artifact.

```{r interaction-stability}
ix <- ladder %>% filter(!is.na(apIQcoef)) %>% mutate(facet = paste(stage, sector, sep=" — "))
if (nrow(ix)) {
  ggplot(ix, aes(reorder(rung, apIQcoef), apIQcoef, fill = tier)) +
    geom_col() + coord_flip() + facet_wrap(~facet, scales="free_x", nrow=1) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey40") +
    scale_fill_manual(values = tierColor) +
    labs(x = NULL, y = "AP × IQ interaction coefficient", fill = "Tier",
         title = "Interaction coefficient across robustness rungs") +
    theme_minimal(base_size = 11)
}
```

# 6. Leave-One-Region-Out (LORO)

A *spatial* leave-one-out (ADR 0013): the model is re-estimated 54 times, each dropping one
region and refitting on the remaining 53 (driver scaling re-frozen on the 53). It is **distinct
from** the temporal split-panel jackknife used for Nickell-bias correction. Two things are reported:
**(A)** how stable each theory coefficient is when any single region is removed — and whether the
Theory Tier survives; and **(B)** genuine out-of-sample prediction for the held-out region, which is
possible because the deliverable's *block* fixed effects keep the held-out region's block effect
identified by its block-mates. A held-out region whose block becomes a singleton is dropped from the
OOS metric (kept in A); OOS-stringency is defined only over held-out adopter region-years and is the
thinner, more provisional number.

```{r loro-body, results='asis'}
loco <- rb$loco
if (is.null(loco) || !length(loco)) {
  cat("_LORO artifact not present — run `pfm::runRobustness(group)` (it runs unless `loro = FALSE`)._\n")
} else {
  fmtn <- function(x, d = 3) ifelse(is.na(x), "—", formatC(round(x, d), format = "f", digits = d))

  cat("\n## 6.1 Coefficient stability across the 54 refits\n\n")
  for (key in names(loco)) {
    L <- loco[[key]]; if (is.null(L)) next
    gb <- unname(L$tier$counts["Green"] + L$tier$counts["Blue"]); tot <- sum(L$tier$counts)
    cat("\n### ", gsub("_", " — ", key), "\n\nFull-sample tier: **", L$tier$full,
        "**. Tier stays Green/Blue in **", gb, "/", tot, "** folds.\n\n", sep = "")
    d <- L$coef
    tbl <- data.frame(Term = d$term, Group = d$group, `Full est.` = round(d$full, 3),
                      `LOO mean` = round(d$looMean, 3), `LOO SD` = round(d$looSD, 3),
                      `Sign-stable` = paste0(round(100 * d$signStable), "%"),
                      `Most influential` = d$mostInfluential, `max |Δβ|` = round(d$maxAbsDelta, 3),
                      check.names = FALSE)
    print(kable(tbl, format = "html") %>%
            kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 11))
    cat("\n\n")
  }

  cat("\n## 6.2 Out-of-sample prediction on the held-out region\n\n")
  adoptKeys <- Filter(function(k) loco[[k]]$meta$stage == "adoption", names(loco))
  strKeys   <- Filter(function(k) loco[[k]]$meta$stage == "stringency", names(loco))
  if (length(adoptKeys)) {
    cat("\n**Adoption** — pooled held-out predictions, benchmarked against the in-sample fit and the naive base rate (lower Brier / higher AUC is better; calibration slope 1 = perfect, < 1 = overconfident out-of-sample).\n\n")
    da <- do.call(rbind, lapply(adoptKeys, function(k){ o <- loco[[k]]$oos
      data.frame(Spec = gsub("_"," — ",k), n = o$n, `Excl.` = o$nExcludedSingletonBlock,
                 `OOS Brier` = fmtn(o$brier), `In-sample` = fmtn(o$inSample$brier), `Base-rate` = fmtn(o$brierBase),
                 `OOS AUC` = fmtn(o$auc), `In-sample AUC` = fmtn(o$inSample$auc),
                 `OOS calib.` = fmtn(o$calibrationSlope), check.names = FALSE) }))
    print(kable(da, format="html") %>% kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=TRUE, font_size=11))
    cat("\n\n")
  }
  if (length(strKeys)) {
    cat("\n**Stringency** — pooled held-out predictions on the `log(1+ECP)` scale (RMSE, observed–predicted correlation). Defined only over held-out adopter region-years — read as provisional.\n\n")
    ds <- do.call(rbind, lapply(strKeys, function(k){ o <- loco[[k]]$oos
      data.frame(Spec = gsub("_"," — ",k), n = o$n, `OOS regions` = o$nRegions, `Excl.` = o$nExcludedSingletonBlock,
                 `OOS RMSE` = fmtn(o$rmse), `In-sample RMSE` = fmtn(o$inSample$rmse),
                 `OOS cor` = fmtn(o$corObsPred), `In-sample cor` = fmtn(o$inSample$corObsPred), check.names = FALSE) }))
    print(kable(ds, format="html") %>% kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=TRUE, font_size=11))
    cat("\n\n")
  }

  # adoption OOS vs in-sample vs base-rate Brier
  if (length(adoptKeys)) {
    bd <- do.call(rbind, lapply(adoptKeys, function(k){ o <- loco[[k]]$oos
      data.frame(Spec = gsub("_"," — ",k),
                 Brier = c(o$brier, o$inSample$brier, o$brierBase),
                 Kind = c("Out-of-sample (LORO)","In-sample","Naive base-rate")) }))
    bd$Kind <- factor(bd$Kind, levels = c("In-sample","Out-of-sample (LORO)","Naive base-rate"))
    print(
      ggplot(bd, aes(Spec, Brier, fill = Kind)) + geom_col(position = "dodge") +
        scale_fill_manual(values = c("In-sample"="#4575b4","Out-of-sample (LORO)"="#d6a000","Naive base-rate"="grey70")) +
        labs(x = NULL, y = "Brier score (lower = better)", fill = NULL,
             title = "Adoption: out-of-sample vs in-sample vs base-rate (lower is better)") +
        theme_minimal(base_size = 12) + theme(legend.position = "top")
    )
  }
}
```

# 7. Why we prefer the deliverable

```{r preference, results='asis'}
cat(
"The deliverable (levels, split Actor Power, three IQ channels) is retained as the coupling model for four reasons:\n\n",
"1. **It is the theory-identifying specification.** Adoption reaches Green tier in both sectors and the AP×IQ interaction is identified at both stages — the parsimony and single-channel rungs lose one or more channels (dropping to Blue/Yellow), i.e. they buy simplicity by discarding the very interaction the module exists to estimate.\n",
"2. **First-difference is a diagnostic, not the deliverable.** Its low pseudo-R² (§2) correctly shows the levels fit is mostly persistence, but FD also strips the cross-region governance variation that identifies the persistent institutional/actor channels — so it cannot answer *which* governance levels make pricing feasible, the question the IAM needs.\n",
"3. **Lagged-DV absorbs the signal.** Adding the lagged price lifts pseudo-R² but drives ΔR²(theory) toward zero — the lag explains policy by its own inertia, leaving nothing for the theory channels; useful as a serial-correlation check, not as a feasibility model.\n",
"4. **The choice is control-robust and BIC-disciplined.** The interaction's sign survives the control specification-curve (§4), and the BIC parsimony tie-break (ε=0.05) already pruned the deliverable toward the lowest-BIC near-tie; the parsimony frontier (§3) makes the simpler alternatives explicit for readers who weight parsimony more heavily.\n\n",
"**Caveat retained:** the Diffuse-stringency level of fit is high on a small adopter panel (N=258) and is best read as a *relative* feasibility signal. The first-difference and parsimony rungs above bound that risk transparently.\n", sep="")
```

# 8. Alternative selection: Difference-First (Dynamic Identification First)

An alternative to the default **levels-first** selection (ADR 0014): rank the `hybridFD`
specs with the same Maximin rule, keep only those that pass a **Falsification Gate** —
re-fit under `pureFD`, where Actor Power (Politics) must *persist* (significant, correct
sign) while Institutional Quality (Polity) must *vanish* (the theoretically-expected
dynamic-vs-structural split) — then re-estimate the winner **in levels**, with the region
fixed effect chosen by the Projection Sanity gate. It never overwrites the levels-first
deliverable; both configs ship.

```{r dif-compare, results='asis'}
if (is.null(dif)) {
  cat("_Difference-First artifact not present — run `pfm::runDifferenceFirst(group)` (or `start.R --steps=difference-first`)._\n")
} else {
  fmtn   <- function(x, d = 3) ifelse(is.na(x), "—", formatC(round(x, d), format = "f", digits = d))
  tierOf <- function(ap, iq, int) ifelse(ap > 0 & iq > 0 & int > 0, "Green",
                                         ifelse(ap > 0 | iq > 0 | int > 0, "Blue", "Yellow"))
  for (stg in c("Adoption", "Stringency")) {
    d <- dif[[stg]]; if (is.null(d)) next
    cat("\n## ", stg, "\n\n")
    tab <- list()
    lfName <- if (!is.null(wf)) wf$selected[[stg]] else NULL
    rowsL <- if (!is.null(wf) && !is.null(lfName))
      wf$results[wf$results$model == lfName & wf$results$stage == stg, , drop = FALSE] else NULL
    if (!is.null(rowsL) && nrow(rowsL)) for (i in seq_len(nrow(rowsL))) { r <- rowsL[i, ]
      tab[[length(tab) + 1L]] <- data.frame(Method = "Levels-first", Spec = sub(" .*", "", lfName),
        Sector = r$sector, Tier = tierOf(r$sigActorPower, r$sigInstQual, r$sigInteractions),
        `ΔR²(theory)` = fmtn(r$deltaR2Theory), `Pseudo-R²` = fmtn(r$pseudoR2),
        maxVIF = fmtn(r$maxVIF, 2), check.names = FALSE) }
    if (!is.na(d$chosen) && !is.null(d$levelsFit)) { lf <- d$levelsFit
      for (i in seq_len(nrow(lf))) { r <- lf[i, ]
        tab[[length(tab) + 1L]] <- data.frame(Method = paste0("Difference-first (FE ", d$chosenFE, ")"),
          Spec = sub(" .*", "", d$chosenBase), Sector = r$sector,
          Tier = tierOf(r$sigActorPower, r$sigInstQual, r$sigInteractions),
          `ΔR²(theory)` = fmtn(r$deltaR2Theory), `Pseudo-R²` = fmtn(r$pseudoR2),
          maxVIF = fmtn(r$maxVIF, 2), check.names = FALSE) } }
    if (length(tab)) {
      print(kable(do.call(rbind, tab), format = "html") %>%
        kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE, font_size = 11))
      cat("\n\n")
    }
    g1 <- d$levelsFirstGate
    if (!is.null(g1)) {
      dd <- g1$detail
      apS <- paste(dd$sector, ifelse(dd$apPass, "yes", "no"), sep = ": ", collapse = "; ")
      iqS <- paste(dd$sector, ifelse(dd$iqPass, "yes", "no"), sep = ": ", collapse = "; ")
      cat("> **Falsification Gate on the levels-first deliverable** (`", sub(" .*", "", d$levelsFirstName),
          "`): ", if (g1$pass) "**passes**" else "**fails**", " — ", g1$reason,
          ". Under pureFD, *AP persists* (", apS, "); *IQ vanishes* (", iqS,
          "). A fail means the deployed levels model's identification is cross-sectional, not dynamic.\n\n", sep = "")
    }
    if (is.na(d$chosen)) {
      n <- if (!is.null(d$falsification)) nrow(d$falsification) else 0
      cat("> **Difference-First found no spec that survives the Falsification Gate for ", stg,
          ".** Across the top ", n, " hybridFD candidates, none kept Actor Power significant ",
          "(correct sign) while Institutional Quality vanished under pureFD — an informative null: ",
          "the ", tolower(stg), " theory signal does not survive pure differencing.\n\n", sep = "")
    } else {
      fr <- d$falsification
      reason <- if (!is.null(fr)) fr$reason[fr$model == d$chosenBase][1] else ""
      rnk <- if (!is.null(fr)) fr$rank[fr$model == d$chosenBase][1] else NA
      cat("> **Difference-First winner:** `", d$chosenBase, "` — hybridFD maximin rank ", rnk,
          ", passed the Falsification Gate (", reason, "), re-estimated in levels with the FE ",
          "chosen by Projection Sanity (**", d$chosenFE, "**).\n\n", sep = "")
    }
  }
  cat("\n**Reading the contrast.** Levels-first maximises *cross-sectional* theory identification ",
      "(highest tier in levels); Difference-First keeps only specs whose Actor-Power signal survives ",
      "*pure differencing* while Institutional Quality washes out, then deploys that spec in levels. ",
      "The two deliverables can therefore differ in channel set and tier — difference-first typically ",
      "trades some levels-tier for a stronger dynamic-identification guarantee. Neither overwrites the ",
      "other (ADR 0014); both YAMLs ship.\n", sep = "")
}
```

# 9. Temporal split: out-of-time (hindcast) validation

The spatial leave-one-region-out test (§6) asks whether the model generalises across
*regions*; this asks whether it generalises across *time* — would a model trained on
the past have predicted the recent adoption wave? (ADR 0015.) We train on 2000–2017,
test on 2018–2023 (single chronological split), and also run a rolling-origin
expanding window (train ≤ t, predict t+1). **Persistence trap:** adoption status is
sticky, so most "adopted" labels in the test years belong to regions that adopted
*before* the cutoff; scoring those is trivial. The **headline adoption metrics are on
the at-risk set** — regions not yet adopted at the cutoff (the genuine "next-wave"
test) — with the persistence-inflated full-set number shown alongside.

```{r temporal-body, results='asis'}
if (is.null(temporal) || !length(temporal)) {
  cat("_Temporal-split artifact not present — run `pfm::runTemporalSplit(group)` (or `start.R --steps=temporal`)._\n")
} else {
  fmtn <- function(x, d = 3) ifelse(is.null(x) || is.na(x), "—", formatC(round(x, d), format = "f", digits = d))
  ak <- Filter(function(k) temporal[[k]]$meta$stage == "adoption", names(temporal))
  sk <- Filter(function(k) temporal[[k]]$meta$stage == "stringency", names(temporal))

  if (length(ak)) {
    cat("\n## 9.1 Adoption — out-of-time discrimination (at-risk set)\n\n")
    da <- do.call(rbind, lapply(ak, function(k){ t <- temporal[[k]]; ar <- t$single$atRisk; fl <- t$single$full; ins <- t$single$inSample
      data.frame(`Spec/Sector` = gsub("_"," — ",k), `at-risk n` = ar$n,
                 `OOS AUC (at-risk)` = fmtn(ar$auc), `OOS AUC (full, persist.)` = fmtn(fl$auc),
                 `In-sample AUC` = fmtn(ins$auc), `OOS Brier` = fmtn(ar$brier), `Base-rate Brier` = fmtn(ar$brierBase),
                 `Rolling AUC (1-yr pooled)` = fmtn(t$rolling$pooled$auc), check.names = FALSE) }))
    print(kable(da, format="html") %>% kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=TRUE, font_size=11))
    cat("\nDiscrimination (AUC) on the genuinely uncertain *at-risk* cases is well below the in-sample and the spatial leave-one-region-out values, and the at-risk Brier exceeds the base rate — i.e. out of time the model **ranks** the next adopters above non-adopters only moderately and is **over-confident**, so its output is a feasibility *ordering*, not a calibrated forecast. The gap between the at-risk AUC and the persistence-inflated full-set AUC shows how much of an apparent out-of-time fit is mere stickiness.\n\n")
    # forecast-horizon curve
    hz <- do.call(rbind, lapply(ak, function(k){ h <- temporal[[k]]$horizon; if (is.null(h)) return(NULL); h$spec <- gsub("_"," — ",k); h }))
    if (!is.null(hz) && nrow(hz)) {
      print(ggplot(hz, aes(horizon, auc, color = spec)) + geom_line() + geom_point() +
        geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey50") +
        scale_x_continuous(breaks = sort(unique(hz$horizon))) + ylim(0.4, 1) +
        labs(x = "Forecast horizon (years ahead of 2017)", y = "Out-of-time AUC (at-risk)", color = NULL,
             title = "Adoption: out-of-time discrimination by forecast horizon") +
        theme_minimal(base_size = 12) + theme(legend.position = "top"))
    }
  }
  if (length(sk)) {
    cat("\n## 9.2 Stringency — out-of-time price prediction\n\n")
    ds <- do.call(rbind, lapply(sk, function(k){ t <- temporal[[k]]; s <- t$single$atRisk; ins <- t$single$inSample
      data.frame(`Spec/Sector` = gsub("_"," — ",k), n = s$n,
                 `OOS RMSE` = fmtn(s$rmse), `In-sample RMSE` = fmtn(ins$rmse),
                 `OOS cor` = fmtn(s$corObsPred), `In-sample cor` = fmtn(ins$corObsPred),
                 `Rolling cor (1-yr pooled)` = fmtn(t$rolling$pooled$corObsPred), check.names = FALSE) }))
    print(kable(ds, format="html") %>% kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=TRUE, font_size=11))
    cat("\nOut-of-time price error roughly doubles vs in-sample, though rank correlation is partly retained (more so for Diffuse) — consistent with the stringency stage being a provisional, within-pioneer signal (§6.6, §6.7).\n\n")
  }
  # coefficient stability (train vs full): share of theory coefficients keeping their sign
  cat("\n## 9.3 Coefficient stability (train 2000–2017 vs full sample)\n\n")
  cs <- do.call(rbind, lapply(names(temporal), function(k){ c <- temporal[[k]]$coef; if (is.null(c)) return(NULL)
    data.frame(`Spec/Sector` = gsub("_"," — ",k), `theory coefs` = nrow(c),
               `sign-stable` = paste0(sum(c$signSame), "/", nrow(c)), check.names = FALSE) }))
  if (!is.null(cs)) print(kable(cs, format="html") %>% kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=TRUE, font_size=11))
  cat("\nSigns are largely stable when the last six years are withheld, but magnitudes shift materially on the shorter training panel — the expected small-sample sensitivity, and another reason to read the level of fit, not just the sign, with caution.\n")
}
```

---

*Generated from the Run-Group artifacts `robustness.rds` (`pfm::runRobustness`), `sweep.rds` (`pfm::runSweep`), `difference-first.rds` (`pfm::runDifferenceFirst`), and `temporal-split.rds` (`pfm::runTemporalSplit`) under `results/<group>/` (ADR 0018). Robustness Ladder & Parsimony Frontier per ADR 0012; Difference-First per ADR 0014; Temporal split per ADR 0015.*
