---
title: "PFM Model Selection — `r params$reportName`"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 1
    toc_float:
      collapsed: true
    number_sections: true
    code_folding: hide
params:
  workflowRds: "results/exhaustive/sweep.rds"
  reportName: "default"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 5.5)
library(ggplot2)
library(dplyr)

root <- getwd()
library(pfmreports)

rds_path <- params$workflowRds
if (!grepl("^([A-Za-z]:|/|\\\\)", rds_path)) rds_path <- file.path(root, rds_path)
if (!file.exists(rds_path)) stop("Workflow RDS not found: ", rds_path,
                                 " — run pfm::runChannelsWorkflow() first.")
wf <- readRDS(rds_path)

results <- wf$results
results$tier <- pfm::computeTheoryTier(results$sigActorPower, results$sigInstQual,
                                       results$sigInteractions)
coefs <- wf$coefficients
stages <- names(wf$maximin)
tierColours <- c(Green = "#c8e6c9", Blue = "#bbdefb", Yellow = "#fff9c4")

# Formatted equation from the coefficients table
formatEquation <- function(model, sector, stage, maxTerms = 14) {
  cc <- coefs %>% filter(model == !!model, sector == !!sector, stage == !!stage)
  if (nrow(cc) == 0) return("(no coefficients recorded)")
  cc <- cc %>% filter(!grepl("^regionFE", term) | abs(estimate) > 0)
  fe <- cc %>% filter(grepl("^regionFE", term))
  main <- cc %>% filter(!grepl("^regionFE", term))
  terms <- sprintf("%+.3f·%s%s", main$estimate,
                   ifelse(main$term == "(Intercept)", "1", main$term),
                   sigStars(main$pValue))
  eq <- paste(terms, collapse = " ")
  if (nrow(fe) > 0) eq <- paste0(eq, sprintf(" + [%d region FE]", nrow(fe)))
  eq
}

forestPlot <- function(model, stage, title = NULL) {
  cc <- coefs %>%
    filter(model == !!model, stage == !!stage, !grepl("^regionFE|^\\(Intercept", term)) %>%
    mutate(lo = estimate - 1.96 * stdError, hi = estimate + 1.96 * stdError,
           sig = pValue < 0.05)
  if (nrow(cc) == 0) return(invisible(NULL))
  ggplot(cc, aes(x = estimate, y = reorder(term, estimate), colour = sig)) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = "grey50") +
    geom_pointrange(aes(xmin = lo, xmax = hi), fatten = 2) +
    facet_wrap(~sector, ncol = 2) +
    scale_colour_manual(values = c(`TRUE` = "#d62728", `FALSE` = "grey60"),
                        labels = c(`TRUE` = "p < 0.05", `FALSE` = "n.s.")) +
    theme_report() +
    theme(axis.text.y = element_text(size = 7.5)) +
    labs(title = title %||% model, x = "Coefficient (robust 95% CI)", y = NULL,
         colour = NULL)
}

howForest <- function() howToRead(
  "Each dot is a coefficient with its robust 95% confidence interval, per sector ",
  "(region fixed effects and intercept omitted). Red = significant at 5%. Drivers ",
  "are **standardized**, so coefficients are per **one standard deviation** of the ",
  "driver on the log-odds (adoption) or log(1+price) (stringency) scale — magnitudes ",
  "are directly comparable across drivers (largest = most influential).")
```

# Executive Summary

```{r exec, results='asis'}
nSpec <- length(unique(results$model))
cat(sprintf(paste0(
  "- **Sweep:** `%s` mode — %d specifications × 4 sector/stages = %d fits (generated %s).\n",
  "- **Selection rule:** Maximin (worse-sector tier, then mean ΔR²(theory)), hard gates ",
  "VIF < 10 / convergence / no lagged terms, followed by the **Projection Sanity gate** ",
  "(expanding batches; severe rules: price > 2000 USD/tCO2, invalid prices, dead region ",
  "blocks. All-region adoption saturation is a *warning*, not severe — expected under an ",
  "ambitious scenario).\n"
), wf$mode, nSpec, nrow(results), format(wf$generated, "%Y-%m-%d %H:%M")))
for (stg in stages) {
  sel <- wf$selected[[stg]]
  sn <- wf$sanity[[stg]]
  cat(sprintf("- **%s — selected:** `%s`%s\n", stg,
              sel %||% "(none)",
              if (!is.null(sn) && isTRUE(sn$forced)) " ⚠ least-flagged fallback (no candidate passed sanity)"
              else if (!is.null(sn)) " (passed Projection Sanity)" else " (maximin only — no scenario data)"))
}
```

# Chosen Models

```{r chosen, results='asis', fig.height=6}
for (stg in stages) {
  sel <- wf$selected[[stg]]
  if (is.null(sel)) { cat(sprintf("## %s\n\n> No model selected.\n\n", stg)); next }
  mm <- wf$maximin[[stg]]
  row <- mm[mm$model == sel, , drop = FALSE]
  cat(sprintf("## %s — `%s`\n\n", stg, sel))
  cat(sprintf(paste0(
    "- **Worse-sector tier:** %s | **mean ΔR²(theory):** %.3f | **worse-sector ΔR²:** %.3f\n",
    "- **Per sector:** %s | %s\n",
    "- **Maximin rank:** %d of %d gate-passing specs\n\n"
  ), row$minTier, row$meanDeltaR2, row$minDeltaR2, row$tierBySector, row$deltaR2BySector,
     row$rank, sum(mm$gatePass)))
  sn <- wf$sanity[[stg]]
  if (!is.null(sn) && !is.null(sn$trace)) {
    tr <- sn$trace[sn$trace$model == sel, , drop = FALSE]
    if (nrow(tr) == 1) {
      cat(sprintf("- **Projection Sanity:** evaluated #%d (batch %d) — %s severe / %s warnings%s\n\n",
                  tr$order, tr$batch, tr$nSevere, tr$nWarning,
                  if (isTRUE(sn$forced)) " — **least-flagged fallback**" else " — **PASS**"))
    }
  }
  for (sec in sort(unique(results$sector))) {
    r <- results %>% filter(model == sel, stage == stg, sector == sec)
    cat(sprintf("**%s equation** (AIC %.1f, pseudo-R² %.3f, maxVIF %.1f, N %d):\n\n```\n%s ~ %s\n```\n\n",
                sec, r$aic, r$pseudoR2, r$maxVIF, r$nObs, tolower(stg),
                formatEquation(sel, sec, stg)))
  }
  p <- forestPlot(sel, stg, title = paste0(stg, " — ", sel))
  if (!is.null(p)) { print(p); cat("\n\n") }
  howForest()
}
```

# Selection Algorithm {.tabset}

## How selection works

```{r how-selection, results='asis'}
# Live parameters from the run where present (falls back to the documented defaults).
sel_meta  <- wf$selectionMeta %||% list()
vifGate   <- sel_meta$vifGate   %||% 10
nearEps   <- sel_meta$nearTieEps %||% 0.025
feSet     <- sel_meta$selectFE   %||% c("H12", "OECDp", "Mundlak")
priceSev  <- sel_meta$priceSevere %||% 2000
clampSev  <- sel_meta$clampSevereFrac %||% 0.25
nSpec     <- length(unique(results$model))
cat(sprintf(paste0(
"The deliverable is chosen by a fixed, pre-registered pipeline — no per-result discretion. ",
"For this run: **%d specifications** (`%s` sweep), gates and thresholds as shown.\n\n",
"1. **Stage 0 — Channel Screen.** Before any model is fit, candidate institutional channels are ",
"screened per sector/stage: pairs that are too collinear (|r| above threshold) are marked ",
"*forbidden* (may never co-enter a spec), and each slot's candidates are ranked by partial ",
"correlation with the outcome. *(see the Stage-0 tab)*\n",
"2. **The sweep (fit grid).** Every surviving spec is fit for all four sector×stage cells ",
"(Adoption/Stringency × Bulk/Diffuse). Saturating-price stringency twins are fit on the ",
"stringency stage only (ADR 0028).\n",
"3. **Hard gates (pass/fail, both sectors).** maxVIF < **%g**, model converged, no unintended ",
"lagged term, ΔR² ≤ 1. A spec failing any gate in either sector is set `gatePass = FALSE` and ",
"ranks last. *(see Gate attrition)*\n",
"4. **Maximin tier.** Specs are ranked by their **worse sector's** theory Tier ",
"(Green > Blue > Yellow) — a spec is only as good as its weaker sector.\n",
"5. **ΔR²(theory) tie-break.** Within a tier, higher mean ΔR²(theory) wins (the unique ",
"incremental theory signal).\n",
"6. **Within-band tie-breaks** (specs within ε = **%g** of the leader): drop-idle-control → ",
"VIF/temporal fragility → low trend share → FE-discounted BIC → name. Temporal sign-stability ",
"is **not** a selection key (it is a robustness exhibit, ADR 0027).\n",
"7. **Projection Sanity gate.** The maximin leader is projection-tested on the scenario; severe ",
"violations (price > **%g** USD/tCO2, invalid prices, a dead region block, > **%g%%** ",
"clamp-pinned region-years) disqualify it and the next candidate is tried (expanding batches). ",
"The first candidate that passes is the **deliverable**. *(see Projection Sanity trace & flags)*\n\n",
"Region fixed effects swept: **%s**. Everything labelled *report-only* below (AUC, Brier, RMSE, ",
"calibration) is shown for transparency and plays **no** role in selection.\n"),
  nSpec, wf$mode, vifGate, nearEps, priceSev, round(100 * clampSev),
  paste(feSet, collapse = ", ")))
```

## Stage 0 — Channel Screen {.tabset}

The **Channel Screen** is the pre-fit filter (Stage 0 of ADR 0004): for each sector×stage it
flags *forbidden* channel pairs (too collinear to co-enter a spec) and ranks each slot's
candidate channels by their partial correlation with the outcome — the data-driven definition of
the "best" accountability/state-capacity index, computed *before* any model is estimated. One
sub-tab per sector×stage below.

```{r screen, results='asis'}
# Order the four screens stage-first to match the requested Adoption_Bulk / Adoption_Diffuse /
# Stringency_Bulk / Stringency_Diffuse tabs (wf$screens is keyed <sector>_<stage>).
wantOrder <- c("Bulk_Adoption", "Diffuse_Adoption", "Bulk_Stringency", "Diffuse_Stringency")
tabLabel  <- c(Bulk_Adoption = "Adoption_Bulk", Diffuse_Adoption = "Adoption_Diffuse",
               Bulk_Stringency = "Stringency_Bulk", Diffuse_Stringency = "Stringency_Diffuse")
keys <- intersect(wantOrder, names(wf$screens))
if (length(keys) == 0) keys <- names(wf$screens)
if (length(keys) == 0) {
  cat("> No screen results in this workflow RDS.\n")
} else {
  for (nm in keys) {
    scr <- wf$screens[[nm]]
    cat(sprintf("### %s\n\n", tabLabel[[nm]] %||% nm))
    if (is.null(scr)) { cat("*No screen computed.*\n\n"); next }
    if (nrow(scr$forbiddenPairs) > 0) {
      print(knitr::kable(scr$forbiddenPairs, digits = 3,
                         caption = paste0("Forbidden pairs (|r| > ", scr$corThreshold, ")")))
      cat("\n\n")
    } else cat("*No forbidden channel pairs.*\n\n")
    if (nrow(scr$slotRankings) > 0) {
      print(knitr::kable(scr$slotRankings, digits = 3,
                         caption = "Slot rankings (partial correlation with the outcome)"))
      cat("\n\n")
    }
    howToRead(
      "Forbidden pairs may never co-enter one spec (too collinear). Slot rankings give each ",
      "candidate channel's partial correlation with the outcome controlling for the other ",
      "slots — higher = the stronger candidate for that slot in this sector/stage.", type = "table")
  }
}
```

## Maximin rankings

Every spec ranked by the full Maximin Selection Rule (worse-sector tier → mean ΔR²(theory) →
within-band tie-breaks). This is the master table the deliverable is read off; one block per
stage. `gatePass = FALSE` rows failed a hard gate in at least one sector and sort last.

```{r maximin}
tabledList(stats::setNames(lapply(stages, function(stg) wf$maximin[[stg]]), stages),
           pageLength = 15, filter = "top")
```

```{r maximin-cap, results='asis'}
howToRead(
  "Worse-sector tier first (Green > Blue > Yellow), tie-broken by mean ΔR²(theory), then the ",
  "within-band keys. `rank` is the maximin position; `gatePass = FALSE` rows failed VIF/",
  "convergence/lagged gates in a sector and rank last. Sort by meanDeltaR2 / minDeltaR2 / sumBIC ",
  "to see the alternative orderings.", type = "table")
```

## Projection Sanity trace {.tabset}

The **expanding-batch walk** down the maximin ranking: starting from the top, each candidate is
projection-tested and its severe/warning flag counts recorded, until one passes — that one is
selected. Split by stage below.

```{r sanity-trace, results='asis'}
any_trace <- FALSE
for (stg in stages) {
  sn <- wf$sanity[[stg]]
  cat(sprintf("### %s\n\n", stg))
  if (is.null(sn) || is.null(sn$trace)) {
    cat("> Sanity gate not run for this stage (no scenario data) — selection was pure maximin.\n\n")
    next
  }
  any_trace <- TRUE
  print(knitr::kable(sn$trace, caption = paste0(stg, " — expanding-batch evaluation order")))
  cat("\n\n")
  howToRead(
    "`order` is the maximin position evaluated; `nSevere`/`nWarning` are the flag counts from ",
    "the Projection Sanity rules. The first row with 0 severe flags is selected. 'Not evaluable' ",
    "rows are FD specs (no scenario projection method yet, ADR 0005) or projection failures.",
    type = "table")
}
if (!any_trace) cat("\n> Sanity gate was not run in any stage — selection was pure maximin.\n")
```

## Gate attrition

```{r gate-attrition, results='asis'}
# Funnel of how many specs survive each hard gate, per stage (sweep-derivable).
funnel <- do.call(rbind, lapply(stages, function(stg) {
  rs <- results[results$stage == stg, ]
  byModel <- split(rs, rs$model)
  passAll <- function(f) sum(vapply(byModel, function(d) all(f(d)), logical(1)))
  data.frame(
    Stage = stg,
    `Specifications` = length(byModel),
    `Converged (both sectors)` = passAll(function(d) isTRUE(all(d$converged))),
    `+ maxVIF < 10` = passAll(function(d) all(d$converged) && all(d$maxVIF < 10, na.rm = TRUE)),
    `+ ΔR² ≤ 1 (gate-passing pool)` = sum(vapply(split(wf$maximin[[stg]], wf$maximin[[stg]]$model),
                                                  function(d) isTRUE(any(d$gatePass)), logical(1))),
    `Sanity-evaluated` = if (!is.null(wf$sanity[[stg]]$trace)) nrow(wf$sanity[[stg]]$trace) else NA_integer_,
    check.names = FALSE, stringsAsFactors = FALSE)
}))
print(knitr::kable(funnel, caption = "Gate attrition — specs surviving each hard gate, per stage"))
cat("\n\n")
howToRead(
  "Each column is the count of specifications still standing after the named hard gate is ",
  "applied (cumulatively, both sectors). The drop between columns is how aggressively that gate ",
  "filters; the last gate-passing pool is what the maximin ranking and the Projection Sanity ",
  "gate then operate on.", type = "table")
```

# Alternative Selections {.tabset}

The deliverable follows one rule (the Maximin tab). These are the **same sweep ranked under
different objectives** — to show how much the choice depends on the rule, and what a
predictability-first, parsimony-first, theory-first or per-sector analyst would pick instead.
None of these change the deliverable.

## Best per sector/stage

```{r best-sector}
if (!is.null(wf$bestPerSector)) {
  interactiveTable(wf$bestPerSector, caption = "Independently best model per sector/stage")
}
```

```{r best-sector-cap, results='asis'}
howToRead(
  "The per-sector optimum ignoring the Shared Specification constraint — what you ",
  "would pick if each sector/stage could have its own formula. Comparing this with ",
  "the Chosen Models quantifies the price of cross-sector consistency.", type = "table")
```

## Best by predictability

```{r best-pred}
pred_ad <- results %>%
  filter(stage == "Adoption", converged, maxVIF < 10) %>%
  group_by(model) %>%
  summarise(meanAUC = mean(auc, na.rm = TRUE), meanBrier = mean(brier, na.rm = TRUE),
            meanCalibSlope = mean(calibrationSlope, na.rm = TRUE),
            minTier = min(tier), .groups = "drop") %>%
  arrange(desc(meanAUC)) %>% head(15)
pred_st <- results %>%
  filter(stage == "Stringency", converged, maxVIF < 10) %>%
  group_by(model, panelTransform) %>%
  summarise(meanRMSE = mean(rmse, na.rm = TRUE), .groups = "drop") %>%
  arrange(meanRMSE) %>% head(15)
tabledList(list("Adoption — top 15 by mean AUC (both sectors)" = pred_ad,
                "Stringency — top 15 by mean in-sample RMSE" = pred_st))
```

```{r best-pred-cap, results='asis'}
howToRead(
  "Predictive Diagnostics view (report-only — these play no role in selection). ",
  "Adoption: AUC = discrimination (1 = perfect), Brier = probability error (lower ",
  "better), calibration slope ≈ 1 = well calibrated. Stringency: RMSE on the ",
  "estimation scale — only comparable within the same panelTransform (levels RMSE ",
  "is in log-price, FD RMSE in Δlog-price).", type = "table")
```

## Best by parsimony

```{r best-parsimony}
# Simplest-adequate lens: among gate-passing specs, rank by FE-discounted BIC (the maximin
# tie-break), summed across sectors so it is comparable to the deliverable rule.
parsi <- lapply(stages, function(stg) {
  rs <- results %>% filter(stage == stg, converged, maxVIF < 10)
  if (!"bic" %in% names(rs) || !nrow(rs)) return(NULL)
  rs %>% group_by(model, panelTransform) %>%
    summarise(sumBIC = sum(bic, na.rm = TRUE), nFE = max(nFE, na.rm = TRUE),
              adjBIC = sum(bic, na.rm = TRUE) - sum(nFE * log(pmax(nObs, 1)), na.rm = TRUE),
              minTier = min(tier), meanDeltaR2 = mean(deltaR2Theory, na.rm = TRUE),
              .groups = "drop") %>%
    arrange(adjBIC) %>% head(15)
})
names(parsi) <- stages
tabledList(Filter(Negate(is.null), parsi))
```

```{r best-parsimony-cap, results='asis'}
howToRead(
  "Ranked by FE-discounted BIC (`adjBIC = ΣBIC − Σ nFE·log n`) — the parsimony tie-break the ",
  "maximin rule uses, so Region-FE granularity doesn't decide on raw parameter count. Lower ",
  "adjBIC = the simplest spec that still fits. Compare the top row here against the deliverable: ",
  "if they differ, the deliverable is buying theory tier / ΔR² at a parsimony cost.", type = "table")
```

## Best by theory content

```{r best-theory}
theoryRank <- lapply(stages, function(stg) {
  results %>% filter(stage == stg, converged, maxVIF < 10) %>%
    group_by(model, panelTransform) %>%
    summarise(meanDeltaR2 = mean(deltaR2Theory, na.rm = TRUE),
              minDeltaR2 = min(deltaR2Theory, na.rm = TRUE),
              meanTheoryFrac = mean(theoryFrac, na.rm = TRUE),
              minTier = min(tier), .groups = "drop") %>%
    arrange(desc(meanDeltaR2)) %>% head(15)
})
names(theoryRank) <- stages
tabledList(theoryRank)
```

```{r best-theory-cap, results='asis'}
howToRead(
  "Ranked purely on institutional-theory content — mean ΔR²(theory) (the unique incremental ",
  "signal from the theory terms), with worse-sector ΔR² and Theory Fraction shown. This ignores ",
  "the tier gating, so it surfaces specs with the most theory signal even if a group misses 5% ",
  "significance. ΔR² leads; Theory Fraction is mechanically low under strong Region FE.",
  type = "table")
```

## Stringency: log1p vs saturating

```{r price-form}
# ADR 0028 head-to-head: same channel spec, both price forms. The saturating twin is named
# "… | satP"; pair on the base name and lay the two forms side by side.
stForm <- results %>%
  filter(stage == "Stringency", panelTransform == "levels", converged) %>%
  mutate(form = ifelse(grepl("\\| satP$", model), "saturating", "log1p"),
         base = sub(" \\| satP$", "", model)) %>%
  group_by(base, form) %>%
  summarise(meanRMSE = mean(rmse, na.rm = TRUE), maxVIF = max(maxVIF, na.rm = TRUE),
            meanDeltaR2 = mean(deltaR2Theory, na.rm = TRUE), .groups = "drop")
has_sat <- any(stForm$form == "saturating")
if (has_sat) {
  cols <- c("meanRMSE", "maxVIF", "meanDeltaR2")
  l <- as.data.frame(stForm[stForm$form == "log1p", c("base", cols)])
  s <- as.data.frame(stForm[stForm$form == "saturating", c("base", cols)])
  names(l)[-1] <- paste0(cols, "_log1p")
  names(s)[-1] <- paste0(cols, "_saturating")
  wide <- merge(l, s, by = "base")
  wide <- wide[order(wide$meanRMSE_saturating), ]
  interactiveTable(wide, caption = "Levels-stringency specs — log1p vs saturating (Pmax=1000)",
                   pageLength = 20, filter = "top")
}
```

```{r price-form-cap, results='asis'}
if (has_sat) {
  howToRead(
    "Each row is one channel spec fit both ways (ADR 0028): `_log1p` = unbounded log(1+ECP), ",
    "`_saturating` = Pmax-bounded. RMSE is on different scales between forms (not directly ",
    "comparable as a fit contest) — the point is that the saturating form **cannot** produce the ",
    "explosive out-of-sample price tails the log1p form can, at little ΔR²/VIF cost. Use this to ",
    "decide whether the bounded form is worth adopting for projection.", type = "table")
} else {
  cat("> No saturating (`… | satP`) stringency specs in this sweep — re-run after wiring ADR 0028 ",
      "(`channelSpecs()` now emits the twins) to populate this comparison.\n")
}
```

## Stage focus — top 10 per stage

```{r stage-focus}
focus <- lapply(stages, function(stg) wf$maximin[[stg]] %>% filter(gatePass) %>% head(10))
names(focus) <- stages
tabledList(focus)
```

```{r stage-focus-cap, results='asis'}
howToRead("The shortlist per stage under the deliverable rule — the gate-passing specs worth a ",
          "confirmation pass. The top row is the maximin winner before the Projection Sanity gate.",
          type = "table")
```

# Top Model Forest Plots {.tabset}

```{r top-forests, results='asis', fig.height=6}
for (stg in stages) {
  cat(sprintf("## %s {.tabset}\n\n", stg))
  top5 <- head(wf$maximin[[stg]]$model[wf$maximin[[stg]]$gatePass], 5)
  bench <- grep("incumbent", unique(results$model), value = TRUE)
  show <- unique(c(top5, wf$selected[[stg]], bench))
  show <- show[show %in% unique(coefs$model)]
  for (m in show) {
    cat(sprintf("### %s\n\n", m))
    p <- forestPlot(m, stg)
    if (!is.null(p)) { print(p); cat("\n\n") }
  }
}
howForest()
```

# Projection Sanity Flags

**What the sanity flags are.** After the Maximin ranking picks a leader, that spec is *projected*
forward on the scenario and screened by the **Projection Sanity gate** — a set of rules that catch
projections no carbon-price path should produce. Each rule that trips records a **flag** with a
**severity**:

- **severe** — *disqualifying*. The spec is rejected and the next maximin candidate is tried.
  Severe rules: projected price > 2000 USD/tCO2, invalid/negative prices, a region block that never
  adopts, or > 25% of region-years pinned at the price clamp (the clamp is load-bearing).
- **warning** — *recorded but not disqualifying*. e.g. all-region adoption saturation (expected
  under an ambitious scenario), seam jumps at the history→scenario join, isolated price spikes.

The table below lists **every** flag raised while evaluating candidates (not just the selected
one), so you can see which specs were rejected and why.

```{r sanity-flags, results='asis'}
flag_rows <- list()
for (stg in stages) {
  sn <- wf$sanity[[stg]]
  if (is.null(sn) || length(sn$flags) == 0) next
  for (m in names(sn$flags)) {
    f <- sn$flags[[m]]
    f$model <- m
    f$stage <- stg
    flag_rows[[length(flag_rows) + 1]] <- f
  }
}
if (length(flag_rows) > 0) {
  all_flags <- do.call(rbind, flag_rows)
  interactiveTable(all_flags, caption = "All Projection Sanity flags across evaluated candidates",
                   pageLength = 25)
} else {
  cat("> No sanity flags recorded (gate skipped, or every evaluated candidate was clean).\n")
}
```

```{r sanity-flags-cap, results='asis'}
howToRead(
  "Every rule violation found while evaluating candidates: severity 'severe' ",
  "disqualifies a spec, 'warning' does not. value is the offending quantity ",
  "(USD/tCO2 for price rules, probability for prob rules). Sort by rule or model to ",
  "see failure patterns — recurring price-explosion flags in one region usually point ",
  "at a driver going out of its historical range there, not at the model.", type = "table")
```

# All Models

```{r all-models}
interactiveTable(
  results %>%
    select(model, sector, stage, panelTransform,
           dplyr::any_of("priceLink"), tier, deltaR2Theory, theoryFrac,
           aic, dplyr::any_of("bic"), pseudoR2, nObs, dplyr::any_of(c("nFE", "trendShare")),
           maxVIF, converged, sigActorPower, sigInstQual,
           sigInteractions, auc, brier, calibrationSlope, rmse),
  caption = "Every fit in the sweep — all measures", pageLength = 25, filter = "top"
)
```

```{r all-models-cap, results='asis'}
howToRead(
  "One row per model × sector × stage with every selection and diagnostic measure. ",
  "Use the column filters: e.g. tier = Green + stage = Stringency reproduces the ",
  "Green-tier shortlist; sort by deltaR2Theory for the theory-content ranking.",
  type = "table")
```

# Measure Definitions

```{r defs}
# Single source of truth shared across all reports (pfmreports::measureDefinitions); grouped so
# the reference card covers every statistic shown in this report AND the others in the suite.
interactiveTable(measureDefinitions(), caption = "Measure / statistic glossary (all PFM reports)",
                 pageLength = 40, filter = "top")
```

```{r defs-cap, results='asis'}
howToRead("The canonical reference card for every measure used anywhere in the PFM report suite ",
          "(filter the `Group` / `Shown in` columns to narrow it). 'Report-only' measures never ",
          "influence which model is chosen.", type = "table")
```
