---
title: "PFM Selection Stability — `r params$reportName`"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 1
    toc_float:
      collapsed: true
    number_sections: true
    code_folding: hide
params:
  bootstrapRds: "results/exhaustive/selection-bootstrap.rds"
  reportName: "default"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 4.8)
library(ggplot2)
library(dplyr)
library(pfmreports)

root <- getwd()
boot_path <- params$bootstrapRds
if (!grepl("^([A-Za-z]:|/|\\\\)", boot_path)) boot_path <- file.path(root, boot_path)
boot <- if (file.exists(boot_path)) readRDS(boot_path) else NULL

# IQ-channel token from a spec name ("X-0385 WGIge|noRoL|DiagAcc ..." -> "WGIge|noRoL|DiagAcc").
chOf <- function(m) vapply(strsplit(as.character(m), " ", fixed = TRUE),
                           function(x) if (length(x) >= 2) x[[2]] else "", character(1))
# Smallest set of channel sets covering >= cover of the resample mass (the selection confidence set).
confSet <- function(freq, cover = 0.90) {
  if (is.null(freq) || !length(freq)) return(character(0))
  f <- sort(as.numeric(freq), decreasing = TRUE); names(f) <- names(sort(freq, decreasing = TRUE))
  names(f)[seq_len(which(cumsum(f) >= cover)[1])]
}
freqDF <- function(x, key) {
  if (is.null(x) || !length(x)) return(NULL)
  data.frame(setNames(list(names(x), round(as.numeric(x), 3)), c(key, "frequency")),
             stringsAsFactors = FALSE, check.names = FALSE)
}
barFreq <- function(df, key, title, fill = "#1f77b4") {
  ggplot(utils::head(df[order(-df$frequency), ], 15),
         aes(x = frequency, y = reorder(.data[[key]], frequency))) +
    geom_col(fill = fill, alpha = 0.85) +
    geom_text(aes(label = sprintf("%.0f%%", 100 * frequency)), hjust = -0.1, size = 3) +
    scale_x_continuous(labels = scales::percent, expand = expansion(mult = c(0, 0.12))) +
    theme_report() + labs(title = title, x = "share of resamples", y = NULL)
}
stages <- if (!is.null(boot)) names(boot$stages) else character(0)
```

# Overview

*What this report is for.* The **selection-uncertainty bootstrap** (ADR 0025): a region-block
(cluster) resample of the *model selection*. It answers "**how stable is the chosen channel set to
the specification search and to which regions happen to be in the sample?**" — the single best
defence against the overfitting / specification-search critique.

```{r overview, results='asis'}
if (is.null(boot)) {
  cat("> **Selection-bootstrap artifact not found.** Run the `selection-bootstrap` step ",
      "(`run-sweep-slurm.sh --paper`, or `start.R --bootstrap=<N>`) to populate this report (ADR 0025).\n")
  knitr::knit_exit()
}
cat(sprintf(paste0(
  "- **Method:** the %d regions are resampled with replacement (region-block bootstrap), the top-**%d** ",
  "gate-passing candidate pool is refit on each resample, re-ranked by the Maximin rule, and the ",
  "**winning channel set** is recorded. **%d resamples**, detail = **%s**.\n",
  "- **Deployed deliverable (point estimate):** %s.\n"),
  54L, boot$topK %||% NA_integer_, boot$nResamples %||% NA_integer_, boot$detail %||% "?",
  paste(sprintf("%s = `%s`", names(boot$specNames), boot$specNames), collapse = "; ")))
```

> **Read this as *specification stability*, not a confidence interval on effects.** The bootstrap
> resamples *which model is selected*, not the coefficients or predictions — so it quantifies how
> robust the **channel choice** is, not the size of any effect. With only 54 region-clusters the
> resamples are correlated and small-N, so treat the frequencies as a robustness *indicator*, not a
> formal frequentist guarantee. The frequencies reflect the **maximin ranking**; the deployed spec is
> the maximin member that also passes Projection-Sanity (see the Selection report).

# Stability by stage {.tabset}

```{r by-stage, results='asis', fig.height=4.6}
for (stg in stages) {
  s <- boot$stages[[stg]]
  cat(sprintf("## %s\n\n", stg))
  csf <- s$channelSetFreq
  if (is.null(csf) || !length(csf)) { cat("> *No successful resamples for this stage.*\n\n"); next }
  deliv <- chOf(boot$specNames[[stg]])
  topSet <- names(csf)[which.max(as.numeric(csf))]; topPct <- round(100 * max(as.numeric(csf)))
  cs <- confSet(csf, 0.90)

  ### 1. Channel-set frequency
  cat("### Channel-set selection frequency\n\n")
  cat(sprintf("**%d effective resamples.** The top channel set **`%s`** wins **%d%%**, across **%d** distinct sets.\n\n",
              s$nEffective %||% NA_integer_, topSet, topPct, length(csf)))
  dfc <- freqDF(csf, "channel_set")
  print(barFreq(dfc, "channel_set", paste0(stg, " — winning channel-set frequency"))); cat("\n\n")
  print(knitr::kable(dfc[order(-dfc$frequency), ], row.names = FALSE,
                     caption = paste0(stg, " — channel-set selection frequency"))); cat("\n\n")
  howToRead("How often each institutional-channel combination wins the Maximin selection across the ",
            "region resamples. A single set dominating (high %) = a robust channel choice; mass spread ",
            "across many sets = the exact channel combination is partly sample-driven.", type = "table")

  ### 2. Per-channel inclusion
  cat("### Per-channel inclusion frequency\n\n")
  tf <- s$channelTokenFreq
  if (!is.null(tf) && length(tf)) {
    dft <- freqDF(tf, "channel")
    print(barFreq(dft, "channel", paste0(stg, " — per-channel survival"), fill = "#2ca02c")); cat("\n\n")
    print(knitr::kable(dft[order(-dft$frequency), ], row.names = FALSE,
                       caption = paste0(stg, " — share of winning specs containing each channel"))); cat("\n\n")
    howToRead("The share of bootstrap winners that include each individual channel. Channels near ",
              "100% are *robustly* part of the story; low-frequency channels are fragile and should ",
              "be reported cautiously.", type = "table")
  } else cat("*Per-channel frequency unavailable.*\n\n")

  ### 3. Exact-spec frequency
  cat("### Exact-spec frequency\n\n")
  if (!is.null(s$specFreq) && length(s$specFreq)) {
    dsf <- freqDF(s$specFreq, "spec"); dsf <- dsf[order(-dsf$frequency), ]
    dsf$deployed <- ifelse(dsf$spec == boot$specNames[[stg]], "◄ deployed", "")
    print(knitr::kable(utils::head(dsf, 15), row.names = FALSE,
                       caption = paste0(stg, " — exact winning-spec frequency (top 15)"))); cat("\n\n")
    howToRead("How often the *exact* specification (X-id) wins — always lower than the channel-set ",
              "frequency, because tied specs within a channel set trade the win. The deployed spec is ",
              "flagged; a low exact-spec % with a high channel-set % is the expected, healthy pattern.",
              type = "table")
  } else cat("*Exact-spec frequency needs `detail = \"full\"` (set `--bootstrapDetail=full`).*\n\n")

  ### 4. Selection confidence set
  cat("### Selection confidence set (90%)\n\n")
  cat(sprintf("The smallest group of channel sets covering **≥ 90%%** of resamples: **%d set(s)** — %s.\n\n",
              length(cs), paste(sprintf("`%s`", cs), collapse = ", ")))
  howToRead("The channel sets the data cannot reliably distinguish from the winner at the 90% level. ",
            "One set = an unambiguous channel choice; several = report them as a set, not a single pick.")

  ### 5. Deliverable check
  cat("### Deliverable check\n\n")
  matchModal <- identical(deliv, topSet)
  delivPct <- round(100 * (as.numeric(csf[deliv]) %||% 0))
  cat(sprintf(paste0("Deployed deliverable channel set: **`%s`**. It is %s the modal bootstrap winner, ",
              "and wins **%d%%** of resamples%s.\n\n"),
              deliv, if (matchModal) "**the same as**" else "**different from**", delivPct,
              if (!matchModal) sprintf(" (modal = `%s` at %d%%)", topSet, topPct) else ""))
  howToRead("Validation: if the deployed deliverable is the modal bootstrap winner, the point estimate ",
            "sits where the resamples concentrate (reassuring). If not, the deliverable is in a ",
            "contested region — disclose it and lean on the channel set, not the exact spec.")
}
```

# Paper summary

```{r paper-summary, results='asis'}
rows <- lapply(stages, function(stg) {
  s <- boot$stages[[stg]]; csf <- s$channelSetFreq
  if (is.null(csf) || !length(csf)) return(NULL)
  tf <- s$channelTokenFreq
  data.frame(
    Stage = stg,
    `Top channel set` = names(csf)[which.max(as.numeric(csf))],
    `Top set %` = round(100 * max(as.numeric(csf))),
    `Confidence set (#sets, 90%)` = length(confSet(csf, 0.90)),
    `Channels ≥ 80%` = if (!is.null(tf)) paste(names(tf)[as.numeric(tf) >= 0.8], collapse = ", ") else "",
    check.names = FALSE, stringsAsFactors = FALSE)
})
tab <- do.call(rbind, Filter(Negate(is.null), rows))
if (!is.null(tab) && nrow(tab)) {
  for (i in seq_len(nrow(tab))) {
    r <- tab[i, ]
    cat(sprintf(paste0("> In a 200-region-block bootstrap of the model selection, the %s-stage ",
        "deliverable's channel set (`%s`) was selected in **%d%%** of resamples; the 90%% selection ",
        "confidence set contains **%d** channel set(s)%s. This indicates the channel choice is %s to ",
        "the specification search and to sample composition.\n\n"),
        tolower(r$Stage), r$`Top channel set`, r$`Top set %`, r$`Confidence set (#sets, 90%)`,
        if (nzchar(r$`Channels ≥ 80%`)) sprintf(", with %s present in ≥80%% of winners", r$`Channels ≥ 80%`) else "",
        if (r$`Top set %` >= 50) "robust" else "moderately sensitive"))
  }
  print(knitr::kable(tab, row.names = FALSE, caption = "Selection-stability summary (paper-ready)"))
  cat("\n\n")
  howToRead("Drop-in robustness paragraph(s) + table for the paper's model-selection section. The ",
            "headline number is the channel-set selection frequency; the confidence-set size is the ",
            "honest uncertainty statement.", type = "table")
}
```

# Measure Definitions

```{r defs}
interactiveTable(measureDefinitions(groups = c("Selection uncertainty", "Selection", "Tier")),
                 caption = "Selection-stability glossary (full card in the selection report)",
                 pageLength = 20, filter = "top")
```
