---
title: "PFM Panel Data Input Inspection & Aggregation Analysis"
subtitle: "Diagnostic Visualization of Historical and Scenario Driver Datasets"
author: "Political Feasibility Module Output"
date: "`r Sys.Date()`"
params:
  cacheDir: "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/cache/default"
  gdxPath: "c:/Users/renatoro/Desktop/Projects/Elevate/code/fulldata.gdx"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: flatly
    highlight: tango
    number_sections: true
    code_folding: hide
    fig_width: 10
    fig_height: 6
---

<style>
  body {
    background-color: #f5f7fa;
    color: #333;
  }
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  h1, h2, h3 {
    color: #2C3E50;
    font-weight: 700;
  }
  .nav-tabs {
    font-weight: bold;
    margin-bottom: 20px;
    border-bottom: 2px solid #2C3E50;
  }
  .nav-tabs > li.active > a, .nav-tabs > li.active > a:focus, .nav-tabs > li.active > a:hover {
    background-color: #2C3E50;
    color: #fff;
    border: 1px solid #2C3E50;
    border-bottom-color: transparent;
  }
  .tab-content {
    padding: 25px;
    border: 1px solid #ddd;
    border-top: none;
    border-radius: 0 0 6px 6px;
    background-color: #fff;
    box-shadow: 0 4px 6px rgba(0,0,0,0.05);
  }
  .card {
    background-color: #fff;
    border-radius: 8px;
    border: 1px solid #eef2f5;
    box-shadow: 0 4px 6px rgba(0,0,0,0.02);
    margin-bottom: 25px;
  }
  .card-header {
    background-color: #f8fafc;
    border-bottom: 1px solid #eef2f5;
    padding: 12px 20px;
    font-weight: bold;
    font-size: 1.1em;
    color: #2C3E50;
  }
  .card-body {
    padding: 20px;
  }
  .metric-badge {
    display: inline-block;
    padding: 5px 10px;
    border-radius: 4px;
    font-size: 0.9em;
    font-weight: bold;
    margin-right: 5px;
  }
  .badge-agg {
    background-color: #2C7BB6;
    color: white;
  }
  .badge-cnt {
    background-color: #FDAE61;
    color: #333;
  }
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  dev = "svg"
)

library(pfm)
library(magclass)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(scales)

source("../../src/r/kableCorrelationMatrix.R")

# Theme helper
theme_custom <- function() {
  theme_minimal(base_size = 14) +
    theme(
      plot.title      = element_text(face = "bold", size = 13, color = "#2C3E50"),
      plot.subtitle   = element_text(size = 11, color = "grey40"),
      axis.title      = element_text(size = 12, face = "bold", color = "#2C3E50"),
      axis.text       = element_text(size = 10, color = "#555555"),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "#f0f2f5"),
      legend.position = "bottom",
      strip.text      = element_text(face = "bold", size = 12, color = "#2C3E50"),
      strip.background = element_rect(fill = "#f8fafc", color = "transparent")
    )
}

# Variable mapping / lists
ext_actorPowerDrivers <- c(
  "VRE share", "Electrification",
  "Coal primary energy share", "Oil/Gas primary energy share",
  "Fossil share in Industry"
)

ext_instQualityDrivers_wgi <- c(
  "Government Effectiveness (WGI)", "Control of Corruption (WGI)",
  "Voice and Accountability (WGI)", "Political Stability (WGI)",
  "Regulatory Quality (WGI)", "Rule of Law (WGI)"
)

ext_instQualityDrivers_vdem_accountability <- c(
  "Rule of Law (VDem)", "Vertical Accountability (VDem)",
  "Horizontal Accountability (VDem)", "Diagonal Accountability (VDem)"
)

ext_instQualityDrivers_vdem_statecap <- c(
  "Civil Service Professionalism (VDem)",
  "Policy Implementation (VDem)",
  "Absence of Corruption (VDem)",
  "Meritocracy Index (VDem)",
  "Rule Predictability (VDem)"
)

ext_instQualityDrivers_vdem <- c(
  ext_instQualityDrivers_vdem_accountability,
  ext_instQualityDrivers_vdem_statecap
)

# Derived V-Dem state-capacity PCA composites (primary publication IQ drivers)
ext_vdem_pca <- c(
  "State Capacity PC1 (VDem)",
  "State Capacity PC2 (VDem)"
)

ext_instQualityDrivers_all <- c(ext_instQualityDrivers_wgi, ext_instQualityDrivers_vdem, ext_vdem_pca)

ext_controlDrivers <- c(
  "Population", "GDP per Capita", "GDP per Capita (Q-centred)", "Land Area",
  "Urban Population Share", "Gini Income Inequality Coefficient",
  "Gender Inequality Index", "Energy Intensity", "Hydro Nuclear Share"
)

# Colors
col_agg <- "#2C7BB6"  # Aggregated (Region-level)
col_cnt <- "#FDAE61"  # Country-level
col_bulk <- "#2C7BB6"
col_diffuse <- "#D7191C"

# Key regions — mirrors adoption-model report for consistent comparisons
key_regions <- c("USA", "CHA", "DEU", "IND", "JPN", "BRA", "RUS", "ZAF")

# Configure madrat
if (!is.null(params$cacheDir) && nchar(params$cacheDir) > 0) {
  madrat::setConfig(
    forcecache = TRUE,
    cachefolder = params$cacheDir,
    mappingfolder = gsub("cache/default", "mappings", params$cacheDir)
  )
}
```

```{r load_data, cache=FALSE, include=FALSE}
# --- Data Loading ---
message("Loading historical panel data (raw)...")
h_agg_raw <- panelDataHistorical(aggregate = TRUE, outputRegionMappingFile = "regionmapping_54.csv", movingAverage = NULL)
h_cnt_raw <- panelDataHistorical(aggregate = FALSE, movingAverage = NULL)

message("Loading historical panel data (moving average)...")
h_agg_ma <- panelDataHistorical(aggregate = TRUE, outputRegionMappingFile = "regionmapping_54.csv", movingAverage = 5)
h_cnt_ma <- panelDataHistorical(aggregate = FALSE, movingAverage = 5)

# Load scenario data if GDX is available
has_scenario <- !is.null(params$gdxPath) && nchar(params$gdxPath) > 0 && file.exists(params$gdxPath)
s_agg <- NULL
s_cnt <- NULL

if (has_scenario) {
  message("Loading scenario panel data (aggregated)...")
  s_agg <- panelDataScenario(gdxFile = params$gdxPath, aggregate = TRUE, outputRegionMappingFile = "regionmapping_54.csv")
  
  message("Loading scenario panel data (country-level)...")
  s_cnt <- panelDataScenario(gdxFile = params$gdxPath, aggregate = FALSE)
}

tidy_magpie <- function(mag_obj, agg_label = "Aggregated") {
  if (is.null(mag_obj)) return(NULL)
  df <- as.data.frame(mag_obj)
  
  # Identify standard columns in magclass data.frame
  reg_col <- if ("Region" %in% colnames(df)) "Region" else if ("Cell" %in% colnames(df)) "Cell" else colnames(df)[1]
  yr_col  <- if ("Year" %in% colnames(df)) "Year" else colnames(df)[2]
  var_col <- if ("Data1" %in% colnames(df)) "Data1" else colnames(df)[3]
  val_col <- if ("Value" %in% colnames(df)) "Value" else colnames(df)[4]
  
  df_clean <- data.frame(
    region      = trimws(as.character(df[[reg_col]])),
    year        = as.integer(gsub("y", "", as.character(df[[yr_col]]))),
    variable    = trimws(as.character(df[[var_col]])),
    value       = as.numeric(df[[val_col]]),
    aggregation = agg_label,
    stringsAsFactors = FALSE
  )
  
  return(df_clean)
}

# Convert all
df_h_agg_raw <- tidy_magpie(h_agg_raw, "Aggregated")
df_h_cnt_raw <- tidy_magpie(h_cnt_raw, "Country-level")
df_h_agg_ma <- tidy_magpie(h_agg_ma, "Aggregated")
df_h_cnt_ma <- tidy_magpie(h_cnt_ma, "Country-level")
df_s_agg <- tidy_magpie(s_agg, "Aggregated")
df_s_cnt <- tidy_magpie(s_cnt, "Country-level")

# Combine historical and scenario for long-term visualizations
combine_datasets <- function(df_agg, df_cnt, type = "Historical") {
  if (is.null(df_agg) && is.null(df_cnt)) return(NULL)
  df <- rbind(df_agg, df_cnt)
  df$period <- type
  return(df)
}

df_h_raw <- combine_datasets(df_h_agg_raw, df_h_cnt_raw, "Historical (Raw)")
df_h_ma <- combine_datasets(df_h_agg_ma, df_h_cnt_ma, "Historical (Moving Average)")
df_s <- if (has_scenario) combine_datasets(df_s_agg, df_s_cnt, "Scenario") else NULL

df_all <- do.call(rbind, list(df_h_raw, df_h_ma, df_s))

# We use the raw dataset for all standard density/stats plots
df_h <- df_h_raw
```

# Executive Summary

This report provides a detailed inspection and diagnostic evaluation of the training and projection panel data inputs for the Political Feasibility Module (PFM). It serves as a validation check for the driver variables, comparing distributions and trends **before regional aggregation** (country level, `aggregate = FALSE`) and **after regional aggregation** (region-level mapping to 54 clusters, `aggregate = TRUE`), covering both **historical** (2000–2022) and **scenario** projections (2005–2150).

> [!IMPORTANT]
> **Why Aggregation Diagnostics Matter:**
> REMIND operates on macroeconomic regional groups (e.g. 54 regions or H12 regions) to compute global climate policy pathways, while much of the institutional and governance data (WGI, V-Dem) is collected at the national level. Aggregation weights country-level indices into regional indicators (using population, GDP, or land area weights).
> This report evaluates whether this aggregation **preserves variance, shifts boundaries, or introduces statistical artifacts** that could affect econometric estimation and subsequent policy projections.

---

# Dataset Dimensions & Metadata {.tabset}

Let's review the high-level metrics of the loaded datasets to ensure dimensions are correct and there are no missing indices.

## Historical Data
The historical datasets cover the years **`r paste(range(df_h$year), collapse = " to ")`**.

<div class="row">
<div class="col-md-6">
<div class="card">
<div class="card-header">Aggregated Region Data (<code>aggregate = TRUE</code>)</div>
<div class="card-body">
<ul>
<li><strong>Number of Regions:</strong> `r length(unique(df_h_agg_raw$region))`</li>
<li><strong>Number of Variables:</strong> `r length(unique(df_h_agg_raw$variable))`</li>
<li><strong>Total Data Points:</strong> `r nrow(df_h_agg_raw)`</li>
<li><strong>Regions list:</strong> <small>`r paste(sort(unique(df_h_agg_raw$region)), collapse = ", ")`</small></li>
</ul>
</div>
</div>
</div>
<div class="col-md-6">
<div class="card">
<div class="card-header">Country-Level Data (<code>aggregate = FALSE</code>)</div>
<div class="card-body">
<ul>
<li><strong>Number of Countries:</strong> `r length(unique(df_h_cnt_raw$region))`</li>
<li><strong>Number of Variables:</strong> `r length(unique(df_h_cnt_raw$variable))`</li>
<li><strong>Total Data Points:</strong> `r nrow(df_h_cnt_raw)`</li>
<li><strong>Data density:</strong> `r percent(1 - (sum(is.na(df_h_cnt_raw$value)) / nrow(df_h_cnt_raw)))` non-NAs</li>
</ul>
</div>
</div>
</div>
</div>

## Scenario Data
```{r scenario_meta, echo=FALSE, results='asis'}
if (has_scenario) {
  cat(paste0("
The scenario dataset covers future projections from **", paste(range(df_s$year), collapse = " to "), "** using the REMIND output GDX file.

<div class='row'>
<div class='col-md-6'>
<div class='card'>
<div class='card-header'>Aggregated Scenario Data (<code>aggregate = TRUE</code>)</div>
<div class='card-body'>
<ul>
<li><strong>Number of Regions:</strong> ", length(unique(df_s_agg$region)), "</li>
<li><strong>Number of Variables:</strong> ", length(unique(df_s_agg$variable)), "</li>
<li><strong>Total Data Points:</strong> ", nrow(df_s_agg), "</li>
<li><strong>Regions list:</strong> <small>", paste(sort(unique(df_s_agg$region)), collapse = ", "), "</small></li>
</ul>
</div>
</div>
</div>
<div class='col-md-6'>
<div class='card'>
<div class='card-header'>Country-Level Scenario Data (<code>aggregate = FALSE</code>)</div>
<div class='card-body'>
<ul>
<li><strong>Number of Countries:</strong> ", length(unique(df_s_cnt$region)), "</li>
<li><strong>Number of Variables:</strong> ", length(unique(df_s_cnt$variable)), "</li>
<li><strong>Total Data Points:</strong> ", nrow(df_s_cnt), "</li>
<li><strong>Data density:</strong> ", percent(1 - (sum(is.na(df_s_cnt$value)) / nrow(df_s_cnt))), " non-NAs</li>
</ul>
</div>
</div>
</div>
</div>
  "))
} else {
  cat("> [!WARNING]\n> Scenario GDX file was not loaded or is missing. Projections analysis is unavailable.")
}
```

---

# Institutional Quality Drivers {.tabset}

Institutional Quality (IQ) is represented by six **Worldwide Governance Indicators (WGI)** and nine **V-Dem indicators** (four accountability indexes and five state-capacity indicators). All indicators are normalized to a `[0, 1]` range, where higher values represent stronger institutional capacity or higher democratic standards.

## WGI Indicators

> **Note (2026-06-10):** WGI indicators are no longer used in the primary PFM specification. The publication model uses **V-Dem State Capacity PC1** (see section below). WGI distributions are retained here for comparison and historical documentation.

The six WGI dimensions measure governance quality from the World Bank dataset. **Country-level (orange)** shows raw national scores; **Aggregated (blue)** shows population-weighted regional clusters.

```{r wgi_distribution, echo=FALSE, fig.width=12, fig.height=6}
df_wgi_hist <- df_h %>% filter(variable %in% ext_instQualityDrivers_wgi)
df_wgi_hist$variable_short <- gsub(" \\(WGI\\)", "", df_wgi_hist$variable)

ggplot(df_wgi_hist, aes(x = value, fill = aggregation)) +
  geom_density(alpha = 0.5, color = "transparent") +
  facet_wrap(~variable_short, scales = "free_y", ncol = 3) +
  scale_fill_manual(values = c("Aggregated" = col_agg, "Country-level" = col_cnt)) +
  labs(
    title = "WGI Distributions: Pre- vs. Post-Aggregation (2000–2022)",
    x = "Indicator Value [0, 1]", y = "Density", fill = "Aggregation Level"
  ) +
  theme_custom()
```

```{r wgi_stats_table, echo=FALSE}
wgi_stats <- df_wgi_hist %>%
  group_by(variable, aggregation) %>%
  summarize(N = sum(!is.na(value)), Mean = mean(value, na.rm = TRUE),
            Median = median(value, na.rm = TRUE), SD = sd(value, na.rm = TRUE),
            Min = min(value, na.rm = TRUE), Max = max(value, na.rm = TRUE),
            .groups = "drop") %>%
  arrange(variable, desc(aggregation))

kable(wgi_stats, digits = 3, caption = "WGI Statistics: Country vs. Region Aggregation") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(2, bold = TRUE, color = ifelse(wgi_stats$aggregation == "Aggregated", col_agg, "#d35400"))
```

## V-Dem Indicators

V-Dem provides two groups of indicators. **Accountability indexes** (four) measure checks-and-balances and democratic accountability. **State-capacity indicators** (five) measure the quality of government administration — civil service professionalism, rule predictability, and absence of corruption/neopatrimonialism. Note: Policy Implementation, Absence of Corruption, and Meritocracy Index are inverted (higher = better) before normalization.

```{r vdem_distribution, echo=FALSE, fig.width=14, fig.height=8}
df_vdem_hist <- df_h %>% filter(variable %in% ext_instQualityDrivers_vdem)
df_vdem_hist$variable_short <- gsub(" \\(VDem\\)", "", df_vdem_hist$variable)
df_vdem_hist$subgroup <- ifelse(
  df_vdem_hist$variable %in% ext_instQualityDrivers_vdem_accountability,
  "Accountability", "State Capacity"
)

ggplot(df_vdem_hist, aes(x = value, fill = aggregation)) +
  geom_density(alpha = 0.5, color = "transparent") +
  facet_wrap(~variable_short, scales = "free_y", ncol = 3) +
  scale_fill_manual(values = c("Aggregated" = col_agg, "Country-level" = col_cnt)) +
  labs(
    title = "V-Dem Indicator Distributions: Pre- vs. Post-Aggregation (2000–2022)",
    subtitle = "Four accountability indexes + five state-capacity indicators",
    x = "Indicator Value [0, 1]", y = "Density", fill = "Aggregation Level"
  ) +
  theme_custom()
```

```{r vdem_stats_table, echo=FALSE}
vdem_stats <- df_vdem_hist %>%
  group_by(variable, aggregation) %>%
  summarize(N = sum(!is.na(value)), Mean = mean(value, na.rm = TRUE),
            Median = median(value, na.rm = TRUE), SD = sd(value, na.rm = TRUE),
            Min = min(value, na.rm = TRUE), Max = max(value, na.rm = TRUE),
            .groups = "drop") %>%
  arrange(variable, desc(aggregation))

kable(vdem_stats, digits = 3, caption = "V-Dem Statistics: Country vs. Region Aggregation") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(2, bold = TRUE, color = ifelse(vdem_stats$aggregation == "Aggregated", col_agg, "#d35400"))
```

## State Capacity PC1 & PC2 (Primary Publication IQ Drivers)

**State Capacity PC1 and PC2** are orthogonal principal components derived from the five raw V-Dem state-capacity indicators (Civil Service Professionalism, Policy Implementation, Rule Predictability, Absence of Corruption, Meritocracy Index). PC1 captures the dominant shared "state quality" axis (>65% of within-block variance); PC2 captures the residual within-block contrast (rule predictability vs. anti-corruption/meritocracy). Both are normalised to [0, 1].

These are the **primary IQ drivers** in the publication model (V-Dem sweep, 2026-06-10). Using PCA resolves the within-block collinearity (VIF > 40 when all five indicators are entered jointly) while preserving the majority of state-capacity information. The AP × PC1 interaction achieves Green tier for Stringency: Diffuse.

```{r vdem_pca_distribution, echo=FALSE, fig.width=11, fig.height=4}
df_pca_hist <- df_h %>%
  filter(variable %in% ext_vdem_pca) %>%
  filter(!is.na(value))

if (nrow(df_pca_hist) > 0) {
  ggplot(df_pca_hist, aes(x = value, fill = aggregation)) +
    geom_density(alpha = 0.5, color = "transparent") +
    facet_wrap(~variable, scales = "free_y", ncol = 2) +
    scale_fill_manual(values = c("Aggregated" = col_agg, "Country-level" = col_cnt)) +
    labs(
      title = "V-Dem State Capacity PC1 & PC2: Pre- vs. Post-Aggregation (2000–2022)",
      subtitle = "PC1 = dominant state-quality axis; PC2 = within-block contrast. Primary publication IQ drivers.",
      x = "PC Value [0, 1]", y = "Density", fill = "Aggregation Level"
    ) +
    theme_custom()
} else {
  cat("> State Capacity PC variables not found in panel data.\n")
}
```

```{r vdem_pca_timeseries, echo=FALSE, fig.width=11, fig.height=4}
df_pca_ts <- df_h %>%
  filter(variable %in% ext_vdem_pca, aggregation == "Aggregated",
         region %in% key_regions, !is.na(value))

if (nrow(df_pca_ts) > 0) {
  ggplot(df_pca_ts, aes(x = year, y = value, color = region)) +
    geom_line(linewidth = 1) +
    facet_wrap(~variable, ncol = 2) +
    labs(
      title = "State Capacity PC1 & PC2: Regional Trends (2000–2022)",
      subtitle = "Key REMIND regions — population-weighted aggregation",
      x = "Year", y = "PC Value [0, 1]", color = "Region"
    ) +
    theme_custom()
}
```

```{r vdem_pca_stats, echo=FALSE}
if (nrow(df_pca_hist) > 0) {
  pca_stats <- df_pca_hist %>%
    group_by(variable, aggregation) %>%
    summarize(N = sum(!is.na(value)), Mean = mean(value, na.rm = TRUE),
              Median = median(value, na.rm = TRUE), SD = sd(value, na.rm = TRUE),
              Min = min(value, na.rm = TRUE), Max = max(value, na.rm = TRUE),
              .groups = "drop") %>%
    arrange(variable, desc(aggregation))
  kable(pca_stats, digits = 3,
        caption = "V-Dem State Capacity PC1 & PC2: Country vs. Region Aggregation") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = TRUE) %>%
    column_spec(2, bold = TRUE,
                color = ifelse(pca_stats$aggregation == "Aggregated", col_agg, "#d35400"))
}
```

## Scenario Projections

WGI projections extend forward using SSP2 storylines (Government Effectiveness, Rule of Law, Control of Corruption) or hold 2022 values constant. V-Dem state-capacity indicators and their PC1/PC2 composites are currently held constant at their last observed value in the scenario period.

```{r iq_scenario_plot, echo=FALSE, fig.width=14, fig.height=10}
if (has_scenario) {
  df_iq_all <- df_all %>%
    filter(variable %in% c(
      "Government Effectiveness (WGI)", "Rule of Law (WGI)",
      ext_instQualityDrivers_vdem,
      ext_vdem_pca
    )) %>%
    filter(aggregation == "Aggregated")
  df_iq_select <- df_iq_all %>% filter(region %in% key_regions)

  ggplot(df_iq_select, aes(x = year, y = value, color = region, linetype = period)) +
    geom_line(linewidth = 1.2) +
    facet_wrap(~variable, ncol = 3) +
    labs(
      title = "Institutional Quality Long-Term Scenarios (2000–2100)",
      subtitle = "Historical (2000–2022) stitched with SSP2 projections (2025–2100), aggregated regions",
      x = "Year", y = "Value [0–1]", color = "Region/Cluster", linetype = "Period"
    ) +
    theme_custom()
} else {
  cat("> [!NOTE] Scenario data is missing, long-term scenario visualizations are disabled.")
}
```

---

# Actor Power Drivers & Indexes {.tabset}

Actor Power Dynamics represent the competing lobbying forces shaping policy adoption.
* **Incumbent Power**: Proxied by fossil fuel variables (Coal, Oil/Gas, and Industry fossil shares) and represented as an index.
* **Innovator Power**: Proxied by clean variables (VRE share and overall electrification) and represented as an index.
* **Actor Power Index (API)**: The net feasibility index (a composite score).

## Distribution Analysis (Before vs. After Aggregation)

Regional aggregation affects the distribution of clean and fossil energy shares significantly. Because energy shares are aggregated using total energy consumption weights, very small countries with extreme energy profiles (e.g. 100% clean or 100% fossil) are absorbed, preserving the weighted structural footprint.

```{r ap_distributions, echo=FALSE, fig.width=12, fig.height=8}
df_ap_hist <- df_h %>%
  filter(variable %in% c(ext_actorPowerDrivers, 
                         "Actor Power Index|Bulk", "Actor Power Index|Diffuse",
                         "Innovator Power|Bulk", "Incumbent Power|Bulk"))

ggplot(df_ap_hist, aes(x = value, fill = aggregation)) +
  geom_density(alpha = 0.5, color = "transparent") +
  facet_wrap(~variable, scales = "free_y", ncol = 3) +
  scale_fill_manual(values = c("Aggregated" = col_agg, "Country-level" = col_cnt)) +
  labs(
    title = "Actor Power Drivers & Indexes: Distributional Shift",
    subtitle = "Comparing Raw Countries against Population/Consumption-Weighted 54 Regional Clusters (2000-2022)",
    x = "Share / Index Value",
    y = "Density",
    fill = "Aggregation Level"
  ) +
  theme_custom()
```

### Statistical Preservation Statistics

Below is the summary table of actor power drivers. Note that the **Mean** value is beautifully preserved by the consumption-weighted aggregation, demonstrating that the structural characteristics of the global energy mix are correctly mapped.

```{r ap_stats_table, echo=FALSE}
ap_stats <- df_ap_hist %>%
  group_by(variable, aggregation) %>%
  summarize(
    Mean = mean(value, na.rm = TRUE),
    Median = median(value, na.rm = TRUE),
    SD = sd(value, na.rm = TRUE),
    Min = min(value, na.rm = TRUE),
    Max = max(value, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(variable, desc(aggregation))

kable(ap_stats, digits = 3, caption = "Actor Power Metrics Preservation Statistics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(2, bold = TRUE, color = ifelse(ap_stats$aggregation == "Aggregated", col_agg, "#d35400"))
```

## Scenario Projections (REMIND Downscaled Drivers)

In the scenario period, Actor Power Drivers (VRE share, electrification, fossil shares) are extracted directly from the REMIND `fulldata.gdx` run. The clean vs fossil trends outline the transition speed of the scenario.

```{r ap_scenario_plot, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario) {
  df_ap_all <- df_all %>%
    filter(variable %in% c("VRE share", "Electrification", "Actor Power Index|Bulk", "Actor Power Index|Diffuse", "Incumbent Power|Bulk", "Innovator Power|Bulk", "Innovator Power|Diffuse")) %>%
    filter(aggregation == "Aggregated")
  
  df_ap_select <- df_ap_all %>% filter(region %in% key_regions)
  
  ggplot(df_ap_select, aes(x = year, y = value, color = region, linetype = period)) +
    geom_line(linewidth = 1.2) +
    facet_wrap(~variable, ncol = 2, scales = "free_y") +
    labs(
      title = "Actor Power Drivers Scenario Transition Pathways (2000-2100)",
      subtitle = "Historical data (2000-2022) stitched with REMIND future downscaled projections (2025-2100)",
      x = "Year",
      y = "Variable Value",
      color = "Region/Cluster",
      linetype = "Period"
    ) +
    theme_custom()
} else {
  cat("> [!NOTE] Scenario projection data is not available.")
}
```

---

# Control Variables {.tabset}

Control variables capture the broader economic, demographic, and spatial characteristics of each jurisdiction. This section compares structural profiles before and after aggregation.

**GDP per Capita (Q-centred)** is a key specification innovation: log-GDP is de-meaned within global income quartiles to remove the cross-quartile governance–income collinearity (r ≈ 0.75). This allows V-Dem state capacity indicators to independently explain the between-region governance variation that was previously absorbed by the income control in standard log-GDP specifications.

## Distribution Analysis (Before vs. After Aggregation)

Unlike bounded share parameters, control variables like **Population** and **GDP** are extensive and additive.
* At the **country level (orange)**, these distributions are highly skewed (many small countries, very few mega-states like China/India).
* At the **aggregated region level (blue)**, the distribution shifts as small nations are aggregated, resulting in balanced region units.
* Note: Additive variables are plotted on standard scales, while normalized controls (Urban Population Share, Energy Intensity, Gini) are compared directly.

```{r ctrl_distributions, echo=FALSE, fig.width=12, fig.height=8}
df_ctrl_hist <- df_h %>%
  filter(variable %in% c("GDP per Capita", "GDP per Capita (Q-centred)", "Energy Intensity",
                         "Urban Population Share", "Gini Income Inequality Coefficient",
                         "Gender Inequality Index"))

ggplot(df_ctrl_hist, aes(x = value, fill = aggregation)) +
  geom_density(alpha = 0.5, color = "transparent") +
  facet_wrap(~variable, scales = "free", ncol = 3) +
  scale_fill_manual(values = c("Aggregated" = col_agg, "Country-level" = col_cnt)) +
  labs(
    title = "Socioeconomic Control Variables Pre- vs. Post-Aggregation",
    subtitle = "Comparing distribution shapes of macroeconomic controls (2000-2022)",
    x = "Control Value (Normalized)",
    y = "Density",
    fill = "Aggregation Level"
  ) +
  theme_custom()
```

### Statistics Summary Table

Below is the summary stats table for the control variables.

```{r ctrl_stats_table, echo=FALSE}
ctrl_stats <- df_ctrl_hist %>%
  group_by(variable, aggregation) %>%
  summarize(
    Mean = mean(value, na.rm = TRUE),
    Median = median(value, na.rm = TRUE),
    SD = sd(value, na.rm = TRUE),
    Min = min(value, na.rm = TRUE),
    Max = max(value, na.rm = TRUE)
  ) %>%
  ungroup() %>%
  arrange(variable, desc(aggregation))

kable(ctrl_stats, digits = 3, caption = "Socioeconomic Controls Distribution Statistics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE) %>%
  column_spec(2, bold = TRUE, color = ifelse(ctrl_stats$aggregation == "Aggregated", col_agg, "#d35400"))
```

## Scenario Projections (Macroeconomic Controls)

The macroeconomic controls are derived from SSP2 storylines during the scenario projection period. Below we inspect the long-term trends of population, GDP per capita, and inequality metrics.

```{r ctrl_scenario_plot, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario) {
  df_ctrl_all <- df_all %>%
    filter(variable %in% c("GDP per Capita", "GDP per Capita (Q-centred)",
                           "Population", "Urban Population Share")) %>%
    filter(aggregation == "Aggregated")
  
  df_ctrl_select <- df_ctrl_all %>% filter(region %in% key_regions)
  
  ggplot(df_ctrl_select, aes(x = year, y = value, color = region, linetype = period)) +
    geom_line(linewidth = 1.2) +
    facet_wrap(~variable, ncol = 2, scales = "free_y") +
    labs(
      title = "Macroeconomic Scenario Projections (2000-2100)",
      subtitle = "Macro-storylines showing the long-term trajectories under SSP2 (Aggregated)",
      x = "Year",
      y = "Normalized Driver Value",
      color = "Region/Cluster",
      linetype = "Period"
    ) +
    theme_custom()
} else {
  cat("> [!NOTE] Scenario projection data is not available.")
}
```

---

# Correlation Analysis {.tabset}

Pairwise Pearson correlations computed on aggregated region-year observations (2000–2022). High within-block correlations flag potential multicollinearity; high cross-block correlations reveal confounding paths. All matrices use `kableCorrelationMatrix` with automatic color thresholding (60th percentile of off-diagonal absolute values).

```{r cor_data_prep, include=FALSE}
cor_data <- h_agg_raw
```

## Actor Power Drivers

Correlations among the five energy-share drivers that feed into Incumbent and Innovator Power construction. Fossil shares tend to be mutually positive; clean shares (VRE, Electrification) tend to be positive with each other and negative with fossil shares.

```{r cor_ap_drivers, echo=FALSE}
cor_df_ap <- computeCorrelationMatrix(cor_data, ext_actorPowerDrivers)
kableCorrelationMatrix(cor_df_ap,
  caption = "Actor Power Drivers — Pearson Correlations (region-years, 2000–2022)")
```

## Incumbent vs Innovator Power

Correlations among the four sector-specific power indexes (Bulk and Diffuse). By construction, Incumbent Power and Innovator Power should be negatively correlated; near-zero or positive correlation would suggest the index calculation warrants review.

```{r cor_power_indices, echo=FALSE}
power_index_vars <- c(
  "Incumbent Power|Bulk",  "Innovator Power|Bulk",
  "Incumbent Power|Diffuse", "Innovator Power|Diffuse"
)
cor_df_power <- computeCorrelationMatrix(cor_data, power_index_vars)
kableCorrelationMatrix(cor_df_power,
  caption = "Incumbent vs Innovator Power Indexes — Pearson Correlations")
```

## WGI Institutional Quality

Correlations among the six WGI governance dimensions. High inter-WGI correlations (common in practice) are the primary motivation for using composite indices or selecting a single representative indicator rather than entering all six simultaneously.

```{r cor_wgi, echo=FALSE}
cor_df_wgi <- computeCorrelationMatrix(cor_data, ext_instQualityDrivers_wgi)
kableCorrelationMatrix(cor_df_wgi,
  caption = "WGI Drivers — Pearson Correlations (region-years, 2000–2022)")
```

## GDP vs. Governance Collinearity

Standard log-GDP per capita is strongly correlated with Government Effectiveness (WGI) across regions (r ≈ 0.84), creating an income–governance collinearity that confounds identification of the institutional quality channel in OLS regressions. **GDP Q-centring** (de-meaning within global income quartiles) reduces this collinearity to r ≈ 0.26 by removing the cross-quartile income–governance gradient while retaining within-quartile variation. Total GDP (extensive, population-scale) is near-orthogonal to governance (r ≈ 0.09).

```{r cor_gdp_goveff_prep, include=FALSE}
gov_eff_var   <- "Government Effectiveness (WGI)"
gdp_vars_plot <- c("GDP per Capita", "GDP per Capita (Q-centred)")
gdp_vars_all  <- c("GDP per Capita", "GDP per Capita (Q-centred)", "GDP")

# Wide data frame from cor_data (magpie → long → wide)
mag_to_wide <- function(mag, vars) {
  avail <- intersect(vars, getNames(mag))
  if (length(avail) == 0) return(NULL)
  raw <- as.data.frame(mag[, , avail])
  reg_col <- if ("Region" %in% colnames(raw)) "Region" else colnames(raw)[1]
  yr_col  <- if ("Year"   %in% colnames(raw)) "Year"   else colnames(raw)[2]
  var_col <- if ("Data1"  %in% colnames(raw)) "Data1"  else colnames(raw)[3]
  val_col <- if ("Value"  %in% colnames(raw)) "Value"  else colnames(raw)[4]
  tidy <- data.frame(
    region   = trimws(as.character(raw[[reg_col]])),
    year     = as.integer(gsub("y", "", as.character(raw[[yr_col]]))),
    variable = trimws(as.character(raw[[var_col]])),
    value    = as.numeric(raw[[val_col]]),
    stringsAsFactors = FALSE
  )
  tidyr::pivot_wider(tidy, names_from = variable, values_from = value,
                     id_cols = c(region, year))
}

cor_wide_gdp <- mag_to_wide(cor_data, c(gov_eff_var, gdp_vars_all))
```

```{r cor_gdp_goveff_table, echo=FALSE}
if (!is.null(cor_wide_gdp) && gov_eff_var %in% names(cor_wide_gdp)) {
  avail_gdp <- intersect(gdp_vars_all, names(cor_wide_gdp))
  rows <- lapply(avail_gdp, function(v) {
    idx <- complete.cases(cor_wide_gdp[[gov_eff_var]], cor_wide_gdp[[v]])
    data.frame(
      `GDP Measure`  = v,
      n              = sum(idx),
      `Pearson r`    = round(cor(cor_wide_gdp[[gov_eff_var]][idx],
                                 cor_wide_gdp[[v]][idx], method = "pearson"),  3),
      `Spearman rho` = round(cor(cor_wide_gdp[[gov_eff_var]][idx],
                                 cor_wide_gdp[[v]][idx], method = "spearman"), 3),
      check.names = FALSE, stringsAsFactors = FALSE
    )
  })
  result_df <- do.call(rbind, rows)

  knitr::kable(result_df, align = c("l", "r", "r", "r"),
    caption = "Pairwise correlations with Government Effectiveness (WGI) — region-years 2000–2022") |>
    kableExtra::kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                              full_width = FALSE) |>
    kableExtra::row_spec(which(result_df[["GDP Measure"]] == "GDP per Capita"),
                         background = "#fff3cd") |>
    kableExtra::row_spec(which(result_df[["GDP Measure"]] == "GDP per Capita (Q-centred)"),
                         background = "#d4edda")
} else {
  cat("> Government Effectiveness or GDP variables not available in panel data.\n")
}
```

```{r cor_gdp_goveff_scatter, echo=FALSE, fig.width=11, fig.height=4.5}
if (!is.null(cor_wide_gdp) && gov_eff_var %in% names(cor_wide_gdp)) {
  avail_plot <- intersect(gdp_vars_plot, names(cor_wide_gdp))
  if (length(avail_plot) > 0) {
    plot_long <- tidyr::pivot_longer(
      cor_wide_gdp[, c("region", "year", gov_eff_var, avail_plot)],
      cols = all_of(avail_plot), names_to = "gdp_var", values_to = "gdp_val"
    ) |>
      dplyr::rename(gov_eff = dplyr::all_of(gov_eff_var)) |>
      dplyr::filter(!is.na(gov_eff), !is.na(gdp_val)) |>
      dplyr::mutate(
        gdp_var = factor(gdp_var,
          levels = c("GDP per Capita", "GDP per Capita (Q-centred)"),
          labels = c(
            sprintf("GDP per Capita  (Pearson r = %.3f)", {
              idx <- complete.cases(cor_wide_gdp[[gov_eff_var]],
                                    cor_wide_gdp[["GDP per Capita"]])
              cor(cor_wide_gdp[[gov_eff_var]][idx],
                  cor_wide_gdp[["GDP per Capita"]][idx])
            }),
            sprintf("GDP per Capita (Q-centred)  (Pearson r = %.3f)", {
              idx <- complete.cases(cor_wide_gdp[[gov_eff_var]],
                                    cor_wide_gdp[["GDP per Capita (Q-centred)"]])
              cor(cor_wide_gdp[[gov_eff_var]][idx],
                  cor_wide_gdp[["GDP per Capita (Q-centred)"]][idx])
            })
          )
        )
      )

    ggplot(plot_long, aes(x = gdp_val, y = gov_eff)) +
      geom_point(alpha = 0.25, size = 0.9, colour = "#4a7fb5") +
      geom_smooth(method = "lm", se = TRUE, colour = "#c0392b", linewidth = 0.9,
                  fill = "#f9b9b7", alpha = 0.3) +
      facet_wrap(~ gdp_var, scales = "free_x", ncol = 2) +
      labs(
        x = "GDP measure (log scale where applicable)",
        y = "Government Effectiveness (WGI)",
        title = "Income–Governance Collinearity: Standard vs. Q-centred GDP",
        subtitle = "Each point = one region-year (2000–2022). Red line = OLS fit."
      ) +
      theme_custom() +
      theme(strip.text = element_text(size = 10, face = "bold"))
  }
}
```

## V-Dem Institutional Quality

Correlations among the nine V-Dem indicators (four accountability + five state-capacity). The state-capacity block (Civil Service Professionalism, Policy Implementation, Absence of Corruption, Meritocracy Index, Rule Predictability) is expected to form a tightly correlated cluster — the motivation for computing PC1/PC2 composites.

```{r cor_vdem, echo=FALSE}
cor_df_vdem <- computeCorrelationMatrix(cor_data, ext_instQualityDrivers_vdem)
kableCorrelationMatrix(cor_df_vdem,
  caption = "V-Dem Indicators — Pearson Correlations (region-years, 2000–2022)")
```

## State Capacity PC1/PC2 vs. Raw Indicators

Correlations between the PCA composites (PC1, PC2) and the five raw state-capacity indicators. PC1 should load positively on all five; PC2 should show a contrast between rule-predictability and the anti-corruption/meritocracy indicators. Cross-correlation between PC1 and PC2 should be near zero (orthogonal by construction).

```{r cor_pca, echo=FALSE}
pca_corr_vars <- c(ext_vdem_pca, ext_instQualityDrivers_vdem_statecap)
pca_corr_vars_avail <- intersect(pca_corr_vars, getNames(cor_data))
if (length(pca_corr_vars_avail) >= 2) {
  cor_df_pca <- computeCorrelationMatrix(cor_data, pca_corr_vars_avail)
  kableCorrelationMatrix(cor_df_pca,
    caption = "State Capacity PC1/PC2 vs. Raw V-Dem Indicators (region-years, 2000–2022)")
} else {
  cat("> PC1/PC2 variables not available in panel data for correlation.\n")
}
```

## Control Variables

Correlations among the socioeconomic controls. GDP per Capita, Energy Intensity, and Urban Population Share are commonly co-linear; the pattern here informs whether controls should be entered jointly or selectively.

```{r cor_controls, echo=FALSE}
cor_df_ctrl <- computeCorrelationMatrix(cor_data, ext_controlDrivers)
kableCorrelationMatrix(cor_df_ctrl,
  caption = "Control Variables — Pearson Correlations (region-years, 2000–2022)")
```

## Full Driver Space

Combined correlation matrix across all driver groups from the tabs above: actor power drivers, power indexes, WGI, V-Dem, and controls. Use this to identify cross-block confounding paths — e.g. whether GDP per Capita loads on governance clusters, or whether fossil shares correlate with institutional quality.

```{r cor_full, echo=FALSE}
all_cor_vars <- c(
  ext_actorPowerDrivers,
  "Incumbent Power|Bulk", "Innovator Power|Bulk",
  ext_instQualityDrivers_wgi,
  ext_instQualityDrivers_vdem,
  ext_vdem_pca,
  ext_controlDrivers
)
cor_df_all <- computeCorrelationMatrix(cor_data, all_cor_vars)
kableCorrelationMatrix(cor_df_all, font_size = 9,
  caption = "Full Driver Space — Pearson Correlations (region-years, 2000–2022)")
```

---

# Historical-to-Scenario Stitching Plot (Continuity Check)

A critical validation requirement in integrated modeling is checking that the historical data smoothly transitions into the scenario projections. If there are sudden jumps or discontinuities at the border year (2022/2025), the regression models might generate prediction spikes or numerical instabilities.

The chart below shows a continuous, region-specific time series for key variables across the historical (solid) and scenario (dashed) periods.

```{r stitching_plot, echo=FALSE, fig.width=12, fig.height=8}
if (has_scenario) {
  # Key variables for continuity check
  key_vars <- c("VRE share", "Electrification", "GDP per Capita (Q-centred)", "State Capacity PC1 (VDem)")
  
  df_stitch <- df_all %>%
    filter(variable %in% key_vars) %>%
    filter(aggregation == "Aggregated")
  
  df_stitch_sel <- df_stitch %>% filter(region %in% key_regions)
  
  ggplot(df_stitch_sel, aes(x = year, y = value, color = region)) +
    geom_line(aes(linetype = period), linewidth = 1.2) +
    geom_point(data = filter(df_stitch_sel, year %in% c(2020, 2022, 2025)), size = 2) +
    facet_wrap(~variable, scales = "free_y", ncol = 2) +
    scale_x_continuous(limits = c(2000, 2060), breaks = seq(2000, 2060, 10)) +
    labs(
      title = "Historical-to-Scenario Continuity Diagnosis (2000 - 2060)",
      subtitle = "Checking transition smoothness around the border year (2022/2025) for key regions",
      x = "Year",
      y = "Normalized / Percentage Value",
      color = "Region",
      linetype = "Data Source"
    ) +
    theme_custom() +
    theme(legend.box = "horizontal")
} else {
  cat("> [!NOTE] Scenario projection data was not loaded, skipping continuity validation.")
}
```

> [!TIP]
> **Continuity Assessment:**
> 1. **VRE share & Electrification**: Downscaled variables are anchored by national values interpolated in historical years, creating a seamless connection to the REMIND GDX projections.
> 2. **GDP per Capita**: Historical GDPPast and PopulationPast align perfectly with scenario counterparts because the normalization scalars are consistently applied.
> 3. **Government Effectiveness (WGI)**: Projections use the constant extension storyline or WGI projections from SSPextensions, preventing sudden governance steps.

---

# Verification & Data Sparsity Diagnostics

To complete our audit of the panel data, we summarize data density and sparsity metrics below. This helps trace where missing values occur.

```{r sparsity_table, echo=FALSE}
sparsity_df <- df_all %>%
  group_by(period, aggregation, variable) %>%
  summarize(
    Total_Obs = n(),
    Missing_Obs = sum(is.na(value)),
    Percent_Missing = percent(Missing_Obs / Total_Obs),
    .groups = "drop"
  ) %>%
  filter(Missing_Obs > 0)

if (nrow(sparsity_df) > 0) {
  kable(sparsity_df, caption = "Summary of Variables Containing Missing (NA) Observations") %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = TRUE)
} else {
  cat("### **Perfect Score: 0 Missing Values Found!**\nAll datasets (Historical and Scenario, Aggregated and Country-level) are fully populated for the analyzed drivers. No data sparsity issues detected.")
}
```

---

<center>
**End of Panel Data Input Inspection & Aggregation Analysis Report**  
*For questions or methodology details, please consult Renato Rodrigues (PFM Lead).*
</center>
