---
title: "`r paste0('PFM Model Selection — ', tools::file_path_sans_ext(basename(params$modelConfigFile)))`"
author: "PFM Modeling Output"
date: "`r Sys.Date()`"
params:
  modelConfigFile: "model-configs/default.yml"
html_document:
  toc: true
  toc_float:
    collapsed: false
  toc_depth: 3
  theme: flatly
  highlight: tango
  number_sections: true
  code_folding: hide
---

<style>
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dev='svglite',  fig.ext = ".svg", fig.width = 12, fig.height = 8, dpi=100)

library(pfm)
library(magclass)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(rprojroot)

source(file.path(rprojroot::find_rstudio_root_file(), "src/r/configHelper.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/modelHelpers.R"))

# Configure madrat
cache_dir   <- getPfmConfig("cacheDir", "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/cache/default")
mapping_dir <- gsub("cache/default", "mappings", cache_dir)
source_dir  <- gsub("cache/default", "sources",  cache_dir)
# Prime any missing VDem caches from source before locking to forcecache=TRUE.
# All other sources (EDGAR, population, etc.) must already be cached.
madrat::setConfig(
  forcecache   = FALSE,
  cachefolder  = cache_dir,
  mappingfolder = mapping_dir,
  sourcefolder = source_dir
)
invisible(tryCatch(madrat::calcOutput("VDem", subtype = "all",           aggregate = FALSE), error = function(e) NULL))
invisible(tryCatch(madrat::calcOutput("VDem", subtype = "stateCapacity", aggregate = FALSE), error = function(e) NULL))
madrat::setConfig(forcecache = TRUE)

# Resolve modelDir as an absolute path relative to the project root to prevent relative directory collisions
model_dir <- getPfmConfig("modelDir", "models")
if (!is_absolute_path(model_dir)) {
  model_dir <- file.path(rprojroot::find_rstudio_root_file(), model_dir)
}
options(pfm.modelDir = model_dir)

```


```{r defining_models, include=FALSE}

# ── Load model configurations from external YAML file ─────────────────────────
model_config_file <- params$modelConfigFile
if (!is_absolute_path(model_config_file)) {
  model_config_file <- file.path(
    rprojroot::find_rstudio_root_file(),
    "reports/model-selection",
    model_config_file
  )
}
if (!file.exists(model_config_file)) {
  stop("Model config file not found: ", model_config_file)
}

model_config_name <- tools::file_path_sans_ext(basename(model_config_file))

# YAML lists read as R lists; unlist() converts them to character vectors so
# they work correctly with pfm functions that expect character vector inputs.
normalize_model_config <- function(cfg) {
  vec_fields <- c("actorPowerDrivers", "actorPowerIndex",
                  "instQualityDrivers", "controlDrivers")
  for (f in vec_fields) {
    if (!is.null(cfg[[f]])) cfg[[f]] <- unlist(cfg[[f]])
  }
  # Ensure logical fields are actually logical (YAML may read as character)
  cfg$includeLagged      <- isTRUE(cfg$includeLagged)
  cfg$interactRegionFE   <- isTRUE(cfg$interactRegionFE)
  cfg$logisticTimeTrend  <- isTRUE(cfg$logisticTimeTrend)
  cfg$useMundlak         <- isTRUE(cfg$useMundlak)
  cfg$gdpGovInteraction  <- isTRUE(cfg$gdpGovInteraction)
  # NULL ridgeInteractions defaults to FALSE (library default)
  cfg$ridgeInteractions  <- isTRUE(cfg$ridgeInteractions)
  cfg
}

model_configs <- lapply(yaml::read_yaml(model_config_file), normalize_model_config)

```


```{r loading data, include=FALSE}

# Enable PFM model store so all fits are cached to disk with deterministic keys
# (formula + data hash). Adoption/stringency reports load from this store so
# their forest plots match model-selection exactly (same fit, same Ridge λ).
.model_store <- file.path(rprojroot::find_rstudio_root_file(),
                           getPfmConfig("modelDir", "models"))
dir.create(.model_store, showWarnings = FALSE, recursive = TRUE)
options(pfm.modelDir = .model_store)

# 1. Load Data
panelData <- getPanelDataHistoricalCached(
  aggregate = TRUE, y = 2000:2022,
  outputRegionMappingFile = "regionmapping_54.csv"
)

# 2. Add squared GDP term
panelData <- mbind(
  panelData,
  setNames(panelData[, , "GDP per Capita"]^2, "GDP per Capita Sq")
)


```


```{r auxiliar functions, include=FALSE}

# Global list to track diagnostics for the final comparison table
all_aics <- data.frame()
all_contribs <- data.frame()
all_ames <- list()   # stores computeAME() results keyed by "Sector_Stage|Model"
all_vifs  <- list()  # stores VIF results keyed by "Sector_Stage|Model"

# Collect AMEs (adoption stage only) for the cross-model comparison table
add_ame <- function(fit, sector_stage, model_name, cfg) {
  if (!grepl("Adoption", sector_stage, ignore.case = TRUE)) return(invisible(NULL))
  if (is.null(fit) || is.null(fit$model)) return(invisible(NULL))
  ame_df <- computeAME(
    fit,
    actorPowerDrivers  = c(cfg$actorPowerDrivers, cfg$actorPowerIndex),
    actorPowerIndex    = cfg$actorPowerIndex,
    instQualityDrivers = cfg$instQualityDrivers,
    controlDrivers     = cfg$controlDrivers
  )
  if (!is.null(ame_df)) {
    rmap <- cfg$regionMappingFixedEffects
    mapping_label <- if (is.null(rmap))                                    "None"
      else if (grepl("H12",  rmap, ignore.case = TRUE))                    "H12"
      else if (grepl("54",   rmap))                                         "54 Regions"
      else if (grepl("OECD", rmap, ignore.case = TRUE))                    "OECD+"
      else                                                                   rmap
    key <- paste0(sector_stage, "|", model_name)
    all_ames[[key]] <<- cbind(
      data.frame(Sector_Stage = sector_stage, Model = model_name,
                 mapping_type = mapping_label,
                 stringsAsFactors = FALSE),
      ame_df
    )
  }
}

# Collect VIF results (already computed inside fitAndDiagnose) for cross-model display.
add_vif <- function(fit, sector_stage, model_name) {
  if (is.null(fit) || is.null(fit$vifRaw) || length(fit$vifRaw) == 0) return(invisible(NULL))
  key <- paste0(sector_stage, "|", model_name)
  all_vifs[[key]] <<- list(
    Sector_Stage = sector_stage,
    Model        = model_name,
    vif_values   = fit$vifRaw,
    maxVIF       = fit$maxVIF %||% NA_real_,
    highVIF      = isTRUE(fit$highVIF),
    flagged      = fit$vifFlagged %||% character(0)
  )
}

# Render a two-part VIF display: (1) per-model summary table; (2) per-term heatmap.
# Region FE dummies and _x_ interaction terms are excluded from VIF by computeVIF()
# and therefore do not appear as columns here.
render_vif_tab <- function(sec_stg, all_vifs_data) {
  entries <- Filter(function(x) x$Sector_Stage == sec_stg, all_vifs_data)
  if (length(entries) == 0) {
    cat("> No VIF data available for this sector/stage.\n\n")
    return(invisible(NULL))
  }

  # ── Part 1: per-model summary ──────────────────────────────────────────────
  cat("**How to read:** VIF measures how much variance of a coefficient estimate is inflated by collinearity with other predictors. Values ≤ 5 are acceptable; 5–10 indicate moderate concern; > 10 indicate serious multicollinearity that destabilises estimates. Region FE dummies and interaction terms (`_x_`) are excluded from VIF computation since they require separate identification logic.\n\n")

  summary_df <- data.frame(
    Model          = sapply(entries, `[[`, "Model"),
    `Max VIF`      = round(sapply(entries, `[[`, "maxVIF"), 1),
    `High VIF`     = sapply(entries, `[[`, "highVIF"),
    `N flagged`    = sapply(entries, function(e) length(e$flagged)),
    `Flagged (>10)`= sapply(entries, function(e) {
      f <- e$flagged; if (length(f) == 0) "—" else paste(f, collapse = ", ")
    }),
    stringsAsFactors = FALSE, check.names = FALSE
  )

  is_high <- summary_df[["High VIF"]] == TRUE
  mv      <- summary_df[["Max VIF"]]
  kbl_sum <- kable(summary_df, format = "html",
                   caption = paste("VIF Model Summary —", sec_stg)) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed","responsive"),
                  font_size = 11, full_width = FALSE) %>%
    column_spec(1, bold = TRUE, width = "22em") %>%
    column_spec(2,
      background = ifelse(mv > 10, "#ffcdd2", ifelse(mv > 5, "#fff9c4", "#e8f5e9")),
      bold = mv > 10) %>%
    column_spec(3,
      background = ifelse(is_high, "#ffcdd2", "white"),
      color      = ifelse(is_high, "#c62828", "black"),
      bold = is_high) %>%
    column_spec(4, width = "5em") %>%
    column_spec(5, width = "30em") %>%
    scroll_box(width = "100%")
  print(kbl_sum)
  cat("\n<br>\n")

  # ── Part 2: per-term heatmap ───────────────────────────────────────────────
  all_terms <- unique(unlist(lapply(entries, function(e) names(e$vif_values))))
  if (length(all_terms) == 0) return(invisible(NULL))

  # Build wide table: rows = models, columns = predictor terms
  wide <- data.frame(Model = sapply(entries, `[[`, "Model"),
                     stringsAsFactors = FALSE)
  for (tm in all_terms) {
    wide[[tm]] <- sapply(entries, function(e) {
      v <- e$vif_values[tm]
      if (length(v) == 0 || is.na(v)) NA_real_ else round(v, 1)
    })
  }

  kbl_heat <- kable(wide, format = "html", digits = 1, na = "—",
                    caption = paste("VIF per Predictor —", sec_stg)) %>%
    kable_styling(bootstrap_options = c("condensed","hover","responsive"),
                  font_size = 10, full_width = TRUE) %>%
    column_spec(1, bold = TRUE, width = "18em",
                extra_css = "white-space:nowrap;") %>%
    row_spec(0, extra_css = "white-space:normal;word-break:break-word;font-size:0.85em;vertical-align:bottom;")

  for (j in seq_along(all_terms)) {
    vals <- wide[[all_terms[j]]]
    bg <- ifelse(is.na(vals),  "white",
          ifelse(vals > 10,    "#ffcdd2",
          ifelse(vals > 5,     "#fff9c4",
                               "#e8f5e9")))
    txt_col <- ifelse(!is.na(vals) & vals > 10, "#c62828", "black")
    bld     <- !is.na(vals) & vals > 10
    kbl_heat <- kbl_heat %>%
      column_spec(j + 1L, background = bg, color = txt_col, bold = bld,
                  width = "4.5em")
  }
  kbl_heat <- kbl_heat %>% scroll_box(width = "100%")
  print(kbl_heat)
  cat("\n\n")

  # Colour legend
  cat('<p style="font-size:0.88em;margin-top:4px;">',
      '<span style="background:#e8f5e9;padding:2px 7px;border-radius:3px;border:1px solid #c8e6c9;">VIF ≤ 5 — acceptable</span>&nbsp;',
      '<span style="background:#fff9c4;padding:2px 7px;border-radius:3px;border:1px solid #f9e79f;">5 &lt; VIF ≤ 10 — moderate concern</span>&nbsp;',
      '<span style="background:#ffcdd2;color:#c62828;font-weight:bold;padding:2px 7px;border-radius:3px;border:1px solid #ef9a9a;">VIF &gt; 10 — high collinearity</span>',
      '</p>\n\n')
}

# get_term_group() is superseded by pfm::classifyTermGroups().
# Kept as a thin wrapper so any downstream code that still calls it continues to work.
get_term_group <- function(nm, cfg) {
  classifyTermGroups(
    nm,
    actorPowerDrivers  = c(cfg$actorPowerDrivers, cfg$actorPowerIndex),
    actorPowerIndex    = cfg$actorPowerIndex,
    instQualityDrivers = cfg$instQualityDrivers,
    controlDrivers     = cfg$controlDrivers
  )
}

# Compute mean β×x contribution per term group and append to all_contribs.
# Theory Score and Theory Frac. are now computed by pfm::computeGroupContributions().
add_contributions <- function(fit, sector_stage, model_name, cfg) {
  if (is.null(fit) || is.null(fit$model)) return(invisible(NULL))

  gc <- computeGroupContributions(
    fit,
    actorPowerDrivers  = c(cfg$actorPowerDrivers, cfg$actorPowerIndex),
    actorPowerIndex    = cfg$actorPowerIndex,
    instQualityDrivers = cfg$instQualityDrivers,
    controlDrivers     = cfg$controlDrivers
  )
  if (is.null(gc)) return(invisible(NULL))

  group_levels <- c("Intercept", "Actor Power", "Inst. Quality", "Interaction",
                    "Controls", "Time Trend", "Path Dep.", "Region FE", "Other")

  row <- list(Sector_Stage = sector_stage, Model = model_name)
  for (g in group_levels) row[[g]] <- if (!is.null(gc[[g]])) gc[[g]] else NA_real_
  row[["Total η"]]     <- gc[["Total η"]]
  row[["Mean P"]]      <- if (grepl("Adoption", sector_stage, ignore.case = TRUE))
    gc[["Mean P"]] else NA_real_
  row[["Theory Score"]] <- gc[["Theory Score"]]
  row[["Theory Frac."]] <- gc[["Theory Frac."]]

  # FE / Theory ratio: how much of the linear predictor is driven by FE vs theory.
  # Values > 1 mean FE dominates theory; < 1 means theory dominates.
  fe_abs     <- abs(if (!is.null(gc[["Region FE"]]) && is.numeric(gc[["Region FE"]]))
                      gc[["Region FE"]] else 0)
  theory_abs <- abs(gc[["Theory Score"]])
  row[["FE / Theory"]] <- if (!is.na(theory_abs) && theory_abs > 1e-10)
                            round(fe_abs / theory_abs, 2) else NA_real_

  all_contribs <<- rbind(all_contribs, as.data.frame(row, stringsAsFactors = FALSE))

  # Propagate Theory Frac. and FE / Theory to the matching all_aics row
  match_idx <- which(all_aics$Sector_Stage == sector_stage & all_aics$Model == model_name)
  if (length(match_idx) > 0) {
    if (!"Theory Frac." %in% names(all_aics)) all_aics[["Theory Frac."]] <<- NA_real_
    all_aics[["Theory Frac."]][match_idx] <<- row[["Theory Frac."]]
    if (!"FE / Theory" %in% names(all_aics)) all_aics[["FE / Theory"]] <<- NA_real_
    all_aics[["FE / Theory"]][match_idx] <<- row[["FE / Theory"]]
  }
}

# Incremental R²: how much theory channels add beyond controls + FE.
# Uses the GLM's stored complete-case model frame (m$model) exclusively —
# this is exactly the data the model was fitted on, with correct factor levels
# and no NA rows that could make the baseline design matrix rank-deficient.
add_delta_r2 <- function(fit, sec_stg, model_name, cfg) {
  if (is.null(fit) || is.null(fit$model)) return(invisible(NULL))
  m <- fit$model

  # For logistf models use their stored model slot; for GLM use m$model.
  df <- if (inherits(m, "logistf")) {
    tryCatch(m$model, error = function(e) NULL)
  } else {
    m$model   # always stored by glm() since model = TRUE is the default
  }
  if (is.null(df) || !is.data.frame(df) || nrow(df) == 0) return(invisible(NULL))

  # Drop unused factor levels that would produce rank-deficient baselines
  df <- droplevels(df)

  # Full formula from the fitted model
  fml_full <- tryCatch(
    if (inherits(m, "logistf")) m$formula
    else                         stats::formula(m$terms),
    error = function(e) NULL
  )
  if (is.null(fml_full)) return(invisible(NULL))

  dep_var <- tryCatch(as.character(fml_full[[2]]), error = function(e) NULL)
  if (is.null(dep_var) || !dep_var %in% colnames(df)) return(invisible(NULL))

  # Identify theory terms to strip: Actor Power, IQ, and interactions (_x_)
  api_safe  <- make.names(c(cfg$actorPowerDrivers, cfg$actorPowerIndex))
  iq_safe   <- make.names(cfg$instQualityDrivers)
  is_theory <- function(term) {
    grepl("_x_", term) ||
    any(sapply(c(api_safe, iq_safe), function(v) grepl(paste0("^", v, "$"), term)))
  }

  all_terms  <- attr(terms(fml_full), "term.labels")
  base_terms <- all_terms[!sapply(all_terms, is_theory)]
  if (length(base_terms) == 0) base_terms <- "1"

  fml_base <- tryCatch(
    as.formula(paste(dep_var, "~", paste(base_terms, collapse = " + "))),
    error = function(e) NULL
  )
  if (is.null(fml_base)) return(invisible(NULL))

  stage <- if (grepl("Adoption", sec_stg, ignore.case = TRUE)) "Adoption" else "Stringency"
  fam   <- if (stage == "Adoption") binomial(link = "logit") else Gamma(link = "log")

  base_fit <- tryCatch(
    suppressWarnings(glm(fml_base, data = df, family = fam,
                          control = list(maxit = 500))),
    error = function(e) {
      message("[ΔR²] Baseline GLM failed for '", model_name, "': ", conditionMessage(e))
      NULL
    }
  )
  if (is.null(base_fit)) return(invisible(NULL))

  # logistf stores log-likelihoods; GLM uses deviance (= -2 * loglik) — both give McFadden's R²
  if (inherits(m, "logistf")) {
    ll_full <- m$loglik["full"]
    ll_null <- m$loglik["null"]
    if (is.null(ll_full) || is.null(ll_null) || !is.finite(ll_full) || !is.finite(ll_null) || ll_null == 0) {
      message("[ΔR²] loglik invalid for '", model_name, "'")
      return(invisible(NULL))
    }
    r2_full <- 1 - ll_full / ll_null
  } else {
    null_dev <- m$null.deviance
    if (is.null(null_dev) || !is.finite(null_dev) || null_dev <= 0) {
      message("[ΔR²] null.deviance invalid for '", model_name, "': ", null_dev)
      return(invisible(NULL))
    }
    r2_full <- 1 - m$deviance / null_dev
  }

  r2_base <- 1 - base_fit$deviance / base_fit$null.deviance
  if (!is.finite(r2_full) || !is.finite(r2_base)) return(invisible(NULL))

  delta <- round(r2_full - r2_base, 3)

  match_idx <- which(all_aics$Sector_Stage == sec_stg & all_aics$Model == model_name)
  if (length(match_idx) > 0) {
    if (!"ΔR²(theory)" %in% names(all_aics)) all_aics[["ΔR²(theory)"]] <<- NA_real_
    all_aics[["ΔR²(theory)"]][match_idx] <<- delta
  }
}

add_aic <- function(fit, sector_stage, model_name, cfg) {
  m <- fit$model
  is_logistf <- inherits(m, "logistf")
  k <- length(coef(m))
  
  # N obs
  nObs <- if (is_logistf) m$n else nobs(m)
  
  # Log-likelihood
  loglik <- if (is_logistf) as.numeric(m$loglik["full"]) else as.numeric(logLik(m))
  
  # ICs
  aic <- if (is_logistf) -2 * loglik + 2 * k else m$aic
  bic <- if (is_logistf) -2 * loglik + k * log(nObs) else BIC(m)
  aicc <- if (!is.na(aic) && nObs > k + 1) aic + (2 * k * (k + 1)) / (nObs - k - 1) else NA
  hqic <- if (!is.na(loglik) && nObs > 0) -2 * loglik + 2 * k * log(log(nObs)) else NA
  
  # Pseudo-R2
  pseudoR2 <- if (is_logistf) {
    if (is.null(m$loglik)) NA else round(1 - (m$loglik["full"] / m$loglik["null"]), 3)
  } else if (!is.null(m$deviance) && !is.null(m$null.deviance)) {
    round(1 - (m$deviance / m$null.deviance), 3)
  } else NA
  
  # Coef testing
  pvals <- if (!is.null(fit$coeftest)) fit$coeftest[, 4] else NULL
  zvals <- if (!is.null(fit$coeftest)) fit$coeftest[, 3] else NULL
  nSig <- if (!is.null(pvals)) sum(pvals < 0.05, na.rm = TRUE) else 0
  
  # Categorize Significant Variables
  sig_vars <- if (!is.null(pvals)) rownames(fit$coeftest)[!is.na(pvals) & pvals < 0.05] else character(0)
  sig_interaction <- sum(grepl(":|_x_", sig_vars))
  sig_base <- sig_vars[!grepl(":|_x_", sig_vars)]
  
  count_matches <- function(vars, patterns) {
    if (length(vars) == 0 || is.null(patterns) || length(patterns) == 0) return(0)
    pats_valid <- make.names(patterns)
    sum(sapply(vars, function(v) any(sapply(pats_valid, function(p) grepl(p, v, fixed = TRUE)))))
  }
  
  sig_iq <- count_matches(sig_base, cfg$instQualityDrivers)
  sig_ap <- count_matches(sig_base, c(cfg$actorPowerDrivers, cfg$actorPowerIndex))
  sig_ctrl <- count_matches(sig_base, cfg$controlDrivers)
  sig_time <- sum(grepl("year|lagged|timeTrend", sig_base, ignore.case = TRUE))
  
  maxAbsZ <- if (!is.null(zvals)) max(abs(zvals), na.rm = TRUE) else NA
  highZ <- if (!is.na(maxAbsZ)) maxAbsZ > 15 else NA
  
  # Flags
  converged <- if (is_logistf) !is.null(m$coefficients) && !any(is.na(m$coefficients)) else m$converged
  kOverN <- k / nObs
  overfitting <- kOverN > 0.1
  separation <- highZ # Approximation for separation based on extreme Z

  # VIF — populated by estimateAdoptionModel / estimatePriceStringencyModel
  # after the pfm package fix that adds vifRaw to the returned result.
  highVIF <- isTRUE(fit$highVIF)                           # any term > 10
  maxVIF  <- fit$maxVIF %||% NA_real_

  # Compile row
  row_df <- data.frame(
    Sector_Stage = sector_stage,
    Model = model_name,
    AIC = round(as.numeric(aic), 1),
    AICc = round(as.numeric(aicc), 1),
    BIC = round(as.numeric(bic), 1),
    HQIC = round(as.numeric(hqic), 1),
    `Log-lik` = round(as.numeric(loglik), 1),
    `Pseudo-R2` = as.numeric(pseudoR2),
    `N obs` = nObs,
    `N predictors` = k,
    `N significant` = nSig,
    `Sig. Interactions` = sig_interaction,
    `Sig. InstQual` = sig_iq,
    `Sig. ActorPower` = sig_ap,
    `Sig. Controls` = sig_ctrl,
    `Sig. TimeTrend` = sig_time,
    `k / N` = round(kOverN, 3),
    Overfitting = overfitting,
    Separation = separation,
    `High-Z` = highZ,
    `Max |Z|` = round(as.numeric(maxAbsZ), 1),
    Converged       = converged,
    `High VIF`      = highVIF,
    `Max VIF`       = round(as.numeric(maxVIF), 1),
    `Theory Frac.`  = NA_real_,
    `FE / Theory`   = NA_real_,
    `ΔR²(theory)`   = NA_real_,
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
  
  all_aics <<- rbind(all_aics, row_df)
}

# Load Shared Helpers for plotting
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/theme.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/forestPlot.R"))

# Wrapper to use the shared rendering function
plot_forest <- function(fit, title = "Coefficient Forest Plot") {
  if (is.null(fit) || is.null(fit$coeftest)) return(NULL)
  df <- pfm::coeftestToDataFrame(fit$coeftest)
  if (nrow(df) == 0) return(NULL)
  render_forest_plot(df, title = title)
}

# Renders a styled kableExtra diagnostics table and prints it directly.
# topic_name == "Phase 1: Base Model Assumptions" triggers baseline-relative scoring;
# any other name uses global scoring across all rows.
render_model_table <- function(sub_df, topic_name, caption_text) {
  if (nrow(sub_df) == 0) return(invisible(NULL))
  idx_all_three    <- which(sub_df$`Sig. Interactions` > 0 & sub_df$`Sig. InstQual` > 0 & sub_df$`Sig. ActorPower` > 0)
  idx_at_least_one <- which(sub_df$`Sig. Interactions` > 0 | sub_df$`Sig. InstQual` > 0 | sub_df$`Sig. ActorPower` > 0)
  get_col_fmt <- function(values, minimize = TRUE) {
    n   <- length(values)
    bg  <- rep("white", n)
    bld <- rep(FALSE, n)
    finite_vals <- values[is.finite(values)]
    if (length(finite_vals) < 1) return(list(bg = bg, color = rep("black", n), bold = bld))
    N_STEPS     <- 10L
    green_ramp  <- grDevices::colorRampPalette(c("#f1f8e9", "#43a047"))(N_STEPS)
    blue_ramp   <- grDevices::colorRampPalette(c("#e3f2fd", "#1e88e5"))(N_STEPS)
    yellow_ramp <- grDevices::colorRampPalette(c("#fffde7", "#ffb300"))(N_STEPS)
    apply_score <- function(j, score) {
      if (is.na(score) || !is.finite(score) || score <= 0) return(invisible())
      step   <- max(1L, min(N_STEPS, as.integer(ceiling(score * N_STEPS))))
      bg[j] <<- if (j %in% idx_all_three)        green_ramp[step]
                else if (j %in% idx_at_least_one) blue_ramp[step]
                else                              yellow_ramp[step]
      if (score >= 0.5) bld[j] <<- TRUE
    }
    if (topic_name == "Phase 1: Base Model Assumptions") {
      baseline_val <- values[1L]
      better <- which(seq_len(n) > 1L & !is.na(values) & !is.na(baseline_val) &
                      (if (minimize) values < baseline_val else values > baseline_val))
      if (length(better) > 0) {
        imps    <- if (minimize) baseline_val - values[better] else values[better] - baseline_val
        max_imp <- max(imps, na.rm = TRUE)
        for (j in better) {
          imp   <- if (minimize) baseline_val - values[j] else values[j] - baseline_val
          score <- if (max_imp > 0) imp / max_imp else 1
          apply_score(j, score)
        }
      }
    } else {
      val_min <- min(finite_vals); val_max <- max(finite_vals); rng <- val_max - val_min
      for (j in seq_len(n)) {
        if (is.na(values[j]) || !is.finite(values[j])) next
        score <- if (rng > 0) { if (minimize) (val_max - values[j]) / rng else (values[j] - val_min) / rng } else 1
        apply_score(j, score)
      }
    }
    list(bg = bg, color = rep("black", n), bold = bld)
  }
  fmt_aic  <- get_col_fmt(sub_df$AIC,          minimize = TRUE)
  fmt_aicc <- get_col_fmt(sub_df$AICc,         minimize = TRUE)
  fmt_bic  <- get_col_fmt(sub_df$BIC,          minimize = TRUE)
  fmt_hqic <- get_col_fmt(sub_df$HQIC,         minimize = TRUE)
  fmt_ll   <- get_col_fmt(sub_df$`Log-lik`,    minimize = FALSE)
  fmt_r2   <- get_col_fmt(sub_df$`Pseudo-R2`,  minimize = FALSE)
  fmt_tf   <- if ("Theory Frac." %in% names(sub_df)) get_col_fmt(sub_df[["Theory Frac."]],  minimize = FALSE) else NULL
  fmt_fet  <- if ("FE / Theory"  %in% names(sub_df)) get_col_fmt(sub_df[["FE / Theory"]],   minimize = TRUE)  else NULL
  fmt_dr2  <- if ("ΔR²(theory)" %in% names(sub_df)) get_col_fmt(sub_df[["ΔR²(theory)"]], minimize = FALSE) else NULL
  high_vif_flag    <- if ("High VIF" %in% names(sub_df)) sub_df$`High VIF` == TRUE else FALSE
  is_warning       <- sub_df$Overfitting == TRUE | sub_df$Separation == TRUE | sub_df$`High-Z` == TRUE | sub_df$Converged == FALSE | isTRUE(high_vif_flag)
  model_name_bg    <- ifelse(is_warning, "#ffebee", "white")
  model_name_color <- ifelse(is_warning, "#c62828", "black")
  model_name_bold  <- is_warning
  k <- kable(sub_df, format = "html", caption = caption_text) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 11)
  k <- k %>%
    column_spec(1, background = model_name_bg, color = model_name_color, bold = model_name_bold, width = "20em") %>%
    column_spec(2, background = fmt_aic$bg,  color = fmt_aic$color,  bold = fmt_aic$bold)  %>%
    column_spec(3, background = fmt_aicc$bg, color = fmt_aicc$color, bold = fmt_aicc$bold) %>%
    column_spec(4, background = fmt_bic$bg,  color = fmt_bic$color,  bold = fmt_bic$bold)  %>%
    column_spec(5, background = fmt_hqic$bg, color = fmt_hqic$color, bold = fmt_hqic$bold) %>%
    column_spec(6, background = fmt_ll$bg,   color = fmt_ll$color,   bold = fmt_ll$bold)   %>%
    column_spec(7, background = fmt_r2$bg,   color = fmt_r2$color,   bold = fmt_r2$bold)
  if (!is.null(fmt_tf)) {
    col_tf <- which(names(sub_df) == "Theory Frac.")
    k <- k %>% column_spec(col_tf, background = fmt_tf$bg, color = fmt_tf$color, bold = fmt_tf$bold)
  }
  if (!is.null(fmt_fet)) {
    col_fet <- which(names(sub_df) == "FE / Theory")
    k <- k %>% column_spec(col_fet, background = fmt_fet$bg, color = fmt_fet$color, bold = fmt_fet$bold)
  }
  if (!is.null(fmt_dr2)) {
    col_dr2 <- which(names(sub_df) == "ΔR²(theory)")
    k <- k %>% column_spec(col_dr2, background = fmt_dr2$bg, color = fmt_dr2$color, bold = fmt_dr2$bold)
  }
  flag_bg    <- function(vals, bad = TRUE) ifelse(!is.na(vals) & vals == bad, "#ffcdd2", "white")
  flag_color <- function(vals, bad = TRUE) ifelse(!is.na(vals) & vals == bad, "#c62828", "black")
  flag_bold  <- function(vals, bad = TRUE) !is.na(vals) & vals == bad
  col_sep  <- which(names(sub_df) == "Separation")
  col_ovf  <- which(names(sub_df) == "Overfitting")
  col_hz   <- which(names(sub_df) == "High-Z")
  col_conv <- which(names(sub_df) == "Converged")
  col_hvif <- which(names(sub_df) == "High VIF")
  col_mvif <- which(names(sub_df) == "Max VIF")
  if (length(col_sep)  > 0) k <- k %>% column_spec(col_sep,  background = flag_bg(sub_df$Separation),        color = flag_color(sub_df$Separation),        bold = flag_bold(sub_df$Separation))
  if (length(col_ovf)  > 0) k <- k %>% column_spec(col_ovf,  background = flag_bg(sub_df$Overfitting),       color = flag_color(sub_df$Overfitting),       bold = flag_bold(sub_df$Overfitting))
  if (length(col_hz)   > 0) k <- k %>% column_spec(col_hz,   background = flag_bg(sub_df$`High-Z`),          color = flag_color(sub_df$`High-Z`),          bold = flag_bold(sub_df$`High-Z`))
  if (length(col_conv) > 0) k <- k %>% column_spec(col_conv, background = flag_bg(sub_df$Converged, FALSE),  color = flag_color(sub_df$Converged, FALSE),  bold = flag_bold(sub_df$Converged, FALSE))
  if (length(col_hvif) > 0) k <- k %>% column_spec(col_hvif, background = flag_bg(sub_df$`High VIF`, TRUE),  color = flag_color(sub_df$`High VIF`, TRUE),  bold = flag_bold(sub_df$`High VIF`, TRUE))
  if (length(col_mvif) > 0) {
    mvif_vals <- suppressWarnings(as.numeric(sub_df$`Max VIF`))
    bg_vif <- ifelse(!is.na(mvif_vals) & mvif_vals > 10, "#ffcdd2",
              ifelse(!is.na(mvif_vals) & mvif_vals > 5,  "#fff9c4", "white"))
    k <- k %>% column_spec(col_mvif, background = bg_vif,
                            color = ifelse(!is.na(mvif_vals) & mvif_vals > 10, "#c62828", "black"),
                            bold  = !is.na(mvif_vals) & mvif_vals > 10)
  }
  k <- k %>% scroll_box(width = "100%")
  print(k)
  cat("\n<br>\n")
}
```

# Introduction

This report details a progressive testing approach to improve the Political Feasibility Module (PFM) two-stage hurdle model. We begin with the standard baseline formulation and add progressive econometric improvements. Success mterics are presented in the Summary and Best Model Choice tab..


```{r model-specs, echo=FALSE, results='asis'}
library(tibble)
library(purrr)

model_specs <- map_df(model_configs, ~ tibble(
  Model = .x$name,
  actorPowerDrivers = paste(.x$actorPowerDrivers, collapse = "<br>"),
  actorPowerIndex = paste(.x$actorPowerIndex, collapse = "<br>"),
  instQualityDrivers = paste(.x$instQualityDrivers, collapse = "<br>"),
  controlDrivers = paste(.x$controlDrivers, collapse = "<br>"),
  `includeLagged*` = as.character(.x$includeLagged),
  logisticTrend = as.character(isTRUE(.x$logisticTimeTrend)),
  useMundlak = as.character(isTRUE(.x$useMundlak)),
  gdpGovInteraction = as.character(isTRUE(.x$gdpGovInteraction)),
  interactRegionFE = as.character(.x$interactRegionFE),
  regionMappingFE = if(is.null(.x$regionMappingFixedEffects)) "NULL" else .x$regionMappingFixedEffects
))

model_specs_formatted <- model_specs
for (col in names(model_specs_formatted)) {
  if (col != "Model") {
    baseline_val <- model_specs_formatted[[col]][1]
    model_specs_formatted[[col]] <- ifelse(
      model_specs_formatted[[col]] != baseline_val,
      paste0("<strong>", model_specs_formatted[[col]], "</strong>"),
      model_specs_formatted[[col]]
    )
  }
}

# Build base model descriptions for the Reference section
desc_df <- data.frame(
  Model = character(),
  Description = character(),
  Assumption = character(),
  stringsAsFactors = FALSE
)

for (m_idx in seq_along(model_configs)) {
  cfg <- model_configs[[m_idx]]
  cfg_desc <- cfg$description
  clean_desc <- gsub("^\\*\\*What it does:\\*\\* ", "", cfg_desc)
  parts <- strsplit(clean_desc, "\\n\\n\\*\\*Assumption:\\*\\* ")[[1]]
  desc_val <- parts[1]
  ass_val <- if (length(parts) > 1) parts[2] else "Inherited from individual components."
  
  desc_df <- rbind(desc_df, data.frame(
    Model = paste0(m_idx, ". ", cfg$name), 
    Description = desc_val, 
    Assumption = ass_val, 
    stringsAsFactors = FALSE
  ))
}

```

# Results {.tabset}

```{r generate-results, results='asis', echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}

sectors <- c("Diffuse", "Bulk")
stages <- c("Adoption", "Stringency")

for (stg in stages) {
  for (sec in sectors) {
    sec_stg <- paste0(stg, ": ", sec)
    cat(sprintf("\n## %s {.tabset}\n\n", sec_stg))

    plots_list <- list()

    # Run all models defined in the YAML config
    for (i in seq_along(model_configs)) {
      cfg <- model_configs[[i]]
      model_name <- paste0(i, ". ", cfg$name)

      fit <- tryCatch({
        if (stg == "Adoption") {
          estimateAdoptionModel(
            data = panelData, sector = sec,
            actorPowerDrivers        = cfg$actorPowerDrivers,
            actorPowerIndex          = cfg$actorPowerIndex,
            instQualityDrivers       = cfg$instQualityDrivers,
            controlDrivers           = cfg$controlDrivers,
            includeLaggedAdoption    = cfg$includeLagged,
            interactRegionFE         = isTRUE(cfg$interactRegionFE),
            regionMappingFixedEffects = cfg$regionMappingFixedEffects,
            useMundlak               = isTRUE(cfg$useMundlak),
            gdpGovInteraction        = isTRUE(cfg$gdpGovInteraction),
            logisticTimeTrend        = isTRUE(cfg$logisticTimeTrend),
            ridgeInteractions        = isTRUE(cfg$ridgeInteractions)
          )
        } else {
          estimatePriceStringencyModel(
            data = panelData, sector = sec,
            actorPowerDrivers        = cfg$actorPowerDrivers,
            actorPowerIndex          = cfg$actorPowerIndex,
            instQualityDrivers       = cfg$instQualityDrivers,
            controlDrivers           = cfg$controlDrivers,
            includeLaggedECP         = cfg$includeLagged,
            interactRegionFE         = isTRUE(cfg$interactRegionFE),
            regionMappingFixedEffects = cfg$regionMappingFixedEffects,
            useMundlak               = isTRUE(cfg$useMundlak),
            gdpGovInteraction        = isTRUE(cfg$gdpGovInteraction),
            logisticTimeTrend        = isTRUE(cfg$logisticTimeTrend),
            ridgeInteractions        = isTRUE(cfg$ridgeInteractions)
          )
        }
      }, error = function(e) {
        message(sprintf("  [SKIP] %s / %s — %s: %s", sec_stg, cfg$name, stg, conditionMessage(e)))
        NULL
      })

      if (!is.null(fit)) {
        add_aic(fit, sec_stg, model_name, cfg)
        add_contributions(fit, sec_stg, model_name, cfg)
        add_delta_r2(fit, sec_stg, model_name, cfg)
        add_ame(fit, sec_stg, model_name, cfg)
        add_vif(fit, sec_stg, model_name)
        plots_list[[i]] <- plot_forest(fit, paste0(stg, " ", sec, " - ", cfg$name))
      }
    }

    # ── Models tab: all models in one table ──────────────────────────────────
    cat("\n### Models\n\n")
    {
      sub_df_all <- all_aics %>%
        filter(Sector_Stage == sec_stg) %>%
        select(-Sector_Stage)
      if ("Theory Frac." %in% names(sub_df_all))
        sub_df_all <- dplyr::relocate(sub_df_all, `Theory Frac.`, .after = `Pseudo-R2`)
      if ("FE / Theory" %in% names(sub_df_all))
        sub_df_all <- dplyr::relocate(sub_df_all, `FE / Theory`, .after = `Theory Frac.`)
      if ("ΔR²(theory)" %in% names(sub_df_all))
        sub_df_all <- dplyr::relocate(sub_df_all, `ΔR²(theory)`, .after = `FE / Theory`)
      if ("ModelNum" %in% names(sub_df_all))
        sub_df_all <- sub_df_all %>% select(-ModelNum)
      render_model_table(sub_df_all, "All Models", paste("All Models —", sec_stg))
    }

    # ── Forest Plots sub-tab ─────────────────────────────────────────────────
    cat("\n### Forest Plots {.tabset}\n\n")
    for (i in seq_along(model_configs)) {
      if (!is.null(plots_list[[i]])) {
        cat(sprintf("\n#### %s. %s\n\n", i, model_configs[[i]]$name))
        cat(model_configs[[i]]$description, "\n\n")
        print(plots_list[[i]])
        cat("\n\n")
      }
    }

    # ── VIF Analysis sub-tab ─────────────────────────────────────────────────
    cat("\n### VIF Analysis\n\n")
    render_vif_tab(sec_stg, all_vifs)

    # ── Term Contributions sub-tab ───────────────────────────────────────────
    cat("\n### Term Contributions\n\n")
    units_lbl <- if (stg == "Adoption")
      "Mean &beta;&times;x contribution to log-odds (&eta;). Positive = pro-adoption; negative = contra-adoption."
    else
      "Mean &beta;&times;x contribution to linear predictor (&eta;). Positive = higher price; negative = lower price."
    cat(sprintf("*%s Each cell is the cross-panel mean of the partial linear predictor for that term group. Adoption rows also show Mean P — the sample mean of P(Adoption) = plogis(&eta;).*\n\n", units_lbl))

    if (nrow(all_contribs) > 0 && "Sector_Stage" %in% names(all_contribs)) {
      sub_c <- all_contribs %>%
        filter(Sector_Stage == sec_stg) %>%
        select(-Sector_Stage)

      if (nrow(sub_c) > 0) {
        sub_c <- sub_c[, !sapply(sub_c, function(x) all(is.na(x))), drop = FALSE]
        tail_cols <- intersect(c("Theory Score", "Theory Frac.", "Total η", "Mean P"), names(sub_c))
        sub_c     <- sub_c[, c(setdiff(names(sub_c), tail_cols), tail_cols), drop = FALSE]

        best_theory_row <- if ("Theory Frac." %in% names(sub_c)) {
          idx <- which.max(sub_c[["Theory Frac."]])
          if (length(idx) > 0 && is.finite(sub_c[["Theory Frac."]][idx])) idx else integer(0)
        } else integer(0)

        kbl <- kable(sub_c, format = "html", digits = 3,
                     caption = paste("Term Contributions —", sec_stg)) %>%
          kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                        font_size = 11, full_width = TRUE) %>%
          scroll_box(width = "100%")

        num_cols <- which(sapply(sub_c, is.numeric))
        for (j in num_cols) {
          vals    <- sub_c[[j]]
          abs_max <- suppressWarnings(max(abs(vals), na.rm = TRUE))
          if (is.finite(abs_max) && abs_max > 0) {
            bg <- ifelse(!is.na(vals) & abs(vals) == abs_max, "#c8e6c9", "white")
            kbl <- kbl %>% column_spec(j, background = bg)
          }
        }

        if (length(best_theory_row) > 0)
          kbl <- kbl %>% row_spec(best_theory_row, underline = TRUE, bold = TRUE)

        print(kbl)
        cat("\n<br>\n")
      }
    } else {
      cat("> Term contributions not available for this sector/stage.\n\n")
    }

  }
}
```

# Best Model Summary {.tabset}

```{r best-models-compute, include=FALSE}
# Build summary data.frame from all_aics (populated during Results section)
# Note: all_aics uses check.names = FALSE so column names contain spaces and special chars
build_best_summary <- tryCatch({
  function(sort_by = "theory") {
    if (is.null(all_aics) || nrow(all_aics) == 0) return(data.frame())
    df <- all_aics

    # Coerce numeric columns using actual column names (check.names = FALSE in all_aics)
    for (col in c("Theory Frac.", "FE / Theory", "ΔR²(theory)", "AIC", "Pseudo-R2",
                  "Sig. Interactions", "Sig. InstQual", "Sig. ActorPower", "Max VIF")) {
      if (col %in% names(df)) df[[col]] <- suppressWarnings(as.numeric(df[[col]]))
    }

    # VIF flag — High VIF (max > 10) is treated as an equivalent hard gate to Overfitting
    hvif <- if ("High VIF" %in% names(df)) df[["High VIF"]] else FALSE
    hvif[is.na(hvif)] <- FALSE

    # Theory tier
    si  <- if ("Sig. Interactions" %in% names(df)) df[["Sig. Interactions"]] else 0
    siq <- if ("Sig. InstQual"     %in% names(df)) df[["Sig. InstQual"]]     else 0
    sap <- if ("Sig. ActorPower"   %in% names(df)) df[["Sig. ActorPower"]]   else 0
    df$Tier <- ifelse(
      (!is.na(si) & si > 0) & (!is.na(siq) & siq > 0) & (!is.na(sap) & sap > 0), "Green",
      ifelse((!is.na(si) & si > 0) | (!is.na(siq) & siq > 0) | (!is.na(sap) & sap > 0), "Blue", "Yellow")
    )

    # Apply VIF flag: models with High VIF are moved to bottom of ranking (treated as flagged)
    df$VIF_Flagged <- hvif

    keep_cols <- c("Sector_Stage", "Model", "Tier", "VIF_Flagged",
                   "Theory Frac.", "FE / Theory", "ΔR²(theory)",
                   "AIC", "Pseudo-R2", "Max VIF", "N obs")
    df_sel <- df[, intersect(keep_cols, names(df)), drop = FALSE]

    vf <- if ("VIF_Flagged" %in% names(df_sel)) as.integer(df_sel$VIF_Flagged) else 0L
    # Max VIF as a soft tiebreaker: among models with equal tier/AIC, prefer lower Max VIF
    mvif_sort <- if ("Max VIF" %in% names(df_sel))
      suppressWarnings(as.numeric(df_sel[["Max VIF"]])) else rep(NA_real_, nrow(df_sel))
    if (sort_by == "theory") {
      tc <- "Theory Frac."
      if (tc %in% names(df_sel))
        df_sel <- df_sel[order(vf, -df_sel[[tc]], mvif_sort, na.last = TRUE), ]
    } else {
      if ("AIC" %in% names(df_sel))
        df_sel <- df_sel[order(vf, df_sel[["AIC"]], mvif_sort, na.last = TRUE), ]
    }
    df_sel
  }
}, error = function(e) function(sort_by = "theory") data.frame())

best_theory_df     <- tryCatch(build_best_summary("theory"),     error = function(e) data.frame())
best_prediction_df <- tryCatch(build_best_summary("prediction"), error = function(e) data.frame())
```

## By Theory Score

Best models ranked by Theory Fraction — the share of model variation driven by Actor Power, Institutional Quality, and their interactions. Green = all three channels significant; Blue = at least one; Yellow = none.

```{r best-theory-table, echo=FALSE, results='asis'}
tryCatch({
  if (!is.null(best_theory_df) && nrow(best_theory_df) > 0) {
    tier_col <- which(names(best_theory_df) == "Tier")
    kbl <- kable(head(best_theory_df, 30), format = "html", digits = 3,
                 caption = "Top 30 Models by Theory Fraction (all sector/stage combinations)") %>%
      kable_styling(bootstrap_options = c("striped","hover","condensed","responsive"),
                    font_size = 11, full_width = TRUE)
    if (length(tier_col) > 0) {
      tiers <- head(best_theory_df$Tier, 30)
      kbl <- kbl %>% column_spec(tier_col,
        background = ifelse(tiers == "Green", "#c8e6c9",
                     ifelse(tiers == "Blue",  "#bbdefb", "#fff9c4")),
        bold = tiers == "Green")
    }
    print(kbl)
  } else {
    cat("> No model results available yet.\n")
  }
}, error = function(e) cat("> Error rendering theory table:", conditionMessage(e), "\n"))
```

## By Prediction Score

Best models ranked by AIC — lower AIC indicates better out-of-sample predictive performance after penalising for model complexity.

```{r best-prediction-table, echo=FALSE, results='asis'}
tryCatch({
  if (!is.null(best_prediction_df) && nrow(best_prediction_df) > 0) {
    tier_col <- which(names(best_prediction_df) == "Tier")
    kbl <- kable(head(best_prediction_df, 30), format = "html", digits = 3,
                 caption = "Top 30 Models by AIC (all sector/stage combinations)") %>%
      kable_styling(bootstrap_options = c("striped","hover","condensed","responsive"),
                    font_size = 11, full_width = TRUE)
    if (length(tier_col) > 0) {
      tiers <- head(best_prediction_df$Tier, 30)
      kbl <- kbl %>% column_spec(tier_col,
        background = ifelse(tiers == "Green", "#c8e6c9",
                     ifelse(tiers == "Blue",  "#bbdefb", "#fff9c4")),
        bold = tiers == "Green")
    }
    print(kbl)
  } else {
    cat("> No model results available yet.\n")
  }
}, error = function(e) cat("> Error rendering prediction table:", conditionMessage(e), "\n"))
```

## Per Sector/Stage Top-5

Best theory model and best prediction model for each sector/stage combination.

```{r best-per-sector-table, echo=FALSE, results='asis'}
tryCatch({
  if (!is.null(best_theory_df) && nrow(best_theory_df) > 0 && "Sector_Stage" %in% names(best_theory_df)) {
    for (ss in unique(best_theory_df[["Sector_Stage"]])) {
      sub_th <- head(best_theory_df[best_theory_df[["Sector_Stage"]] == ss, ], 5)
      sub_pr <- if (!is.null(best_prediction_df) && nrow(best_prediction_df) > 0 &&
                    "Sector_Stage" %in% names(best_prediction_df))
                  head(best_prediction_df[best_prediction_df[["Sector_Stage"]] == ss, ], 5)
                else data.frame()
      cat(sprintf("\n### %s\n\n", ss))
      cat("**Top 5 by Theory Fraction:**\n\n")
      if (nrow(sub_th) > 0) print(kable(sub_th, format = "html", digits = 3) %>%
        kable_styling(bootstrap_options = c("condensed","hover"), font_size = 10, full_width = FALSE))
      if (nrow(sub_pr) > 0) {
        cat("\n\n**Top 5 by AIC (prediction):**\n\n")
        print(kable(sub_pr, format = "html", digits = 3) %>%
          kable_styling(bootstrap_options = c("condensed","hover"), font_size = 10, full_width = FALSE))
      }
    }
  } else {
    cat("> Run model-selection Results sections first to populate rankings.\n")
  }
}, error = function(e) cat("> Error rendering per-sector table:", conditionMessage(e), "\n"))
```

---

# Reference {.tabset}

## Diagnostic Metrics

* **AIC (Akaike Information Criterion):** *Lower is better.* Measures goodness of fit while penalizing for the number of parameters.
* **AICc:** *Lower is better.* AIC corrected for smaller sample sizes.
* **BIC (Bayesian Information Criterion):** *Lower is better.* Similar to AIC but with a harsher penalty for model complexity.
* **HQIC (Hannan-Quinn Information Criterion):** *Lower is better.* An alternative criterion that falls between AIC and BIC in strictness.
* **Log-lik:** *Higher is better (less negative).* The log-likelihood of the fitted model.
* **Pseudo-R2:** *Higher is better.* A measure of how much variance the model explains relative to a null model (McFadden's Pseudo-R²).
* **N obs:** Number of observations used to fit the model.
* **N predictors (k):** Number of coefficients/parameters estimated by the model.
* **N significant:** Count of coefficients that are statistically significant ($p < 0.05$).
* **Sig. Interactions:** Count of significant interaction terms.
* **Sig. InstQual:** Count of significant Institutional Quality variables.
* **Sig. ActorPower:** Count of significant Actor Power variables.
* **Sig. Controls:** Count of significant macroeconomic control variables.
* **Sig. TimeTrend:** Count of significant time-trend or path-dependency variables.
* **Theory Frac.:** *Higher is better (0–1).* Fraction of the non-intercept linear predictor attributable to the three theoretical channels: Actor Power, Institutional Quality, and their Interaction. Computed as |Actor Power + Inst. Quality + Interaction| / Σ|all non-intercept group contributions|. A value of 0.6 means 60% of the non-constant model variation is driven by theory rather than controls, time trend, path dependency, or regional fixed effects.
* **FE / Theory:** *Lower is better.* Ratio of |Region FE contribution| to |Theory Score|. Values > 1 mean region dummies dominate theory; values < 1 mean theory dominates. A value of 0 (Mundlak models) means no FE is used. Models with low FE / Theory are less susceptible to FE-driven overfitting and have more theory-driven predictions.
* **ΔR²(theory):** *Higher is better.* Incremental McFadden Pseudo-R² uniquely attributable to the theory channels (Actor Power, Institutional Quality, and their Interaction). Computed as R²_full − R²_baseline, where the baseline retains all controls, region FE, time trend, and lagged terms but strips all theory variables. A value of 0.10 means the theory variables add 10 percentage points of explained variance beyond what controls and FE already explain. This is the cleanest measure of how much the theory is actually contributing — a high Pseudo-R² combined with a low ΔR²(theory) signals that controls or FE are driving the fit, not the theory.
* **k / N:** Ratio of predictors to observations. *Lower is better.* Values above 0.1 may indicate overfitting.
* **Overfitting:** *FALSE is better.* Flags if the k/N ratio exceeds 0.1.
* **Separation:** *FALSE is better.* Flags potential complete or quasi-complete separation.
* **High-Z:** *FALSE is better.* Flags extremely high Z-scores ($|Z| > 15$), usually a symptom of numerical instability.
* **Max |Z|:** *Lower is better.* The maximum absolute Z-score among the predictors.
* **Converged:** *TRUE is better.* Indicates if the optimization algorithm reached a solution.
* **High VIF:** *FALSE is better.* Hard gate — TRUE means Max VIF > 10; model is ranked after all VIF-clean models regardless of tier or AIC (Step 2a).
* **Max VIF:** *Lower is better.* Continuous tiebreaker (Step 5b) — among VIF-clean models with equal AIC, prefer lower Max VIF. Cells are colour-coded: green ≤ 5, yellow 5–10, red > 10.

### VIF Analysis tab

Each sector/stage tab contains a **VIF Analysis** sub-tab with two panels:

* **VIF Model Summary** — one row per model showing the maximum VIF, a flag if any predictor exceeds 10, and the names of flagged predictors.
* **VIF per Predictor heatmap** — rows are models, columns are individual predictor terms. Cells are colour-coded: <span style="background:#e8f5e9;padding:1px 5px;border-radius:2px;">green ≤ 5</span>, <span style="background:#fff9c4;padding:1px 5px;border-radius:2px;">yellow 5–10</span>, <span style="background:#ffcdd2;color:#c62828;padding:1px 5px;border-radius:2px;">red > 10</span>. Region FE dummies and `_x_` interaction terms are excluded from VIF computation (they have separate identification structures and their collinearity is handled by the model's design rather than by VIF trimming). VIF is computed from the OLS R² of each predictor regressed on all remaining predictors in the design matrix, without requiring external packages.

## Colour Guide

### Colour Families

Each cell is coloured based on the **theoretical robustness** of that model row — determined by whether the three theoretical channels (Actor Power, Institutional Quality, Interaction) produce statistically significant terms:

<table style="border-collapse:collapse; margin-bottom:16px;">
<tr>
  <td style="padding:6px 12px;">
    <span style="background:linear-gradient(to right,#f1f8e9,#43a047); display:inline-block; width:140px; height:20px; border-radius:3px; border:1px solid #ccc;"></span>
  </td>
  <td style="padding:6px 12px; font-weight:bold;">Green</td>
  <td style="padding:6px 12px;">Model has <strong>at least one significant term in all three</strong> theory groups (Actor Power, Institutional Quality, and Interaction).</td>
</tr>
<tr>
  <td style="padding:6px 12px;">
    <span style="background:linear-gradient(to right,#e3f2fd,#1e88e5); display:inline-block; width:140px; height:20px; border-radius:3px; border:1px solid #ccc;"></span>
  </td>
  <td style="padding:6px 12px; font-weight:bold;">Blue</td>
  <td style="padding:6px 12px;">Model has <strong>at least one significant term in at least one</strong> theory group, but not all three.</td>
</tr>
<tr>
  <td style="padding:6px 12px;">
    <span style="background:linear-gradient(to right,#fffde7,#ffb300); display:inline-block; width:140px; height:20px; border-radius:3px; border:1px solid #ccc;"></span>
  </td>
  <td style="padding:6px 12px; font-weight:bold;">Yellow</td>
  <td style="padding:6px 12px;">Model has <strong>no significant theory terms</strong>. Improvements are driven by controls, FE, or path dependency rather than theory.</td>
</tr>
</table>

### Gradient Intensity

Within each colour family, the **shade intensity** encodes how good the value is relative to the other models in the same column:

* **Darkest shade** → best value in the column for that metric
* **Lightest shade** → value closest to the worst in the column
* **White** → missing value or (Base models Phase 1 only) worse than baseline

Intensity is computed as a normalised score:
$$\text{score} = \frac{\text{best} - \text{value}}{\text{best} - \text{worst}} \quad \text{(lower-is-better metrics)}$$

Scores are mapped to 10 discrete gradient steps. **Bold text** = score ≥ 0.5 (top half of the performance range).

### Scoring behaviour

* **Models tab:** every row is scored globally — all models compete in the same column, so the full gradient is visible across base models and combos simultaneously.
* **Base models tab — Phase 1:** only models that *beat the baseline* are coloured; intensity reflects the magnitude of improvement. Models equal to or worse than baseline remain white.
* **Base models tab — Phases 2–3:** every row is scored globally within the phase table.

### Other formatting

* <span style="background-color: #ffebee; color: #c62828; padding: 2px 6px; font-weight:bold; border-radius:3px;">Red model name</span> — **Warning flag**: model has Overfitting = TRUE, Separation = TRUE, High-Z = TRUE, or Converged = FALSE. Treat results from flagged models with caution.
* <span style="background-color: #ffcdd2; color: #c62828; padding: 2px 6px; border-radius:3px;">Red cell</span> — **Flag cell**: the individual Separation, Overfitting, High-Z, or Converged cell that triggered the warning.
* **Mundlak models** have FE / Theory = 0 (no region FE contribution) and will always receive the darkest shade in the FE / Theory column.
* **ΔR²(theory)** is gradient-coloured with higher = better (darker). A model with ΔR²(theory) = 0 would mean the theory variables add nothing beyond controls — it appears white regardless of group.
* **Green overwrites Blue** and **Blue overwrites Yellow** when a model qualifies for multiple groups — the most theory-rich group always takes precedence.

## Model Configurations

```{r ref-model-configurations, echo=FALSE, results='asis'}
print(kable(model_specs_formatted, format = "html", escape = FALSE,
            caption = "Summary of Tested Model Configurations") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"),
                full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = TRUE))
```

## Model Descriptions

```{r ref-model-descriptions, echo=FALSE, results='asis'}
print(kable(desc_df, format = "html", escape = FALSE,
            caption = "Base Model Assumptions and Descriptions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"),
                font_size = 11, full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "15em") %>%
  column_spec(2, width = "35em") %>%
  column_spec(3, width = "35em"))

```

## Selection Criteria

The decision procedure for identifying the best model candidate(s), applied in this order:

<table style="width:100%;border-collapse:collapse;font-size:0.92em;">
<colgroup>
  <col style="width:4%"/>
  <col style="width:14%"/>
  <col style="width:16%"/>
  <col style="width:66%"/>
</colgroup>
<thead>
<tr style="background:#f5f5f5;">
  <th style="padding:6px 8px;border:1px solid #ddd;">Step</th>
  <th style="padding:6px 8px;border:1px solid #ddd;">Criterion</th>
  <th style="padding:6px 8px;border:1px solid #ddd;">Direction</th>
  <th style="padding:6px 8px;border:1px solid #ddd;">Rule</th>
</tr>
</thead>
<tbody>
<tr><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;">1–2</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>Converged + no flags</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">TRUE / FALSE required</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>Hard gates.</strong> Eliminate models that did not converge or are flagged for Separation, High-Z, or Overfitting. Flagged models sink to the bottom of any ranking.</td></tr>
<tr style="background:#fdf2f2;"><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;"><strong>2a</strong></td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>VIF collinearity gate</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">Max VIF &lt; 10 required</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>Hard gate (new).</strong> Models with any predictor VIF &gt; 10 are flagged as <code>High VIF = TRUE</code> and ranked after all clean models, regardless of tier or AIC. VIF &gt; 10 inflates standard errors by √VIF ≥ 3.2×, making individual coefficient estimates unreliable and significance tests uninformative. VIF &gt; 5 (moderate) is highlighted but not disqualifying. This gate prevents collinear model families (e.g., WGI Full Suite, VDem State Cap Full) from appearing as top selections despite invalid inference.</td></tr>
<tr style="background:#fafafa;"><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;">3</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>Theory significance tier</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">All three &gt; at least one &gt; none</td><td style="padding:6px 8px;border:1px solid #ddd;">Prefer 🟢 green (Interactions + InstQual + ActorPower all significant) over 🔵 blue (at least one) over 🟡 yellow (none). Theory must be present to win.</td></tr>
<tr><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;">4</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>ΔR²(theory)</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">Higher is better</td><td style="padding:6px 8px;border:1px solid #ddd;">Within the best tier: theory channels must add unique explanatory power beyond controls and FE. A high Pseudo-R² with low ΔR²(theory) signals controls are doing the work, not theory.</td></tr>
<tr style="background:#fafafa;"><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;">5</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>AIC / AICc / BIC / HQIC</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">Lower is better</td><td style="padding:6px 8px;border:1px solid #ddd;">Fit-vs-complexity penalty. BIC is strictest on parsimony; prefer models where multiple criteria agree. A model that wins all four is a strong candidate.</td></tr>
<tr style="background:#fdf6f6;"><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;"><strong>5b</strong></td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>Max VIF (ranking)</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">Lower is better (continuous tiebreaker)</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>Soft ranking criterion.</strong> Among VIF-clean models (Max VIF &lt; 10) with equal or near-equal AIC, prefer lower Max VIF. This prevents borderline-collinear models (e.g., VIF 8–9) from ranking above structurally cleaner alternatives with similar fit. Unlike step 2a (which is a hard gate), this step only re-orders models that have already passed the gate — it does not eliminate any model. A model at VIF = 8.5 remains eligible but ranks below an otherwise equivalent model at VIF = 3.0.</td></tr>
<tr><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;">6</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>Pseudo-R²</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">Higher is better</td><td style="padding:6px 8px;border:1px solid #ddd;">Overall model fit against the null model (McFadden). Combined with ΔR²(theory) this shows whether the total fit and the theory contribution move together.</td></tr>
<tr style="background:#fafafa;"><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;">7</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>Theory Frac.</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">Higher is better (≥ 0.25)</td><td style="padding:6px 8px;border:1px solid #ddd;">Theory channels should drive ≥ 25% of the non-intercept linear predictor. Values below 0.25 indicate the model is mostly explained by controls, time trend, or FE.</td></tr>
<tr><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;">8</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>FE / Theory</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">Lower is better (0 = Mundlak)</td><td style="padding:6px 8px;border:1px solid #ddd;">Ratio of region-FE contribution to theory score. Values &gt; 1 mean dummies dominate theory. Mundlak models are always 0. Prefer models where theory dominates FE.</td></tr>
<tr style="background:#fafafa;"><td style="padding:6px 8px;border:1px solid #ddd;text-align:center;">9</td><td style="padding:6px 8px;border:1px solid #ddd;"><strong>k / N</strong></td><td style="padding:6px 8px;border:1px solid #ddd;">Lower is better (&lt; 0.1)</td><td style="padding:6px 8px;border:1px solid #ddd;">Parsimony guard: ratio of predictors to observations. Values above 0.1 risk overfitting; models exceeding this threshold are already flagged in Step 2.</td></tr>
</tbody>
</table>

<br>Steps 1–2 and 2a are eliminatory hard gates. Step 3 defines the tier. Within a tier, steps 4–8 resolve ties in priority order: a model with high ΔR²(theory) and slightly worse AIC beats one with the best AIC but minimal theory contribution. **VIF step 2a** is applied simultaneously with steps 1–2: models with Max VIF > 10 are ranked last within every tier, below all VIF-clean models. **Step 5b** (Max VIF ranking) acts as a continuous tiebreaker among VIF-clean models — it does not eliminate any model but consistently favours structurally cleaner specifications when fit statistics are equivalent.

# Model Rankings {.tabset}

All fitted models ranked from best to worst per sector/stage using the Selection Criteria sort key: flagged models last → theory tier (🟢 green first) → ΔR²(theory) descending → AIC ascending → **Max VIF ascending** (tiebreaker: among models with equal AIC, prefer lower Max VIF).

**Rank columns (Rk.\*)** show each model's position on that individual criterion across all models in the table — **1 = best**. When a model's Rk.\* value is much lower than its overall position `#`, that criterion favours it more than the composite sort does; when it is much higher, that criterion works against it. A consistent pattern across multiple Rk.\* columns confirms the ranking; diverging ranks signal trade-offs worth scrutinising.

```{r model-rankings, echo=FALSE, results='asis', warning=FALSE}
rank_asc  <- function(x) rank( x, ties.method = "min", na.last = TRUE)
rank_desc <- function(x) rank(-x, ties.method = "min", na.last = TRUE)

tier_label <- function(inter, iq, ap) {
  ifelse(inter > 0 & iq > 0 & ap > 0, "\U0001f7e2 Green",
    ifelse(inter > 0 | iq > 0 | ap > 0, "\U0001f535 Blue", "\U0001f7e1 Yellow"))
}

for (stg in c("Adoption", "Stringency")) {
  for (sec in c("Diffuse", "Bulk")) {
    sec_stg <- paste0(stg, ": ", sec)
    cat(sprintf("\n## %s\n\n", sec_stg))

    sub_df <- all_aics %>% filter(Sector_Stage == sec_stg) %>% select(-Sector_Stage)
    if (nrow(sub_df) == 0) { cat("> No models available.\n\n"); next }
    if (!"ModelNum" %in% names(sub_df)) sub_df$ModelNum <- NA_integer_
    sub_df <- sub_df %>% select(-any_of("ModelNum"))

    high_vif_col <- if ("High VIF" %in% names(sub_df)) sub_df$`High VIF` == TRUE else FALSE
    is_flagged   <- sub_df$Overfitting == TRUE | sub_df$Separation == TRUE |
                    sub_df$`High-Z` == TRUE | sub_df$Converged == FALSE |
                    isTRUE(high_vif_col)
    theory_tier  <- ifelse(sub_df$`Sig. Interactions` > 0 & sub_df$`Sig. InstQual` > 0 &
                             sub_df$`Sig. ActorPower` > 0, 1L,
                    ifelse(sub_df$`Sig. Interactions` > 0 | sub_df$`Sig. InstQual` > 0 |
                             sub_df$`Sig. ActorPower` > 0, 2L, 3L))
    dr2_vals     <- if ("ΔR²(theory)" %in% names(sub_df)) sub_df[["ΔR²(theory)"]] else rep(NA_real_, nrow(sub_df))
    aic_vals     <- sub_df$AIC
    mvif_col     <- if ("Max VIF" %in% names(sub_df)) suppressWarnings(as.numeric(sub_df$`Max VIF`)) else rep(NA_real_, nrow(sub_df))

    ord <- order(as.integer(is_flagged), theory_tier, -dr2_vals, aic_vals, mvif_col, na.last = TRUE)
    sub_df       <- sub_df[ord, ]
    is_flagged   <- is_flagged[ord]
    theory_tier  <- theory_tier[ord]
    dr2_vals     <- dr2_vals[ord]
    aic_vals     <- aic_vals[ord]

    n <- nrow(sub_df)

    # Per-criterion ranks (1 = best for that criterion)
    rk_flag <- rank(as.integer(is_flagged), ties.method = "min")
    rk_tier <- rank(theory_tier,            ties.method = "min")
    rk_dr2  <- rank_desc(dr2_vals)
    rk_aic  <- rank_asc(aic_vals)
    rk_bic  <- rank_asc(sub_df$BIC)
    rk_hqic <- rank_asc(sub_df$HQIC)
    rk_r2   <- rank_desc(sub_df$`Pseudo-R2`)
    rk_tf   <- if ("Theory Frac." %in% names(sub_df)) rank_desc(sub_df[["Theory Frac."]]) else rep(NA_integer_, n)
    rk_fet  <- if ("FE / Theory"  %in% names(sub_df)) rank_asc(sub_df[["FE / Theory"]])   else rep(NA_integer_, n)
    rk_kn   <- rank_asc(sub_df$`k / N`)
    rk_vif  <- rank_asc(mvif_col)

    # Helper: rank (value) — rank outside, value in parentheses
    vr <- function(vals, ranks, digits = 2) {
      mapply(function(v, r) {
        vstr <- if (is.na(v) || !is.finite(v)) "—" else formatC(round(v, digits), digits = digits, format = "f")
        paste0(r, " (", vstr, ")")
      }, vals, ranks, SIMPLIFY = TRUE)
    }

    tier_names <- c("1" = "Green", "2" = "Blue", "3" = "Yellow")

    out <- data.frame(
      `#`      = seq_len(n),
      Model    = sub_df$Model,
      Flag     = paste0(rk_flag, " (", ifelse(is_flagged, "Yes", "No"), ")"),
      Tier     = paste0(rk_tier, " (", tier_names[as.character(theory_tier)], ")"),
      `ΔR²` = vr(dr2_vals,           rk_dr2,  3),
      AIC      = vr(aic_vals,                   rk_aic,  1),
      BIC      = vr(sub_df$BIC,                 rk_bic,  1),
      HQIC     = vr(sub_df$HQIC,                rk_hqic, 1),
      `R²`= vr(sub_df$`Pseudo-R2`,         rk_r2,   3),
      TFrac    = vr(if ("Theory Frac." %in% names(sub_df)) sub_df[["Theory Frac."]] else rep(NA_real_, n), rk_tf, 3),
      `FE/Th`  = vr(if ("FE / Theory"  %in% names(sub_df)) sub_df[["FE / Theory"]]  else rep(NA_real_, n), rk_fet, 3),
      MaxVIF   = vr(mvif_col, rk_vif, 1),
      `k/N`    = vr(sub_df$`k / N`,             rk_kn,   3),
      check.names = FALSE,
      stringsAsFactors = FALSE
    )

    # Colour maps
    ramp_green <- grDevices::colorRampPalette(c("#a5d6a7", "white"))(n)
    tier_bg  <- c("Green" = "#c8e6c9", "Blue" = "#bbdefb", "Yellow" = "#fff9c4")

    make_bg <- function(ranks) {
      ifelse(is.na(ranks), "white", ramp_green[pmin(pmax(as.integer(ranks), 1L), n)])
    }

    k <- kable(out, format = "html", caption = paste("Model Rankings —", sec_stg),
               escape = FALSE) %>%
      kable_styling(bootstrap_options = c("hover", "condensed", "responsive"),
                    font_size = 12, full_width = TRUE) %>%
      row_spec(0, extra_css = "white-space:normal;word-break:break-word;padding:3px 5px;font-size:0.9em;") %>%
      column_spec(1, bold = TRUE, width = "2em",  extra_css = "padding:2px 4px;") %>%
      column_spec(2, width = "26em", extra_css = "padding:2px 4px;") %>%
      column_spec(3,  # Flag
        background = ifelse(is_flagged, "#ffcdd2", make_bg(rk_flag)),
        color      = ifelse(is_flagged, "#c62828", "black")) %>%
      column_spec(4,  # Tier
        background = tier_bg[sub("^\\d+\\s*\\((.+)\\)$", "\\1", out$Tier)]) %>%
      column_spec(5,  background = make_bg(rk_dr2))  %>%
      column_spec(6,  background = make_bg(rk_aic))  %>%
      column_spec(7,  background = make_bg(rk_bic))  %>%
      column_spec(8,  background = make_bg(rk_hqic)) %>%
      column_spec(9,  background = make_bg(rk_r2))   %>%
      column_spec(10, background = make_bg(rk_tf))   %>%
      column_spec(11, background = make_bg(rk_fet))  %>%
      column_spec(12, # MaxVIF — red if VIF flagged, gradient otherwise
        background = ifelse(!is.na(mvif_col) & mvif_col > 10, "#ffcdd2",
                     ifelse(!is.na(mvif_col) & mvif_col > 5,  "#fff9c4",
                            make_bg(rk_vif))),
        color = ifelse(!is.na(mvif_col) & mvif_col > 10, "#c62828", "black"),
        bold  = !is.na(mvif_col) & mvif_col > 10) %>%
      column_spec(13, background = make_bg(rk_kn))

    k <- k %>% scroll_box(width = "100%")
    print(k)
    cat("\n<br>\n")
  }
}
```

# Average Marginal Effects (AME) Comparison {.tabset}

AMEs express each predictor's effect on the **probability scale**: how many percentage points does Pr(Adoption) change, on average across the sample, for a one-unit increase in that predictor? This complements AIC-based model selection by revealing whether theory variables have *practically meaningful* effect sizes — a model with low AIC but near-zero theory AMEs may be statistically preferred but theoretically hollow.

**Reading the table:** rows are theory-group predictors; columns are model specifications. Cells show AME ± SE. Bold = |AME| / SE > 1.96 (significant at 5%). Only the adoption stage is shown (logit); AMEs for the stringency stage (log-link GLM) are on the log scale and are not directly comparable.

```{r ame-comparison, echo=FALSE, results='asis'}
# ── helpers ────────────────────────────────────────────────────────────────────

# Colour-coded HTML cell. Fixes NA-SE display ("±NA" previously shown).
ame_cell_html <- function(ame, se, p, max_abs) {
  if (is.na(ame)) return('<span style="color:#bbb;font-size:0.85em;">—</span>')
  intensity <- min(abs(ame) / max(max_abs, 1e-6), 1)
  fade      <- round(255 - intensity * 185)
  bg        <- if (ame >= 0) sprintf("rgb(255,%d,%d)", fade, fade)
               else          sprintf("rgb(%d,%d,255)", fade, fade)
  sig       <- !is.na(p) && p < 0.05
  val       <- if (sig) sprintf("<b>%.4f</b>", ame) else sprintf("%.4f", ame)
  err  <- if (!is.na(se) && is.finite(se))
            sprintf('<span style="font-size:0.78em;color:#555;">&plusmn;%.4f</span>', se)
          else ""
  body <- if (nchar(err) > 0) paste0(val, "<br>", err) else val
  sprintf(
    '<div style="background:%s;padding:3px 5px;border-radius:4px;text-align:center;line-height:1.4;">%s</div>',
    bg, body
  )
}

# Render one transposed AME table (rows = models, cols = terms).
# No scroll_box — columns are narrow with wrapping headers to fit the page.
render_ame_table <- function(df_ss, term_filter_fn, caption_txt) {
  df_sub <- df_ss[term_filter_fn(df_ss), ]
  if (nrow(df_sub) == 0) {
    cat("> No matching predictors.\n\n"); return(invisible(NULL))
  }
  models_here <- unique(df_ss$Model)
  all_terms   <- unique(df_sub$term)

  grp_order <- c("Actor Power", "Inst. Quality", "Interaction", "Controls",
                 "Time Trend", "Path Dep.", "Region FE", "Other")
  term_grp  <- df_sub$term_group[match(all_terms, df_sub$term)]

  # For Region FE-only tables, sub-group columns by mapping type (H12 / OECD+ / 54 Regions)
  is_rfe_tab <- length(unique(term_grp)) == 1L && unique(term_grp) == "Region FE" &&
                "mapping_type" %in% colnames(df_sub)

  if (is_rfe_tab) {
    mapping_order <- c("H12", "OECD+", "54 Regions", "None", "Other")
    term_mapping  <- vapply(all_terms, function(t) {
      mt <- df_sub$mapping_type[df_sub$term == t & !is.na(df_sub$ame)]
      if (length(mt) == 0) return("Other")
      names(sort(table(mt), decreasing = TRUE))[1L]
    }, character(1))
    ord          <- order(match(term_mapping, mapping_order), all_terms)
    all_terms    <- all_terms[ord]
    term_grp     <- term_grp[ord]
    term_mapping <- term_mapping[ord]
    mt_runs      <- rle(term_mapping)
    header_spec  <- c(setNames(1L, " "),
                      setNames(as.list(mt_runs$lengths),
                               paste0(mt_runs$values, " regions")))
  } else {
    ord       <- order(match(term_grp, grp_order), all_terms)
    all_terms <- all_terms[ord]
    term_grp  <- term_grp[ord]
    grp_runs    <- rle(term_grp)
    header_spec <- c(setNames(1L, " "),
                     setNames(as.list(grp_runs$lengths), grp_runs$values))
  }

  max_abs <- max(abs(df_sub$ame), na.rm = TRUE)

  wide <- data.frame(Model = models_here, stringsAsFactors = FALSE)
  for (t in all_terms) {
    wide[[t]] <- vapply(models_here, function(mod) {
      row <- df_sub[df_sub$Model == mod & df_sub$term == t, ]
      if (nrow(row) == 0) return('<span style="color:#bbb;">—</span>')
      ame_cell_html(row$ame, row$se, row$p, max_abs)
    }, character(1))
  }

  # Readable HTML labels (dots → spaces, <br> wrapping)
  display_names <- c("Model", vapply(all_terms, clean_term, character(1)))

  n_data_cols <- length(all_terms)
  kb <- kable(wide, format = "html", escape = FALSE, caption = caption_txt,
              col.names = display_names) %>%
    kable_styling(
      bootstrap_options = c("condensed", "bordered"),
      font_size = 10, full_width = TRUE
    ) %>%
    add_header_above(header_spec, escape = FALSE) %>%
    # Wide model-name column; data columns let CSS handle wrapping
    column_spec(1, bold = TRUE, width = "22em",
                extra_css = "vertical-align:middle;") %>%
    column_spec(seq(2, n_data_cols + 1), width = "5em",
                extra_css = "min-width:4em;") %>%
    # Allow header cells to wrap (prevents horizontal overflow)
    row_spec(0, extra_css = "white-space:normal;word-break:break-word;vertical-align:bottom;padding:4px 3px;font-size:0.9em;")
  print(kb)
  cat("\n\n")
}

# ── main rendering ─────────────────────────────────────────────────────────────
if (length(all_ames) == 0) {
  cat("> No AME results available.\n")
} else {
  ame_all <- do.call(rbind, all_ames)
  sector_stages_ame <- unique(ame_all$Sector_Stage)
  sector_stages_ame <- sector_stages_ame[
    grepl("Adoption", sector_stages_ame, ignore.case = TRUE)
  ]

  for (sec_stg in sector_stages_ame) {
    cat("\n## ", sec_stg, " {.tabset}\n\n")
    df_ss <- ame_all[ame_all$Sector_Stage == sec_stg, ]

    # ── Tab 1: Core theory — Actor Power, Inst. Quality, API×IQ interactions ──
    cat("\n### Core Theory\n\n")
    cat("Actor Power drivers, Institutional Quality indicators, and their interactions",
        "with the API. Background = |AME| (red = positive, blue = negative).",
        "**Bold** = p < 0.05.\n\n")
    render_ame_table(
      df_ss,
      term_filter_fn = function(d) {
        is_rfe_interaction <- grepl("regionFE", d$term, ignore.case = TRUE) &
                             (grepl("_x_", d$term) | grepl(":", d$term))
        is_pure_rfe        <- d$term_group == "Region FE" &
                             !grepl("_x_|:", d$term)
        d$term != "(Intercept)" &
          !is_pure_rfe & !is_rfe_interaction &
          d$term_group %in% c("Actor Power", "Inst. Quality", "Interaction")
      },
      caption_txt = paste(sec_stg, "— Core Theory AMEs")
    )

    # ── Tab 2: Controls, time trend, path dependency ──────────────────────────
    cat("\n### Controls & Trend\n\n")
    render_ame_table(
      df_ss,
      term_filter_fn = function(d) {
        d$term_group %in% c("Controls", "Time Trend", "Path Dep.") |
          d$term == "(Intercept)"
      },
      caption_txt = paste(sec_stg, "— Controls & Trend AMEs")
    )

    # ── Tab 3: Pure region fixed effects (dummies) ────────────────────────────
    cat("\n### Region Fixed Effects\n\n")
    cat("AME for each region dummy: average change in Pr(Adoption) for a region",
        "belonging to that group vs. the reference region.\n\n")
    render_ame_table(
      df_ss,
      term_filter_fn = function(d) {
        d$term_group == "Region FE" & !grepl("_x_|:", d$term)
      },
      caption_txt = paste(sec_stg, "— Region FE AMEs")
    )

    # ── Tab 4: Region × API interaction terms (interactRegionFE models) ───────
    cat("\n### Region × API Interactions\n\n")
    cat("Present only in models with `interactRegionFE = TRUE`.",
        "Captures heterogeneous Actor Power effects across regions.\n\n")
    render_ame_table(
      df_ss,
      term_filter_fn = function(d) {
        grepl("regionFE", d$term, ignore.case = TRUE) &
          (grepl("_x_", d$term) | grepl(":", d$term))
      },
      caption_txt = paste(sec_stg, "— Region × API Interaction AMEs")
    )
  }
}
```

```{r save-results, include=FALSE}
# Save all_aics to CSV so downstream scripts can read model rankings.
results_dir <- file.path(rprojroot::find_rstudio_root_file(), "data")
dir.create(results_dir, showWarnings = FALSE, recursive = TRUE)

results_csv <- file.path(results_dir,
  paste0("model_selection_results_", model_config_name, ".csv"))
# Ensure High VIF and Max VIF columns are present even if no model computed VIF
if (!"High VIF" %in% names(all_aics)) all_aics[["High VIF"]] <- NA
if (!"Max VIF"  %in% names(all_aics)) all_aics[["Max VIF"]]  <- NA
write.csv(all_aics, results_csv, row.names = FALSE)
message("Model selection results saved to: ", results_csv)
```
