---
title: "Progressive PFM Model Formulations and Improvements"
author: "PFM Modeling Output"
date: "`r Sys.Date()`"
html_document:
  toc: true
  toc_float:
    collapsed: false
  toc_depth: 3
  theme: flatly
  highlight: tango
  number_sections: true
  code_folding: hide
---

<style>
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dev='svglite',  fig.ext = ".svg", fig.width = 12, fig.height = 8, dpi=100)

library(pfm)
library(magclass)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(rprojroot)

source(file.path(rprojroot::find_rstudio_root_file(), "src/r/configHelper.R"))

# Configure madrat
cache_dir <- getPfmConfig("cacheDir", "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/cache/default")
mapping_dir <- gsub("cache/default", "mappings", cache_dir)
madrat::setConfig(
  forcecache = TRUE,
  cachefolder = cache_dir,
  mappingfolder = mapping_dir
)

# Resolve modelDir as an absolute path relative to the project root to prevent relative directory collisions
model_dir <- getPfmConfig("modelDir", "models")
if (!is_absolute_path(model_dir)) {
  model_dir <- file.path(rprojroot::find_rstudio_root_file(), model_dir)
}
options(pfm.modelDir = model_dir)

```


```{r defining_models, include=FALSE}

# Global Model Configurations
model_configs <- list(
  list(
    name = "Baseline",
    phases = c("Phase 1: Base Model Assumptions", "Phase 2: Institutional Quality", "Phase 3: Regional Fixed Effects"),
    description = "**What it does:** Uses a single composite Actor Power Index.\n\n**Assumption:** Actor power operates symmetrically (a 1-unit increase in Innovator Power perfectly cancels out a 1-unit increase in Incumbent Power). Institutional quality interacts globally with this net power.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Split Variables",
    phases = "Phase 1: Base Model Assumptions",
    description = "**What it does:** Replaces the combined index with discrete components.\n\n**Assumption:** Incumbents and Innovators exert asymmetric power dynamics over policy adoption.",
    actorPowerDrivers = c("Innovator Power", "Incumbent Power"), 
    actorPowerIndex = c("Innovator Power", "Incumbent Power"),
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Advanced Controls",
    phases = "Phase 1: Base Model Assumptions",
    description = "**What it does:** Expands the structural controls.\n\n**Assumption:** Broader macroeconomic and demographic factors drive feasibility.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = c("GDP per Capita", "Energy Intensity", "Urban Population Share"),
    includeLagged = FALSE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Non-Linear Dynamics",
    phases = "Phase 1: Base Model Assumptions",
    description = "**What it does:** Includes a squared term for GDP per Capita.\n\n**Assumption:** The Environmental Kuznets Curve hypothesis applies to pricing adoption.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = c("GDP per Capita Sq"),
    includeLagged = FALSE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Heterogeneous Regional Effects",
    phases = "Phase 1: Base Model Assumptions",
    description = "**What it does:** Interacts Region Fixed Effects with structural drivers.\n\n**Assumption:** The impact of actor power and institutional quality varies intrinsically by region.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = TRUE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Path Dependency",
    phases = "Phase 1: Base Model Assumptions",
    description = "**What it does:** Introduces lagged adoption/price stringency.\n\n**Assumption:** Policy adoption strongly correlates with historical status (inertia).",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = TRUE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Institutional Quality - Rule of Law",
    phases = "Phase 2: Institutional Quality",
    description = "**What it does:** Introduces VDem institutional quality Rule of Law indicator.\n\n**Assumption:** Test alternative institutional quality indicators.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = c("Government Effectiveness (WGI)", "Rule of Law (VDem)"), 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Institutional Quality - Diagonal Accountability",
    phases = "Phase 2: Institutional Quality",
    description = "**What it does:** Introduces VDem Diagonal Accountability indicator (Media & CSOs monitoring).\n\n**Assumption:** Non-electoral public scrutiny exposes incumbent lobbying.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = c("Government Effectiveness (WGI)", "Diagonal Accountability (VDem)"), 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Institutional Quality - Diagonal & Vertical Accountability",
    phases = "Phase 2: Institutional Quality",
    description = "**What it does:** Introduces VDem Diagonal & Vertical Accountability indicators.\n\n**Assumption:** Combines media scrutiny with electoral pressure at the ballot box.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = c("Government Effectiveness (WGI)", "Diagonal Accountability (VDem)", "Vertical Accountability (VDem)"), 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Institutional Quality - Full Accountability",
    phases = "Phase 2: Institutional Quality",
    description = "**What it does:** Introduces all three VDem Accountability indicators (Vertical, Horizontal, Diagonal).\n\n**Assumption:** Test full interactive accountability system.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = c("Government Effectiveness (WGI)", "Vertical Accountability (VDem)", "Horizontal Accountability (VDem)", "Diagonal Accountability (VDem)"), 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Institutional Quality - VDem",
    phases = "Phase 2: Institutional Quality",
    description = "**What it does:** Introduces all VDem institutional quality indicators.\n\n**Assumption:** Test alternative institutional quality indicators.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = c("Government Effectiveness (WGI)", "Rule of Law (VDem)", "Vertical Accountability (VDem)", "Horizontal Accountability (VDem)", "Diagonal Accountability (VDem)"), 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmappingH12.csv"
  ),
  list(
    name = "Region FE - EU/OECD+",
    phases = "Phase 3: Regional Fixed Effects",
    description = "**What it does:** Uses a specialized region mapping for fixed effects (EU, OECD+, Rest of World).\n\n**Assumption:** Policy dynamics are clustered in these macro-regions.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = TRUE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmapping_EU_OECDp.csv"
  ),
  list(
    name = "Region FE - None",
    phases = "Phase 3: Regional Fixed Effects",
    description = "**What it does:** Disables region fixed effects.\n\n**Assumption:** Eliminating region fixed effects highlights global drivers.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = TRUE, interactRegionFE = FALSE,
    regionMappingFixedEffects = NULL
  ),
  list(
    name = "Region FE - 54 Regions",
    phases = "Phase 3: Regional Fixed Effects",
    description = "**What it does:** Uses 54 native regions for fixed effects.\n\n**Assumption:** Highly granular regional effects isolate country-level variations better.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = TRUE, interactRegionFE = FALSE,
    regionMappingFixedEffects = "regionmapping_54.csv"
  )
)


```


```{r loading data, include=FALSE}

# 1. Load Data
panelData <- getPanelDataHistoricalCached(
  aggregate = TRUE, y = 2000:2022,
  outputRegionMappingFile = "regionmapping_54.csv"
)

# 2. Add squared GDP term
panelData <- mbind(
  panelData,
  setNames(panelData[, , "GDP per Capita"]^2, "GDP per Capita Sq")
)


```


```{r auxiliar functions, include=FALSE}

# Global list to track diagnostics for the final comparison table
all_aics <- data.frame()
combo_summary_df <- data.frame()

add_aic <- function(fit, sector_stage, model_name, cfg) {
  m <- fit$model
  is_logistf <- inherits(m, "logistf")
  k <- length(coef(m))
  
  # N obs
  nObs <- if (is_logistf) m$n else nobs(m)
  
  # Log-likelihood
  loglik <- if (is_logistf) as.numeric(m$loglik["full"]) else as.numeric(logLik(m))
  
  # ICs
  aic <- if (is_logistf) -2 * loglik + 2 * k else m$aic
  bic <- if (is_logistf) -2 * loglik + k * log(nObs) else BIC(m)
  aicc <- if (!is.na(aic) && nObs > k + 1) aic + (2 * k * (k + 1)) / (nObs - k - 1) else NA
  hqic <- if (!is.na(loglik) && nObs > 0) -2 * loglik + 2 * k * log(log(nObs)) else NA
  
  # Pseudo-R2
  pseudoR2 <- if (is_logistf) {
    if (is.null(m$loglik)) NA else round(1 - (m$loglik["full"] / m$loglik["null"]), 3)
  } else if (!is.null(m$deviance) && !is.null(m$null.deviance)) {
    round(1 - (m$deviance / m$null.deviance), 3)
  } else NA
  
  # Coef testing
  pvals <- if (!is.null(fit$coeftest)) fit$coeftest[, 4] else NULL
  zvals <- if (!is.null(fit$coeftest)) fit$coeftest[, 3] else NULL
  nSig <- if (!is.null(pvals)) sum(pvals < 0.05, na.rm = TRUE) else 0
  
  # Categorize Significant Variables
  sig_vars <- if (!is.null(pvals)) rownames(fit$coeftest)[!is.na(pvals) & pvals < 0.05] else character(0)
  sig_interaction <- sum(grepl(":|_x_", sig_vars))
  sig_base <- sig_vars[!grepl(":|_x_", sig_vars)]
  
  count_matches <- function(vars, patterns) {
    if (length(vars) == 0 || is.null(patterns) || length(patterns) == 0) return(0)
    pats_valid <- make.names(patterns)
    sum(sapply(vars, function(v) any(sapply(pats_valid, function(p) grepl(p, v, fixed = TRUE)))))
  }
  
  sig_iq <- count_matches(sig_base, cfg$instQualityDrivers)
  sig_ap <- count_matches(sig_base, c(cfg$actorPowerDrivers, cfg$actorPowerIndex))
  sig_ctrl <- count_matches(sig_base, cfg$controlDrivers)
  sig_time <- sum(grepl("year|lagged|timeTrend", sig_base, ignore.case = TRUE))
  
  maxAbsZ <- if (!is.null(zvals)) max(abs(zvals), na.rm = TRUE) else NA
  highZ <- if (!is.na(maxAbsZ)) maxAbsZ > 15 else NA
  
  # Flags
  converged <- if (is_logistf) !is.null(m$coefficients) && !any(is.na(m$coefficients)) else m$converged
  kOverN <- k / nObs
  overfitting <- kOverN > 0.1
  separation <- highZ # Approximation for separation based on extreme Z
  
  # Compile row
  row_df <- data.frame(
    Sector_Stage = sector_stage,
    Model = model_name,
    AIC = round(as.numeric(aic), 1),
    AICc = round(as.numeric(aicc), 1),
    BIC = round(as.numeric(bic), 1),
    HQIC = round(as.numeric(hqic), 1),
    `Log-lik` = round(as.numeric(loglik), 1),
    `Pseudo-R2` = as.numeric(pseudoR2),
    `N obs` = nObs,
    `N predictors` = k,
    `N significant` = nSig,
    `Sig. Interactions` = sig_interaction,
    `Sig. InstQual` = sig_iq,
    `Sig. ActorPower` = sig_ap,
    `Sig. Controls` = sig_ctrl,
    `Sig. TimeTrend` = sig_time,
    `k / N` = round(kOverN, 3),
    Overfitting = overfitting,
    Separation = separation,
    `High-Z` = highZ,
    `Max |Z|` = round(as.numeric(maxAbsZ), 1),
    Converged = converged,
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
  
  all_aics <<- rbind(all_aics, row_df)
}

get_colored_models <- function(phase_name, sec_stg, all_aics_data, topics_list) {
  target_models <- topics_list[[phase_name]]
  sub_df <- all_aics_data %>% filter(Sector_Stage == sec_stg, Model %in% target_models)
  baseline_name <- target_models[1] # fallback
  if(nrow(sub_df) == 0) return(list(green=baseline_name, blue=baseline_name, yellow=baseline_name))
  
  metrics <- c("AIC", "AICc", "BIC", "HQIC", "Log-lik", "Pseudo-R2")
  minim <- c(TRUE, TRUE, TRUE, TRUE, FALSE, FALSE)
  
  idx_all_three <- which(sub_df$`Sig. Interactions` > 0 & sub_df$`Sig. InstQual` > 0 & sub_df$`Sig. ActorPower` > 0)
  idx_at_least_one <- which(sub_df$`Sig. Interactions` > 0 | sub_df$`Sig. InstQual` > 0 | sub_df$`Sig. ActorPower` > 0)
  
  greens <- c()
  blues <- c()
  yellows <- c()
  
  if (phase_name == "Phase 1: Base Model Assumptions") {
    baseline_row <- sub_df[1, ]
    for (j in 2:nrow(sub_df)) {
      beats <- FALSE
      for (m in 1:length(metrics)) {
         if (!is.na(sub_df[[metrics[m]]][j]) && !is.na(baseline_row[[metrics[m]]])) {
            if (minim[m] && sub_df[[metrics[m]]][j] < baseline_row[[metrics[m]]]) beats <- TRUE
            if (!minim[m] && sub_df[[metrics[m]]][j] > baseline_row[[metrics[m]]]) beats <- TRUE
         }
      }
      if (beats) {
         if (j %in% idx_all_three) greens <- c(greens, sub_df$Model[j])
         else if (j %in% idx_at_least_one) blues <- c(blues, sub_df$Model[j])
         else yellows <- c(yellows, sub_df$Model[j])
      }
    }
  } else {
    for (m in 1:length(metrics)) {
      values <- sub_df[[metrics[m]]]
      if (all(is.na(values))) next
      
      best_overall <- if(minim[m]) min(values, na.rm=TRUE) else max(values, na.rm=TRUE)
      val_all_three <- values[idx_all_three]
      best_all_three <- if(length(val_all_three)>0) { if(minim[m]) min(val_all_three, na.rm=TRUE) else max(val_all_three, na.rm=TRUE) } else NA
      val_at_least_one <- values[idx_at_least_one]
      best_at_least_one <- if(length(val_at_least_one)>0) { if(minim[m]) min(val_at_least_one, na.rm=TRUE) else max(val_at_least_one, na.rm=TRUE) } else NA
      
      if (!is.na(best_all_three)) {
         greens <- c(greens, sub_df$Model[which(values == best_all_three & (1:nrow(sub_df) %in% idx_all_three))])
      }
      if (!is.na(best_at_least_one)) {
         blues <- c(blues, sub_df$Model[which(values == best_at_least_one & (1:nrow(sub_df) %in% idx_at_least_one))])
      }
      if (!is.na(best_overall)) {
         yellows <- c(yellows, sub_df$Model[which(values == best_overall)])
      }
    }
  }
  
  greens <- unique(setdiff(greens, baseline_name))
  blues <- unique(setdiff(blues, c(baseline_name, greens)))
  yellows <- unique(setdiff(yellows, c(baseline_name, greens, blues)))
  
  if (length(greens) == 0) greens <- baseline_name
  if (length(blues) == 0) blues <- baseline_name
  if (length(yellows) == 0) yellows <- baseline_name
  
  list(green = greens, blue = blues, yellow = yellows)
}


# Load Shared Helpers for plotting
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/theme.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/forestPlot.R"))

# Wrapper to use the shared rendering function
plot_forest <- function(fit, title = "Coefficient Forest Plot") {
  if (is.null(fit) || is.null(fit$coeftest)) return(NULL)
  
  df <- pfm::coeftestToDataFrame(fit$coeftest)
  
  if(nrow(df) == 0) return(NULL)
  
  render_forest_plot(df, title = title)
}
```

# Introduction

This report details a progressive testing approach to improve the Political Feasibility Module (PFM) two-stage hurdle model. We begin with the standard baseline formulation and add progressive econometric improvements. Success mterics are presented in the Summary and Best Model Choice tab..


```{r model-specs, echo=FALSE, results='asis'}
library(tibble)
library(purrr)

model_specs <- map_df(model_configs, ~ tibble(
  Model = .x$name,
  actorPowerDrivers = paste(.x$actorPowerDrivers, collapse = "<br>"),
  actorPowerIndex = paste(.x$actorPowerIndex, collapse = "<br>"),
  instQualityDrivers = paste(.x$instQualityDrivers, collapse = "<br>"),
  controlDrivers = paste(.x$controlDrivers, collapse = "<br>"),
  `includeLagged*` = as.character(.x$includeLagged),
  interactRegionFE = as.character(.x$interactRegionFE),
  regionMappingFE = if(is.null(.x$regionMappingFixedEffects)) "NULL" else .x$regionMappingFixedEffects
))

model_specs_formatted <- model_specs
for (col in names(model_specs_formatted)) {
  if (col != "Model") {
    baseline_val <- model_specs_formatted[[col]][1]
    model_specs_formatted[[col]] <- ifelse(
      model_specs_formatted[[col]] != baseline_val,
      paste0("<strong>", model_specs_formatted[[col]], "</strong>"),
      model_specs_formatted[[col]]
    )
  }
}

print(kable(model_specs_formatted, format = "html", escape = FALSE, caption = "Summary of Tested Model Configurations") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"), full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = TRUE))

# Print model descriptions and assumptions table for the base models
desc_df <- data.frame(
  Model = character(),
  Description = character(),
  Assumption = character(),
  stringsAsFactors = FALSE
)

for (m_idx in seq_along(model_configs)) {
  cfg <- model_configs[[m_idx]]
  cfg_desc <- cfg$description
  clean_desc <- gsub("^\\*\\*What it does:\\*\\* ", "", cfg_desc)
  parts <- strsplit(clean_desc, "\\n\\n\\*\\*Assumption:\\*\\* ")[[1]]
  desc_val <- parts[1]
  ass_val <- if (length(parts) > 1) parts[2] else "Inherited from individual components."
  
  desc_df <- rbind(desc_df, data.frame(
    Model = paste0(m_idx, ". ", cfg$name), 
    Description = desc_val, 
    Assumption = ass_val, 
    stringsAsFactors = FALSE
  ))
}

print(kable(desc_df, format = "html", escape = FALSE, caption = "Model Assumptions and Descriptions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"), font_size = 11, full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "15em") %>%
  column_spec(2, width = "35em") %>%
  column_spec(3, width = "35em"))

```

---

# Results {.tabset}

```{r generate-results, results='asis', echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
# Define the topics and the exact models that should compete in each heat dynamically
topics_template <- list()
for (idx in seq_along(model_configs)) {
  cfg <- model_configs[[idx]]
  if (!is.null(cfg$phases)) {
    model_str <- paste0(idx, ". ", cfg$name)
    for (ph in cfg$phases) {
      if (is.null(topics_template[[ph]])) {
        topics_template[[ph]] <- c()
      }
      topics_template[[ph]] <- c(topics_template[[ph]], model_str)
    }
  }
}

sectors <- c("Diffuse", "Bulk")
stages <- c("Adoption", "Stringency")

for (stg in stages) {
  for (sec in sectors) {
    sec_stg <- paste0(stg, ": ", sec)
    cat(sprintf("\n## %s {.tabset}\n\n", sec_stg))
    
    # 1. Base model configs for this iteration
    local_configs <- model_configs
    plots_list <- list()
    
    # 2. Run Base Models
    for (i in seq_along(local_configs)) {
      cfg <- local_configs[[i]]
      
      if (stg == "Adoption") {
        fit <- estimateAdoptionModel(
          data = panelData, sector = sec,
          actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
          instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
          includeLaggedAdoption = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE,
          regionMappingFixedEffects = cfg$regionMappingFixedEffects
        )
      } else {
        fit <- estimatePriceStringencyModel(
          data = panelData, sector = sec,
          actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
          instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
          includeLaggedECP = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE,
          regionMappingFixedEffects = cfg$regionMappingFixedEffects
        )
      }
      
      add_aic(fit, sec_stg, paste0(i, ". ", cfg$name), cfg)
      plot_title <- paste0(stg, " ", sec, " - ", cfg$name)
      plots_list[[i]] <- plot_forest(fit, plot_title)
    }
    
    topics <- topics_template
    # 3. Determine winners and construct dynamic combos
    c_p1 <- get_colored_models("Phase 1: Base Model Assumptions", sec_stg, all_aics, topics)
    c_p2 <- get_colored_models("Phase 2: Institutional Quality", sec_stg, all_aics, topics)
    c_p3 <- get_colored_models("Phase 3: Regional Fixed Effects", sec_stg, all_aics, topics)
    
    grid_G <- expand.grid(P1=c_p1$green, P2=c_p2$green, P3=c_p3$green, stringsAsFactors=FALSE)
    
    p1_gb <- unique(c(c_p1$green, c_p1$blue))
    p2_gb <- unique(c(c_p2$green, c_p2$blue))
    p3_gb <- unique(c(c_p3$green, c_p3$blue))
    grid_B <- expand.grid(P1=p1_gb, P2=p2_gb, P3=p3_gb, stringsAsFactors=FALSE)
    
    p1_gby <- unique(c(p1_gb, c_p1$yellow))
    p2_gby <- unique(c(p2_gb, c_p2$yellow))
    p3_gby <- unique(c(p3_gb, c_p3$yellow))
    grid_Y <- expand.grid(P1=p1_gby, P2=p2_gby, P3=p3_gby, stringsAsFactors=FALSE)
    
    all_combos <- unique(rbind(grid_G, grid_B, grid_Y))
    
    # Filter out pure baseline combination
    is_base_p1 <- all_combos$P1 == topics[["Phase 1: Base Model Assumptions"]][1]
    is_base_p2 <- all_combos$P2 == topics[["Phase 2: Institutional Quality"]][1]
    is_base_p3 <- all_combos$P3 == topics[["Phase 3: Regional Fixed Effects"]][1]
    all_combos <- all_combos[!(is_base_p1 & is_base_p2 & is_base_p3), ]
    
    combo_start_idx <- length(model_configs) + 1
    combo_names <- c()
    
    if (nrow(all_combos) > 0) {
      for(k in seq_len(nrow(all_combos))) {
        p1_name <- all_combos$P1[k]
        p2_name <- all_combos$P2[k]
        p3_name <- all_combos$P3[k]
        
        idx_p1 <- as.integer(sub("\\..*", "", p1_name))
        idx_p2 <- as.integer(sub("\\..*", "", p2_name))
        idx_p3 <- as.integer(sub("\\..*", "", p3_name))
        
        base_cfg <- local_configs[[idx_p1]]
        cfg_p2 <- local_configs[[idx_p2]]
        cfg_p3 <- local_configs[[idx_p3]]
        
        base_name <- sub("^[0-9]+\\. ", "", p1_name)
        inst_short <- sub("^[0-9]+\\. Institutional Quality - ", "", p2_name)
        if(inst_short == "1. Baseline") inst_short <- "Base Inst"
        reg_short <- sub("^[0-9]+\\. Region FE - ", "", p3_name)
        if(reg_short == "1. Baseline") reg_short <- "Base Reg"
        
        combo_name <- paste0(combo_start_idx + k - 1, ". Combo ", k, " (", base_name, " + ", inst_short, " + ", reg_short, ")")
        combo_names <- c(combo_names, combo_name)
        
        local_configs[[combo_start_idx + k - 1]] <- list(
          name = combo_name,
          description = paste0("**What it does:** Combines Base Winner '", base_name, "' with best institution (", cfg_p2$name, ") and best region (", cfg_p3$name, ")."),
          actorPowerDrivers = base_cfg$actorPowerDrivers, 
          actorPowerIndex = base_cfg$actorPowerIndex,
          instQualityDrivers = cfg_p2$instQualityDrivers, 
          controlDrivers = base_cfg$controlDrivers,
          includeLagged = base_cfg$includeLagged, 
          interactRegionFE = cfg_p3$interactRegionFE,
          regionMappingFixedEffects = cfg_p3$regionMappingFixedEffects
        )
        
        combo_summary_df <<- rbind(combo_summary_df, data.frame(
          Sector_Stage = sec_stg,
          Combo = combo_name,
          Base_Model = base_name,
          Institution = cfg_p2$name,
          Region = cfg_p3$name,
          Actor_Power = paste(base_cfg$actorPowerDrivers, collapse=", "),
          Controls = paste(base_cfg$controlDrivers, collapse=", "),
          Lagged = as.character(base_cfg$includeLagged),
          Region_Map = if(is.null(cfg_p3$regionMappingFixedEffects)) "None" else cfg_p3$regionMappingFixedEffects,
          stringsAsFactors = FALSE
        ))
      }
    }
    
    topics[["Phase 4: The Final Combinations"]] <- c("1. Baseline", combo_names)
    
    # 4. Run Models
    if (nrow(all_combos) > 0) {
      for (i in seq(combo_start_idx, combo_start_idx + nrow(all_combos) - 1)) {
        cfg <- local_configs[[i]]
        
        if (stg == "Adoption") {
          fit <- estimateAdoptionModel(
            data = panelData, sector = sec,
            actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
            instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
            includeLaggedAdoption = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE,
            regionMappingFixedEffects = cfg$regionMappingFixedEffects
          )
        } else {
          fit <- estimatePriceStringencyModel(
            data = panelData, sector = sec,
            actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
            instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
            includeLaggedECP = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE,
            regionMappingFixedEffects = cfg$regionMappingFixedEffects
          )
        }
        
        add_aic(fit, sec_stg, cfg$name, cfg)
        plot_title <- paste0(stg, " ", sec, " - ", cfg$name)
        plots_list[[i]] <- plot_forest(fit, plot_title)
      }
    }
    
    # 2. Render Summary Tables Sub-tab
    cat("\n### Summary Tables\n\n")
    
    for (topic_name in names(topics)) {
      target_models <- topics[[topic_name]]
      
      sub_df <- all_aics %>%
        filter(Sector_Stage == sec_stg, Model %in% target_models) %>%
        arrange(match(Model, target_models)) %>%
        select(-Sector_Stage) # Drop redundant columns
        
      if("ModelNum" %in% names(sub_df)) {
        sub_df <- sub_df %>% select(-ModelNum)
      }
        
      if(nrow(sub_df) == 0) next
      
      # Identify best values for styling
      best_aic  <- min(sub_df$AIC, na.rm = TRUE)
      best_aicc <- min(sub_df$AICc, na.rm = TRUE)
      best_bic  <- min(sub_df$BIC, na.rm = TRUE)
      best_hqic <- min(sub_df$HQIC, na.rm = TRUE)
      best_ll   <- max(sub_df$`Log-lik`, na.rm = TRUE)
      best_r2   <- max(sub_df$`Pseudo-R2`, na.rm = TRUE)
      best_nsig <- max(sub_df$`N significant`, na.rm = TRUE)
      
      # Identify models matching criteria for theoretical robustness
      idx_all_three <- which(sub_df$`Sig. Interactions` > 0 & sub_df$`Sig. InstQual` > 0 & sub_df$`Sig. ActorPower` > 0)
      idx_at_least_one <- which(sub_df$`Sig. Interactions` > 0 | sub_df$`Sig. InstQual` > 0 | sub_df$`Sig. ActorPower` > 0)
    
      # Function to generate formatting vectors per column based on priorities
      get_col_fmt <- function(values, minimize = TRUE) {
        bg_colors <- rep("white", length(values))
        bold_flags <- rep(FALSE, length(values))
        
        if (topic_name == "Phase 1: Base Model Assumptions") {
          baseline_val <- values[1]
          for (j in 2:length(values)) {
            if (!is.na(values[j]) && !is.na(baseline_val)) {
              is_better <- if(minimize) values[j] < baseline_val else values[j] > baseline_val
              if (is_better) {
                if (j %in% idx_all_three) {
                  bg_colors[j] <- "#c8e6c9" # Light Green
                } else if (j %in% idx_at_least_one) {
                  bg_colors[j] <- "#bbdefb" # Light Blue
                } else {
                  bg_colors[j] <- "#fff9c4" # Light Yellow
                }
                bold_flags[j] <- TRUE
              }
            }
          }
        } else {
          find_best <- function(idx) {
            if(length(idx) == 0) return(NA)
            val_subset <- values[idx]
            if (minimize) min(val_subset, na.rm = TRUE) else max(val_subset, na.rm = TRUE)
          }
          
          best_overall <- if(minimize) min(values, na.rm = TRUE) else max(values, na.rm = TRUE)
          best_all_three <- find_best(idx_all_three)
          best_at_least_one <- find_best(idx_at_least_one)
          
          if(!is.na(best_overall)) {
            m <- which(values == best_overall)
            bg_colors[m] <- "#fff9c4" # Light Yellow
            bold_flags[m] <- TRUE
          }
          if(!is.na(best_at_least_one)) {
            m <- which(values == best_at_least_one)
            bg_colors[m] <- "#bbdefb" # Light Blue
            bold_flags[m] <- TRUE
          }
          if(!is.na(best_all_three)) {
            m <- which(values == best_all_three)
            bg_colors[m] <- "#c8e6c9" # Light Green
            bold_flags[m] <- TRUE
          }
        }
        
        list(bg = bg_colors, color = rep("black", length(values)), bold = bold_flags)
      }
      
      fmt_aic  <- get_col_fmt(sub_df$AIC, minimize = TRUE)
      fmt_aicc <- get_col_fmt(sub_df$AICc, minimize = TRUE)
      fmt_bic  <- get_col_fmt(sub_df$BIC, minimize = TRUE)
      fmt_hqic <- get_col_fmt(sub_df$HQIC, minimize = TRUE)
      fmt_ll   <- get_col_fmt(sub_df$`Log-lik`, minimize = FALSE)
      fmt_r2   <- get_col_fmt(sub_df$`Pseudo-R2`, minimize = FALSE)
      
      # Warnings (Model Name Highlight)
      is_warning <- sub_df$Overfitting == TRUE | sub_df$Separation == TRUE | sub_df$`High-Z` == TRUE | sub_df$Converged == FALSE
      model_name_bg <- ifelse(is_warning, "#ffebee", "white")
      model_name_color <- ifelse(is_warning, "#c62828", "black")
      model_name_bold <- is_warning
        
      # Build base table with scroll_box for wide columns
      k <- kable(sub_df, format = "html", caption = paste("Model Diagnostics:", topic_name)) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 11)
        
      # Apply conditional highlighting for the most important metric columns
      k <- k %>%
        column_spec(1, background = model_name_bg, color = model_name_color, bold = model_name_bold) %>%
        column_spec(2, background = fmt_aic$bg, color = fmt_aic$color, bold = fmt_aic$bold) %>%
        column_spec(3, background = fmt_aicc$bg, color = fmt_aicc$color, bold = fmt_aicc$bold) %>%
        column_spec(4, background = fmt_bic$bg, color = fmt_bic$color, bold = fmt_bic$bold) %>%
        column_spec(5, background = fmt_hqic$bg, color = fmt_hqic$color, bold = fmt_hqic$bold) %>%
        column_spec(6, background = fmt_ll$bg, color = fmt_ll$color, bold = fmt_ll$bold) %>%
        column_spec(7, background = fmt_r2$bg, color = fmt_r2$color, bold = fmt_r2$bold) %>%
        scroll_box(width = "100%")
          
      # Render the table
      cat(sprintf("\n#### %s\n\n", topic_name))
      print(k)
      cat("\n<br>\n")
      
      # Print model descriptions (only for Phase 4: The Final Combinations, with no Assumption column)
      if (topic_name == "Phase 4: The Final Combinations") {
        desc_df <- data.frame(
          Model = character(),
          Description = character(),
          stringsAsFactors = FALSE
        )
        
        for (tm in target_models) {
          cfg_desc <- ""
          for (m_idx in seq_along(local_configs)) {
            cfg <- local_configs[[m_idx]]
            expected_name <- if (m_idx <= length(model_configs)) paste0(m_idx, ". ", cfg$name) else cfg$name
            if (expected_name == tm) {
              cfg_desc <- cfg$description
              break
            }
          }
          if (cfg_desc != "") {
            clean_desc <- gsub("^\\*\\*What it does:\\*\\* ", "", cfg_desc)
            parts <- strsplit(clean_desc, "\\n\\n\\*\\*Assumption:\\*\\* ")[[1]]
            desc_val <- parts[1]
            
            desc_df <- rbind(desc_df, data.frame(
              Model = tm, 
              Description = desc_val, 
              stringsAsFactors = FALSE
            ))
          }
        }
        
        if (nrow(desc_df) > 0) {
          print(kable(desc_df, format = "html", escape = FALSE) %>%
            kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"), font_size = 11, full_width = TRUE) %>%
            column_spec(1, bold = TRUE, width = "15em") %>%
            column_spec(2, width = "70em"))
        }
        cat("\n<br>\n")
      }
    }
    
    # 3. Render Forest Plots Sub-tab
    cat("\n### Forest Plots {.tabset}\n\n")
    
    for (i in seq_along(local_configs)) {
      if (!is.null(plots_list[[i]])) {
        cat(sprintf("\n#### %s. %s\n\n", i, local_configs[[i]]$name))
        cat(local_configs[[i]]$description, "\n\n")
        print(plots_list[[i]])
        cat("\n\n")
      }
    }
    
    # 4. Render Combo Configuration Sub-tab for this Sector/Stage
    cat("\n### Combo Configuration\n\n")
    local_combos <- combo_summary_df[combo_summary_df$Sector_Stage == sec_stg, ]
    if(nrow(local_combos) > 0) {
      print(kable(local_combos, format = "html") %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"), font_size = 11) %>%
        scroll_box(width = "100%"))
    } else {
      cat("No combos were generated for this sector/stage.")
    }
    cat("\n\n")
    
  }
}
```

# Diagnostic Metrics Guide

* **AIC (Akaike Information Criterion):** *Lower is better.* Measures goodness of fit while penalizing for the number of parameters.
* **AICc:** *Lower is better.* AIC corrected for smaller sample sizes.
* **BIC (Bayesian Information Criterion):** *Lower is better.* Similar to AIC but with a harsher penalty for model complexity.
* **HQIC (Hannan-Quinn Information Criterion):** *Lower is better.* An alternative criterion that falls between AIC and BIC in strictness.
* **Log-lik:** *Higher is better (less negative).* The log-likelihood of the fitted model.
* **Pseudo-R2:** *Higher is better.* A measure of how much variance the model explains relative to a null model (McFadden's Pseudo-R²).
* **N obs:** Number of observations used to fit the model.
* **N predictors (k):** Number of coefficients/parameters estimated by the model.
* **N significant:** Count of coefficients that are statistically significant ($p < 0.05$).
* **Sig. Interactions:** Count of significant interaction terms.
* **Sig. InstQual:** Count of significant Institutional Quality variables.
* **Sig. ActorPower:** Count of significant Actor Power variables.
* **Sig. Controls:** Count of significant macroeconomic control variables.
* **Sig. TimeTrend:** Count of significant time-trend or path-dependency variables.
* **k / N:** Ratio of predictors to observations. *Lower is better.* Values above 0.1 may indicate overfitting.
* **Overfitting:** *FALSE is better.* Flags if the k/N ratio exceeds 0.1.
* **Separation:** *FALSE is better.* Flags potential complete or quasi-complete separation.
* **High-Z:** *FALSE is better.* Flags extremely high Z-scores ($|Z| > 15$), usually a symptom of numerical instability.
* **Max |Z|:** *Lower is better.* The maximum absolute Z-score among the predictors.
* **Converged:** *TRUE is better.* Indicates if the optimization algorithm reached a solution.

## Highlight Legend (Ranking Columns)

* <span style="background-color: #c8e6c9; padding: 2px 4px; font-weight: bold;">Light Green</span>: **Best Model with All Three Theories** (Lowest AIC among models that have at least one significant interaction, one significant institutional quality, and one significant actor power term).
* <span style="background-color: #bbdefb; padding: 2px 4px; font-weight: bold;">Light Blue</span>: **Best Model with At Least One Theory** (Lowest AIC among models that have significant coefficients for at least one of those three theoretical clusters).
* <span style="background-color: #fff9c4; padding: 2px 4px; font-weight: bold;">Light Yellow</span>: **Best Model Overall** (Lowest AIC in the table).
* <span style="background-color: #ffebee; color: #c62828; padding: 2px 4px; font-weight: bold;">Red (Model Name)</span>: **Warning Flag** indicates a model suffering from Overfitting, Separation, abnormally High Z-scores, or failure to Converge.

*(Note: Ranking colors overwrite each other in priority. Green overwrites Blue and Yellow, Blue overwrites Yellow).*

# Key Conclusions
1. **Split Variables** better capture the asymmetric lobbying pressures of Incumbents vs. Innovators.
2. **Advanced Controls** (Energy Intensity, Urban Share) consistently reveal cleaner estimates by controlling for macro-structural confounders.
3. **Heterogeneous Region Effects** (`interactRegionFE`) show that Actor Power operates fundamentally differently inside the EU compared to the OECD or the rest of the world.
4. **Path Dependency** (`lagged_adoption` and `lagged_ecp`) provides a massive AIC drop, proving that political inertia is the single most powerful predictor of policy status.

The **Fully Advanced Model (Model 6: Path Dependency)** drastically outperforms the Baseline and represents the strongest overall specification for both Adoption and Stringency.
