---
title: "`r paste0('PFM Model Selection — ', tools::file_path_sans_ext(basename(params$modelConfigFile)))`"
author: "PFM Modeling Output"
date: "`r Sys.Date()`"
params:
  modelConfigFile: "model-configs/default.yml"
html_document:
  toc: true
  toc_float:
    collapsed: false
  toc_depth: 3
  theme: flatly
  highlight: tango
  number_sections: true
  code_folding: hide
---

<style>
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dev='svglite',  fig.ext = ".svg", fig.width = 12, fig.height = 8, dpi=100)

library(pfm)
library(magclass)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)
library(rprojroot)

source(file.path(rprojroot::find_rstudio_root_file(), "src/r/configHelper.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/modelHelpers.R"))

# Configure madrat
cache_dir <- getPfmConfig("cacheDir", "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/cache/default")
mapping_dir <- gsub("cache/default", "mappings", cache_dir)
madrat::setConfig(
  forcecache = TRUE,
  cachefolder = cache_dir,
  mappingfolder = mapping_dir
)

# Resolve modelDir as an absolute path relative to the project root to prevent relative directory collisions
model_dir <- getPfmConfig("modelDir", "models")
if (!is_absolute_path(model_dir)) {
  model_dir <- file.path(rprojroot::find_rstudio_root_file(), model_dir)
}
options(pfm.modelDir = model_dir)

```


```{r defining_models, include=FALSE}

# ── Load model configurations from external YAML file ─────────────────────────
model_config_file <- params$modelConfigFile
if (!is_absolute_path(model_config_file)) {
  model_config_file <- file.path(
    rprojroot::find_rstudio_root_file(),
    "reports/model-selection",
    model_config_file
  )
}
if (!file.exists(model_config_file)) {
  stop("Model config file not found: ", model_config_file)
}

model_config_name <- tools::file_path_sans_ext(basename(model_config_file))

# YAML lists read as R lists; unlist() converts them to character vectors so
# they work correctly with pfm functions that expect character vector inputs.
normalize_model_config <- function(cfg) {
  vec_fields <- c("phases", "actorPowerDrivers", "actorPowerIndex",
                  "instQualityDrivers", "controlDrivers")
  for (f in vec_fields) {
    if (!is.null(cfg[[f]])) cfg[[f]] <- unlist(cfg[[f]])
  }
  # Ensure logical fields are actually logical (YAML may read as character)
  cfg$includeLagged    <- isTRUE(cfg$includeLagged)
  cfg$interactRegionFE <- isTRUE(cfg$interactRegionFE)
  cfg
}

model_configs <- lapply(yaml::read_yaml(model_config_file), normalize_model_config)

```


```{r loading data, include=FALSE}

# 1. Load Data
panelData <- getPanelDataHistoricalCached(
  aggregate = TRUE, y = 2000:2022,
  outputRegionMappingFile = "regionmapping_54.csv"
)

# 2. Add squared GDP term
panelData <- mbind(
  panelData,
  setNames(panelData[, , "GDP per Capita"]^2, "GDP per Capita Sq")
)


```


```{r auxiliar functions, include=FALSE}

# Global list to track diagnostics for the final comparison table
all_aics <- data.frame()
combo_summary_df <- data.frame()
all_contribs <- data.frame()
all_ames <- list()   # stores computeAME() results keyed by "Sector_Stage|Model"

# Collect AMEs (adoption stage only) for the cross-model comparison table
add_ame <- function(fit, sector_stage, model_name, cfg) {
  if (!grepl("Adoption", sector_stage, ignore.case = TRUE)) return(invisible(NULL))
  if (is.null(fit) || is.null(fit$model)) return(invisible(NULL))
  ame_df <- computeAME(
    fit,
    actorPowerDrivers  = c(cfg$actorPowerDrivers, cfg$actorPowerIndex),
    actorPowerIndex    = cfg$actorPowerIndex,
    instQualityDrivers = cfg$instQualityDrivers,
    controlDrivers     = cfg$controlDrivers
  )
  if (!is.null(ame_df)) {
    rmap <- cfg$regionMappingFixedEffects
    mapping_label <- if (is.null(rmap))                                    "None"
      else if (grepl("H12",  rmap, ignore.case = TRUE))                    "H12"
      else if (grepl("54",   rmap))                                         "54 Regions"
      else if (grepl("OECD", rmap, ignore.case = TRUE))                    "OECD+"
      else                                                                   rmap
    key <- paste0(sector_stage, "|", model_name)
    all_ames[[key]] <<- cbind(
      data.frame(Sector_Stage = sector_stage, Model = model_name,
                 mapping_type = mapping_label,
                 stringsAsFactors = FALSE),
      ame_df
    )
  }
}

# get_term_group() is superseded by pfm::classifyTermGroups().
# Kept as a thin wrapper so any downstream code that still calls it continues to work.
get_term_group <- function(nm, cfg) {
  classifyTermGroups(
    nm,
    actorPowerDrivers  = c(cfg$actorPowerDrivers, cfg$actorPowerIndex),
    actorPowerIndex    = cfg$actorPowerIndex,
    instQualityDrivers = cfg$instQualityDrivers,
    controlDrivers     = cfg$controlDrivers
  )
}

# Compute mean β×x contribution per term group and append to all_contribs.
# Theory Score and Theory Frac. are now computed by pfm::computeGroupContributions().
add_contributions <- function(fit, sector_stage, model_name, cfg) {
  if (is.null(fit) || is.null(fit$model)) return(invisible(NULL))

  gc <- computeGroupContributions(
    fit,
    actorPowerDrivers  = c(cfg$actorPowerDrivers, cfg$actorPowerIndex),
    actorPowerIndex    = cfg$actorPowerIndex,
    instQualityDrivers = cfg$instQualityDrivers,
    controlDrivers     = cfg$controlDrivers
  )
  if (is.null(gc)) return(invisible(NULL))

  group_levels <- c("Intercept", "Actor Power", "Inst. Quality", "Interaction",
                    "Controls", "Time Trend", "Path Dep.", "Region FE", "Other")

  row <- list(Sector_Stage = sector_stage, Model = model_name)
  for (g in group_levels) row[[g]] <- if (!is.null(gc[[g]])) gc[[g]] else NA_real_
  row[["Total η"]]     <- gc[["Total η"]]
  row[["Mean P"]]      <- if (grepl("Adoption", sector_stage, ignore.case = TRUE))
    gc[["Mean P"]] else NA_real_
  row[["Theory Score"]] <- gc[["Theory Score"]]
  row[["Theory Frac."]] <- gc[["Theory Frac."]]

  all_contribs <<- rbind(all_contribs, as.data.frame(row, stringsAsFactors = FALSE))

  # Propagate Theory Frac. to the matching all_aics row so Summary Tables can highlight it
  match_idx <- which(all_aics$Sector_Stage == sector_stage & all_aics$Model == model_name)
  if (length(match_idx) > 0) {
    if (!"Theory Frac." %in% names(all_aics)) all_aics[["Theory Frac."]] <<- NA_real_
    all_aics[["Theory Frac."]][match_idx] <<- row[["Theory Frac."]]
  }
}

add_aic <- function(fit, sector_stage, model_name, cfg) {
  m <- fit$model
  is_logistf <- inherits(m, "logistf")
  k <- length(coef(m))
  
  # N obs
  nObs <- if (is_logistf) m$n else nobs(m)
  
  # Log-likelihood
  loglik <- if (is_logistf) as.numeric(m$loglik["full"]) else as.numeric(logLik(m))
  
  # ICs
  aic <- if (is_logistf) -2 * loglik + 2 * k else m$aic
  bic <- if (is_logistf) -2 * loglik + k * log(nObs) else BIC(m)
  aicc <- if (!is.na(aic) && nObs > k + 1) aic + (2 * k * (k + 1)) / (nObs - k - 1) else NA
  hqic <- if (!is.na(loglik) && nObs > 0) -2 * loglik + 2 * k * log(log(nObs)) else NA
  
  # Pseudo-R2
  pseudoR2 <- if (is_logistf) {
    if (is.null(m$loglik)) NA else round(1 - (m$loglik["full"] / m$loglik["null"]), 3)
  } else if (!is.null(m$deviance) && !is.null(m$null.deviance)) {
    round(1 - (m$deviance / m$null.deviance), 3)
  } else NA
  
  # Coef testing
  pvals <- if (!is.null(fit$coeftest)) fit$coeftest[, 4] else NULL
  zvals <- if (!is.null(fit$coeftest)) fit$coeftest[, 3] else NULL
  nSig <- if (!is.null(pvals)) sum(pvals < 0.05, na.rm = TRUE) else 0
  
  # Categorize Significant Variables
  sig_vars <- if (!is.null(pvals)) rownames(fit$coeftest)[!is.na(pvals) & pvals < 0.05] else character(0)
  sig_interaction <- sum(grepl(":|_x_", sig_vars))
  sig_base <- sig_vars[!grepl(":|_x_", sig_vars)]
  
  count_matches <- function(vars, patterns) {
    if (length(vars) == 0 || is.null(patterns) || length(patterns) == 0) return(0)
    pats_valid <- make.names(patterns)
    sum(sapply(vars, function(v) any(sapply(pats_valid, function(p) grepl(p, v, fixed = TRUE)))))
  }
  
  sig_iq <- count_matches(sig_base, cfg$instQualityDrivers)
  sig_ap <- count_matches(sig_base, c(cfg$actorPowerDrivers, cfg$actorPowerIndex))
  sig_ctrl <- count_matches(sig_base, cfg$controlDrivers)
  sig_time <- sum(grepl("year|lagged|timeTrend", sig_base, ignore.case = TRUE))
  
  maxAbsZ <- if (!is.null(zvals)) max(abs(zvals), na.rm = TRUE) else NA
  highZ <- if (!is.na(maxAbsZ)) maxAbsZ > 15 else NA
  
  # Flags
  converged <- if (is_logistf) !is.null(m$coefficients) && !any(is.na(m$coefficients)) else m$converged
  kOverN <- k / nObs
  overfitting <- kOverN > 0.1
  separation <- highZ # Approximation for separation based on extreme Z
  
  # Compile row
  row_df <- data.frame(
    Sector_Stage = sector_stage,
    Model = model_name,
    AIC = round(as.numeric(aic), 1),
    AICc = round(as.numeric(aicc), 1),
    BIC = round(as.numeric(bic), 1),
    HQIC = round(as.numeric(hqic), 1),
    `Log-lik` = round(as.numeric(loglik), 1),
    `Pseudo-R2` = as.numeric(pseudoR2),
    `N obs` = nObs,
    `N predictors` = k,
    `N significant` = nSig,
    `Sig. Interactions` = sig_interaction,
    `Sig. InstQual` = sig_iq,
    `Sig. ActorPower` = sig_ap,
    `Sig. Controls` = sig_ctrl,
    `Sig. TimeTrend` = sig_time,
    `k / N` = round(kOverN, 3),
    Overfitting = overfitting,
    Separation = separation,
    `High-Z` = highZ,
    `Max |Z|` = round(as.numeric(maxAbsZ), 1),
    Converged = converged,
    `Theory Frac.` = NA_real_,
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
  
  all_aics <<- rbind(all_aics, row_df)
}

get_colored_models <- function(phase_name, sec_stg, all_aics_data, topics_list) {
  target_models <- topics_list[[phase_name]]
  sub_df <- all_aics_data %>% filter(Sector_Stage == sec_stg, Model %in% target_models)
  baseline_name <- target_models[1] # fallback
  if(nrow(sub_df) == 0) return(list(green=baseline_name, blue=baseline_name, yellow=baseline_name))
  
  metrics <- c("AIC", "AICc", "BIC", "HQIC", "Log-lik", "Pseudo-R2")
  minim <- c(TRUE, TRUE, TRUE, TRUE, FALSE, FALSE)
  
  idx_all_three <- which(sub_df$`Sig. Interactions` > 0 & sub_df$`Sig. InstQual` > 0 & sub_df$`Sig. ActorPower` > 0)
  idx_at_least_one <- which(sub_df$`Sig. Interactions` > 0 | sub_df$`Sig. InstQual` > 0 | sub_df$`Sig. ActorPower` > 0)
  
  greens <- c()
  blues <- c()
  yellows <- c()
  
  if (phase_name == "Phase 1: Base Model Assumptions") {
    baseline_row <- sub_df[1, ]
    for (j in 2:nrow(sub_df)) {
      beats <- FALSE
      for (m in 1:length(metrics)) {
         if (!is.na(sub_df[[metrics[m]]][j]) && !is.na(baseline_row[[metrics[m]]])) {
            if (minim[m] && sub_df[[metrics[m]]][j] < baseline_row[[metrics[m]]]) beats <- TRUE
            if (!minim[m] && sub_df[[metrics[m]]][j] > baseline_row[[metrics[m]]]) beats <- TRUE
         }
      }
      if (beats) {
         if (j %in% idx_all_three) greens <- c(greens, sub_df$Model[j])
         else if (j %in% idx_at_least_one) blues <- c(blues, sub_df$Model[j])
         else yellows <- c(yellows, sub_df$Model[j])
      }
    }
  } else {
    for (m in 1:length(metrics)) {
      values <- sub_df[[metrics[m]]]
      if (all(is.na(values))) next
      
      best_overall <- if(minim[m]) min(values, na.rm=TRUE) else max(values, na.rm=TRUE)
      val_all_three <- values[idx_all_three]
      best_all_three <- if(length(val_all_three)>0) { if(minim[m]) min(val_all_three, na.rm=TRUE) else max(val_all_three, na.rm=TRUE) } else NA
      val_at_least_one <- values[idx_at_least_one]
      best_at_least_one <- if(length(val_at_least_one)>0) { if(minim[m]) min(val_at_least_one, na.rm=TRUE) else max(val_at_least_one, na.rm=TRUE) } else NA
      
      if (!is.na(best_all_three)) {
         greens <- c(greens, sub_df$Model[which(values == best_all_three & (1:nrow(sub_df) %in% idx_all_three))])
      }
      if (!is.na(best_at_least_one)) {
         blues <- c(blues, sub_df$Model[which(values == best_at_least_one & (1:nrow(sub_df) %in% idx_at_least_one))])
      }
      if (!is.na(best_overall)) {
         yellows <- c(yellows, sub_df$Model[which(values == best_overall)])
      }
    }
  }
  
  greens <- unique(setdiff(greens, baseline_name))
  blues <- unique(setdiff(blues, c(baseline_name, greens)))
  yellows <- unique(setdiff(yellows, c(baseline_name, greens, blues)))
  
  if (length(greens) == 0) greens <- baseline_name
  if (length(blues) == 0) blues <- baseline_name
  if (length(yellows) == 0) yellows <- baseline_name
  
  list(green = greens, blue = blues, yellow = yellows)
}


# Load Shared Helpers for plotting
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/theme.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/forestPlot.R"))

# Wrapper to use the shared rendering function
plot_forest <- function(fit, title = "Coefficient Forest Plot") {
  if (is.null(fit) || is.null(fit$coeftest)) return(NULL)
  
  df <- pfm::coeftestToDataFrame(fit$coeftest)
  
  if(nrow(df) == 0) return(NULL)
  
  render_forest_plot(df, title = title)
}
```

# Introduction

This report details a progressive testing approach to improve the Political Feasibility Module (PFM) two-stage hurdle model. We begin with the standard baseline formulation and add progressive econometric improvements. Success mterics are presented in the Summary and Best Model Choice tab..


```{r model-specs, echo=FALSE, results='asis'}
library(tibble)
library(purrr)

model_specs <- map_df(model_configs, ~ tibble(
  Model = .x$name,
  actorPowerDrivers = paste(.x$actorPowerDrivers, collapse = "<br>"),
  actorPowerIndex = paste(.x$actorPowerIndex, collapse = "<br>"),
  instQualityDrivers = paste(.x$instQualityDrivers, collapse = "<br>"),
  controlDrivers = paste(.x$controlDrivers, collapse = "<br>"),
  `includeLagged*` = as.character(.x$includeLagged),
  interactRegionFE = as.character(.x$interactRegionFE),
  regionMappingFE = if(is.null(.x$regionMappingFixedEffects)) "NULL" else .x$regionMappingFixedEffects
))

model_specs_formatted <- model_specs
for (col in names(model_specs_formatted)) {
  if (col != "Model") {
    baseline_val <- model_specs_formatted[[col]][1]
    model_specs_formatted[[col]] <- ifelse(
      model_specs_formatted[[col]] != baseline_val,
      paste0("<strong>", model_specs_formatted[[col]], "</strong>"),
      model_specs_formatted[[col]]
    )
  }
}

print(kable(model_specs_formatted, format = "html", escape = FALSE, caption = "Summary of Tested Model Configurations") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"), full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = TRUE))

# Print model descriptions and assumptions table for the base models
desc_df <- data.frame(
  Model = character(),
  Description = character(),
  Assumption = character(),
  stringsAsFactors = FALSE
)

for (m_idx in seq_along(model_configs)) {
  cfg <- model_configs[[m_idx]]
  cfg_desc <- cfg$description
  clean_desc <- gsub("^\\*\\*What it does:\\*\\* ", "", cfg_desc)
  parts <- strsplit(clean_desc, "\\n\\n\\*\\*Assumption:\\*\\* ")[[1]]
  desc_val <- parts[1]
  ass_val <- if (length(parts) > 1) parts[2] else "Inherited from individual components."
  
  desc_df <- rbind(desc_df, data.frame(
    Model = paste0(m_idx, ". ", cfg$name), 
    Description = desc_val, 
    Assumption = ass_val, 
    stringsAsFactors = FALSE
  ))
}

print(kable(desc_df, format = "html", escape = FALSE, caption = "Model Assumptions and Descriptions") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"), font_size = 11, full_width = TRUE) %>%
  column_spec(1, bold = TRUE, width = "15em") %>%
  column_spec(2, width = "35em") %>%
  column_spec(3, width = "35em"))

```

---

# Results {.tabset}

```{r generate-results, results='asis', echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
# Define the topics and the exact models that should compete in each heat dynamically
topics_template <- list()
for (idx in seq_along(model_configs)) {
  cfg <- model_configs[[idx]]
  if (!is.null(cfg$phases)) {
    model_str <- paste0(idx, ". ", cfg$name)
    for (ph in cfg$phases) {
      if (is.null(topics_template[[ph]])) {
        topics_template[[ph]] <- c()
      }
      topics_template[[ph]] <- c(topics_template[[ph]], model_str)
    }
  }
}

sectors <- c("Diffuse", "Bulk")
stages <- c("Adoption", "Stringency")

for (stg in stages) {
  for (sec in sectors) {
    sec_stg <- paste0(stg, ": ", sec)
    cat(sprintf("\n## %s {.tabset}\n\n", sec_stg))
    
    # 1. Base model configs for this iteration
    local_configs <- model_configs
    plots_list <- list()
    
    # 2. Run Base Models
    for (i in seq_along(local_configs)) {
      cfg <- local_configs[[i]]
      
      if (stg == "Adoption") {
        fit <- estimateAdoptionModel(
          data = panelData, sector = sec,
          actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
          instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
          includeLaggedAdoption = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE,
          regionMappingFixedEffects = cfg$regionMappingFixedEffects
        )
      } else {
        fit <- estimatePriceStringencyModel(
          data = panelData, sector = sec,
          actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
          instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
          includeLaggedECP = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE,
          regionMappingFixedEffects = cfg$regionMappingFixedEffects
        )
      }
      
      add_aic(fit, sec_stg, paste0(i, ". ", cfg$name), cfg)
      add_contributions(fit, sec_stg, paste0(i, ". ", cfg$name), cfg)
      add_ame(fit, sec_stg, paste0(i, ". ", cfg$name), cfg)
      plot_title <- paste0(stg, " ", sec, " - ", cfg$name)
      plots_list[[i]] <- plot_forest(fit, plot_title)
    }
    
    topics <- topics_template
    # 3. Determine winners and construct dynamic combos
    c_p1 <- get_colored_models("Phase 1: Base Model Assumptions", sec_stg, all_aics, topics)
    c_p2 <- get_colored_models("Phase 2: Institutional Quality", sec_stg, all_aics, topics)
    c_p3 <- get_colored_models("Phase 3: Regional Fixed Effects", sec_stg, all_aics, topics)
    
    grid_G <- expand.grid(P1=c_p1$green, P2=c_p2$green, P3=c_p3$green, stringsAsFactors=FALSE)
    
    p1_gb <- unique(c(c_p1$green, c_p1$blue))
    p2_gb <- unique(c(c_p2$green, c_p2$blue))
    p3_gb <- unique(c(c_p3$green, c_p3$blue))
    grid_B <- expand.grid(P1=p1_gb, P2=p2_gb, P3=p3_gb, stringsAsFactors=FALSE)
    
    p1_gby <- unique(c(p1_gb, c_p1$yellow))
    p2_gby <- unique(c(p2_gb, c_p2$yellow))
    p3_gby <- unique(c(p3_gb, c_p3$yellow))
    grid_Y <- expand.grid(P1=p1_gby, P2=p2_gby, P3=p3_gby, stringsAsFactors=FALSE)
    
    all_combos <- unique(rbind(grid_G, grid_B, grid_Y))
    
    # Filter out pure baseline combination
    is_base_p1 <- all_combos$P1 == topics[["Phase 1: Base Model Assumptions"]][1]
    is_base_p2 <- all_combos$P2 == topics[["Phase 2: Institutional Quality"]][1]
    is_base_p3 <- all_combos$P3 == topics[["Phase 3: Regional Fixed Effects"]][1]
    all_combos <- all_combos[!(is_base_p1 & is_base_p2 & is_base_p3), ]
    
    combo_start_idx <- length(model_configs) + 1
    combo_names <- c()
    
    if (nrow(all_combos) > 0) {
      for(k in seq_len(nrow(all_combos))) {
        p1_name <- all_combos$P1[k]
        p2_name <- all_combos$P2[k]
        p3_name <- all_combos$P3[k]
        
        idx_p1 <- as.integer(sub("\\..*", "", p1_name))
        idx_p2 <- as.integer(sub("\\..*", "", p2_name))
        idx_p3 <- as.integer(sub("\\..*", "", p3_name))
        
        base_cfg <- local_configs[[idx_p1]]
        cfg_p2 <- local_configs[[idx_p2]]
        cfg_p3 <- local_configs[[idx_p3]]
        
        base_name <- sub("^[0-9]+\\. ", "", p1_name)
        inst_short <- sub("^[0-9]+\\. Institutional Quality - ", "", p2_name)
        if(inst_short == "1. Baseline") inst_short <- "Base Inst"
        reg_short <- sub("^[0-9]+\\. Region FE - ", "", p3_name)
        if(reg_short == "1. Baseline") reg_short <- "Base Reg"
        
        combo_name <- paste0(combo_start_idx + k - 1, ". Combo ", k, " (", base_name, " + ", inst_short, " + ", reg_short, ")")
        combo_names <- c(combo_names, combo_name)
        
        local_configs[[combo_start_idx + k - 1]] <- list(
          name = combo_name,
          description = paste0("**What it does:** Combines Base Winner '", base_name, "' with best institution (", cfg_p2$name, ") and best region (", cfg_p3$name, ")."),
          actorPowerDrivers = base_cfg$actorPowerDrivers, 
          actorPowerIndex = base_cfg$actorPowerIndex,
          instQualityDrivers = cfg_p2$instQualityDrivers, 
          controlDrivers = base_cfg$controlDrivers,
          includeLagged = base_cfg$includeLagged, 
          interactRegionFE = cfg_p3$interactRegionFE,
          regionMappingFixedEffects = cfg_p3$regionMappingFixedEffects
        )
        
        combo_summary_df <<- rbind(combo_summary_df, data.frame(
          Sector_Stage = sec_stg,
          Combo = combo_name,
          Base_Model = base_name,
          Institution = cfg_p2$name,
          Region = cfg_p3$name,
          Actor_Power = paste(base_cfg$actorPowerDrivers, collapse=", "),
          Controls = paste(base_cfg$controlDrivers, collapse=", "),
          Lagged = as.character(base_cfg$includeLagged),
          Region_Map = if(is.null(cfg_p3$regionMappingFixedEffects)) "None" else cfg_p3$regionMappingFixedEffects,
          stringsAsFactors = FALSE
        ))
      }
    }
    
    topics[["Phase 4: The Final Combinations"]] <- c("1. Baseline", combo_names)
    
    # 4. Run Models
    if (nrow(all_combos) > 0) {
      for (i in seq(combo_start_idx, combo_start_idx + nrow(all_combos) - 1)) {
        cfg <- local_configs[[i]]
        
        if (stg == "Adoption") {
          fit <- estimateAdoptionModel(
            data = panelData, sector = sec,
            actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
            instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
            includeLaggedAdoption = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE,
            regionMappingFixedEffects = cfg$regionMappingFixedEffects
          )
        } else {
          fit <- estimatePriceStringencyModel(
            data = panelData, sector = sec,
            actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
            instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
            includeLaggedECP = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE,
            regionMappingFixedEffects = cfg$regionMappingFixedEffects
          )
        }
        
        add_aic(fit, sec_stg, cfg$name, cfg)
        add_contributions(fit, sec_stg, cfg$name, cfg)
        add_ame(fit, sec_stg, cfg$name, cfg)
        plot_title <- paste0(stg, " ", sec, " - ", cfg$name)
        plots_list[[i]] <- plot_forest(fit, plot_title)
      }
    }
    
    # 2. Render Summary Tables Sub-tab
    cat("\n### Summary Tables\n\n")
    
    for (topic_name in names(topics)) {
      target_models <- topics[[topic_name]]
      
      sub_df <- all_aics %>%
        filter(Sector_Stage == sec_stg, Model %in% target_models) %>%
        arrange(match(Model, target_models)) %>%
        select(-Sector_Stage)
      if ("Theory Frac." %in% names(sub_df))
        sub_df <- dplyr::relocate(sub_df, `Theory Frac.`, .after = `Pseudo-R2`)
        
      if("ModelNum" %in% names(sub_df)) {
        sub_df <- sub_df %>% select(-ModelNum)
      }
        
      if(nrow(sub_df) == 0) next
      
      # Identify best values for styling
      best_aic  <- min(sub_df$AIC, na.rm = TRUE)
      best_aicc <- min(sub_df$AICc, na.rm = TRUE)
      best_bic  <- min(sub_df$BIC, na.rm = TRUE)
      best_hqic <- min(sub_df$HQIC, na.rm = TRUE)
      best_ll   <- max(sub_df$`Log-lik`, na.rm = TRUE)
      best_r2   <- max(sub_df$`Pseudo-R2`, na.rm = TRUE)
      best_nsig <- max(sub_df$`N significant`, na.rm = TRUE)
      
      # Identify models matching criteria for theoretical robustness
      idx_all_three <- which(sub_df$`Sig. Interactions` > 0 & sub_df$`Sig. InstQual` > 0 & sub_df$`Sig. ActorPower` > 0)
      idx_at_least_one <- which(sub_df$`Sig. Interactions` > 0 | sub_df$`Sig. InstQual` > 0 | sub_df$`Sig. ActorPower` > 0)
    
      # Function to generate formatting vectors per column based on priorities
      get_col_fmt <- function(values, minimize = TRUE) {
        bg_colors <- rep("white", length(values))
        bold_flags <- rep(FALSE, length(values))
        
        if (topic_name == "Phase 1: Base Model Assumptions") {
          baseline_val <- values[1]
          for (j in 2:length(values)) {
            if (!is.na(values[j]) && !is.na(baseline_val)) {
              is_better <- if(minimize) values[j] < baseline_val else values[j] > baseline_val
              if (is_better) {
                if (j %in% idx_all_three) {
                  bg_colors[j] <- "#c8e6c9" # Light Green
                } else if (j %in% idx_at_least_one) {
                  bg_colors[j] <- "#bbdefb" # Light Blue
                } else {
                  bg_colors[j] <- "#fff9c4" # Light Yellow
                }
                bold_flags[j] <- TRUE
              }
            }
          }
        } else {
          find_best <- function(idx) {
            if(length(idx) == 0) return(NA)
            val_subset <- values[idx]
            if (minimize) min(val_subset, na.rm = TRUE) else max(val_subset, na.rm = TRUE)
          }
          
          best_overall <- if(minimize) min(values, na.rm = TRUE) else max(values, na.rm = TRUE)
          best_all_three <- find_best(idx_all_three)
          best_at_least_one <- find_best(idx_at_least_one)
          
          if(!is.na(best_overall)) {
            m <- which(values == best_overall)
            bg_colors[m] <- "#fff9c4" # Light Yellow
            bold_flags[m] <- TRUE
          }
          if(!is.na(best_at_least_one)) {
            m <- which(values == best_at_least_one)
            bg_colors[m] <- "#bbdefb" # Light Blue
            bold_flags[m] <- TRUE
          }
          if(!is.na(best_all_three)) {
            m <- which(values == best_all_three)
            bg_colors[m] <- "#c8e6c9" # Light Green
            bold_flags[m] <- TRUE
          }
        }
        
        list(bg = bg_colors, color = rep("black", length(values)), bold = bold_flags)
      }
      
      fmt_aic  <- get_col_fmt(sub_df$AIC, minimize = TRUE)
      fmt_aicc <- get_col_fmt(sub_df$AICc, minimize = TRUE)
      fmt_bic  <- get_col_fmt(sub_df$BIC, minimize = TRUE)
      fmt_hqic <- get_col_fmt(sub_df$HQIC, minimize = TRUE)
      fmt_ll   <- get_col_fmt(sub_df$`Log-lik`, minimize = FALSE)
      fmt_r2   <- get_col_fmt(sub_df$`Pseudo-R2`, minimize = FALSE)
      fmt_tf   <- if ("Theory Frac." %in% names(sub_df))
        get_col_fmt(sub_df[["Theory Frac."]], minimize = FALSE) else NULL

      # Warnings (Model Name Highlight)
      is_warning <- sub_df$Overfitting == TRUE | sub_df$Separation == TRUE | sub_df$`High-Z` == TRUE | sub_df$Converged == FALSE
      model_name_bg <- ifelse(is_warning, "#ffebee", "white")
      model_name_color <- ifelse(is_warning, "#c62828", "black")
      model_name_bold <- is_warning
        
      # Build base table with scroll_box for wide columns
      k <- kable(sub_df, format = "html", caption = paste("Model Diagnostics:", topic_name)) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 11)
        
      # Apply conditional highlighting for the most important metric columns
      k <- k %>%
        column_spec(1, background = model_name_bg, color = model_name_color, bold = model_name_bold) %>%
        column_spec(2, background = fmt_aic$bg, color = fmt_aic$color, bold = fmt_aic$bold) %>%
        column_spec(3, background = fmt_aicc$bg, color = fmt_aicc$color, bold = fmt_aicc$bold) %>%
        column_spec(4, background = fmt_bic$bg, color = fmt_bic$color, bold = fmt_bic$bold) %>%
        column_spec(5, background = fmt_hqic$bg, color = fmt_hqic$color, bold = fmt_hqic$bold) %>%
        column_spec(6, background = fmt_ll$bg, color = fmt_ll$color, bold = fmt_ll$bold) %>%
        column_spec(7, background = fmt_r2$bg, color = fmt_r2$color, bold = fmt_r2$bold)
      if (!is.null(fmt_tf)) {
        col_tf <- which(names(sub_df) == "Theory Frac.")
        k <- k %>% column_spec(col_tf, background = fmt_tf$bg, color = fmt_tf$color, bold = fmt_tf$bold)
      }
      k <- k %>% scroll_box(width = "100%")
          
      # Render the table
      cat(sprintf("\n#### %s\n\n", topic_name))
      print(k)
      cat("\n<br>\n")
      
      # Print model descriptions (only for Phase 4: The Final Combinations, with no Assumption column)
      if (topic_name == "Phase 4: The Final Combinations") {
        desc_df <- data.frame(
          Model = character(),
          Description = character(),
          stringsAsFactors = FALSE
        )
        
        for (tm in target_models) {
          cfg_desc <- ""
          for (m_idx in seq_along(local_configs)) {
            cfg <- local_configs[[m_idx]]
            expected_name <- if (m_idx <= length(model_configs)) paste0(m_idx, ". ", cfg$name) else cfg$name
            if (expected_name == tm) {
              cfg_desc <- cfg$description
              break
            }
          }
          if (cfg_desc != "") {
            clean_desc <- gsub("^\\*\\*What it does:\\*\\* ", "", cfg_desc)
            parts <- strsplit(clean_desc, "\\n\\n\\*\\*Assumption:\\*\\* ")[[1]]
            desc_val <- parts[1]
            
            desc_df <- rbind(desc_df, data.frame(
              Model = tm, 
              Description = desc_val, 
              stringsAsFactors = FALSE
            ))
          }
        }
        
        if (nrow(desc_df) > 0) {
          print(kable(desc_df, format = "html", escape = FALSE) %>%
            kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"), font_size = 11, full_width = TRUE) %>%
            column_spec(1, bold = TRUE, width = "15em") %>%
            column_spec(2, width = "70em"))
        }
        cat("\n<br>\n")
      }
    }
    
    # 3. Render Term Contributions Sub-tab
    cat("\n### Term Contributions\n\n")
    units_lbl <- if (stg == "Adoption")
      "Mean &beta;&times;x contribution to log-odds (&eta;). Positive = pro-adoption; negative = contra-adoption."
    else
      "Mean &beta;&times;x contribution to linear predictor (&eta;). Positive = higher price; negative = lower price."
    cat(sprintf("*%s Each cell is the cross-panel mean of the partial linear predictor for that term group. Adoption rows also show Mean P — the sample mean of P(Adoption) = plogis(&eta;).*\n\n", units_lbl))

    for (topic_name in names(topics)) {
      target_models <- topics[[topic_name]]
      if (nrow(all_contribs) == 0 || !"Sector_Stage" %in% names(all_contribs)) {
        cat("> Term contributions not available for this sector/stage.\n\n")
        break
      }
      sub_c <- all_contribs %>%
        filter(Sector_Stage == sec_stg, Model %in% target_models) %>%
        arrange(match(Model, target_models)) %>%
        select(-Sector_Stage)

      if (nrow(sub_c) == 0) next

      # Drop columns that are entirely NA (term group absent in all models of this phase)
      sub_c <- sub_c[, !sapply(sub_c, function(x) all(is.na(x))), drop = FALSE]

      # Theory Score and Theory Frac. are pre-computed in add_contributions.
      # Reorder: group contributions first, then summary columns.
      tail_cols <- intersect(c("Theory Score", "Theory Frac.", "Total η", "Mean P"), names(sub_c))
      sub_c     <- sub_c[, c(setdiff(names(sub_c), tail_cols), tail_cols), drop = FALSE]

      # Underline the row with the highest Theory Fraction
      best_theory_row <- if ("Theory Frac." %in% names(sub_c)) {
        idx <- which.max(sub_c[["Theory Frac."]])
        if (length(idx) > 0 && is.finite(sub_c[["Theory Frac."]][idx])) idx else integer(0)
      } else integer(0)

      # Highlight: for each numeric column, shade the largest absolute contribution
      kbl <- kable(sub_c, format = "html", digits = 3,
                   caption = paste("Term Contributions:", topic_name)) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                      font_size = 11, full_width = TRUE) %>%
        scroll_box(width = "100%")

      # Colour-code numeric columns: green = largest contribution (abs) in column
      num_cols <- which(sapply(sub_c, is.numeric))
      for (j in num_cols) {
        vals    <- sub_c[[j]]
        abs_max <- suppressWarnings(max(abs(vals), na.rm = TRUE))
        if (is.finite(abs_max) && abs_max > 0) {
          bg <- ifelse(!is.na(vals) & abs(vals) == abs_max, "#c8e6c9", "white")
          kbl <- kbl %>% column_spec(j, background = bg)
        }
      }

      # Underline + bold the row with the highest combined theory contribution
      if (length(best_theory_row) > 0)
        kbl <- kbl %>% row_spec(best_theory_row, underline = TRUE, bold = TRUE)

      cat(sprintf("\n#### %s\n\n", topic_name))
      print(kbl)
      cat("\n<br>\n")
    }

    # 4. Render Forest Plots Sub-tab
    cat("\n### Forest Plots {.tabset}\n\n")
    
    for (i in seq_along(local_configs)) {
      if (!is.null(plots_list[[i]])) {
        cat(sprintf("\n#### %s. %s\n\n", i, local_configs[[i]]$name))
        cat(local_configs[[i]]$description, "\n\n")
        print(plots_list[[i]])
        cat("\n\n")
      }
    }

    # 5. Combo Configuration for this Sector/Stage
    cat("\n### Combo Configuration\n\n")
    local_combos <- combo_summary_df[combo_summary_df$Sector_Stage == sec_stg, ]
    if (nrow(local_combos) > 0) {
      print(kable(local_combos, format = "html") %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"),
                      font_size = 11) %>%
        scroll_box(width = "100%"))
    } else {
      cat("> No combos recorded for this sector/stage.\n")
    }
    cat("\n\n")

  }
}
```

# Legend {.tabset}

## Diagnostic Metrics

* **AIC (Akaike Information Criterion):** *Lower is better.* Measures goodness of fit while penalizing for the number of parameters.
* **AICc:** *Lower is better.* AIC corrected for smaller sample sizes.
* **BIC (Bayesian Information Criterion):** *Lower is better.* Similar to AIC but with a harsher penalty for model complexity.
* **HQIC (Hannan-Quinn Information Criterion):** *Lower is better.* An alternative criterion that falls between AIC and BIC in strictness.
* **Log-lik:** *Higher is better (less negative).* The log-likelihood of the fitted model.
* **Pseudo-R2:** *Higher is better.* A measure of how much variance the model explains relative to a null model (McFadden's Pseudo-R²).
* **N obs:** Number of observations used to fit the model.
* **N predictors (k):** Number of coefficients/parameters estimated by the model.
* **N significant:** Count of coefficients that are statistically significant ($p < 0.05$).
* **Sig. Interactions:** Count of significant interaction terms.
* **Sig. InstQual:** Count of significant Institutional Quality variables.
* **Sig. ActorPower:** Count of significant Actor Power variables.
* **Sig. Controls:** Count of significant macroeconomic control variables.
* **Sig. TimeTrend:** Count of significant time-trend or path-dependency variables.
* **k / N:** Ratio of predictors to observations. *Lower is better.* Values above 0.1 may indicate overfitting.
* **Overfitting:** *FALSE is better.* Flags if the k/N ratio exceeds 0.1.
* **Separation:** *FALSE is better.* Flags potential complete or quasi-complete separation.
* **High-Z:** *FALSE is better.* Flags extremely high Z-scores ($|Z| > 15$), usually a symptom of numerical instability.
* **Max |Z|:** *Lower is better.* The maximum absolute Z-score among the predictors.
* **Converged:** *TRUE is better.* Indicates if the optimization algorithm reached a solution.
* **Theory Frac.:** *Higher is better (0–1).* Fraction of the non-intercept linear predictor attributable to the three theoretical channels: Actor Power, Institutional Quality, and their Interaction. Computed as |Actor Power + Inst. Quality + Interaction| / Σ|all non-intercept group contributions|. A value of 0.6 means 60% of the non-constant model variation is driven by theory rather than controls, time trend, path dependency, or regional fixed effects. Use this to compare how "theory-driven" each specification is relative to its own scale of prediction.

## Highlight Legend

* <span style="background-color: #c8e6c9; padding: 2px 4px; font-weight: bold;">Light Green</span>: **Best Model with All Three Theories** (Lowest AIC among models that have at least one significant interaction, one significant institutional quality, and one significant actor power term).
* <span style="background-color: #bbdefb; padding: 2px 4px; font-weight: bold;">Light Blue</span>: **Best Model with At Least One Theory** (Lowest AIC among models that have significant coefficients for at least one of those three theoretical clusters).
* <span style="background-color: #fff9c4; padding: 2px 4px; font-weight: bold;">Light Yellow</span>: **Best Model Overall** (Lowest AIC in the table).
* <span style="background-color: #ffebee; color: #c62828; padding: 2px 4px; font-weight: bold;">Red (Model Name)</span>: **Warning Flag** indicates a model suffering from Overfitting, Separation, abnormally High Z-scores, or failure to Converge.

*(Note: Ranking colors overwrite each other in priority. Green overwrites Blue and Yellow, Blue overwrites Yellow).*

# Average Marginal Effects (AME) Comparison {.tabset}

AMEs express each predictor's effect on the **probability scale**: how many percentage points does Pr(Adoption) change, on average across the sample, for a one-unit increase in that predictor? This complements AIC-based model selection by revealing whether theory variables have *practically meaningful* effect sizes — a model with low AIC but near-zero theory AMEs may be statistically preferred but theoretically hollow.

**Reading the table:** rows are theory-group predictors; columns are model specifications. Cells show AME ± SE. Bold = |AME| / SE > 1.96 (significant at 5%). Only the adoption stage is shown (logit); AMEs for the stringency stage (log-link GLM) are on the log scale and are not directly comparable.

```{r ame-comparison, echo=FALSE, results='asis'}
# ── helpers ────────────────────────────────────────────────────────────────────

# Colour-coded HTML cell. Fixes NA-SE display ("±NA" previously shown).
ame_cell_html <- function(ame, se, p, max_abs) {
  if (is.na(ame)) return('<span style="color:#bbb;font-size:0.85em;">—</span>')
  intensity <- min(abs(ame) / max(max_abs, 1e-6), 1)
  fade      <- round(255 - intensity * 185)
  bg        <- if (ame >= 0) sprintf("rgb(255,%d,%d)", fade, fade)
               else          sprintf("rgb(%d,%d,255)", fade, fade)
  sig       <- !is.na(p) && p < 0.05
  val       <- if (sig) sprintf("<b>%.4f</b>", ame) else sprintf("%.4f", ame)
  err  <- if (!is.na(se) && is.finite(se))
            sprintf('<span style="font-size:0.78em;color:#555;">&plusmn;%.4f</span>', se)
          else ""
  body <- if (nchar(err) > 0) paste0(val, "<br>", err) else val
  sprintf(
    '<div style="background:%s;padding:3px 5px;border-radius:4px;text-align:center;line-height:1.4;">%s</div>',
    bg, body
  )
}

# Render one transposed AME table (rows = models, cols = terms).
# No scroll_box — columns are narrow with wrapping headers to fit the page.
render_ame_table <- function(df_ss, term_filter_fn, caption_txt) {
  df_sub <- df_ss[term_filter_fn(df_ss), ]
  if (nrow(df_sub) == 0) {
    cat("> No matching predictors.\n\n"); return(invisible(NULL))
  }
  models_here <- unique(df_ss$Model)
  all_terms   <- unique(df_sub$term)

  grp_order <- c("Actor Power", "Inst. Quality", "Interaction", "Controls",
                 "Time Trend", "Path Dep.", "Region FE", "Other")
  term_grp  <- df_sub$term_group[match(all_terms, df_sub$term)]

  # For Region FE-only tables, sub-group columns by mapping type (H12 / OECD+ / 54 Regions)
  is_rfe_tab <- length(unique(term_grp)) == 1L && unique(term_grp) == "Region FE" &&
                "mapping_type" %in% colnames(df_sub)

  if (is_rfe_tab) {
    mapping_order <- c("H12", "OECD+", "54 Regions", "None", "Other")
    term_mapping  <- vapply(all_terms, function(t) {
      mt <- df_sub$mapping_type[df_sub$term == t & !is.na(df_sub$ame)]
      if (length(mt) == 0) return("Other")
      names(sort(table(mt), decreasing = TRUE))[1L]
    }, character(1))
    ord          <- order(match(term_mapping, mapping_order), all_terms)
    all_terms    <- all_terms[ord]
    term_grp     <- term_grp[ord]
    term_mapping <- term_mapping[ord]
    mt_runs      <- rle(term_mapping)
    header_spec  <- c(setNames(1L, " "),
                      setNames(as.list(mt_runs$lengths),
                               paste0(mt_runs$values, " regions")))
  } else {
    ord       <- order(match(term_grp, grp_order), all_terms)
    all_terms <- all_terms[ord]
    term_grp  <- term_grp[ord]
    grp_runs    <- rle(term_grp)
    header_spec <- c(setNames(1L, " "),
                     setNames(as.list(grp_runs$lengths), grp_runs$values))
  }

  max_abs <- max(abs(df_sub$ame), na.rm = TRUE)

  wide <- data.frame(Model = models_here, stringsAsFactors = FALSE)
  for (t in all_terms) {
    wide[[t]] <- vapply(models_here, function(mod) {
      row <- df_sub[df_sub$Model == mod & df_sub$term == t, ]
      if (nrow(row) == 0) return('<span style="color:#bbb;">—</span>')
      ame_cell_html(row$ame, row$se, row$p, max_abs)
    }, character(1))
  }

  # Readable HTML labels (dots → spaces, <br> wrapping)
  display_names <- c("Model", vapply(all_terms, clean_term, character(1)))

  n_data_cols <- length(all_terms)
  kb <- kable(wide, format = "html", escape = FALSE, caption = caption_txt,
              col.names = display_names) %>%
    kable_styling(
      bootstrap_options = c("condensed", "bordered"),
      font_size = 10, full_width = TRUE
    ) %>%
    add_header_above(header_spec, escape = FALSE) %>%
    # Wide model-name column; data columns let CSS handle wrapping
    column_spec(1, bold = TRUE, width = "22em",
                extra_css = "vertical-align:middle;") %>%
    column_spec(seq(2, n_data_cols + 1), width = "5em",
                extra_css = "min-width:4em;") %>%
    # Allow header cells to wrap (prevents horizontal overflow)
    row_spec(0, extra_css = "white-space:normal;word-break:break-word;vertical-align:bottom;padding:4px 3px;font-size:0.9em;")
  print(kb)
  cat("\n\n")
}

# ── main rendering ─────────────────────────────────────────────────────────────
if (length(all_ames) == 0) {
  cat("> No AME results available.\n")
} else {
  ame_all <- do.call(rbind, all_ames)
  sector_stages_ame <- unique(ame_all$Sector_Stage)
  sector_stages_ame <- sector_stages_ame[
    grepl("Adoption", sector_stages_ame, ignore.case = TRUE)
  ]

  for (sec_stg in sector_stages_ame) {
    cat("\n## ", sec_stg, " {.tabset}\n\n")
    df_ss <- ame_all[ame_all$Sector_Stage == sec_stg, ]

    # ── Tab 1: Core theory — Actor Power, Inst. Quality, API×IQ interactions ──
    cat("\n### Core Theory\n\n")
    cat("Actor Power drivers, Institutional Quality indicators, and their interactions",
        "with the API. Background = |AME| (red = positive, blue = negative).",
        "**Bold** = p < 0.05.\n\n")
    render_ame_table(
      df_ss,
      term_filter_fn = function(d) {
        is_rfe_interaction <- grepl("regionFE", d$term, ignore.case = TRUE) &
                             (grepl("_x_", d$term) | grepl(":", d$term))
        is_pure_rfe        <- d$term_group == "Region FE" &
                             !grepl("_x_|:", d$term)
        d$term != "(Intercept)" &
          !is_pure_rfe & !is_rfe_interaction &
          d$term_group %in% c("Actor Power", "Inst. Quality", "Interaction")
      },
      caption_txt = paste(sec_stg, "— Core Theory AMEs")
    )

    # ── Tab 2: Controls, time trend, path dependency ──────────────────────────
    cat("\n### Controls & Trend\n\n")
    render_ame_table(
      df_ss,
      term_filter_fn = function(d) {
        d$term_group %in% c("Controls", "Time Trend", "Path Dep.") |
          d$term == "(Intercept)"
      },
      caption_txt = paste(sec_stg, "— Controls & Trend AMEs")
    )

    # ── Tab 3: Pure region fixed effects (dummies) ────────────────────────────
    cat("\n### Region Fixed Effects\n\n")
    cat("AME for each region dummy: average change in Pr(Adoption) for a region",
        "belonging to that group vs. the reference region.\n\n")
    render_ame_table(
      df_ss,
      term_filter_fn = function(d) {
        d$term_group == "Region FE" & !grepl("_x_|:", d$term)
      },
      caption_txt = paste(sec_stg, "— Region FE AMEs")
    )

    # ── Tab 4: Region × API interaction terms (interactRegionFE models) ───────
    cat("\n### Region × API Interactions\n\n")
    cat("Present only in models with `interactRegionFE = TRUE`.",
        "Captures heterogeneous Actor Power effects across regions.\n\n")
    render_ame_table(
      df_ss,
      term_filter_fn = function(d) {
        grepl("regionFE", d$term, ignore.case = TRUE) &
          (grepl("_x_", d$term) | grepl(":", d$term))
      },
      caption_txt = paste(sec_stg, "— Region × API Interaction AMEs")
    )
  }
}
```
