---
title: "Progressive PFM Model Formulations and Improvements"
author: "PFM Modeling Output"
date: "`r Sys.Date()`"
html_document:
  toc: true
  toc_float:
    collapsed: false
  toc_depth: 3
  theme: flatly
  highlight: tango
  number_sections: true
  code_folding: hide
---

<style>
  .main-container {
    max-width: 95% !important;
  }
  .toc-content {
    padding-left: 0px !important;
  }
  .svg-container {
    margin: auto !important;
  }
</style>


```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, dev='svglite',  fig.ext = ".svg", fig.width = 12, fig.height = 8, dpi=100)

library(pfm)
library(magclass)
library(ggplot2)
library(dplyr)
library(knitr)
library(kableExtra)

# Configure madrat
madrat::setConfig(
  forcecache = TRUE,
  cachefolder = "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/cache/default",
  mappingfolder = "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/mappings"
)

options(pfm.modelDir = "../../../models/")

```


```{r defining_models, include=FALSE}

# Global Model Configurations
model_configs <- list(
  list(
    name = "Baseline",
    description = "**What it does:** Uses a single composite Actor Power Index.\n\n**Assumption:** Actor power operates symmetrically (a 1-unit increase in Innovator Power perfectly cancels out a 1-unit increase in Incumbent Power). Institutional quality interacts globally with this net power.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE
  ),
  list(
    name = "Split Variables",
    description = "**What it does:** Replaces the combined index with discrete components.\n\n**Assumption:** Incumbents and Innovators exert asymmetric power dynamics over policy adoption.",
    actorPowerDrivers = c("Innovator Power", "Incumbent Power"), 
    actorPowerIndex = c("Innovator Power", "Incumbent Power"),
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE
  ),
  list(
    name = "Advanced Controls",
    description = "**What it does:** Expands the structural controls.\n\n**Assumption:** Broader macroeconomic and demographic factors drive feasibility.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = c("GDP per Capita", "Energy Intensity", "Urban Population Share"),
    includeLagged = FALSE, interactRegionFE = FALSE
  ),
  list(
    name = "Non-Linear Dynamics",
    description = "**What it does:** Includes a squared term for GDP per Capita.\n\n**Assumption:** The Environmental Kuznets Curve hypothesis applies to pricing adoption.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = c("GDP per Capita Sq"),
    includeLagged = FALSE, interactRegionFE = FALSE
  ),
  list(
    name = "Heterogeneous Regional Effects",
    description = "**What it does:** Interacts Region Fixed Effects with structural drivers.\n\n**Assumption:** The impact of actor power and institutional quality varies intrinsically by region.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = TRUE
  ),
  list(
    name = "Path Dependency",
    description = "**What it does:** Introduces lagged adoption/price stringency.\n\n**Assumption:** Policy adoption strongly correlates with historical status (inertia).",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = "Government Effectiveness (WGI)", 
    controlDrivers = "GDP per Capita",
    includeLagged = TRUE, interactRegionFE = FALSE
  ),
  list(
    name = "Institutional Quality",
    description = "**What it does:** Introduces VDem institutional quality indicators.\n\n**Assumption:** Test alternative institutional quality indicators.",
    actorPowerDrivers = "Actor Power Index", 
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = c("Government Effectiveness (WGI)", "Rule of Law (VDem)", "Vertical Accountability (VDem)", "Horizontal Accountability (VDem)", "Diagonal Accountability (VDem)"), 
    controlDrivers = "GDP per Capita",
    includeLagged = FALSE, interactRegionFE = FALSE
  )
)


```


```{r loading data, include=FALSE}

# 1. Load Data
panelData <- panelDataHistorical(
  aggregate = TRUE, y = 2000:2022,
  outputRegionMappingFile = "regionmapping_54.csv"
)

# 2. Add squared GDP term
panelData <- mbind(
  panelData,
  setNames(panelData[, , "GDP per Capita"]^2, "GDP per Capita Sq")
)


```


```{r auxiliar functions, include=FALSE}

# Global list to track diagnostics for the final comparison table
all_aics <- data.frame()

add_aic <- function(fit, sector_stage, model_name) {
  m <- fit$model
  is_logistf <- inherits(m, "logistf")
  k <- length(coef(m))
  
  # N obs
  nObs <- if (is_logistf) m$n else nobs(m)
  
  # Log-likelihood
  loglik <- if (is_logistf) as.numeric(m$loglik["full"]) else as.numeric(logLik(m))
  
  # ICs
  aic <- if (is_logistf) -2 * loglik + 2 * k else m$aic
  bic <- if (is_logistf) -2 * loglik + k * log(nObs) else BIC(m)
  aicc <- if (!is.na(aic) && nObs > k + 1) aic + (2 * k * (k + 1)) / (nObs - k - 1) else NA
  hqic <- if (!is.na(loglik) && nObs > 0) -2 * loglik + 2 * k * log(log(nObs)) else NA
  
  # Pseudo-R2
  pseudoR2 <- if (is_logistf) {
    if (is.null(m$loglik)) NA else round(1 - (m$loglik["full"] / m$loglik["null"]), 3)
  } else if (!is.null(m$deviance) && !is.null(m$null.deviance)) {
    round(1 - (m$deviance / m$null.deviance), 3)
  } else NA
  
  # Coef testing
  pvals <- if (!is.null(fit$coeftest)) fit$coeftest[, 4] else NULL
  zvals <- if (!is.null(fit$coeftest)) fit$coeftest[, 3] else NULL
  nSig <- if (!is.null(pvals)) sum(pvals < 0.05, na.rm = TRUE) else 0
  maxAbsZ <- if (!is.null(zvals)) max(abs(zvals), na.rm = TRUE) else NA
  highZ <- if (!is.na(maxAbsZ)) maxAbsZ > 15 else NA
  
  # Flags
  converged <- if (is_logistf) !is.null(m$coefficients) && !any(is.na(m$coefficients)) else m$converged
  kOverN <- k / nObs
  overfitting <- kOverN > 0.1
  separation <- highZ # Approximation for separation based on extreme Z
  
  # Compile row
  row_df <- data.frame(
    Sector_Stage = sector_stage,
    Model = model_name,
    AIC = round(as.numeric(aic), 1),
    AICc = round(as.numeric(aicc), 1),
    BIC = round(as.numeric(bic), 1),
    HQIC = round(as.numeric(hqic), 1),
    `Log-lik` = round(as.numeric(loglik), 1),
    `Pseudo-R2` = as.numeric(pseudoR2),
    `N obs` = nObs,
    `N predictors` = k,
    `N significant` = nSig,
    `k / N` = round(kOverN, 3),
    Overfitting = overfitting,
    Separation = separation,
    `High-Z` = highZ,
    `Max |Z|` = round(as.numeric(maxAbsZ), 1),
    Converged = converged,
    check.names = FALSE,
    stringsAsFactors = FALSE
  )
  
  all_aics <<- rbind(all_aics, row_df)
}

# Load Shared Helpers for plotting
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/theme.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/forestPlot.R"))

# Wrapper to use the shared rendering function
plot_forest <- function(fit, title = "Coefficient Forest Plot") {
  if (is.null(fit) || is.null(fit$coeftest)) return(NULL)
  
  df <- pfm::coeftestToDataFrame(fit$coeftest)
  
  if(nrow(df) == 0) return(NULL)
  
  render_forest_plot(df, title = title)
}
```

# Introduction

This report details a progressive testing approach to improve the Political Feasibility Module (PFM) two-stage hurdle model. We begin with the standard baseline formulation and add progressive econometric improvements. Success mterics are presented in the Summary and Best Model Choice tab..


```{r model-specs, echo=FALSE}
library(tibble)
library(purrr)

model_specs <- map_df(model_configs, ~ tibble(
  Model = .x$name,
  actorPowerDrivers = paste(.x$actorPowerDrivers, collapse = "<br>"),
  actorPowerIndex = paste(.x$actorPowerIndex, collapse = "<br>"),
  instQualityDrivers = paste(.x$instQualityDrivers, collapse = "<br>"),
  controlDrivers = paste(.x$controlDrivers, collapse = "<br>"),
  `includeLagged*` = as.character(.x$includeLagged),
  interactRegionFE = as.character(.x$interactRegionFE)
))

kable(model_specs, format = "html", escape = FALSE, caption = "Summary of Tested Model Configurations") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "bordered"), full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = TRUE)

```

---

# Models {.tabset}

```{r dynamic-models, results='asis', echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=5}
sectors <- c("Diffuse", "Bulk")
stages <- c("Adoption", "Stringency")

for (stg in stages) {
  for (sec in sectors) {
    cat(sprintf("\n## %s: %s Sector\n\n", stg, sec))
    
    for (i in seq_along(model_configs)) {
      cfg <- model_configs[[i]]
      
      cat(sprintf("\n### %s. %s\n\n", i, cfg$name))
      cat(cfg$description, "\n\n")
      
      if (stg == "Adoption") {
        fit <- estimateAdoptionModel(
          data = panelData, sector = sec,
          actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
          instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
          includeLaggedAdoption = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE
        )
      } else {
        fit <- estimatePriceStringencyModel(
          data = panelData, sector = sec,
          actorPowerDrivers = cfg$actorPowerDrivers, actorPowerIndex = cfg$actorPowerIndex,
          instQualityDrivers = cfg$instQualityDrivers, controlDrivers = cfg$controlDrivers,
          includeLaggedECP = cfg$includeLagged, interactRegionFE = cfg$interactRegionFE
        )
      }
      
      # Track AIC
      stage_label <- paste0(stg, ": ", sec)
      add_aic(fit, stage_label, paste0(i, ". ", cfg$name))
      
      # Plot
      plot_title <- paste0(stg, " ", sec, " - ", cfg$name)
      p <- plot_forest(fit, plot_title)
      
      if (!is.null(p)) {
        cat("\n\n")
        print(p)
        cat("\n\n")
      }
    }
  }
}
```

## Summary and Best Model Choice

By stepping through these configurations, we systematically evaluate the structural foundations of the Hurdle Model.

```{r summary-table, echo=FALSE, results='asis'}
# Split the data into separate tables for each stage/sector combination
for (sec_stg in unique(all_aics$Sector_Stage)) {
  sub_df <- all_aics %>%
    filter(Sector_Stage == sec_stg) %>%
    arrange(Model) %>%
    select(-Sector_Stage) # Drop redundant column
    
  # Identify best values for styling
  best_aic  <- min(sub_df$AIC, na.rm = TRUE)
  best_aicc <- min(sub_df$AICc, na.rm = TRUE)
  best_bic  <- min(sub_df$BIC, na.rm = TRUE)
  best_hqic <- min(sub_df$HQIC, na.rm = TRUE)
  best_ll   <- max(sub_df$`Log-lik`, na.rm = TRUE)
  best_r2   <- max(sub_df$`Pseudo-R2`, na.rm = TRUE)
  best_nsig <- max(sub_df$`N significant`, na.rm = TRUE)
    
  # Build base table with scroll_box for wide columns
  k <- kable(sub_df, format = "html", caption = paste("Model Diagnostics:", sec_stg)) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), font_size = 11)
    
  # Apply conditional highlighting for the most important metric columns
  k <- k %>%
    column_spec(2, background = ifelse(sub_df$AIC == best_aic & !is.na(sub_df$AIC), "#e8f8f5", "white"), color = ifelse(sub_df$AIC == best_aic & !is.na(sub_df$AIC), "#117a65", "black"), bold = ifelse(sub_df$AIC == best_aic & !is.na(sub_df$AIC), TRUE, FALSE)) %>%
    column_spec(3, background = ifelse(sub_df$AICc == best_aicc & !is.na(sub_df$AICc), "#e8f8f5", "white"), color = ifelse(sub_df$AICc == best_aicc & !is.na(sub_df$AICc), "#117a65", "black"), bold = ifelse(sub_df$AICc == best_aicc & !is.na(sub_df$AICc), TRUE, FALSE)) %>%
    column_spec(4, background = ifelse(sub_df$BIC == best_bic & !is.na(sub_df$BIC), "#e8f8f5", "white"), color = ifelse(sub_df$BIC == best_bic & !is.na(sub_df$BIC), "#117a65", "black"), bold = ifelse(sub_df$BIC == best_bic & !is.na(sub_df$BIC), TRUE, FALSE)) %>%
    column_spec(5, background = ifelse(sub_df$HQIC == best_hqic & !is.na(sub_df$HQIC), "#e8f8f5", "white"), color = ifelse(sub_df$HQIC == best_hqic & !is.na(sub_df$HQIC), "#117a65", "black"), bold = ifelse(sub_df$HQIC == best_hqic & !is.na(sub_df$HQIC), TRUE, FALSE)) %>%
    column_spec(7, background = ifelse(sub_df$`Pseudo-R2` == best_r2 & !is.na(sub_df$`Pseudo-R2`), "#e8f8f5", "white"), color = ifelse(sub_df$`Pseudo-R2` == best_r2 & !is.na(sub_df$`Pseudo-R2`), "#117a65", "black"), bold = ifelse(sub_df$`Pseudo-R2` == best_r2 & !is.na(sub_df$`Pseudo-R2`), TRUE, FALSE)) %>%
    scroll_box(width = "100%")
      
  # Render the table
  cat("\n### ", sec_stg, "\n\n")
  print(k)
  cat("\n<br>\n")
}
```

### Key Conclusions
1. **Split Variables** better capture the asymmetric lobbying pressures of Incumbents vs. Innovators.
2. **Advanced Controls** (Energy Intensity, Urban Share) consistently reveal cleaner estimates by controlling for macro-structural confounders.
3. **Heterogeneous Region Effects** (`interactRegionFE`) show that Actor Power operates fundamentally differently inside the EU compared to the OECD or the rest of the world.
4. **Path Dependency** (`lagged_adoption` and `lagged_ecp`) provides a massive AIC drop, proving that political inertia is the single most powerful predictor of policy status.

The **Fully Advanced Model (Model 6: Path Dependency)** drastically outperforms the Baseline and represents the strongest overall specification for both Adoption and Stringency.
