---
title: "`r params$modelName` — PFM Stringency Model Report"
subtitle: "Price Stringency Estimation, Historical Fit, and Scenario Projections for Bulk & Diffuse Sectors"
author: "Political Feasibility Module"
date: "`r Sys.Date()`"
params:
  cacheDir: "C:/Users/renatoro/Desktop/Input Data/remind_inputdata/cache/default"
  gdxPath: "c:/Users/renatoro/Desktop/Projects/Elevate/code/fulldata.gdx"
  assetDir: "output/stringency_model"
  modelName: "Baseline + Rule of Law"
  reportName: "default"
  instQualityDriversBulk: !r c("Government Effectiveness (WGI)", "Rule of Law (VDem)")
  controlDriversBulk: !r c("GDP per Capita", "Hydro Nuclear Share")
  regionMappingFEBulk: "regionmappingH12.csv"
  includeLaggedBulk: false
  logisticTimeTrendBulk: false
  logisticTimeTrendDiffuse: ~
  useMundlakBulk: false
  instQualityDriversDiffuse: ~
  controlDriversDiffuse: ~
  regionMappingFEDiffuse: ~
  includeLaggedDiffuse: ~
  useMundlakDiffuse: ~
  actorPowerDriversBulk: ~
  actorPowerIndexBulk: "Actor Power Index"
  actorPowerDriversDiffuse: ~
  actorPowerIndexDiffuse: ~
  adoptionInstQualityDrivers: !r c("Government Effectiveness (WGI)", "Rule of Law (VDem)")
  adoptionControlDrivers: !r c("GDP per Capita", "Hydro Nuclear Share")
  adoptionRegionMappingFE: "regionmappingH12.csv"
  adoptionThreshold: 0.5
  adoptionThresholdBulk: 0.30
  adoptionThresholdDiffuse: 0.17
  snapshotYears: !r c(2030L, 2040L, 2050L, 2070L, 2100L)
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
    theme: flatly
    highlight: tango
    number_sections: true
    code_folding: hide
    fig_width: 10
    fig_height: 6
---

<style>
  body { background-color: #f5f7fa; color: #333; }
  .main-container { max-width: 95% !important; }
  .toc-content { padding-left: 0px !important; }
  h1, h2, h3 { color: #2C3E50; font-weight: 700; }
  .nav-tabs { font-weight: bold; margin-bottom: 20px; border-bottom: 2px solid #2C3E50; }
  .nav-tabs > li.active > a,
  .nav-tabs > li.active > a:focus,
  .nav-tabs > li.active > a:hover {
    background-color: #2C3E50; color: #fff;
    border: 1px solid #2C3E50; border-bottom-color: transparent;
  }
  .tab-content {
    padding: 25px; border: 1px solid #ddd; border-top: none;
    border-radius: 0 0 6px 6px; background-color: #fff;
    box-shadow: 0 4px 6px rgba(0,0,0,0.05);
  }
  .card { background-color: #fff; border-radius: 8px; border: 1px solid #eef2f5;
          box-shadow: 0 4px 6px rgba(0,0,0,0.02); margin-bottom: 25px; }
  .card-header { background-color: #f8fafc; border-bottom: 1px solid #eef2f5;
                 padding: 12px 20px; font-weight: bold; font-size: 1.1em; color: #2C3E50; }
  .card-body { padding: 20px; }
  /* Sector selector bar */
  #sectorBar {
    position: sticky; top: 0; z-index: 9999;
    background: #2C3E50; color: white;
    padding: 8px 24px; display: flex; align-items: center; gap: 12px;
    box-shadow: 0 2px 8px rgba(0,0,0,0.3);
    margin-bottom: 10px;
  }
  .sector-btn {
    padding: 5px 22px; border-radius: 4px; font-weight: 700;
    cursor: pointer; font-size: 0.9em; transition: all 0.15s;
  }
  .sector-btn.active {
    background: white !important; color: #2C3E50 !important; border: none !important;
  }
  .sector-btn.inactive {
    background: rgba(255,255,255,0.15) !important; color: white !important;
    border: 1px solid rgba(255,255,255,0.5) !important;
  }
  .chart-description {
    background: #f8fafc; border-left: 4px solid #2C3E50;
    padding: 10px 16px; margin: 10px 0 18px 0;
    font-size: 0.93em; color: #444; line-height: 1.55;
  }
</style>

<div id="sectorBar">
  <span style="font-weight:700;font-size:0.95em;margin-right:4px;">Sector:</span>
  <button id="btn-bulk" class="sector-btn active" onclick="setSector('Bulk')">Bulk</button>
  <button id="btn-diffuse" class="sector-btn inactive" onclick="setSector('Diffuse')">Diffuse</button>
  <span style="font-size:0.8em;opacity:0.65;margin-left:8px;">Click to show only Bulk or Diffuse results throughout the report</span>
</div>

<script>
(function() {
  window.setSector = function(sector) {
    // Update button styles
    document.getElementById('btn-bulk').className    = 'sector-btn ' + (sector === 'Bulk'    ? 'active' : 'inactive');
    document.getElementById('btn-diffuse').className = 'sector-btn ' + (sector === 'Diffuse' ? 'active' : 'inactive');
    // Show / hide sector content blocks directly via CSS
    document.querySelectorAll('.bulk-only').forEach(function(el) {
      el.style.display = sector === 'Bulk' ? '' : 'none';
    });
    document.querySelectorAll('.diffuse-only').forEach(function(el) {
      el.style.display = sector === 'Diffuse' ? '' : 'none';
    });

    // Toggle tab header visibility for duplicate tabs
    var snapshotTabs = [];
    var rankingTabs = [];
    document.querySelectorAll('.nav-tabs li').forEach(function(li) {
      var a = li.querySelector('a');
      if (a) {
        var text = (a.textContent || '').trim();
        // Remove numbers like "8.3 " or "8.4 " to match the name
        var cleanText = text.replace(/^[0-9.]+\s+/, '');
        if (cleanText === 'Snapshot Maps') {
          snapshotTabs.push(li);
        } else if (cleanText === 'Regional Price Ranking') {
          rankingTabs.push(li);
        }
      }
    });

    var updateTabGroup = function(tabsGroup) {
      if (tabsGroup.length === 2) {
        var firstTab = tabsGroup[0];
        var secondTab = tabsGroup[1];
        if (sector === 'Bulk') {
          firstTab.style.display = '';
          secondTab.style.display = 'none';
          if (secondTab.classList.contains('active')) {
            secondTab.classList.remove('active');
            firstTab.querySelector('a').click();
          }
        } else {
          firstTab.style.display = 'none';
          secondTab.style.display = '';
          if (firstTab.classList.contains('active')) {
            firstTab.classList.remove('active');
            secondTab.querySelector('a').click();
          }
        }
      }
    };

    updateTabGroup(snapshotTabs);
    updateTabGroup(rankingTabs);
  };
  // Run after the full page (including rmarkdown's buildTabsets) has loaded
  window.addEventListener('load', function() { setSector('Bulk'); });
})();
</script>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE, warning = FALSE, message = FALSE,
  fig.align = "center", dev = "svg"
)

library(pfm)
library(magclass)
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)
library(scales)
library(ggrepel)
library(patchwork)

source(file.path(rprojroot::find_rstudio_root_file(), "src/r/configHelper.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/theme.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/modelHelpers.R"))
source(file.path(rprojroot::find_rstudio_root_file(), "src/r/forestPlot.R"))

if (!is.null(params$cacheDir) && nchar(params$cacheDir) > 0) {
  madrat::setConfig(
    forcecache    = TRUE,
    cachefolder   = params$cacheDir,
    mappingfolder = gsub("cache/default", "mappings", params$cacheDir)
  )
}

.model_store <- file.path(rprojroot::find_rstudio_root_file(),
                           getPfmConfig("modelDir", "cache"))
dir.create(.model_store, showWarnings = FALSE, recursive = TRUE)
options(pfm.modelDir = .model_store)

plotOutputDir <- file.path(params$assetDir, "plots")
if (!dir.exists(plotOutputDir)) dir.create(plotOutputDir, recursive = TRUE)

# ── Per-sector config resolution (Diffuse falls back to Bulk if NULL) ──────────
resolve_sector <- function(diffuse_val, bulk_val) {
  if (is.null(diffuse_val) || (is.character(diffuse_val) && length(diffuse_val) == 0))
    bulk_val else diffuse_val
}

resolve_actor_index <- function(val, default = "Actor Power Index") {
  if (is.null(val) || !nzchar(as.character(val))) default else val
}

cfg <- list(
  Bulk = list(
    instQuality       = params$instQualityDriversBulk,
    controls          = params$controlDriversBulk,
    regionFE          = if (is.null(params$regionMappingFEBulk) ||
                            nchar(as.character(params$regionMappingFEBulk)) == 0)
                          NULL else params$regionMappingFEBulk,
    lagged            = isTRUE(params$includeLaggedBulk),
    logisticTimeTrend = isTRUE(params$logisticTimeTrendBulk),
    useMundlak        = isTRUE(params$useMundlakBulk),
    actorPower        = params$actorPowerDriversBulk,
    actorIndex        = resolve_actor_index(params$actorPowerIndexBulk)
  ),
  Diffuse = list(
    instQuality       = resolve_sector(params$instQualityDriversDiffuse, params$instQualityDriversBulk),
    controls          = resolve_sector(params$controlDriversDiffuse,     params$controlDriversBulk),
    regionFE          = {
      raw <- resolve_sector(params$regionMappingFEDiffuse, params$regionMappingFEBulk)
      if (is.null(raw) || nchar(as.character(raw)) == 0) NULL else raw
    },
    lagged            = isTRUE(resolve_sector(params$includeLaggedDiffuse,        params$includeLaggedBulk)),
    logisticTimeTrend = isTRUE(resolve_sector(params$logisticTimeTrendDiffuse,    params$logisticTimeTrendBulk)),
    useMundlak        = isTRUE(resolve_sector(params$useMundlakDiffuse,           params$useMundlakBulk)),
    actorPower        = resolve_sector(params$actorPowerDriversDiffuse, params$actorPowerDriversBulk),
    actorIndex        = resolve_actor_index(resolve_sector(params$actorPowerIndexDiffuse, params$actorPowerIndexBulk))
  )
)

adp_cfg <- list(
  instQuality = params$adoptionInstQualityDrivers,
  controls    = params$adoptionControlDrivers,
  regionFE    = if (is.null(params$adoptionRegionMappingFE) ||
                    nchar(as.character(params$adoptionRegionMappingFE)) == 0)
                  NULL else params$adoptionRegionMappingFE
)

adoption_threshold         <- as.numeric(params$adoptionThreshold)
adoption_threshold_bulk    <- as.numeric(params$adoptionThresholdBulk    %||% params$adoptionThreshold)
adoption_threshold_diffuse <- as.numeric(params$adoptionThresholdDiffuse %||% params$adoptionThreshold)
snapshot_years     <- as.integer(params$snapshotYears)
key_regions        <- c("USA", "CHA", "DEU", "IND", "JPN", "BRA", "RUS", "ZAF")
LOG_TRANSFORM      <- TRUE  # always TRUE — matches estimatePriceStringencyModel default

back_transform <- function(x) exp(x) - 1

# ── Resolve GDX path ───────────────────────────────────────────────────────────
gdx_path   <- params$gdxPath
if (!is.null(gdx_path) && nchar(gdx_path) > 0 && !is_absolute_path(gdx_path))
  gdx_path <- file.path(rprojroot::find_rstudio_root_file(), gdx_path)
has_scenario <- !is.null(gdx_path) && nchar(gdx_path) > 0 && file.exists(gdx_path)

tidy_magpie <- function(mag_obj, period_label = "Historical") {
  if (is.null(mag_obj)) return(NULL)
  df      <- as.data.frame(mag_obj)
  reg_col <- if ("Region" %in% colnames(df)) "Region" else colnames(df)[1]
  yr_col  <- if ("Year"   %in% colnames(df)) "Year"   else colnames(df)[2]
  var_col <- if ("Data1"  %in% colnames(df)) "Data1"  else colnames(df)[3]
  val_col <- if ("Value"  %in% colnames(df)) "Value"  else colnames(df)[4]
  data.frame(
    region   = trimws(as.character(df[[reg_col]])),
    year     = as.integer(gsub("y", "", as.character(df[[yr_col]]))),
    variable = trimws(as.character(df[[var_col]])),
    value    = as.numeric(df[[val_col]]),
    period   = period_label,
    stringsAsFactors = FALSE
  )
}
```

```{r load_data, cache=FALSE, include=FALSE}
panelData <- getPanelDataHistoricalCached(
  aggregate = TRUE, y = 2000:2022,
  outputRegionMappingFile = "regionmapping_54.csv"
)

df_hist  <- tidy_magpie(panelData, "Historical")
future_mag <- NULL; df_scen <- NULL

if (has_scenario) {
  future_mag <- getPanelDataScenarioCached(
    gdxFile = gdx_path, aggregate = TRUE,
    outputRegionMappingFile = "regionmapping_54.csv"
  )
  df_scen <- tidy_magpie(future_mag, "Scenario")
}
df_all <- do.call(rbind, Filter(Negate(is.null), list(df_hist, df_scen)))
```

```{r fit_models, cache=FALSE, include=FALSE}
# ── Stringency models ──────────────────────────────────────────────────────────
fit_str <- list(
  Bulk = estimatePriceStringencyModel(
    data                      = panelData, sector = "Bulk",
    actorPowerDrivers         = cfg$Bulk$actorPower,
    actorPowerIndex           = cfg$Bulk$actorIndex,
    instQualityDrivers        = cfg$Bulk$instQuality,
    controlDrivers            = cfg$Bulk$controls,
    includeLaggedECP          = cfg$Bulk$lagged,
    regionMappingFixedEffects = if (isTRUE(cfg$Bulk$useMundlak)) NULL else cfg$Bulk$regionFE,
    logTransform              = LOG_TRANSFORM,
    logisticTimeTrend         = isTRUE(cfg$Bulk$logisticTimeTrend),
    useMundlak                = isTRUE(cfg$Bulk$useMundlak)
  ),
  Diffuse = estimatePriceStringencyModel(
    data                      = panelData, sector = "Diffuse",
    actorPowerDrivers         = cfg$Diffuse$actorPower,
    actorPowerIndex           = cfg$Diffuse$actorIndex,
    instQualityDrivers        = cfg$Diffuse$instQuality,
    controlDrivers            = cfg$Diffuse$controls,
    includeLaggedECP          = cfg$Diffuse$lagged,
    regionMappingFixedEffects = if (isTRUE(cfg$Diffuse$useMundlak)) NULL else cfg$Diffuse$regionFE,
    logTransform              = LOG_TRANSFORM,
    logisticTimeTrend         = isTRUE(cfg$Diffuse$logisticTimeTrend),
    useMundlak                = isTRUE(cfg$Diffuse$useMundlak)
  )
)

# ── Adoption models (for projection gating only) ───────────────────────────────
fit_adp <- list(
  Bulk = estimateAdoptionModel(
    data                      = panelData, sector = "Bulk",
    actorPowerDrivers         = NULL,
    actorPowerIndex           = "Actor Power Index",
    instQualityDrivers        = adp_cfg$instQuality,
    controlDrivers            = adp_cfg$controls,
    regionMappingFixedEffects = adp_cfg$regionFE,
    compute                   = c(ame = FALSE, predictedProbs = FALSE)
  ),
  Diffuse = estimateAdoptionModel(
    data                      = panelData, sector = "Diffuse",
    actorPowerDrivers         = NULL,
    actorPowerIndex           = "Actor Power Index",
    instQualityDrivers        = adp_cfg$instQuality,
    controlDrivers            = adp_cfg$controls,
    regionMappingFixedEffects = adp_cfg$regionFE,
    compute                   = c(ame = FALSE, predictedProbs = FALSE)
  )
)

# ── Prepare future panel data ──────────────────────────────────────────────────
# Driver standardization (ADR 0007) must reuse the mean/sd FROZEN on the
# historical panel; otherwise the scenario design is standardized on its own
# scale and the historical-frozen coefficients no longer apply (biased/exploding
# prices, seam discontinuity). We recompute the frozen scaling from the same
# historical panel the models were fit on and pass it via `driverScaling=`.
future_df <- list(Bulk = NULL, Diffuse = NULL)
if (!is.null(future_mag)) {
  for (sec in c("Bulk", "Diffuse")) {
    ap  <- union(cfg[[sec]]$actorPower, NULL)
    ai  <- union(cfg[[sec]]$actorIndex, "Actor Power Index")
    iq  <- union(cfg[[sec]]$instQuality, adp_cfg$instQuality)
    ctl <- union(cfg[[sec]]$controls, adp_cfg$controls)
    fe  <- if (isTRUE(cfg[[sec]]$useMundlak)) NULL else cfg[[sec]]$regionFE
    histPrep <- preparePanelData(
      data = panelData, sector = sec,
      actorPowerDrivers = ap, actorPowerIndex = ai,
      instQualityDrivers = iq, controlDrivers = ctl,
      regionMappingFixedEffects = fe, useMundlak = isTRUE(cfg[[sec]]$useMundlak), lag = 1
    )
    future_df[[sec]] <- preparePanelData(
      data = future_mag, sector = sec,
      actorPowerDrivers = ap, actorPowerIndex = ai,
      instQualityDrivers = iq, controlDrivers = ctl,
      regionMappingFixedEffects = fe, useMundlak = isTRUE(cfg[[sec]]$useMundlak), lag = 1,
      driverScaling = attr(histPrep, "driverScaling"),
      trendFreezeYear = max(magclass::getYears(panelData, as.integer = TRUE))
    )
  }
}
```

```{r helper_functions, include=FALSE}
# ── Fit statistics ─────────────────────────────────────────────────────────────
compute_str_stats <- function(fit) {
  if (is.null(fit)) return(NULL)
  m    <- fit$model
  k    <- length(coef(m))
  nObs <- nobs(m)
  ll   <- as.numeric(logLik(m))
  r2   <- if (!is.null(m$deviance) && !is.null(m$null.deviance))
            round(1 - m$deviance / m$null.deviance, 3) else NA
  rmse_log <- sqrt(mean(residuals(m, type = "response")^2, na.rm = TRUE))
  list(k = k, nObs = nObs, loglik = round(ll, 2),
       aic = round(AIC(m), 1), bic = round(BIC(m), 1),
       r2 = r2, rmse_log = rmse_log,
       converged = isTRUE(m$converged))
}

# ── Semi-elasticity: exp(beta)-1 with delta-method SE ─────────────────────────
compute_semi_elasticity <- function(fit, instQualityDrivers = NULL, controlDrivers = NULL) {
  if (is.null(fit) || is.null(fit$coeftest)) return(NULL)
  ct   <- fit$coeftest
  beta <- ct[, 1]; se_beta <- ct[, 2]
  semi <- exp(beta) - 1
  se_s <- exp(beta) * se_beta
  df   <- data.frame(
    term         = rownames(ct),
    estimate     = beta,
    semi_elas    = semi,
    se           = se_s,
    lower        = semi - 1.96 * se_s,
    upper        = semi + 1.96 * se_s,
    z            = ct[, 3],
    p            = ct[, 4],
    stringsAsFactors = FALSE
  )
  # Classify term group
  df$group <- classifyTermGroups(
    df$term,
    actorPowerDrivers  = NULL,
    actorPowerIndex    = "Actor Power Index",
    instQualityDrivers = instQualityDrivers,
    controlDrivers     = controlDrivers
  )
  df
}

# ── MEM in USD/tCO2: beta * mean(fitted price) ────────────────────────────────
compute_mem <- function(fit) {
  if (is.null(fit) || is.null(fit$coeftest)) return(NULL)
  m             <- fit$model
  mean_log      <- mean(fitted(m), na.rm = TRUE)
  mean_price    <- back_transform(mean_log)
  ct            <- fit$coeftest
  beta          <- ct[, 1]; se_beta <- ct[, 2]
  mem           <- beta * mean_price
  se_mem        <- se_beta * mean_price
  data.frame(
    term       = rownames(ct),
    mem        = mem,
    se         = se_mem,
    lower      = mem - 1.96 * se_mem,
    upper      = mem + 1.96 * se_mem,
    mean_price = mean_price,
    stringsAsFactors = FALSE
  )
}

# ── 45° Calibration Scatter ────────────────────────────────────────────────────
# ── Shared calibration data extractor ────────────────────────────────────────
get_calib_data <- function(fit, outlier_sd = 2) {
  if (is.null(fit)) return(NULL)
  m <- fit$model
  obs <- m$y; hat <- fitted(m)
  if (is.null(obs) || length(obs) == 0) return(NULL)
  obs_usd <- back_transform(obs); hat_usd <- back_transform(hat)

  # Get region/year from fit$data when available (correct labels; handles cached models)
  data_full <- fit$data
  if (!is.null(data_full) && "region" %in% colnames(data_full) && "year" %in% colnames(data_full)) {
    na_idx  <- as.integer(attr(m$model, "na.action"))
    ok_idx  <- if (length(na_idx) > 0) setdiff(seq_len(nrow(data_full)), na_idx) else seq_len(nrow(data_full))
    region_col <- as.character(data_full$region[ok_idx])
    year_col   <- as.integer(data_full$year[ok_idx])
  } else {
    train <- m$model
    region_col <- if ("region" %in% colnames(train)) as.character(train$region)
                  else rep(NA_character_, nrow(train))
    year_col   <- if ("timeTrend" %in% colnames(train))
                    as.integer(train$timeTrend + 1999L)
                  else seq_along(obs)
  }

  df <- data.frame(obs = obs_usd, hat = hat_usd, resid = obs - hat,
                   region = region_col, year = year_col, stringsAsFactors = FALSE)
  resid_sd <- sd(df$resid, na.rm = TRUE)
  df$outlier <- abs(df$resid) > outlier_sd * resid_sd
  list(df = df, resid_sd = resid_sd, outlier_sd = outlier_sd)
}

make_calibration_scatter <- function(fit, sector_name, outlier_sd = 2) {
  calib <- get_calib_data(fit, outlier_sd)
  if (is.null(calib)) return(NULL)
  df_plot <- calib$df; resid_sd <- calib$resid_sd
  n_out   <- sum(df_plot$outlier, na.rm = TRUE)

  # Axis limits based on NON-OUTLIER range so the main distribution fills the chart
  df_in  <- df_plot[!df_plot$outlier, ]
  ax_max <- max(c(df_in$obs, df_in$hat), na.rm = TRUE) * 1.08

  band_x  <- seq(0, ax_max, length.out = 200)
  band_df <- data.frame(
    x    = band_x,
    y_lo = (band_x + 1) * exp(-outlier_sd * resid_sd) - 1,
    y_hi = (band_x + 1) * exp( outlier_sd * resid_sd) - 1
  )
  col_sec  <- if (sector_name == "Bulk") col_bulk else col_diffuse
  sub_note <- if (n_out > 0)
    sprintf("45° = perfect fit | Shaded = ±%g log-SD | %d outlier(s) excluded from view — see table below", outlier_sd, n_out)
  else
    sprintf("45° = perfect fit | Shaded = ±%g log-SD | No outliers", outlier_sd)

  ggplot(df_in, aes(x = hat, y = obs)) +
    geom_ribbon(data = band_df, aes(x = x, ymin = y_lo, ymax = y_hi),
                inherit.aes = FALSE, fill = col_sec, alpha = 0.12) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "grey40", linewidth = 0.8) +
    geom_point(aes(text = paste0(
      "Country: ", region, "\n",
      "Year: ", year, "\n",
      "Fitted: $", round(hat, 2), "/tCO₂\n",
      "Observed: $", round(obs, 2), "/tCO₂"
    )), color = col_sec, size = 1.8, alpha = 0.7) +
    scale_x_continuous(limits = c(0, ax_max)) +
    scale_y_continuous(limits = c(0, ax_max)) +
    coord_equal() +
    labs(title    = sprintf("%s Sector — Calibration: Observed vs Fitted Price", sector_name),
         subtitle = sub_note,
         x = "Fitted price (USD/tCO₂)", y = "Observed price (USD/tCO₂)") +
    theme_report()
}

make_outlier_table <- function(fit, sector_name, outlier_sd = 2) {
  calib <- get_calib_data(fit, outlier_sd)
  if (is.null(calib)) return(invisible(NULL))
  df_out <- calib$df[calib$df$outlier, ]
  if (nrow(df_out) == 0) return(invisible(NULL))
  df_out <- df_out[order(-df_out$obs), ]
  tbl <- data.frame(
    Region              = df_out$region,
    Year                = df_out$year,
    `Observed ($/tCO₂)` = round(df_out$obs, 1),
    `Fitted ($/tCO₂)`   = round(df_out$hat, 1),
    `Log-residual`      = round(df_out$resid, 3),
    check.names = FALSE
  )
  kable(tbl, format = "html",
        caption = sprintf("%s — Excluded Outliers (|log-residual| > %g SD)", sector_name, outlier_sd)) %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
}

# ── Fitted vs Observed Density (log space) ────────────────────────────────────
make_density_plot <- function(fit, sector_name) {
  if (is.null(fit)) return(NULL)
  m   <- fit$model
  obs <- m$y; hat <- fitted(m)
  if (is.null(obs)) return(NULL)
  # Cap x-axis at 99th percentile to prevent extreme outliers from compressing the main distribution
  x_cap <- quantile(obs, 0.99, na.rm = TRUE)
  df_dens <- rbind(
    data.frame(value = obs, source = "Observed",  stringsAsFactors = FALSE),
    data.frame(value = hat, source = "Fitted",    stringsAsFactors = FALSE)
  )
  col_sec <- if (sector_name == "Bulk") col_bulk else col_diffuse
  ggplot(df_dens, aes(x = value, fill = source, color = source)) +
    geom_density(alpha = 0.35, linewidth = 0.8) +
    scale_fill_manual(values  = c("Observed" = col_sec, "Fitted" = col_warn)) +
    scale_color_manual(values = c("Observed" = col_sec, "Fitted" = col_warn)) +
    coord_cartesian(xlim = c(NA_real_, x_cap)) +
    labs(
      title    = sprintf("%s Sector — Fitted vs Observed Distribution (log scale)", sector_name),
      subtitle = "Dependent variable: log(1 + ECP). Closer overlap = better calibration. X-axis capped at 99th percentile.",
      x = "log(1 + ECP)", y = "Density", fill = NULL, color = NULL
    ) +
    theme_report()
}

# ── Historical Price Trajectories (observed dots + fitted line) ────────────────
make_hist_trajectory <- function(fit, sector_name, key_regs) {
  if (is.null(fit)) return(NULL)
  m <- fit$model
  if (is.null(m)) return(NULL)
  obs <- back_transform(m$y)
  hat <- back_transform(fitted(m))

  # Prefer fit$data (full training set with region/year) over the GLM model frame,
  # which only stores formula variables and never includes the 'region' column.
  data_full <- fit$data
  if (!is.null(data_full) && "region" %in% colnames(data_full) && "year" %in% colnames(data_full)) {
    na_idx   <- as.integer(attr(m$model, "na.action"))
    all_idx  <- seq_len(nrow(data_full))
    ok_idx   <- if (length(na_idx) > 0) setdiff(all_idx, na_idx) else all_idx
    region_col <- as.character(data_full$region[ok_idx])
    year_col   <- as.integer(data_full$year[ok_idx])
  } else {
    # Fallback: try the GLM model frame (region absent when using regionFE)
    train      <- m$model
    region_col <- if ("region" %in% colnames(train)) as.character(train$region)
                  else rep(NA_character_, nrow(train))
    year_col   <- if ("timeTrend" %in% colnames(train))
                    as.integer(train$timeTrend + 1999L)   # timeTrend = year - 1999
                  else seq_len(nrow(train))
  }

  if (all(is.na(region_col))) return(NULL)
  df <- data.frame(region = region_col, year = year_col,
                   observed = obs, fitted = hat, stringsAsFactors = FALSE)
  highlight <- intersect(key_regs, unique(df$region))
  df_key <- df[df$region %in% highlight, ]
  if (nrow(df_key) == 0) return(NULL)
  df_long <- tidyr::pivot_longer(df_key, cols = c("observed", "fitted"),
                                  names_to = "type", values_to = "price")
  ggplot(df_long, aes(x = year, y = price, color = region, linetype = type)) +
    geom_line(data = df_long[df_long$type == "fitted",  ], linewidth = 1) +
    geom_point(data = df_long[df_long$type == "observed",], size = 2, alpha = 0.8) +
    scale_linetype_manual(values = c("observed" = "blank", "fitted" = "solid")) +
    scale_y_continuous() +
    facet_wrap(~region, scales = "free_y", ncol = 4) +
    labs(
      title    = sprintf("%s Sector — Historical Price Trajectories: Observed vs Fitted", sector_name),
      subtitle = "Points = observed ECP | Lines = model-fitted ECP (USD/tCO₂, back-transformed)",
      x = "Year", y = "ECP (USD/tCO₂)", color = "Region"
    ) +
    guides(linetype = "none") +
    theme_report() +
    theme(legend.position = "none")
}

# ── Residual Diagnostics (2 or 3 panels) ──────────────────────────────────────
make_residual_plots <- function(fit, sector_name, instQualityDrivers = NULL) {
  if (is.null(fit)) return(NULL)
  m     <- fit$model
  train <- m$model
  resid <- residuals(m, type = "deviance")
  hat   <- fitted(m)
  col_sec <- if (sector_name == "Bulk") col_bulk else col_diffuse

  df_r <- data.frame(
    resid  = resid,
    fitted = hat,
    year   = if ("timeTrend" %in% colnames(train)) as.integer(train$timeTrend + 1999L) else seq_along(resid),
    api    = if ("Actor.Power.Index" %in% colnames(train)) train$Actor.Power.Index else NA_real_,
    stringsAsFactors = FALSE
  )

  p1 <- ggplot(df_r, aes(x = fitted, y = resid)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(color = col_sec, alpha = 0.5, size = 1.5) +
    geom_smooth(method = "loess", se = TRUE, color = col_bad, linewidth = 0.8) +
    labs(title = "Residuals vs Fitted", x = "Fitted (log scale)", y = "Deviance residual") +
    theme_report()

  p2 <- ggplot(df_r, aes(x = year, y = resid)) +
    geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
    geom_point(color = col_sec, alpha = 0.5, size = 1.5) +
    geom_smooth(method = "loess", se = TRUE, color = col_bad, linewidth = 0.8) +
    labs(title = "Residuals vs Year", x = "Year", y = "Deviance residual") +
    theme_report()

  # 3rd panel: Actor Power Index if in model, else first IQ driver, else omit
  p3 <- if (!all(is.na(df_r$api))) {
    ggplot(df_r, aes(x = api, y = resid)) +
      geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
      geom_point(color = col_sec, alpha = 0.5, size = 1.5) +
      geom_smooth(method = "loess", se = TRUE, color = col_bad, linewidth = 0.8) +
      labs(title = "Residuals vs Actor Power Index", x = "Actor Power Index", y = "Deviance residual") +
      theme_report()
  } else {
    # Fallback: first available IQ driver
    iq_safe <- if (!is.null(instQualityDrivers) && length(instQualityDrivers) > 0)
                 make.names(instQualityDrivers[[1]]) else NULL
    if (!is.null(iq_safe) && iq_safe %in% colnames(train)) {
      df_r$iq_val <- train[[iq_safe]]
      ggplot(df_r, aes(x = iq_val, y = resid)) +
        geom_hline(yintercept = 0, linetype = "dashed", color = "grey50") +
        geom_point(color = col_sec, alpha = 0.5, size = 1.5) +
        geom_smooth(method = "loess", se = TRUE, color = col_bad, linewidth = 0.8) +
        labs(title = sprintf("Residuals vs %s", instQualityDrivers[[1]]),
             x = instQualityDrivers[[1]], y = "Deviance residual") +
        theme_report()
    } else NULL
  }

  if (!is.null(p3)) {
    (p1 | p2 | p3) + patchwork::plot_annotation(
      title    = sprintf("%s Sector — Residual Diagnostics", sector_name),
      subtitle = "Loess smoother shown. A flat line at zero indicates well-behaved residuals."
    )
  } else {
    (p1 | p2) + patchwork::plot_annotation(
      title    = sprintf("%s Sector — Residual Diagnostics", sector_name),
      subtitle = "Loess smoother shown. A flat line at zero indicates well-behaved residuals."
    )
  }
}

# ── Partial Regression (Component + Residual) Plots ───────────────────────────
make_partial_regression_plots <- function(fit, sector_name, instQualityDrivers, controlDrivers) {
  if (is.null(fit)) return(NULL)
  m     <- fit$model
  if (is.null(m)) return(NULL)
  train <- if (!is.null(fit$data)) fit$data else m$model
  ct    <- fit$coeftest
  if (is.null(ct) || is.null(train)) return(NULL)
  beta  <- coef(m)
  resid <- residuals(m, type = "response")
  col_sec <- if (sector_name == "Bulk") col_bulk else col_diffuse

  # Theory predictors: Actor Power + IQ (continuous, non-interaction, non-FE)
  theory_vars <- make.names(c("Actor Power Index", instQualityDrivers))
  avail <- intersect(theory_vars, colnames(train))
  if (length(avail) == 0) return(NULL)

  plots <- lapply(avail, function(v) {
    if (!v %in% names(beta)) return(NULL)
    x_vals  <- train[[v]]
    y_vals  <- beta[v] * x_vals + resid    # component + residual
    df_cr   <- data.frame(x = x_vals, y = y_vals, stringsAsFactors = FALSE)
    # Fitted line: just the component (beta * x)
    x_range <- seq(min(x_vals, na.rm = TRUE), max(x_vals, na.rm = TRUE), length.out = 100)
    df_line  <- data.frame(x = x_range, y = beta[v] * x_range)
    
    lbl <- clean_term_plain(v)
    
    ggplot(df_cr, aes(x = x, y = y)) +
      geom_point(color = col_sec, alpha = 0.4, size = 1.5) +
      geom_line(data = df_line, aes(x = x, y = y), color = col_bad, linewidth = 1) +
      labs(x = lbl, y = "Component + Residual") +
      theme_report() +
      theme(axis.title = element_text(size = 9))
  })
  plots <- Filter(Negate(is.null), plots)
  if (length(plots) == 0) return(NULL)
  wrap_plots(plots, ncol = min(3, length(plots))) +
    patchwork::plot_annotation(
      title    = sprintf("%s Sector — Partial Regression Plots", sector_name),
      subtitle = "Component + residual for each theory predictor. Slope = model coefficient."
    )
}

# ── Adoption-gated price projections ──────────────────────────────────────────
# Helper: align a factor column to model's known levels, mapping unknowns to "Other"
.align_fe <- function(df, col, model_levels) {
  if (is.null(model_levels) || !col %in% colnames(df)) return(df)
  vals <- as.character(df[[col]])
  vals[!vals %in% model_levels] <- "Other"     # map unknowns to reference level
  df[[col]] <- factor(vals, levels = model_levels)
  df
}

get_gated_projections <- function(fit_str, fit_adp, fdf, threshold,
                                  str_regionFE = NULL, adp_regionFE = NULL) {
  if (is.null(fit_str) || is.null(fit_adp) || is.null(fdf) || nrow(fdf) == 0) return(NULL)
  m_str <- fit_str$model; m_adp <- fit_adp$model

  # Use only PREDICTOR variables (exclude response) for complete.cases.
  # all.vars(formula()) includes the LHS response; in future panel data
  # the response (ecp) is NA, which would drop ALL rows if included.
  pred_vars_str <- all.vars(delete.response(terms(m_str)))
  pred_vars_adp <- all.vars(delete.response(terms(m_adp)))
  # Exclude regionFE from complete.cases check — we handle its levels manually below
  needed_str <- setdiff(intersect(pred_vars_str, colnames(fdf)), "regionFE")
  needed_adp <- setdiff(intersect(pred_vars_adp, colnames(fdf)), "regionFE")
  all_needed  <- union(needed_str, needed_adp)
  if (length(all_needed) == 0) return(NULL)
  df_cc <- fdf[complete.cases(fdf[, all_needed, drop = FALSE]), ]
  if (nrow(df_cc) == 0) return(NULL)

  # Dynamic region mapper
  .map_region_fe <- function(df, mapping_file) {
    if (is.null(mapping_file) || nchar(mapping_file) == 0) {
      df$regionFE <- factor(rep("Other", nrow(df)))
      return(df)
    }
    mapping <- madrat::toolGetMapping(mapping_file, type = "regional", where = "mappingfolder")
    regionLookup <- stats::setNames(mapping$RegionCode, mapping$CountryCode)
    regFE <- if (all(df$region %in% mapping$RegionCode)) {
      df$region
    } else {
      regionLookup[df$region]
    }
    regFE <- trimws(as.character(regFE))
    regFE[is.na(regFE) | regFE == "" | regFE == "NA"] <- "Other"
    df$regionFE <- factor(regFE)
    df
  }

  df_cc_str <- .map_region_fe(df_cc, str_regionFE)
  df_cc_adp <- .map_region_fe(df_cc, adp_regionFE)

  # Align regionFE factor levels: unknown levels are mapped to "Other" (reference)
  # so predict() doesn't fail on unseen levels.
  str_fe_levels <- if (!is.null(m_str$xlevels)) m_str$xlevels[["regionFE"]] else NULL
  adp_fe_levels <- if (!is.null(m_adp$model) && "regionFE" %in% colnames(m_adp$model))
                     levels(m_adp$model$regionFE) else str_fe_levels
  
  df_cc_str <- .align_fe(df_cc_str, "regionFE", str_fe_levels)
  df_cc_adp <- .align_fe(df_cc_adp, "regionFE", adp_fe_levels)

  pr_adp   <- tryCatch(plogis(predict(m_adp, newdata = df_cc_adp)), error = function(e) rep(NA_real_, nrow(df_cc)))
  pred_log <- tryCatch(predict(m_str, newdata = df_cc_str),          error = function(e) rep(NA_real_, nrow(df_cc)))

  # Projection Guards (ADR 0007), matching pfm::projectSpecScenario: the legacy
  # report's own predict() path must clamp the log(1+ECP) response to the
  # in-sample fitted maximum + extrapLogMargin (2) and cap the back-transformed
  # price at priceCeiling (5000), or an out-of-sample driver/trend excursion
  # explodes through expm1() (e.g. CHA Diffuse → 3e11 USD/tCO2). The time-trend
  # freeze tames this at the source; the guard is the safety net.
  extrapLogMargin <- 2; priceCeiling <- 5000
  insResp <- tryCatch(as.numeric(fitted(m_str)), error = function(e) NULL)
  capVal  <- if (!is.null(insResp) && length(insResp) > 0) max(insResp, na.rm = TRUE) + extrapLogMargin else Inf
  if (is.finite(capVal)) pred_log <- pmin(pmax(pred_log, 0), capVal)
  price_usd <- back_transform(pred_log)
  price_usd <- pmin(price_usd, priceCeiling)

  df_cc$pr_adoption  <- pr_adp
  df_cc$price_hat_log <- pred_log
  df_cc$price_hat_usd <- price_usd
  df_cc$price_gated   <- ifelse(df_cc$pr_adoption >= threshold, df_cc$price_hat_usd, NA_real_)
  df_cc
}

# ── Hurdle projections: E[ECP] = Pr(Adoption) * exp(stringency) - Pr * 1 ─────
get_hurdle_projections <- function(fit_str, fit_adp, fdf, str_regionFE = NULL, adp_regionFE = NULL) {
  proj <- get_gated_projections(fit_str, fit_adp, fdf, threshold = 0,
                                str_regionFE = str_regionFE, adp_regionFE = adp_regionFE)
  if (is.null(proj)) return(NULL)
  proj$expected_ecp <- proj$pr_adoption * proj$price_hat_usd
  proj
}

# ── Price trajectory plot (future projections, adoption-gated) ────────────────
make_price_trajectory_plot <- function(proj, sector_name, key_regs, full_facet = FALSE) {
  if (is.null(proj) || nrow(proj) == 0) return(NULL)
  df <- proj[!is.na(proj$price_gated) & proj$year <= 2100, ]
  if (nrow(df) == 0) return(NULL)
  col_sec <- if (sector_name == "Bulk") col_bulk else col_diffuse
  all_r   <- unique(df$region)
  hi      <- intersect(key_regs, all_r)
  df$grp  <- ifelse(df$region %in% hi, as.character(df$region), "Other")
  df$grp  <- factor(df$grp, levels = c(hi, "Other"))
  my_cols <- setNames(c(scales::hue_pal()(length(hi)), "grey80"), c(hi, "Other"))

  y_max <- max(250, quantile(df$price_gated, 0.975, na.rm = TRUE))

  if (full_facet) {
    ggplot(df, aes(x = year, y = price_gated, group = region)) +
      geom_line(color = col_sec, linewidth = 0.7, alpha = 0.8) +
      scale_y_continuous() +
      coord_cartesian(ylim = c(0, y_max)) +
      facet_wrap(~region, ncol = 9) +
      labs(title = sprintf("%s — Adoption-Gated Price Trajectories (All Regions)", sector_name),
           x = "Year", y = "ECP (USD/tCO₂)") +
      theme_report() +
      theme(axis.text = element_text(size = 7), strip.text = element_text(size = 7))
  } else {
    ggplot(df, aes(x = year, y = price_gated, group = region, color = grp,
                   alpha = I(ifelse(grp == "Other", 0.3, 1)),
                   linewidth = I(ifelse(grp == "Other", 0.4, 1.2)))) +
      geom_line() +
      scale_color_manual(values = my_cols) +
      scale_y_continuous() +
      coord_cartesian(ylim = c(0, y_max)) +
      labs(title  = sprintf("%s — Adoption-Gated Price Trajectories (Key Regions)", sector_name),
           subtitle = "NA = region has not yet crossed adoption threshold",
           x = "Year", y = "ECP (USD/tCO₂)", color = "Region") +
      theme_report()
  }
}

# ── Price Onset Gantt (year of first non-NA price, coloured by starting level) ─
make_price_gantt <- function(proj, sector_name, end_year = 2100) {
  if (is.null(proj) || nrow(proj) == 0) return(NULL)
  df <- proj[!is.na(proj$price_gated), ]
  if (nrow(df) == 0) return(NULL)
  onset <- df %>%
    group_by(region) %>%
    summarise(onset_year   = min(year),
              onset_price  = price_gated[which.min(year)],
              .groups = "drop") %>%
    arrange(onset_year, onset_price)
  onset$region <- factor(onset$region, levels = onset$region)

  ggplot(onset, aes(xmin = onset_year, xmax = end_year, y = region, fill = onset_price)) +
    geom_rect(aes(ymin = as.numeric(region) - 0.4, ymax = as.numeric(region) + 0.4),
              color = "white", linewidth = 0.3) +
    geom_text(aes(x = onset_year, label = sprintf("$%.0f", onset_price)),
              hjust = -0.1, size = 2.5, color = "grey20") +
    scale_fill_viridis_c(option = "plasma", name = "Starting\nprice ($/tCO₂)") +
    scale_x_continuous(breaks = seq(2025, end_year, 10)) +
    labs(
      title    = sprintf("%s — Price Onset Chart", sector_name),
      subtitle = "Bar starts at first year price crosses adoption threshold. Colour = starting price level.",
      x = "Year", y = NULL
    ) +
    theme_report() +
    theme(axis.text.y = element_text(size = 7))
}

# ── Price snapshot map ─────────────────────────────────────────────────────────
make_price_map <- function(proj, mapping_54, world, snap_year, sector_name) {
  if (is.null(proj) || is.null(mapping_54) || is.null(world)) return(NULL)
  df_yr <- proj[proj$year == snap_year & !is.na(proj$price_gated),
                c("region", "price_gated")]
  if (nrow(df_yr) == 0) return(NULL)
  map_data <- merge(mapping_54, df_yr, by.x = "RegionCode", by.y = "region", all.x = TRUE)
  world_merged <- merge(world, map_data, by.x = "iso_key", by.y = "CountryCode", all.x = TRUE)
  lim_max <- max(250, quantile(df_yr$price_gated, 0.95, na.rm = TRUE))
  ggplot(world_merged) +
    geom_sf(aes(fill = price_gated), color = "white", linewidth = 0.1) +
    scale_fill_viridis_c(option = "plasma", na.value = "grey85",
                         limits = c(0, lim_max), oob = scales::squish,
                         name = "USD/tCO₂") +
    labs(
      title    = sprintf("%s — Projected ECP in %d (Adoption-Gated)", sector_name, snap_year),
      subtitle = "Grey = no adoption projected by this year"
    ) +
    theme_void() +
    theme(legend.position = "right",
          plot.title    = element_text(face = "bold", size = 13),
          plot.subtitle = element_text(size = 10, color = "grey40"))
}

# ── Price ranking table at snapshot years ─────────────────────────────────────
make_price_ranking <- function(proj, snap_years, top_n = 10) {
  if (is.null(proj) || nrow(proj) == 0) return(NULL)
  years_avail <- intersect(snap_years, unique(proj$year[!is.na(proj$price_gated)]))
  if (length(years_avail) == 0) return(NULL)
  lapply(years_avail, function(yr) {
    df_yr <- proj[proj$year == yr & !is.na(proj$price_gated), ]
    df_yr <- df_yr[order(-df_yr$price_gated), ]
    df_yr$rank <- seq_len(nrow(df_yr))
    head(df_yr[, c("rank", "region", "price_gated")], top_n)
  })
}

# ── Price distribution violin ─────────────────────────────────────────────────
make_price_distribution <- function(proj, snap_years, sector_name) {
  if (is.null(proj) || nrow(proj) == 0) return(NULL)
  df <- proj[proj$year %in% snap_years & !is.na(proj$price_gated), ]
  if (nrow(df) == 0) return(NULL)
  df$year_f <- factor(df$year)
  col_sec <- if (sector_name == "Bulk") col_bulk else col_diffuse
  ggplot(df, aes(x = year_f, y = price_gated)) +
    geom_violin(fill = col_sec, alpha = 0.4, color = col_sec) +
    geom_boxplot(width = 0.15, outlier.size = 1.5, fill = "white", color = col_sec) +
    scale_y_continuous() +
    labs(
      title    = sprintf("%s — Cross-Regional Price Distribution at Snapshot Years", sector_name),
      subtitle = "Violin = density | Box = median & IQR | Only adopted regions shown",
      x = "Year", y = "ECP (USD/tCO₂)"
    ) +
    theme_report()
}

# ── Hurdle: E[ECP] trajectories ───────────────────────────────────────────────
make_hurdle_trajectories <- function(hurdle, sector_name, key_regs) {
  if (is.null(hurdle) || nrow(hurdle) == 0) return(NULL)
  df <- hurdle[hurdle$year <= 2100, ]
  if (nrow(df) == 0) return(NULL)
  col_sec <- if (sector_name == "Bulk") col_bulk else col_diffuse
  all_r   <- unique(df$region)
  hi      <- intersect(key_regs, all_r)
  df$grp <- ifelse(df$region %in% hi, as.character(df$region), "Other")
  df$grp <- factor(df$grp, levels = c(hi, "Other"))
  my_cols <- setNames(c(scales::hue_pal()(length(hi)), "grey80"), c(hi, "Other"))
  
  y_max <- max(250, quantile(df$expected_ecp, 0.975, na.rm = TRUE))

  ggplot(df, aes(x = year, y = expected_ecp, group = region, color = grp,
                     alpha = I(ifelse(grp == "Other", 0.3, 1)),
                     linewidth = I(ifelse(grp == "Other", 0.4, 1.2)))) +
    geom_line() +
    scale_color_manual(values = my_cols) +
    scale_y_continuous() +
    coord_cartesian(ylim = c(0, y_max)) +
    labs(
      title    = sprintf("%s — Expected ECP: E[ECP] = Pr(Adoption) × Conditional Price", sector_name),
      subtitle = "Prices fade in gradually as Pr(Adoption) rises — no hard threshold cut",
      x = "Year", y = "Expected ECP (USD/tCO₂)", color = "Region"
    ) +
    theme_report()
}

# ── Hurdle decomposition table ────────────────────────────────────────────────
make_hurdle_decomposition <- function(gated, hurdle, snap_years) {
  if (is.null(gated) || is.null(hurdle) || nrow(gated) == 0) return(NULL)
  years_avail <- intersect(snap_years, unique(hurdle$year))
  if (length(years_avail) == 0) return(NULL)
  rows <- lapply(years_avail, function(yr) {
    h <- hurdle[hurdle$year == yr, c("region", "pr_adoption", "price_hat_usd", "expected_ecp")]
    g <- gated[gated$year  == yr, c("region", "price_gated")]
    m <- merge(h, g, by = "region", all.x = TRUE)
    m$year <- yr; m
  })
  df <- do.call(rbind, rows)
  df[, c("year", "region", "pr_adoption", "conditional_price" = "price_hat_usd",
         "expected_ecp", "gated_price" = "price_gated")]
}

# ── Threshold sensitivity ──────────────────────────────────────────────────────
make_threshold_sensitivity <- function(fit_str, fit_adp, fdf, sector_name, key_regs,
                                       thresholds = c(0.3, 0.5, 0.7),
                                       str_regionFE = NULL, adp_regionFE = NULL) {
  if (is.null(fit_str) || is.null(fit_adp) || is.null(fdf)) return(NULL)
  df_all <- do.call(rbind, lapply(thresholds, function(thr) {
    p <- get_gated_projections(fit_str, fit_adp, fdf, thr,
                               str_regionFE = str_regionFE, adp_regionFE = adp_regionFE)
    if (is.null(p)) return(NULL)
    p$threshold <- factor(sprintf("Threshold = %.1f", thr))
    p
  }))
  if (is.null(df_all) || nrow(df_all) == 0) return(NULL)
  df_key <- df_all[df_all$region %in% key_regs, ]
  if (nrow(df_key) == 0) return(NULL)
  
  df_key <- df_key[df_key$year <= 2100, ]
  if (nrow(df_key) == 0) return(NULL)

  ggplot(df_key, aes(x = year, y = price_gated, color = threshold, group = threshold)) +
    geom_line(linewidth = 1, na.rm = TRUE) +
    scale_y_continuous() +
    scale_color_viridis_d(option = "D") +
    facet_wrap(~region, scales = "free_y", ncol = 4) +
    labs(
      title    = sprintf("%s — Threshold Sensitivity Analysis", sector_name),
      subtitle = "Projected ECP under three adoption thresholds. Key regions only.",
      x = "Year", y = "ECP (USD/tCO₂)", color = "Threshold"
    ) +
    theme_report()
}

# ── Convergence map: difference B - C ─────────────────────────────────────────
make_convergence_map <- function(gated, hurdle, mapping_54, world, snap_year, sector_name) {
  if (is.null(gated) || is.null(hurdle) || is.null(mapping_54) || is.null(world)) return(NULL)
  g <- gated[gated$year == snap_year, c("region", "price_gated")]
  h <- hurdle[hurdle$year == snap_year, c("region", "expected_ecp")]
  df <- merge(g, h, by = "region", all = TRUE)
  price_gated_val <- ifelse(is.na(df$price_gated), 0, df$price_gated)
  df$diff <- price_gated_val - df$expected_ecp
  map_data     <- merge(mapping_54, df, by.x = "RegionCode", by.y = "region", all.x = TRUE)
  world_merged <- merge(world, map_data, by.x = "iso_key", by.y = "CountryCode", all.x = TRUE)
  lim <- max(250, quantile(abs(df$diff), 0.95, na.rm = TRUE))
  ggplot(world_merged) +
    geom_sf(aes(fill = diff), color = "white", linewidth = 0.1) +
    scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#d73027", midpoint = 0,
                         limits = c(-lim, lim), oob = scales::squish, na.value = "grey85",
                         name = "Δ ECP\n(USD/tCO₂)") +
    labs(
      title    = sprintf("%s — Convergence Map: Hard-Gated minus Expected ECP (%d)", sector_name, snap_year),
      subtitle = "Red = hard-gating overestimates | Blue = hard-gating underestimates (e.g. not yet adopted) | Grey = missing data"
    ) +
    theme_void() +
    theme(legend.position = "right",
          plot.title    = element_text(face = "bold", size = 13),
          plot.subtitle = element_text(size = 10, color = "grey40"))
}
```

```{r proj_setup, include=FALSE}
# ── Ensure fit$data is available for historical trajectory / calibration plots ─
# When a stringency model is loaded from cache, $data is NULL. Reconstruct it
# by calling preparePanelData with the same config used during estimation.
ensure_str_data <- function(fit, sec, cfg_sec, panel_data) {
  if (is.null(fit) || (!is.null(fit$data) && "region" %in% colnames(fit$data))) return(fit)
  td <- tryCatch(
    pfm::preparePanelData(
      data                      = panel_data, sector = sec,
      actorPowerDrivers         = cfg_sec$actorPower,
      actorPowerIndex           = cfg_sec$actorIndex,
      instQualityDrivers        = cfg_sec$instQuality,
      controlDrivers            = cfg_sec$controls,
      regionMappingFixedEffects = if (isTRUE(cfg_sec$useMundlak)) NULL else cfg_sec$regionFE,
      useMundlak                = isTRUE(cfg_sec$useMundlak),
      lag = 1
    ),
    error = function(e) NULL
  )
  if (!is.null(td)) fit$data <- td[td$ecp > 0, , drop = FALSE]
  fit
}

fit_str$Bulk    <- ensure_str_data(fit_str$Bulk,    "Bulk",    cfg$Bulk,    panelData)
fit_str$Diffuse <- ensure_str_data(fit_str$Diffuse, "Diffuse", cfg$Diffuse, panelData)

proj_str <- list(
  Bulk    = get_gated_projections(fit_str$Bulk,    fit_adp$Bulk,    future_df$Bulk,    adoption_threshold_bulk,
                                   str_regionFE = if (isTRUE(cfg$Bulk$useMundlak))    NULL else cfg$Bulk$regionFE,
                                   adp_regionFE = adp_cfg$regionFE),
  Diffuse = get_gated_projections(fit_str$Diffuse, fit_adp$Diffuse, future_df$Diffuse, adoption_threshold_diffuse,
                                   str_regionFE = if (isTRUE(cfg$Diffuse$useMundlak)) NULL else cfg$Diffuse$regionFE,
                                   adp_regionFE = adp_cfg$regionFE)
)
hurdle <- list(
  Bulk    = get_hurdle_projections(fit_str$Bulk,    fit_adp$Bulk,    future_df$Bulk,
                                    str_regionFE = if (isTRUE(cfg$Bulk$useMundlak))    NULL else cfg$Bulk$regionFE,
                                    adp_regionFE = adp_cfg$regionFE),
  Diffuse = get_hurdle_projections(fit_str$Diffuse, fit_adp$Diffuse, future_df$Diffuse,
                                    str_regionFE = if (isTRUE(cfg$Diffuse$useMundlak)) NULL else cfg$Diffuse$regionFE,
                                    adp_regionFE = adp_cfg$regionFE)
)

# Maps setup (same pattern as adoption report)
mapping_54 <- NULL; world <- NULL
if (has_scenario) {
  mapping_54 <- tryCatch(
    madrat::toolGetMapping("regionmapping_54.csv", type = "regional", where = "mappingfolder"),
    error = function(e) NULL
  )
  if (!is.null(mapping_54)) {
    suppressPackageStartupMessages(library(sf))
    world         <- rnaturalearth::ne_countries(scale = "small", returnclass = "sf")
    world$iso_key <- as.character(world$adm0_a3)
  }
}

stats_bulk    <- compute_str_stats(fit_str$Bulk)
stats_diffuse <- compute_str_stats(fit_str$Diffuse)

se_bulk    <- compute_semi_elasticity(fit_str$Bulk,    cfg$Bulk$instQuality,    cfg$Bulk$controls)
se_diffuse <- compute_semi_elasticity(fit_str$Diffuse, cfg$Diffuse$instQuality, cfg$Diffuse$controls)
mem_bulk    <- compute_mem(fit_str$Bulk)
mem_diffuse <- compute_mem(fit_str$Diffuse)
```

---

# Executive Summary

<div class="row">
<div class="col-md-4 bulk-only">
<div class="card">
<div class="card-header" style="border-left:4px solid #2980b9;">Bulk Sector Model</div>
<div class="card-body">
<ul>
  <li><strong>Model name:</strong> `r params$modelName`</li>
  <li><strong>Response:</strong> log(1 + ECP_Bulk) — gaussian GLM (identity link)</li>
  <li><strong>Sample:</strong> Region-years with ECP_Bulk &gt; 0</li>
  <li><strong>IQ drivers:</strong> `r paste(cfg$Bulk$instQuality, collapse = ", ")`</li>
  <li><strong>Controls:</strong> `r paste(cfg$Bulk$controls, collapse = ", ")`</li>
  <li><strong>`r if (isTRUE(cfg$Bulk$useMundlak)) "Mundlak correction" else "Region FE"`:</strong> `r if (isTRUE(cfg$Bulk$useMundlak)) "ON (within-region group means)" else if (is.null(cfg$Bulk$regionFE)) "None" else cfg$Bulk$regionFE`</li>
  <li><strong>Lagged ECP:</strong> `r cfg$Bulk$lagged`</li>
</ul>
</div></div></div>
<div class="col-md-4 diffuse-only">
<div class="card">
<div class="card-header" style="border-left:4px solid #27ae60;">Diffuse Sector Model</div>
<div class="card-body">
<ul>
  <li><strong>Model name:</strong> `r params$modelName`</li>
  <li><strong>Response:</strong> log(1 + ECP_Diffuse) — gaussian GLM (identity link)</li>
  <li><strong>Sample:</strong> Region-years with ECP_Diffuse &gt; 0</li>
  <li><strong>IQ drivers:</strong> `r paste(cfg$Diffuse$instQuality, collapse = ", ")`</li>
  <li><strong>Controls:</strong> `r paste(cfg$Diffuse$controls, collapse = ", ")`</li>
  <li><strong>`r if (isTRUE(cfg$Diffuse$useMundlak)) "Mundlak correction" else "Region FE"`:</strong> `r if (isTRUE(cfg$Diffuse$useMundlak)) "ON (within-region group means)" else if (is.null(cfg$Diffuse$regionFE)) "None" else cfg$Diffuse$regionFE`</li>
  <li><strong>Lagged ECP:</strong> `r cfg$Diffuse$lagged`</li>
</ul>
</div></div></div>
<div class="col-md-4">
<div class="card">
<div class="card-header">Projection Settings</div>
<div class="card-body">
<ul>
  <li><strong>Adoption threshold:</strong> `r adoption_threshold` — prices shown only above this probability</li>
  <li><strong>Adoption model IQ:</strong> `r paste(adp_cfg$instQuality, collapse = ", ")`</li>
  <li><strong>Snapshot years:</strong> `r paste(snapshot_years, collapse = ", ")`</li>
  <li><strong>Scenario data:</strong> `r if (has_scenario) paste0("Loaded from ", basename(params$gdxPath)) else "Not provided — projection sections disabled"`</li>
  <li><strong>Historical range:</strong> 2000–2022</li>
</ul>
</div></div></div></div>

```{r exec-summary-table, echo=FALSE}
mem_mean_bulk    <- if (!is.null(mem_bulk))    round(mem_bulk$mean_price[1], 1)    else NA
mem_mean_diffuse <- if (!is.null(mem_diffuse)) round(mem_diffuse$mean_price[1], 1) else NA

summary_df <- data.frame(
  Sector       = c("Bulk", "Diffuse"),
  `R²`         = c(stats_bulk$r2,      stats_diffuse$r2),
  AIC          = c(stats_bulk$aic,     stats_diffuse$aic),
  BIC          = c(stats_bulk$bic,     stats_diffuse$bic),
  `Log-lik`    = c(stats_bulk$loglik,  stats_diffuse$loglik),
  `N obs`      = c(stats_bulk$nObs,    stats_diffuse$nObs),
  `N pred`     = c(stats_bulk$k,       stats_diffuse$k),
  `Mean fitted price (USD/tCO₂)` = c(mem_mean_bulk, mem_mean_diffuse),
  Converged    = c(stats_bulk$converged, stats_diffuse$converged),
  check.names  = FALSE
)
kable(summary_df, format = "html", caption = "Stringency Model Fit Summary") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE) %>%
  column_spec(2, bold = TRUE,
    color = ifelse(summary_df$`R²` >= 0.5, col_good,
      ifelse(summary_df$`R²` >= 0.25, col_warn, col_bad)))
```

---

# Data Quality Check

```{r data-qa, echo=FALSE, results='asis'}
ecp_bulk_name    <- "Effective Carbon Price|Bulk"
ecp_diffuse_name <- "Effective Carbon Price|Diffuse"
for (var_check in c(ecp_bulk_name, ecp_diffuse_name)) {
  df_v <- df_hist[df_hist$variable == var_check & df_hist$value > 0, ]
  cat(sprintf("\n**%s:** %d region-years with positive price across %d regions.\n\n",
              var_check, nrow(df_v), length(unique(df_v$region))))
}

# Check scenario coverage
if (has_scenario) {
  for (sec in c("Bulk", "Diffuse")) {
    fdf <- future_df[[sec]]
    if (!is.null(fdf)) {
      cat(sprintf("**Scenario (%s):** %d region-year rows, %d regions, years %d–%d.\n\n",
                  sec, nrow(fdf), length(unique(fdf$region)),
                  min(fdf$year), max(fdf$year)))
    }
  }
}
```

```{r data-coverage-table, echo=FALSE}
ecp_vars <- c("Effective Carbon Price|Bulk", "Effective Carbon Price|Diffuse")
df_ecp_hist <- df_hist[df_hist$variable %in% ecp_vars & df_hist$value > 0, ]
cov_tbl <- df_ecp_hist %>%
  group_by(variable) %>%
  summarise(
    `Regions with price` = n_distinct(region),
    `Obs (region-years)` = n(),
    `Min ECP (USD/tCO₂)` = round(min(value), 1),
    `Median ECP`         = round(median(value), 1),
    `Max ECP`            = round(max(value), 1),
    `Year range`         = paste0(min(year), "–", max(year)),
    .groups = "drop"
  )
kable(cov_tbl, format = "html", caption = "Historical ECP Coverage (Positive Prices Only)") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)
```

---

# Input Data {.tabset}

## Historical Distributions

```{r hist-dist, echo=FALSE, fig.height=5}
ecp_hist_df <- df_hist[df_hist$variable %in% c("Effective Carbon Price|Bulk",
                                                 "Effective Carbon Price|Diffuse") &
                          df_hist$value > 0, ]
ecp_hist_df$sector <- sub("Effective Carbon Price\\|", "", ecp_hist_df$variable)

ggplot(ecp_hist_df, aes(x = value, fill = sector)) +
  geom_histogram(bins = 40, alpha = 0.7, color = "white", position = "identity") +
  scale_fill_manual(values = c(Bulk = col_bulk, Diffuse = col_diffuse)) +
  scale_x_log10(labels = scales::label_comma()) +
  facet_wrap(~sector, scales = "free_y") +
  labs(title = "Distribution of Historical Carbon Prices (Adopters Only, Log Scale)",
       x = "ECP (USD/tCO₂, log scale)", y = "Count") +
  theme_report() + theme(legend.position = "none")
```

## Historical Trends

```{r hist-trends, echo=FALSE, fig.height=6}
ecp_key <- df_hist[df_hist$variable %in% c("Effective Carbon Price|Bulk",
                                             "Effective Carbon Price|Diffuse") &
                     df_hist$region %in% key_regions & df_hist$value > 0, ]
ecp_key$sector <- sub("Effective Carbon Price\\|", "", ecp_key$variable)
if (nrow(ecp_key) > 0) {
  ggplot(ecp_key, aes(x = year, y = value, color = sector)) +
    geom_line(linewidth = 1) + geom_point(size = 2) +
    scale_color_manual(values = c(Bulk = col_bulk, Diffuse = col_diffuse)) +
    scale_y_continuous() +
    facet_wrap(~region, scales = "free_y", ncol = 4) +
    labs(title = "Historical ECP Trends — Key Regions",
         x = "Year", y = "ECP (USD/tCO₂)", color = "Sector") +
    theme_report()
}
```

## Historical Variable Statistics

```{r var-stats, echo=FALSE}
all_str_vars <- unique(c("Effective Carbon Price|Bulk", "Effective Carbon Price|Diffuse",
                          "Actor Power Index|Bulk", "Actor Power Index|Diffuse",
                          cfg$Bulk$instQuality, cfg$Bulk$controls))
df_stats_hist <- df_hist[df_hist$variable %in% all_str_vars, ]
stat_tbl <- df_stats_hist %>%
  group_by(variable) %>%
  summarise(N = n(), Mean = round(mean(value, na.rm=TRUE),3),
            Median = round(median(value, na.rm=TRUE),3),
            SD = round(sd(value, na.rm=TRUE),3),
            Min = round(min(value, na.rm=TRUE),3),
            Max = round(max(value, na.rm=TRUE),3), .groups="drop")
kable(stat_tbl, format="html",
      caption="Historical Variable Statistics (2000–2022)") %>%
  kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=FALSE)
```

## Scenario Projections

```{r scen-proj-check, echo=FALSE, results='asis'}
if (!has_scenario) cat("> **Scenario GDX not provided.** Re-render with `gdxPath` pointing to `fulldata.gdx`.\n")
```

```{r scen-proj-plot, echo=FALSE, fig.height=5}
if (has_scenario && !is.null(df_scen)) {
  api_scen <- df_scen[df_scen$variable %in% c("Actor Power Index|Bulk", "Actor Power Index|Diffuse") &
                        df_scen$region %in% key_regions, ]
  api_scen$sector <- sub("Actor Power Index\\|", "", api_scen$variable)
  if (nrow(api_scen) > 0) {
    ggplot(api_scen, aes(x = year, y = value, color = region, linetype = sector)) +
      geom_line(linewidth = 1) +
      facet_wrap(~sector, ncol = 2) +
      labs(title = "Scenario Actor Power Index — Key Regions",
           x = "Year", y = "Actor Power Index", color = "Region") +
      theme_report()
  }
}
```

---

# Model Estimation {.tabset}

## Forest Plots

<div class="bulk-only">

```{r forest-bulk, echo=FALSE, fig.height=7}
if (!is.null(fit_str$Bulk) && !is.null(fit_str$Bulk$coeftest)) {
  df_coef <- coeftestToDataFrame(fit_str$Bulk$coeftest)
  p <- render_forest_plot(df_coef, title = "Bulk Stringency — Coefficient Forest Plot (log scale)")
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "forest_bulk_stringency.png"), p, width=10, height=7, dpi=300)
    ggsave(file.path(plotOutputDir, "forest_bulk_stringency.svg"), p, width=10, height=7)
  }
}
```

<div class="chart-description">
**How to read this chart:** Each row is one model coefficient on the log(1 + ECP) scale. The point estimate is the dot; the horizontal bar is the 95% confidence interval (robust clustered SE by region). A coefficient of β means a 1-unit increase in the predictor multiplies the expected carbon price by exp(β). Coefficients whose CI does not cross zero are statistically significant. Positive coefficients increase the expected price; negative coefficients decrease it.
</div>

</div>

<div class="diffuse-only">

```{r forest-diffuse, echo=FALSE, fig.height=7}
if (!is.null(fit_str$Diffuse) && !is.null(fit_str$Diffuse$coeftest)) {
  df_coef <- coeftestToDataFrame(fit_str$Diffuse$coeftest)
  p <- render_forest_plot(df_coef, title = "Diffuse Stringency — Coefficient Forest Plot (log scale)")
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "forest_diffuse_stringency.png"), p, width=10, height=7, dpi=300)
    ggsave(file.path(plotOutputDir, "forest_diffuse_stringency.svg"), p, width=10, height=7)
  }
}
```

<div class="chart-description">
**How to read this chart:** Each row is one model coefficient on the log(1 + ECP) scale. The point estimate is the dot; the horizontal bar is the 95% confidence interval (robust clustered SE by region). A coefficient of β means a 1-unit increase in the predictor multiplies the expected carbon price by exp(β). Coefficients whose CI does not cross zero are statistically significant.
</div>

</div>

## Coefficient Tables

<div class="bulk-only">

```{r est-bulk, echo=FALSE}
if (!is.null(fit_str$Bulk) && !is.null(fit_str$Bulk$coeftest)) {
  ct <- pfm::coeftestToDataFrame(fit_str$Bulk$coeftest) %>%
    mutate(sig = ifelse(pValue < 0.001, "***", ifelse(pValue < 0.01, "**",
                  ifelse(pValue < 0.05, "*", ifelse(pValue < 0.1, ".", "")))),
           across(where(is.numeric), ~round(.x, 4)))
  colnames(ct) <- c("Term", "Estimate", "Std. Error", "z value", "Pr(>|z|)", "Sig.")
  kable(ct, format = "html", caption = "Bulk Stringency Model — Coefficients (log scale)") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE) %>%
    column_spec(6, bold = TRUE)
}
```

</div>

<div class="diffuse-only">

```{r est-diffuse, echo=FALSE}
if (!is.null(fit_str$Diffuse) && !is.null(fit_str$Diffuse$coeftest)) {
  ct <- pfm::coeftestToDataFrame(fit_str$Diffuse$coeftest) %>%
    mutate(sig = ifelse(pValue < 0.001, "***", ifelse(pValue < 0.01, "**",
                  ifelse(pValue < 0.05, "*", ifelse(pValue < 0.1, ".", "")))),
           across(where(is.numeric), ~round(.x, 4)))
  colnames(ct) <- c("Term", "Estimate", "Std. Error", "z value", "Pr(>|z|)", "Sig.")
  kable(ct, format = "html", caption = "Diffuse Stringency Model — Coefficients (log scale)") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE) %>%
    column_spec(6, bold = TRUE)
}
```

</div>

```{r make-comparison-table-fn, include=FALSE}
make_comparison_table <- function(se_b, se_d) {
  if (is.null(se_b) || is.null(se_d)) return(NULL)
  shared <- intersect(se_b$term, se_d$term)
  data.frame(
    Term                 = shared,
    `Bulk β`             = round(se_b$estimate[match(shared, se_b$term)], 3),
    `Bulk semi-elas.`    = sprintf("%.1f%%", 100 * se_b$semi_elas[match(shared, se_b$term)]),
    `Bulk p`             = round(se_b$p[match(shared, se_b$term)], 3),
    `Diffuse β`          = round(se_d$estimate[match(shared, se_d$term)], 3),
    `Diffuse semi-elas.` = sprintf("%.1f%%", 100 * se_d$semi_elas[match(shared, se_d$term)]),
    `Diffuse p`          = round(se_d$p[match(shared, se_d$term)], 3),
    check.names = FALSE
  )
}
```

---

# Model Formula

The stringency model estimates $\mathbb{E}\!\left[\log(1+\text{ECP}) \mid \text{adopted}\right] = X\beta$ — a **gaussian GLM with identity link** on the $\log(1 + \text{ECP})$ dependent variable (back-transformed as $\text{ECP} = e^{X\beta}-1$). This replaces the former Gamma-log link, which double-exponentiated under projection and drove the price explosion (ADR 0007). Drivers are **standardized**, so each coefficient $\beta$ is the additive effect on $\log(1+\text{ECP})$ per **one-standard-deviation** increase in the predictor; $\beta = 0.3^{***}$ means +1 SD multiplies $(1+\text{ECP})$ by $\exp(0.3) \approx 1.35$ (+35%).

<div class="bulk-only">

```{r model-formula-bulk, echo=FALSE, results='asis'}
if (!is.null(fit_str$Bulk) && !is.null(fit_str$Bulk$coeftest)) {
  cat("**Bulk Stringency Model Equation:**\n\n")
  cat(formatModelEquationLatex(
    coefficients       = pfm::coeftestToDataFrame(fit_str$Bulk$coeftest),
    actorPowerIndex    = cfg$Bulk$actorIndex,
    actorPowerDrivers  = cfg$Bulk$actorPower,
    instQualityDrivers = cfg$Bulk$instQuality,
    controlDrivers     = cfg$Bulk$controls,
    depVar             = "\\log\\left(\\mathbb{E}[\\text{ECP}_{\\text{Bulk}} \\mid \\text{adopted}]\\right)"
  ), "\n\n")
  cat("Significance codes: $^{***}\\ p < 0.01$, $^{**}\\ p < 0.05$, $^{*}\\ p < 0.1$.\n")
}
```

</div>

<div class="diffuse-only">

```{r model-formula-diffuse, echo=FALSE, results='asis'}
if (!is.null(fit_str$Diffuse) && !is.null(fit_str$Diffuse$coeftest)) {
  cat("**Diffuse Stringency Model Equation:**\n\n")
  cat(formatModelEquationLatex(
    coefficients       = pfm::coeftestToDataFrame(fit_str$Diffuse$coeftest),
    actorPowerIndex    = cfg$Diffuse$actorIndex,
    actorPowerDrivers  = cfg$Diffuse$actorPower,
    instQualityDrivers = cfg$Diffuse$instQuality,
    controlDrivers     = cfg$Diffuse$controls,
    depVar             = "\\log\\left(\\mathbb{E}[\\text{ECP}_{\\text{Diffuse}} \\mid \\text{adopted}]\\right)"
  ), "\n\n")
  cat("Significance codes: $^{***}\\ p < 0.01$, $^{**}\\ p < 0.05$, $^{*}\\ p < 0.1$.\n")
}
```

</div>

---

## Fit Diagnosis {.tabset}

```{r safe-print-repel-fn, include=FALSE}
safe_print_repel <- function(p, path, w = 9, h = 6) {
  if (is.null(p)) return(invisible(NULL))
  result <- tryCatch(
    { print(p); "ok" },
    error = function(e) {
      p_safe <- p
      p_safe$layers <- p_safe$layers[
        !sapply(p_safe$layers, function(l) inherits(l$geom, "GeomTextRepel"))
      ]
      print(p_safe); "fallback"
    }
  )
  tryCatch(ggsave(path, p, width = w, height = h, dpi = 300),
           error = function(e) message("ggsave skipped: ", conditionMessage(e)))
  invisible(result)
}
```

### 45° Calibration Scatter

<div class="bulk-only">
```{r cal-scatter-bulk, echo=FALSE, fig.height=7, dev="png", dpi=150}
p_cs <- make_calibration_scatter(fit_str$Bulk, "Bulk")
if (!is.null(p_cs)) {
  if (knitr::is_html_output()) {
    p_plotly <- plotly::ggplotly(p_cs, tooltip = "text")
    p_plotly <- plotly::layout(p_plotly, margin = list(t = 50, b = 50, l = 50, r = 50))
    p_plotly
  } else {
    print(p_cs)
  }
  tryCatch(ggsave(file.path(plotOutputDir, "calib_scatter_bulk.png"), p_cs, width=9, height=7, dpi=300), error=function(e)NULL)
}
```
<div class="chart-description">
**How to read this chart:** Each dot is one region-year observation with positive ECP_Bulk. Dots on the 45° dashed line = perfect fit; above = model underestimate; below = model overestimate. The shaded band spans ±2 log-SD. Outliers (beyond ±2 SD) are **excluded from the view** to preserve scale for the main distribution — they are listed in the table below with their region, year, and exact prices.
</div>

```{r calib-outlier-tbl-bulk, echo=FALSE}
make_outlier_table(fit_str$Bulk, "Bulk")
```
</div>

<div class="diffuse-only">
```{r cal-scatter-diffuse, echo=FALSE, fig.height=7, dev="png", dpi=150}
p_cs <- make_calibration_scatter(fit_str$Diffuse, "Diffuse")
if (!is.null(p_cs)) {
  if (knitr::is_html_output()) {
    p_plotly <- plotly::ggplotly(p_cs, tooltip = "text")
    p_plotly <- plotly::layout(p_plotly, margin = list(t = 50, b = 50, l = 50, r = 50))
    p_plotly
  } else {
    print(p_cs)
  }
  tryCatch(ggsave(file.path(plotOutputDir, "calib_scatter_diffuse.png"), p_cs, width=9, height=7, dpi=300), error=function(e)NULL)
}
```
<div class="chart-description">
**How to read this chart:** Same interpretation as Bulk — each dot is a region-year with positive ECP_Diffuse. Outliers are excluded from the axis range and listed below.
</div>

```{r calib-outlier-tbl-diffuse, echo=FALSE}
make_outlier_table(fit_str$Diffuse, "Diffuse")
```
</div>

### Fitted vs Observed Distribution

<div class="bulk-only">
```{r dens-bulk, echo=FALSE, fig.height=4}
p <- make_density_plot(fit_str$Bulk, "Bulk")
if (!is.null(p)) print(p)
```
<div class="chart-description">
**How to read this chart:** Density curves of the observed (sector colour) and model-fitted (orange) log(1 + ECP) values. Closer overlap indicates better calibration. A narrower fitted distribution suggests under-dispersion; a shifted distribution suggests systematic bias.
</div>
</div>

<div class="diffuse-only">
```{r dens-diffuse, echo=FALSE, fig.height=4}
p <- make_density_plot(fit_str$Diffuse, "Diffuse")
if (!is.null(p)) print(p)
```
<div class="chart-description">
**How to read this chart:** Density curves for the Diffuse sector — observed vs model-fitted log(1 + ECP_Diffuse). Close alignment indicates the model captures the observed price distribution well.
</div>
</div>

### Price Trajectories (Historical)

<div class="bulk-only">
```{r hist-traj-bulk, echo=FALSE, fig.height=8}
p <- make_hist_trajectory(fit_str$Bulk, "Bulk", key_regions)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "hist_traj_bulk.png"), p, width=12, height=8, dpi=300)
} else {
  cat("> Historical trajectory not available (requires non-cached model run).\n")
}
```
<div class="chart-description">
**How to read this chart:** Dots = observed ECP; line = model-fitted ECP for key regions. Close alignment indicates a good in-sample fit. Systematic gaps suggest the model is missing a region-specific dynamic. Only regions with at least one year of positive ECP appear.
</div>
</div>

<div class="diffuse-only">
```{r hist-traj-diffuse, echo=FALSE, fig.height=8}
p <- make_hist_trajectory(fit_str$Diffuse, "Diffuse", key_regions)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "hist_traj_diffuse.png"), p, width=12, height=8, dpi=300)
} else {
  cat("> Historical trajectory not available (requires non-cached model run).\n")
}
```
<div class="chart-description">
**How to read this chart:** Dots = observed ECP_Diffuse; line = model-fitted ECP_Diffuse for key regions. Close alignment = good in-sample fit. Persistent gaps indicate unmodelled region-specific effects in the diffuse sector.
</div>
</div>

### Residual Diagnostics

<div class="bulk-only">
```{r resid-bulk, echo=FALSE, fig.width=14, fig.height=5}
p <- make_residual_plots(fit_str$Bulk, "Bulk", cfg$Bulk$instQuality)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "resid_bulk.png"), p, width=14, height=5, dpi=300)
}
```
<div class="chart-description">
**How to read this chart:** (1) Residuals vs Fitted — flat loess near zero is ideal; curvature suggests functional misspecification. (2) Residuals vs Year — no trend indicates no unmodelled temporal dynamics. (3) Residuals vs Actor Power — no pattern confirms the linear functional form is adequate. Deviance residuals shown throughout.
</div>
</div>

<div class="diffuse-only">
```{r resid-diffuse, echo=FALSE, fig.width=14, fig.height=5}
p <- make_residual_plots(fit_str$Diffuse, "Diffuse", cfg$Diffuse$instQuality)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "resid_diffuse.png"), p, width=14, height=5, dpi=300)
}
```
<div class="chart-description">
**How to read this chart:** Same three-panel diagnostic for the Diffuse sector. A flat loess at zero in all three panels indicates well-behaved residuals. Deviations indicate potential misspecification in that dimension.
</div>
</div>

### Partial Regression Plots

<div class="bulk-only">
```{r partial-bulk, echo=FALSE, fig.width=12, fig.height=5}
p <- make_partial_regression_plots(fit_str$Bulk, "Bulk",
                                    cfg$Bulk$instQuality, cfg$Bulk$controls)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "partial_bulk.png"), p, width=12, height=5, dpi=300)
}
```
<div class="chart-description">
**How to read this chart:** The marginal contribution of each theory predictor to log(1 + ECP) after removing the linear effects of all other predictors. The slope of the red line equals the model coefficient. A steeper positive slope = stronger association with higher carbon prices, conditional on all other predictors.
</div>
</div>

<div class="diffuse-only">
```{r partial-diffuse, echo=FALSE, fig.width=12, fig.height=5}
p <- make_partial_regression_plots(fit_str$Diffuse, "Diffuse",
                                    cfg$Diffuse$instQuality, cfg$Diffuse$controls)
if (!is.null(p)) {
  print(p)
  ggsave(file.path(plotOutputDir, "partial_diffuse.png"), p, width=12, height=5, dpi=300)
}
```
<div class="chart-description">
**How to read this chart:** Partial regression plots for Diffuse sector theory predictors. Slope = model coefficient. Confirms whether the linear specification is appropriate for each predictor.
</div>
</div>

---

# Model Interpretation {.tabset}

## Semi-Elasticity Effects

The **semi-elasticity** of a log-link GLM coefficient β is defined as `exp(β) − 1`. It gives the **percentage change in expected carbon price** for a 1-unit increase in the predictor, holding all else equal. For example, β = 0.30 → semi-elasticity = 35%: a 1-unit increase in that variable is associated with a 35% higher expected ECP. Confidence intervals are computed via the delta method: SE(semi) = exp(β) × SE(β). Significance stars are identical to those from the original coefficient test (testing semi = 0 is equivalent to testing β = 0).

<div class="bulk-only">

```{r se-forest-bulk, echo=FALSE, fig.height=6}
if (!is.null(se_bulk)) {
  df_se <- se_bulk %>%
    filter(!grepl("^\\(Intercept\\)|^regionFE|^timeTrend", term)) %>%
    mutate(term = factor(term, levels = rev(term)), sig = p < 0.05)
  ggplot(df_se, aes(x = semi_elas * 100, y = term, color = group)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_errorbarh(aes(xmin = lower * 100, xmax = upper * 100), height = 0.3, linewidth = 0.7) +
    geom_point(aes(shape = sig), size = 3) +
    scale_shape_manual(values = c("TRUE" = 19, "FALSE" = 1), guide = "none") +
    scale_color_viridis_d(option = "D") +
    labs(title = "Bulk — Semi-Elasticity Forest Plot",
         subtitle = "% change in expected price per 1-unit increase. Filled dot = p < 0.05; hollow = not significant.",
         x = "Semi-elasticity (%)", y = NULL, color = "Term Group") +
    theme_report()
}
```

<div class="chart-description">
**How to read this chart:** Each row is one predictor (intercept, region FE, and time trend excluded). The point is the semi-elasticity (% change in expected ECP per 1-unit increase); the horizontal bar is the delta-method 95% CI. Bars extending right of the dashed zero line indicate the variable is associated with higher carbon prices; bars extending left indicate lower prices. **Filled circles** are statistically significant (p &lt; 0.05); hollow circles are not. The semi-elasticity is computed as exp(β) − 1 from the log-link GLM coefficient β, so it captures the multiplicative (percentage) effect rather than the additive (level) effect.
</div>

```{r se-table-bulk, echo=FALSE}
if (!is.null(se_bulk)) {
  tbl <- se_bulk %>%
    filter(!grepl("^\\(Intercept\\)|^regionFE|^timeTrend", term)) %>%
    mutate(across(c(estimate, semi_elas, se, lower, upper), ~round(.x, 4)),
           p = round(p, 3),
           sig = ifelse(p < 0.001, "***", ifelse(p < 0.01, "**",
                   ifelse(p < 0.05, "*", ifelse(p < 0.1, ".", "")))),
           `Semi-elas. (%)` = sprintf("%.1f%%", semi_elas * 100),
           `95% CI`         = sprintf("[%.1f%%, %.1f%%]", lower*100, upper*100)) %>%
    select(term, group, estimate, `Semi-elas. (%)`, `95% CI`, p, sig)
  kable(tbl, format="html", caption="Bulk — Semi-Elasticities (% change in expected ECP per 1-unit increase in predictor)") %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=FALSE)
}
```

</div>

<div class="diffuse-only">

```{r se-forest-diffuse, echo=FALSE, fig.height=6}
if (!is.null(se_diffuse)) {
  df_se <- se_diffuse %>%
    filter(!grepl("^\\(Intercept\\)|^regionFE|^timeTrend", term)) %>%
    mutate(term = factor(term, levels = rev(term)), sig = p < 0.05)
  ggplot(df_se, aes(x = semi_elas * 100, y = term, color = group)) +
    geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
    geom_errorbarh(aes(xmin = lower * 100, xmax = upper * 100), height = 0.3, linewidth = 0.7) +
    geom_point(aes(shape = sig), size = 3) +
    scale_shape_manual(values = c("TRUE" = 19, "FALSE" = 1), guide = "none") +
    scale_color_viridis_d(option = "D") +
    labs(title = "Diffuse — Semi-Elasticity Forest Plot",
         subtitle = "% change in expected price per 1-unit increase. Filled dot = p < 0.05; hollow = not significant.",
         x = "Semi-elasticity (%)", y = NULL, color = "Term Group") +
    theme_report()
}
```

<div class="chart-description">
**How to read this chart:** Semi-elasticities for the Diffuse sector. The interpretation is identical to the Bulk plot: each bar shows the percentage change in expected ECP_Diffuse for a 1-unit increase in the predictor. Filled dots are significant (p &lt; 0.05). Compare with the Bulk plot to see which predictors have similar vs divergent effects across sectors.
</div>

```{r se-table-diffuse, echo=FALSE}
if (!is.null(se_diffuse)) {
  tbl <- se_diffuse %>%
    filter(!grepl("^\\(Intercept\\)|^regionFE|^timeTrend", term)) %>%
    mutate(across(c(estimate, semi_elas, se, lower, upper), ~round(.x, 4)),
           p = round(p, 3),
           sig = ifelse(p < 0.001, "***", ifelse(p < 0.01, "**",
                   ifelse(p < 0.05, "*", ifelse(p < 0.1, ".", "")))),
           `Semi-elas. (%)` = sprintf("%.1f%%", semi_elas * 100),
           `95% CI`         = sprintf("[%.1f%%, %.1f%%]", lower*100, upper*100)) %>%
    select(term, group, estimate, `Semi-elas. (%)`, `95% CI`, p, sig)
  kable(tbl, format="html", caption="Diffuse — Semi-Elasticities (% change in expected ECP per 1-unit increase in predictor)") %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=FALSE)
}
```

</div>

## Marginal Effects in USD/tCO₂

The **Marginal Effect at Mean (MEM)** translates the log-scale coefficient into policy-relevant USD/tCO₂ units: MEM = β × E[ECP|x̄], evaluated at the sample mean of fitted prices. An MEM of $5/tCO₂ means that a 1-unit increase in the predictor is associated with $5 higher expected ECP, at the average price level in the sample.

<div class="bulk-only">

```{r mem-bar-bulk, echo=FALSE, fig.height=5}
if (!is.null(mem_bulk)) {
  df_mem <- mem_bulk %>%
    filter(!grepl("^\\(Intercept\\)|^regionFE|^timeTrend", term)) %>%
    mutate(term = factor(term, levels = term[order(mem)]))
  ggplot(df_mem, aes(x = mem, y = term,
                      fill = ifelse(mem >= 0, col_good, col_bad))) +
    geom_col(alpha = 0.85) +
    geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3, color = "grey30") +
    geom_vline(xintercept = 0, linetype = "dashed") +
    scale_fill_identity() +
    scale_x_continuous() +
    labs(title = "Bulk — Marginal Effects at Mean (USD/tCO₂)",
         subtitle = sprintf("Evaluated at sample mean fitted price: $%.1f/tCO₂", mem_bulk$mean_price[1]),
         x = "MEM (USD/tCO₂)", y = NULL) +
    theme_report()
}
```

<div class="chart-description">
**How to read this chart:** Each bar shows the dollar-value impact on expected ECP_Bulk of a 1-unit increase in the predictor, evaluated at the average fitted price in the sample. Green bars increase the expected price; red bars decrease it. The whiskers are 95% CIs (delta method). This chart is complementary to the semi-elasticity plot: semi-elasticities show the *percentage* change and are independent of price level; MEMs show the *absolute* USD impact, which scales with the baseline price level.
</div>

```{r mem-table-bulk, echo=FALSE}
if (!is.null(mem_bulk)) {
  tbl <- mem_bulk %>%
    filter(!grepl("^\\(Intercept\\)|^regionFE|^timeTrend", term)) %>%
    mutate(across(c(mem, se, lower, upper), ~round(.x, 2)),
           `MEM (USD/tCO₂)` = sprintf("%.2f", mem),
           `95% CI` = sprintf("[%.2f, %.2f]", lower, upper)) %>%
    select(term, `MEM (USD/tCO₂)`, `95% CI`)
  kable(tbl, format="html",
        caption=sprintf("Bulk — Marginal Effects at Mean (USD/tCO₂), evaluated at $%.1f/tCO₂",
                        mem_bulk$mean_price[1])) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=FALSE)
}
```

</div>

<div class="diffuse-only">

```{r mem-bar-diffuse, echo=FALSE, fig.height=5}
if (!is.null(mem_diffuse)) {
  df_mem <- mem_diffuse %>%
    filter(!grepl("^\\(Intercept\\)|^regionFE|^timeTrend", term)) %>%
    mutate(term = factor(term, levels = term[order(mem)]))
  ggplot(df_mem, aes(x = mem, y = term,
                      fill = ifelse(mem >= 0, col_good, col_bad))) +
    geom_col(alpha = 0.85) +
    geom_errorbarh(aes(xmin = lower, xmax = upper), height = 0.3, color = "grey30") +
    geom_vline(xintercept = 0, linetype = "dashed") +
    scale_fill_identity() +
    scale_x_continuous() +
    labs(title = "Diffuse — Marginal Effects at Mean (USD/tCO₂)",
         subtitle = sprintf("Evaluated at sample mean fitted price: $%.1f/tCO₂", mem_diffuse$mean_price[1]),
         x = "MEM (USD/tCO₂)", y = NULL) +
    theme_report()
}
```

<div class="chart-description">
**How to read this chart:** MEM for the Diffuse sector. Each bar shows the USD/tCO₂ impact of a 1-unit increase in the predictor at the average Diffuse ECP level. Comparing this chart to the Bulk MEM chart reveals whether the same predictor has a larger or smaller absolute dollar impact across sectors (driven by differences in both the coefficient magnitude and the baseline price level).
</div>

```{r mem-table-diffuse, echo=FALSE}
if (!is.null(mem_diffuse)) {
  tbl <- mem_diffuse %>%
    filter(!grepl("^\\(Intercept\\)|^regionFE|^timeTrend", term)) %>%
    mutate(across(c(mem, se, lower, upper), ~round(.x, 2)),
           `MEM (USD/tCO₂)` = sprintf("%.2f", mem),
           `95% CI` = sprintf("[%.2f, %.2f]", lower, upper)) %>%
    select(term, `MEM (USD/tCO₂)`, `95% CI`)
  kable(tbl, format="html",
        caption=sprintf("Diffuse — Marginal Effects at Mean (USD/tCO₂), evaluated at $%.1f/tCO₂",
                        mem_diffuse$mean_price[1])) %>%
    kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=FALSE)
}
```

</div>

---

# Future Projections (Adoption-Gated) {.tabset}

```{r proj-gate-check, echo=FALSE, results='asis'}
if (!has_scenario) {
  cat("> [!WARNING]\n> **Scenario GDX not provided.** All projection sections below are disabled.\n\n")
}
```

## Price Trajectories

<div class="bulk-only">

```{r price-traj-bulk, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario && !is.null(proj_str$Bulk)) {
  p <- make_price_trajectory_plot(proj_str$Bulk, "Bulk", key_regions)
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "price_traj_bulk.png"), p, width=12, height=6, dpi=300)
    ggsave(file.path(plotOutputDir, "price_traj_bulk.svg"), p, width=12, height=6)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** Each line shows the projected ECP_Bulk trajectory for one region, **conditional on adoption** (a price line only appears once the region's Pr(Adoption) ≥ `r adoption_threshold`). Key regions are highlighted in distinct colours; all other regions are shown in grey. A line that starts late (or is absent) means that region is not projected to cross the adoption threshold until later in the scenario. Lines that appear mid-way through reflect "late adopters". The y-axis is in USD/tCO₂ (back-transformed from the log-scale model). The full-facet view below shows all regions individually.
</div>

```{r price-traj-bulk-full, echo=FALSE, fig.width=16, fig.height=20, class.output="fold-hide"}
if (has_scenario && !is.null(proj_str$Bulk)) {
  p <- make_price_trajectory_plot(proj_str$Bulk, "Bulk", key_regions, full_facet = TRUE)
  if (!is.null(p)) print(p)
}
```

</div>

<div class="diffuse-only">

```{r price-traj-diffuse, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario && !is.null(proj_str$Diffuse)) {
  p <- make_price_trajectory_plot(proj_str$Diffuse, "Diffuse", key_regions)
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "price_traj_diffuse.png"), p, width=12, height=6, dpi=300)
    ggsave(file.path(plotOutputDir, "price_traj_diffuse.svg"), p, width=12, height=6)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** Projected ECP_Diffuse conditional on adoption. Lines only appear once Pr(Diffuse Adoption) ≥ threshold. Regions absent from the chart are not projected to cross the adoption threshold in the scenario window.
</div>

```{r price-traj-diffuse-full, echo=FALSE, fig.width=16, fig.height=20, class.output="fold-hide"}
if (has_scenario && !is.null(proj_str$Diffuse)) {
  p <- make_price_trajectory_plot(proj_str$Diffuse, "Diffuse", key_regions, full_facet = TRUE)
  if (!is.null(p)) print(p)
}
```

</div>

## Price Onset Chart

<div class="bulk-only">

```{r gantt-bulk, echo=FALSE, fig.width=12, fig.height=16}
if (has_scenario && !is.null(proj_str$Bulk)) {
  p <- make_price_gantt(proj_str$Bulk, "Bulk")
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "gantt_bulk.png"), p, width=12, height=16, dpi=300)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** Each row is one region; the bar starts at the first year the region crosses the adoption threshold. Bar colour = starting price level (warmer = higher). Regions sorted from earliest to latest adopter. Provides a "when and at what level" summary of the Bulk pricing rollout.
</div>

</div>

<div class="diffuse-only">

```{r gantt-diffuse, echo=FALSE, fig.width=12, fig.height=16}
if (has_scenario && !is.null(proj_str$Diffuse)) {
  p <- make_price_gantt(proj_str$Diffuse, "Diffuse")
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "gantt_diffuse.png"), p, width=12, height=16, dpi=300)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** Price onset chart for the Diffuse sector. Compare with Bulk to see whether sectors tend to be adopted at similar times or one leads the other.
</div>

</div>

```{r maps-main, echo=FALSE, results='asis'}
if (has_scenario && !is.null(proj_str$Bulk) && !is.null(mapping_54) && !is.null(world)) {
  # Bulk snapshot maps
  cat('<div class="bulk-only">\n\n')
  cat('## Snapshot Maps {.tabset}\n\n')
  for (yr in snapshot_years) {
    cat(sprintf("\n### %d\n\n", yr))
    p <- make_price_map(proj_str$Bulk, mapping_54, world, yr, "Bulk")
    if (!is.null(p)) {
      print(p)
      ggsave(file.path(plotOutputDir, sprintf("map_price_bulk_%d.png", yr)), p, width=12, height=6, dpi=300)
    }
    cat("\n\n")
  }
  cat('\n</div>\n\n')

  # Diffuse snapshot maps
  cat('<div class="diffuse-only">\n\n')
  cat('## Snapshot Maps {.tabset}\n\n')
  for (yr in snapshot_years) {
    cat(sprintf("\n### %d\n\n", yr))
    p <- make_price_map(proj_str$Diffuse, mapping_54, world, yr, "Diffuse")
    if (!is.null(p)) {
      print(p)
      ggsave(file.path(plotOutputDir, sprintf("map_price_diffuse_%d.png", yr)), p, width=12, height=6, dpi=300)
    }
    cat("\n\n")
  }
  cat('\n</div>\n\n')
} else {
  cat('\n## Snapshot Maps\n\n> Map rendering requires scenario data and regionmapping_54.csv.\n\n')
}
```

```{r ranking-main, echo=FALSE, results='asis'}
# Bulk Regional Price Ranking
if (has_scenario && !is.null(proj_str$Bulk)) {
  cat('<div class="bulk-only">\n\n')
  cat('## Regional Price Ranking {.tabset}\n\n')
  ranks_bulk <- make_price_ranking(proj_str$Bulk, snapshot_years, top_n = 15)
  if (!is.null(ranks_bulk)) {
    for (i in seq_along(ranks_bulk)) {
      yr  <- snapshot_years[i]
      tbl <- ranks_bulk[[i]]
      names(tbl) <- c("Rank", "Region", "ECP (USD/tCO₂)")
      tbl[, 3] <- round(tbl[, 3], 1)
      cat(sprintf("\n### %d\n\n", yr))
      cat(sprintf("\n**Top 15 regions by projected ECP in %d:**\n\n", yr))
      print(kable(tbl, format="html") %>%
        kable_styling(bootstrap_options=c("striped","hover","condensed"),
                      full_width=FALSE))
      cat("\n\n")
    }
  }
  cat('\n</div>\n\n')
}

# Diffuse Regional Price Ranking
if (has_scenario && !is.null(proj_str$Diffuse)) {
  cat('<div class="diffuse-only">\n\n')
  cat('## Regional Price Ranking {.tabset}\n\n')
  ranks_diff <- make_price_ranking(proj_str$Diffuse, snapshot_years, top_n = 15)
  if (!is.null(ranks_diff)) {
    for (i in seq_along(ranks_diff)) {
      yr  <- snapshot_years[i]
      tbl <- ranks_diff[[i]]
      names(tbl) <- c("Rank", "Region", "ECP (USD/tCO₂)")
      tbl[, 3] <- round(tbl[, 3], 1)
      cat(sprintf("\n### %d\n\n", yr))
      cat(sprintf("\n**Top 15 regions by projected ECP in %d:**\n\n", yr))
      print(kable(tbl, format="html") %>%
        kable_styling(bootstrap_options=c("striped","hover","condensed"),
                      full_width=FALSE))
      cat("\n\n")
    }
  }
  cat('\n</div>\n\n')
}

if (!has_scenario) {
  cat('\n## Regional Price Ranking\n\n> Scenario data not available.\n\n')
}
```

## Price Distribution Over Time

<div class="bulk-only">

```{r dist-bulk, echo=FALSE, fig.height=6}
if (has_scenario && !is.null(proj_str$Bulk)) {
  p <- make_price_distribution(proj_str$Bulk, snapshot_years, "Bulk")
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "dist_bulk.png"), p, width=10, height=6, dpi=300)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** Violin + box plots of projected ECP_Bulk across adopted regions at snapshot years. The violin = full distribution; the box = median & IQR. Only adopted regions shown. A widening violin = more price heterogeneity over time; a rising median = overall upward price trend.
</div>

</div>

<div class="diffuse-only">

```{r dist-diffuse, echo=FALSE, fig.height=6}
if (has_scenario && !is.null(proj_str$Diffuse)) {
  p <- make_price_distribution(proj_str$Diffuse, snapshot_years, "Diffuse")
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "dist_diffuse.png"), p, width=10, height=6, dpi=300)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** Distribution of projected ECP_Diffuse at snapshot years (adopted regions only). Compare with the Bulk distribution to assess whether diffuse sector prices are higher or lower on average, and whether they are more or less dispersed.
</div>

</div>

---

# Expected Carbon Price — Full Hurdle Model {.tabset}

This section shows the full two-stage hurdle estimate: **E[ECP] = Pr(Adoption) × Conditional Price**. Unlike Section 6 which hard-gates at the adoption threshold, here the price fades in continuously as adoption probability rises, giving the probabilistically correct expected carbon price.

```{r hurdle-check, echo=FALSE, results='asis'}
if (!has_scenario) cat("> **Scenario GDX not provided.** This section requires scenario data.\n\n")
```

## Expected Price Trajectories

<div class="bulk-only">

```{r hurdle-traj-bulk, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario && !is.null(hurdle$Bulk)) {
  p <- make_hurdle_trajectories(hurdle$Bulk, "Bulk", key_regions)
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "hurdle_traj_bulk.png"), p, width=12, height=6, dpi=300)
    ggsave(file.path(plotOutputDir, "hurdle_traj_bulk.svg"), p, width=12, height=6)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** E[ECP_Bulk] = Pr(Adoption) × Conditional_Price. The price fades in gradually as adoption probability rises — no hard cliff. A region with Pr(Adoption) = 0.4 and conditional price $20/tCO₂ shows expected ECP = $8/tCO₂.
</div>

</div>

<div class="diffuse-only">

```{r hurdle-traj-diffuse, echo=FALSE, fig.width=12, fig.height=6}
if (has_scenario && !is.null(hurdle$Diffuse)) {
  p <- make_hurdle_trajectories(hurdle$Diffuse, "Diffuse", key_regions)
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "hurdle_traj_diffuse.png"), p, width=12, height=6, dpi=300)
    ggsave(file.path(plotOutputDir, "hurdle_traj_diffuse.svg"), p, width=12, height=6)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** E[ECP_Diffuse] = Pr(Diffuse Adoption) × Conditional Diffuse Price. Same gradual fade-in interpretation as Bulk.
</div>

</div>

```{r hurdle-decomp, echo=FALSE, results='asis'}
has_bulk_decomp    <- has_scenario && !is.null(hurdle$Bulk)   && !is.null(proj_str$Bulk)
has_diffuse_decomp <- has_scenario && !is.null(hurdle$Diffuse) && !is.null(proj_str$Diffuse)

if (has_bulk_decomp || has_diffuse_decomp) {
  cat('## Hurdle Decomposition {.tabset}\n\n')
  cat('Shows Pr(Adoption), Conditional Price, Expected ECP, and the hard-gated price side-by-side. Where hard-gated and expected diverge most indicates where the threshold assumption is load-bearing.\n\n')
  for (yr in snapshot_years) {
    cat(sprintf("\n### %d\n\n", yr))
    if (has_bulk_decomp) {
      h <- hurdle$Bulk[hurdle$Bulk$year == yr, ]
      g <- proj_str$Bulk[proj_str$Bulk$year == yr, c("region", "price_gated")]
      if (nrow(h) > 0) {
        df_dec <- merge(h[, c("region", "pr_adoption", "price_hat_usd", "expected_ecp")],
                        g, by = "region", all.x = TRUE)
        df_dec <- df_dec[order(-df_dec$expected_ecp), ]
        df_dec[, 2:5] <- round(df_dec[, 2:5], 2)
        names(df_dec) <- c("Region", "Pr(Adoption)", "Cond. Price (USD/tCO₂)", "Expected ECP (USD/tCO₂)", "Gated Price (USD/tCO₂)")
        cat('<div class="bulk-only">\n\n')
        cat(sprintf("**Bulk Hurdle Decomposition — %d:**\n\n", yr))
        print(kable(df_dec, format="html") %>%
          kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=FALSE))
        cat("\n\n</div>\n\n")
      }
    }
    if (has_diffuse_decomp) {
      h <- hurdle$Diffuse[hurdle$Diffuse$year == yr, ]
      g <- proj_str$Diffuse[proj_str$Diffuse$year == yr, c("region", "price_gated")]
      if (nrow(h) > 0) {
        df_dec <- merge(h[, c("region", "pr_adoption", "price_hat_usd", "expected_ecp")],
                        g, by = "region", all.x = TRUE)
        df_dec <- df_dec[order(-df_dec$expected_ecp), ]
        df_dec[, 2:5] <- round(df_dec[, 2:5], 2)
        names(df_dec) <- c("Region", "Pr(Adoption)", "Cond. Price (USD/tCO₂)", "Expected ECP (USD/tCO₂)", "Gated Price (USD/tCO₂)")
        cat('<div class="diffuse-only">\n\n')
        cat(sprintf("**Diffuse Hurdle Decomposition — %d:**\n\n", yr))
        print(kable(df_dec, format="html") %>%
          kable_styling(bootstrap_options=c("striped","hover","condensed"), full_width=FALSE))
        cat("\n\n</div>\n\n")
      }
    }
  }
}

if (!has_scenario) cat('\n## Hurdle Decomposition\n\n> Scenario data not available.\n\n')
```

## Threshold Sensitivity

<div class="bulk-only">

```{r thresh-bulk, echo=FALSE, fig.width=14, fig.height=8}
if (has_scenario && !is.null(future_df$Bulk)) {
  p <- make_threshold_sensitivity(fit_str$Bulk, fit_adp$Bulk, future_df$Bulk,
                                   "Bulk", key_regions,
                                   str_regionFE = cfg$Bulk$regionFE, adp_regionFE = adp_cfg$regionFE)
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "thresh_sensitivity_bulk.png"), p, width=14, height=8, dpi=300)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** Each panel shows projected ECP_Bulk for one key region under three adoption thresholds (0.3, 0.5, 0.7). Lower threshold → earlier onset. A wide spread between lines = projections are threshold-sensitive; a narrow spread = results are robust.
</div>

</div>

<div class="diffuse-only">

```{r thresh-diffuse, echo=FALSE, fig.width=14, fig.height=8}
if (has_scenario && !is.null(future_df$Diffuse)) {
  p <- make_threshold_sensitivity(fit_str$Diffuse, fit_adp$Diffuse, future_df$Diffuse,
                                   "Diffuse", key_regions,
                                   str_regionFE = cfg$Diffuse$regionFE, adp_regionFE = adp_cfg$regionFE)
  if (!is.null(p)) {
    print(p)
    ggsave(file.path(plotOutputDir, "thresh_sensitivity_diffuse.png"), p, width=14, height=8, dpi=300)
  }
} else cat("> Scenario data not available.")
```

<div class="chart-description">
**How to read this chart:** Threshold sensitivity for the Diffuse sector. Same interpretation as Bulk — a wide spread between threshold lines means results depend heavily on the threshold choice.
</div>

</div>

---

# Comparison {.tabset}

This section compares Bulk and Diffuse sector results side-by-side. Use this section to assess whether the two sectors behave similarly or diverge in their carbon pricing dynamics.

## Model Fit Comparison

```{r fit-comparison, echo=FALSE, fig.height=5}
fit_cmp <- data.frame(
  Sector    = c("Bulk", "Diffuse"),
  `R²`      = c(stats_bulk$r2,   stats_diffuse$r2),
  AIC       = c(stats_bulk$aic,  stats_diffuse$aic),
  BIC       = c(stats_bulk$bic,  stats_diffuse$bic),
  `N obs`   = c(stats_bulk$nObs, stats_diffuse$nObs),
  check.names = FALSE
)

p_r2 <- ggplot(fit_cmp, aes(x = Sector, y = `R²`,
    fill = factor(ifelse(`R²` >= 0.5, "Good", ifelse(`R²` >= 0.25, "Moderate", "Poor")),
                  levels = c("Poor", "Moderate", "Good")))) +
  geom_col(width = 0.5) +
  geom_text(aes(label = sprintf("%.3f", `R²`)), vjust = -0.5, size = 4, fontface = "bold") +
  scale_fill_manual(values = c("Poor" = col_bad, "Moderate" = col_warn, "Good" = col_good),
                    name = "Fit quality") +
  labs(title = "R² by Sector (deviance-based)", y = "R²", x = NULL) +
  ylim(0, 1) + theme_report()

p_aic <- ggplot(fit_cmp, aes(x = Sector, y = AIC, fill = Sector)) +
  geom_col(width = 0.5) +
  geom_text(aes(label = sprintf("%.0f", AIC)), vjust = -0.5, size = 4, fontface = "bold") +
  scale_fill_manual(values = c("Bulk" = col_bulk, "Diffuse" = col_diffuse)) +
  labs(title = "AIC by Sector (lower = better)", y = "AIC", x = NULL) +
  theme_report() + theme(legend.position = "none")

library(patchwork)
p_r2 + p_aic + plot_layout(guides = "collect") & theme(legend.position = "bottom")
```

<div class="chart-description">
**How to read these charts:** R² (left) is the deviance-based pseudo-R² — values above 0.5 indicate a good fit (green), 0.25–0.5 moderate (orange), below 0.25 poor (red). AIC (right) measures relative model quality within each sector — lower is better, but AICs are not comparable across sectors because they are estimated on different samples. Use R² for within-sector fit quality and AIC for model selection across specifications within the same sector.
</div>

```{r fit-comparison-table, echo=FALSE}
kable(fit_cmp, format = "html", caption = "Stringency Model Fit Summary — Bulk vs Diffuse") %>%
  kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
```

## Coefficient Comparison

```{r sector-compare, echo=FALSE}
cmp <- make_comparison_table(se_bulk, se_diffuse)
if (!is.null(cmp)) {
  kable(cmp, format = "html",
        caption = "Sector Comparison — Semi-Elasticities for Shared Predictors") %>%
    kable_styling(bootstrap_options = c("striped","hover","condensed"), full_width = FALSE)
} else {
  cat("> Comparison table not available (one or both sector models are missing).\n")
}
```

<div class="chart-description">
**How to read this table:** Each row is a predictor that appears in both the Bulk and Diffuse model specifications. Columns show the raw log-scale coefficient (β), the semi-elasticity (% change per unit), and the p-value for each sector. Large differences in semi-elasticities between sectors indicate that the predictor has a materially different effect on industrial vs household/transport carbon pricing. Divergent significance (significant in one sector but not the other) may indicate sector-specific political economy dynamics.
</div>

## Convergence Maps {.tabset}

Difference between the hard-gated price (Section 6, threshold-based) and the expected price (Section 8, probability-weighted). **Red** = hard-gating overestimates vs the probabilistic expectation; **blue** = hard-gating underestimates. The largest discrepancies occur near the adoption threshold, where the binary gate creates a cliff that the hurdle model smooths out.

```{r conv-maps, echo=FALSE, results='asis'}
conv_year <- snapshot_years[ceiling(length(snapshot_years) / 2)]
if (has_scenario && !is.null(mapping_54) && !is.null(world)) {
  for (sec in c("Bulk", "Diffuse")) {
    cat(sprintf("\n### %s (%d)\n\n", sec, conv_year))
    p <- make_convergence_map(proj_str[[sec]], hurdle[[sec]],
                               mapping_54, world, conv_year, sec)
    if (!is.null(p)) {
      print(p)
      ggsave(file.path(plotOutputDir, sprintf("conv_map_%s_%d.png", tolower(sec), conv_year)),
             p, width=12, height=6, dpi=300)
    }
    cat("\n\n")
  }
} else {
  cat("> Map rendering requires scenario data and regionmapping_54.csv.\n")
}
```

---

# Appendix {.tabset}

## Adoption Model Formulae (Gating)

```{r appendix-adp-formula, echo=FALSE, results='asis'}
for (sec in c("Bulk", "Diffuse")) {
  fit <- fit_adp[[sec]]
  if (!is.null(fit)) {
    cat(sprintf("\n**%s Adoption (gating only):**\n\n", sec))
    cat("```\n")
    cat(deparse(fit$formula), "\n")
    cat("```\n\n")
  }
}
cat(sprintf("\nAdoption threshold: **%.2f**\n", adoption_threshold))
```

## Interpretation Guide

**Gaussian GLM (identity link) on log(1+ECP):** The model estimates `E[log(1+ECP)] = Xβ` (back-transform `ECP = exp(Xβ)−1`). With standardized drivers, a coefficient β is the additive effect on `log(1+ECP)` per 1-SD increase in X, i.e. it multiplies `(1+ECP)` by `exp(β)` — a multiplicative effect of `(exp(β)−1)×100`%.

**Log-transform of DV:** `ecp → log(1+ecp)` before fitting. Back-transform for display: `exp(fitted) − 1`. The +1 avoids log(0) and ensures non-negative predictions.

**Semi-elasticity:** `exp(β) − 1`. Delta-method SE: `exp(β) × SE(β)`. Testing semi = 0 is equivalent to testing β = 0, so p-values are identical to those from the coefficient table.

**Marginal Effect at Mean (MEM):** `β × E[ECP|x̄]`. Translates the log-scale coefficient into USD/tCO₂ at the sample-mean fitted price.

**Adoption gating:** Stringency models are estimated on `ecp > 0` observations only. In projections, prices are shown only where the adoption model predicts `Pr(Adoption) ≥ threshold`.

**Expected ECP:** `E[ECP] = Pr(Adoption) × Conditional_Price`. The full two-stage hurdle model output — the probabilistically-correct expected carbon price.

**Residual types:** Response residuals (observed − fitted on the `log(1+ECP)` scale) are shown throughout, appropriate for the gaussian (identity-link) model.

## Session Information

```{r session, echo=FALSE}
tryCatch(
  print(sessionInfo()),
  error = function(e) {
    cat("R Version:", R.version.string, "\n")
    cat("Platform:", R.version$platform, "\n")
  }
)
```
