---
title: "PFM Panel Data — Input Diagnostics"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_depth: 1
    toc_float:
      collapsed: true
    number_sections: true
    code_folding: hide
params:
  cacheDir: ""
  gdxPath: "data/fulldata.gdx"
  reportName: "default"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 5.5)
library(ggplot2)
library(dplyr)
library(tidyr)
library(magclass)
library(pfm)
library(rprojroot)

root <- find_rstudio_root_file()
source(file.path(root, "src/r/configHelper.R"))
source(file.path(root, "src/r/theme.R"))
source(file.path(root, "src/r/reportHelpers.R"))

if (nzchar(params$cacheDir)) madrat::setConfig(cachefolder = params$cacheDir)
madrat::setConfig(forcecache = TRUE)

# ── Load panels ────────────────────────────────────────────────────────────────
hist_mag <- getPanelDataHistoricalCached(
  aggregate = TRUE, y = 2000:2022,
  outputRegionMappingFile = "regionmapping_54.csv"
)

gdx_path <- params$gdxPath
if (nzchar(gdx_path) && !is_absolute_path(gdx_path)) gdx_path <- file.path(root, gdx_path)
scen_mag <- NULL
if (nzchar(gdx_path) && file.exists(gdx_path)) {
  scen_mag <- tryCatch(
    getPanelDataScenarioCached(gdxFile = gdx_path, aggregate = TRUE,
                               outputRegionMappingFile = "regionmapping_54.csv"),
    error = function(e) NULL
  )
}
has_scen <- !is.null(scen_mag)

hist_long <- magpieToLong(hist_mag, "Historical")
scen_long <- if (has_scen) magpieToLong(scen_mag, "Scenario") else NULL
all_long  <- rbind(hist_long, scen_long)

# ── Variable groups (dynamic) ──────────────────────────────────────────────────
vars <- getNames(hist_mag)
grp <- list(
  ecp      = grep("^Effective Carbon Price\\|", vars, value = TRUE),
  apIndex  = grep("^(Actor Power Index|Innovator Power|Incumbent Power)\\|", vars, value = TRUE),
  apDriver = intersect(vars, c("VRE share", "Electrification", "Coal primary energy share",
                               "Oil/Gas primary energy share", "Fossil share in Industry")),
  wgi      = grep("\\(WGI\\)$", vars, value = TRUE),
  vdemAcc  = intersect(vars, c("Rule of Law (VDem)", "Vertical Accountability (VDem)",
                               "Horizontal Accountability (VDem)", "Diagonal Accountability (VDem)")),
  vdemSC   = intersect(vars, c("Civil Service Professionalism (VDem)", "Policy Implementation (VDem)",
                               "Rule Predictability (VDem)", "Absence of Corruption (VDem)",
                               "Meritocracy Index (VDem)")),
  scPC     = grep("^State Capacity PC", vars, value = TRUE)
)
grp$iq       <- c(grp$wgi, grp$vdemAcc, grp$vdemSC, grp$scPC)
grp$controls <- setdiff(vars, c(grp$ecp, grp$apIndex, grp$apDriver, grp$iq))

# ── Reusable section builders ──────────────────────────────────────────────────
statsTable <- function(varSet) {
  hist_long %>%
    filter(variable %in% varSet) %>%
    group_by(variable) %>%
    summarise(
      n = sum(is.finite(value)), nNA = sum(!is.finite(value)),
      min = min(value, na.rm = TRUE), p25 = quantile(value, 0.25, na.rm = TRUE),
      median = median(value, na.rm = TRUE), mean = mean(value, na.rm = TRUE),
      p75 = quantile(value, 0.75, na.rm = TRUE), max = max(value, na.rm = TRUE),
      sd = sd(value, na.rm = TRUE), .groups = "drop"
    )
}

densityPlot <- function(varSet, title) {
  d <- hist_long %>% filter(variable %in% varSet, is.finite(value))
  ggplot(d, aes(x = value, fill = variable)) +
    geom_density(alpha = 0.35, colour = NA) +
    facet_wrap(~variable, scales = "free", ncol = 3) +
    theme_report() + theme(legend.position = "none") +
    labs(title = title, x = NULL, y = "Density")
}

trajectoryPlot <- function(varSet, title, capPrices = FALSE) {
  d <- all_long %>% filter(variable %in% varSet, is.finite(value))
  capNote <- ""
  if (capPrices) {
    ci <- capAtP99(d$value)
    d$value <- ci$values
    capNote <- capAnnotation(ci)
  }
  p <- ggplot(d, aes(x = year, y = value, group = interaction(region, source),
                     colour = source)) +
    geom_line(alpha = 0.3, linewidth = 0.3) +
    stat_summary(aes(group = source), fun = median, geom = "line", linewidth = 1.1) +
    facet_wrap(~variable, scales = "free_y", ncol = 3) +
    scale_colour_manual(values = c(Historical = "#1f77b4", Scenario = "#d62728")) +
    geom_vline(xintercept = 2022.5, linetype = "dashed", colour = "grey40") +
    theme_report() +
    labs(title = title, x = NULL, y = NULL, colour = NULL,
         subtitle = capNote)
  p
}

dispersionPlot <- function(varSet, title) {
  d <- all_long %>%
    filter(variable %in% varSet, is.finite(value)) %>%
    group_by(variable, source, year) %>%
    summarise(sdAcrossRegions = sd(value, na.rm = TRUE), .groups = "drop")
  ggplot(d, aes(x = year, y = sdAcrossRegions, colour = source)) +
    geom_line(linewidth = 0.9) +
    facet_wrap(~variable, scales = "free_y", ncol = 3) +
    scale_colour_manual(values = c(Historical = "#1f77b4", Scenario = "#d62728")) +
    geom_vline(xintercept = 2022.5, linetype = "dashed", colour = "grey40") +
    theme_report() +
    labs(title = title, x = NULL, y = "SD across regions", colour = NULL)
}

howDensity <- function() howToRead(
  "Each panel is the distribution of one variable over all region-years of the ",
  "historical panel (2000–2022), after all transformations (most variables are ",
  "normalized to [0,1]). Look for: values escaping the expected range, spikes at ",
  "0/1 (clamping pile-ups), or bimodality that a linear model would miss.")

howTrajectory <- function(unit = "normalized units [0,1]") howToRead(
  "Thin lines are individual regions; thick lines are the cross-region median. ",
  "Blue = historical, red = SSP2 scenario; the dashed vertical line is the ",
  "historical/scenario seam (2022/2025). Units: ", unit, ". Look for: jumps at ",
  "the seam (misalignment), implausible scenario drift, or regions pinned at the bounds.")

howDispersion <- function() howToRead(
  "Standard deviation across regions per year — the cross-sectional variation the ",
  "model identifies from. A falling line means regional convergence (less identifying ",
  "variation); an abrupt level change at the dashed seam indicates a projection ",
  "method inconsistency rather than a real-world change.")

howStats <- function() howToRead(
  "Summary statistics over all historical region-years. n = non-missing values, ",
  "nNA = missing. Most transformed variables should lie in [0,1]; min/max outside ",
  "that range or large nNA counts indicate upstream data problems.", type = "table")
```

# Executive Summary

```{r exec, results='asis'}
cat(sprintf(paste0(
  "- **Historical panel:** %d regions × %d years × %d variables (2000–2022).\n",
  "- **Scenario panel:** %s\n",
  "- **Variable groups:** %d carbon price, %d actor-power indexes, %d actor-power drivers, ",
  "%d institutional quality, %d controls.\n",
  "- Every chart and table below carries a *How to read* note; the **Seam-Continuity ",
  "Table** (last section before sparsity) is the fastest way to spot historical/scenario ",
  "misalignment.\n"
), nregions(hist_mag), nyears(hist_mag), ndata(hist_mag),
   if (has_scen) sprintf("%d regions × %d years × %d variables (SSP2).",
                         nregions(scen_mag), nyears(scen_mag), ndata(scen_mag))
   else "**not loaded** (no gdx) — scenario sections are skipped.",
   length(grp$ecp), length(grp$apIndex), length(grp$apDriver), length(grp$iq),
   length(grp$controls)))
```

# Data Lineage & Transformations {.tabset}

## Transformation methods

```{r methods, results='asis'}
methods <- data.frame(
  Method = c("Min–max normalization", "Log-scale + min–max", "Inversion (bad-when-high)",
             "State-capacity PCA", "GDP quartile-centring", "Driver standardization",
             "Moving average", "Scenario clamping", "Logistic time trend"),
  Definition = c(
    "x' = (x − min)/(max − min), bounds from **country-level** historical data so regional aggregates stay comparable across panels.",
    "x' = (log x − min log)/(max log − min log); applied to Population, GDP, GDP per Capita, Land Area (additive variables use **regional** bounds).",
    "x' = 1 − x before normalization, for Executive Corruption → Policy Implementation, Political Corruption → Absence of Corruption, Neopatrimonialism → Meritocracy Index.",
    "First two principal components of the 5-variable V-Dem state-capacity block, fit on historical data; scenario panels are projected with the **stored rotation** (fit/apply).",
    "x' = x − mean(income-quartile of region); breaks, group means and region assignments **frozen from the historical fit** (Scenario Scope, 2026-06-12).",
    "x'' = (x' − mean)/sd, applied to driver main effects + interaction factors at model time; mean/sd **frozen from the historical fit**. Fit-/significance-/tier-neutral; makes coefficients per-SD comparable (de-scale = ÷ sd). Negatives (e.g. Actor Power Index) handled.",
    "5-year moving average applied at panel load to smooth year-to-year noise.",
    "Scenario values clamped to historical normalization bounds (no extrapolation outside [0,1]).",
    "Raw logistic bounded in [0,1] (midpoint 2030, steepness 0.08, **no rescale**): toe before 2000, saturates ~1 by ~2060 — single common curve, cannot drive unbounded extrapolation."
  ),
  Implementation = c("mrpfm::toolNormalize", "mrpfm::toolNormalize(log(x))",
                     "panelDataHistorical (invert map)", "pfm::computeVDemStateCapacityPC",
                     "pfm:::computeGdpQCentred", "pfm::preparePanelData (driverScaling)",
                     "panelDataHistorical(movingAverage = 5)",
                     "mrpfm::toolNormalize(clamp = TRUE)", "pfm::preparePanelData"),
  stringsAsFactors = FALSE
)
knitr::kable(methods)
howToRead("One row per transformation applied anywhere in the pipeline: the formula, ",
          "where its bounds/parameters come from, and the implementing function. ",
          "If a variable looks wrong elsewhere in this report, check its method here first.",
          type = "table")
```

## Variable lineage

```{r lineage}
lineage <- hist_long %>%
  group_by(variable) %>%
  summarise(minObs = round(min(value, na.rm = TRUE), 3),
            maxObs = round(max(value, na.rm = TRUE), 3),
            nNA = sum(!is.finite(value)), .groups = "drop") %>%
  mutate(
    group = case_when(
      variable %in% grp$ecp ~ "Carbon Price",
      variable %in% grp$apIndex ~ "Actor Power Index",
      variable %in% grp$apDriver ~ "Actor Power Driver",
      variable %in% grp$wgi ~ "Inst. Quality (WGI)",
      variable %in% c(grp$vdemAcc, grp$vdemSC) ~ "Inst. Quality (VDem)",
      variable %in% grp$scPC ~ "Inst. Quality (PCA)",
      TRUE ~ "Control"
    ),
    source = case_when(
      grepl("WGI", variable) ~ "World Bank WGI",
      grepl("VDem", variable) ~ "V-Dem Country-Year",
      grepl("Carbon Price", variable) ~ "WB Carbon Pricing Dashboard",
      grepl("Power|VRE|Electrification|Coal|Oil|Fossil", variable) ~ "IEA/REMIND energy data",
      grepl("GDP|Population", variable) ~ "WDI / SSP",
      grepl("Land", variable) ~ "FAO",
      TRUE ~ "SSP extensions / WDI"
    ),
    transformations = case_when(
      grepl("Q-centred", variable) ~ "log + min–max + quartile-centring (frozen fit)",
      grepl("State Capacity PC", variable) ~ "inversion + min–max + PCA (stored rotation) + rescale [0,1]",
      grepl("Carbon Price", variable) ~ "sector ECP aggregation (USD/tCO2, untransformed)",
      grepl("GDP|Population|Land", variable) ~ "log + min–max (clamped in scenario)",
      TRUE ~ "min–max to [0,1] (country-level bounds) + 5y moving average"
    )
  ) %>%
  select(group, variable, source, transformations, minObs, maxObs, nNA) %>%
  arrange(group, variable)
interactiveTable(lineage, caption = "Per-variable lineage: source, transformation chain, observed range")
```

```{r lineage-caption, results='asis'}
howToRead("One row per final panel variable: where it comes from, every transformation ",
          "between raw source and the regression, and the range actually observed. ",
          "Sort by nNA to find sparse variables; minObs/maxObs outside [0,1] for a ",
          "normalized variable is a bug.", type = "table")
```

# Carbon Price & Adoption {.tabset}

## Price distributions

```{r ecp-dist, fig.height=4}
d <- hist_long %>% filter(variable %in% grp$ecp, is.finite(value), value > 0)
ci <- capAtP99(d$value)
d$value <- ci$values
ggplot(d, aes(x = value, fill = variable)) +
  geom_histogram(bins = 40, alpha = 0.6, position = "identity") +
  facet_wrap(~variable, ncol = 2) +
  theme_report() + theme(legend.position = "none") +
  labs(title = "Positive Effective Carbon Prices (historical)",
       x = "USD/tCO2", y = "Region-years", subtitle = capAnnotation(ci))
```

```{r ecp-dist-cap, results='asis'}
howToRead("Histogram of strictly positive ECP region-years per sector, in USD/tCO2, ",
          "display-capped at the 99th percentile (Price Outlier convention). The mass ",
          "near zero is normal (introductory prices); compare Bulk vs Diffuse support — ",
          "the stringency model is fit on exactly these observations.")
```

## Adoption over time

```{r adoption-time, fig.height=4.5}
ad <- hist_long %>%
  filter(variable %in% grp$ecp) %>%
  mutate(sector = sub("Effective Carbon Price\\|", "", variable),
         adopted = is.finite(value) & value > 0) %>%
  group_by(sector, year) %>%
  summarise(nAdopted = sum(adopted), share = mean(adopted), .groups = "drop")
ggplot(ad, aes(x = year, y = share, colour = sector)) +
  geom_line(linewidth = 1.1) + geom_point(size = 1.4) +
  theme_report() +
  labs(title = "Share of regions with a positive carbon price",
       x = NULL, y = "Share of regions", colour = "Sector")
```

```{r adoption-time-cap, results='asis'}
howToRead("Fraction of the 54 panel regions with ECP > 0 per year and sector — the ",
          "diffusion curve the adoption stage models. The logistic time trend assumes ",
          "this keeps rising in an S-curve with midpoint 2030; if the observed curve ",
          "has already flattened, that assumption deserves scrutiny.")
```

## Statistics

```{r ecp-stats, results='asis'}
knitr::kable(statsTable(grp$ecp), digits = 2)
howStats()
```

# Institutional Quality {.tabset}

## Distributions

```{r iq-dist, fig.height=9}
print(densityPlot(grp$iq, "Institutional quality variables (historical, transformed)"))
```

```{r iq-dist-cap, results='asis'}
howDensity()
```

## Trajectories & convergence

```{r iq-traj, fig.height=10}
print(trajectoryPlot(c(grp$vdemAcc, grp$scPC),
                     "V-Dem accountability + state-capacity PCs: historical and scenario"))
```

```{r iq-traj-cap, results='asis'}
howTrajectory()
if (has_scen) howToRead(
  "Scenario V-Dem indicators follow a logistic convergence to the 75th global ",
  "percentile by 2150 (midpoint 2080) — identical for every SSP (Scenario Scope: ",
  "SSP2-only). The red medians should bend smoothly upward; kinks or seam jumps ",
  "indicate projection bugs, not data.")
```

## Dispersion

```{r iq-disp, fig.height=7}
print(dispersionPlot(c(grp$vdemAcc, grp$scPC), "Cross-region dispersion: institutional quality"))
```

```{r iq-disp-cap, results='asis'}
howDispersion()
```

## Statistics

```{r iq-stats, results='asis'}
interactiveTable(statsTable(grp$iq), caption = "Institutional quality: historical summary statistics")
howStats()
```

# Actor Power {.tabset}

## Distributions

```{r ap-dist, fig.height=7}
print(densityPlot(c(grp$apIndex, grp$apDriver), "Actor power indexes and drivers (historical)"))
```

```{r ap-dist-cap, results='asis'}
howDensity()
```

## Trajectories

```{r ap-traj, fig.height=7}
print(trajectoryPlot(grp$apIndex, "Actor power indexes: historical and scenario"))
```

```{r ap-traj-cap, results='asis'}
howTrajectory()
```

## Dispersion

```{r ap-disp, fig.height=6}
print(dispersionPlot(grp$apIndex, "Cross-region dispersion: actor power"))
```

```{r ap-disp-cap, results='asis'}
howDispersion()
```

## Statistics

```{r ap-stats, results='asis'}
interactiveTable(statsTable(c(grp$apIndex, grp$apDriver)), caption = "Actor power: historical summary statistics")
howStats()
```

# Control Variables {.tabset}

## Distributions

```{r ctl-dist, fig.height=8}
print(densityPlot(grp$controls, "Control variables (historical, transformed)"))
```

```{r ctl-dist-cap, results='asis'}
howDensity()
```

## GDP per Capita (Q-centred): seam continuity

```{r gdpq, fig.height=4.5}
print(trajectoryPlot("GDP per Capita (Q-centred)",
                     "GDP per Capita (Q-centred): frozen historical reference"))
```

```{r gdpq-cap, results='asis'}
howToRead("The Q-centred GDP regressor with the quartile reference **frozen from the ",
          "historical fit** (breaks, group means, region assignments — Scenario Scope ",
          "2026-06-12). Continuity at the dashed seam is the acceptance test for that ",
          "fix: scenario lines must continue where historical lines end, then drift ",
          "upward as regions grow relative to their fixed historical income-class mean.")
```

## Logistic time trend

```{r trend, fig.height=4}
# Single common, bounded, saturating logistic (midpoint 2030, steepness 0.08, no
# rescale) — matches pfm::preparePanelData defaults.
yrs <- 2000:2100
td <- data.frame(year = yrs, trend = 1 / (1 + exp(-0.08 * (yrs - 2030))))
ggplot(td, aes(year, trend)) +
  annotate("rect", xmin = 2000, xmax = 2022, ymin = -Inf, ymax = Inf,
           alpha = 0.12, fill = "#1f77b4") +
  geom_line(linewidth = 1.1, colour = "#d62728") +
  geom_hline(yintercept = c(0, 1), linetype = "dotted") +
  geom_vline(xintercept = 2030, linetype = "dashed", colour = "grey50") +
  annotate("text", x = 2011, y = 0.9, label = "historical window",
           colour = "#1f77b4", size = 3.4) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_report() +
  labs(title = "Logistic time trend (single common, bounded, saturating curve)",
       x = NULL, y = "Trend value (bounded in [0,1])")
```

```{r trend-cap, results='asis'}
howToRead("The S-curve trend regressor: a raw logistic bounded in [0,1] (midpoint ",
          "2030 dashed, steepness 0.08). Its flat toe is before the shaded historical ",
          "window, it rises meaningfully through history (~0.08 → 0.35 over 2000–2022, ",
          "so the coefficient is identified on real movement), and it SATURATES near 1 ",
          "by ~2060 — so it can never drive an unbounded extrapolation. One common ",
          "curve for all scenarios; scenario differences come through the drivers, not ",
          "the trend (avoids double-counting ambition). Midpoint/steepness are tunable ",
          "(empirical calibration parked).")
```

## Statistics

```{r ctl-stats, results='asis'}
interactiveTable(statsTable(grp$controls), caption = "Controls: historical summary statistics")
howStats()
```

# Panel Transform Variants {.tabset}

```{r transform-prep}
pt_data <- list()
for (sec in c("Bulk", "Diffuse")) {
  base_df <- preparePanelData(
    data = hist_mag, sector = sec,
    actorPowerDrivers = c("Innovator Power", "Incumbent Power"),
    actorPowerIndex = "Actor Power Index",
    instQualityDrivers = intersect(c("State Capacity PC1 (VDem)", "Rule of Law (VDem)",
                                     "Vertical Accountability (VDem)"), vars),
    controlDrivers = intersect("GDP per Capita (Q-centred)", vars),
    regionMappingFixedEffects = NULL
  )
  base_df$adoption <- as.integer(base_df$ecp > 0)
  for (tr in c("levels", "hybridFD", "pureFD")) {
    for (stg in c("adoption", "stringency")) {
      df_t <- tryCatch(
        applyPanelTransform(base_df, panelTransform = tr, stage = stg,
                            actorPowerDrivers = c("Innovator Power", "Incumbent Power"),
                            actorPowerIndex = "Actor Power Index",
                            instQualityDrivers = "State Capacity PC1 (VDem)",
                            controlDrivers = "GDP per Capita (Q-centred)",
                            verbose = FALSE),
        error = function(e) NULL
      )
      if (!is.null(df_t)) {
        if (stg == "stringency" && tr == "levels") {
          df_t <- df_t[df_t$ecp > 0, , drop = FALSE]
          df_t$ecp <- log1p(df_t$ecp)
        }
        pt_data[[paste(sec, stg, tr, sep = "|")]] <- data.frame(
          sector = sec, stage = stg, transform = tr, n = nrow(df_t),
          outcome = if (stg == "adoption") mean(df_t$adoption, na.rm = TRUE)
                    else mean(df_t$ecp, na.rm = TRUE),
          stringsAsFactors = FALSE
        )
      }
    }
  }
}
pt_summary <- do.call(rbind, pt_data)
rownames(pt_summary) <- NULL
```

## Sample sizes & outcomes

```{r transform-table, results='asis'}
pt_wide <- pt_summary %>%
  mutate(outcomeLabel = ifelse(stage == "adoption", "event rate", "mean outcome")) %>%
  select(sector, stage, transform, n, outcome, outcomeLabel)
knitr::kable(pt_wide, digits = 3,
             caption = "Panel Transform variants: estimation sample size and outcome level")
howToRead("Each row is one Panel Transform per sector/stage. For adoption, n shrinks ",
          "under FD because the sample becomes the hazard at-risk set (not-yet-adopted ",
          "region-years) and 'event rate' becomes the onset rate. For stringency, n is ",
          "within-spell year pairs and the outcome is Δlog(1+ECP) (mean near 0). ",
          "Use this to judge the power cost of each transform before sweeping it.",
          type = "table")
```

# Correlation Analysis

```{r corr, fig.height=8}
corr_vars <- intersect(c("Actor Power Index|Bulk", "Innovator Power|Bulk", "Incumbent Power|Bulk",
                         grp$scPC, grp$vdemAcc, "Government Effectiveness (WGI)",
                         "GDP per Capita", "GDP per Capita (Q-centred)"), vars)
corr_df <- preparePanelData(
  data = hist_mag, sector = "Bulk",
  actorPowerDrivers = c("Innovator Power", "Incumbent Power"),
  actorPowerIndex = "Actor Power Index",
  instQualityDrivers = intersect(c(grp$scPC, grp$vdemAcc, "Government Effectiveness (WGI)"), vars),
  controlDrivers = intersect(c("GDP per Capita", "GDP per Capita (Q-centred)"), vars),
  regionMappingFixedEffects = NULL
)
num_cols <- setdiff(colnames(corr_df)[vapply(corr_df, is.numeric, logical(1))],
                    c("year", "timeTrend", "logisticTimeTrend", "lagged_ecp", "lagged_adoption"))
num_cols <- num_cols[!grepl("_x_", num_cols)]
cm <- cor(corr_df[, num_cols, drop = FALSE], use = "pairwise.complete.obs")
cm_long <- as.data.frame(as.table(cm))
names(cm_long) <- c("v1", "v2", "r")
ggplot(cm_long, aes(v1, v2, fill = r)) +
  geom_tile() +
  geom_text(aes(label = sprintf("%.2f", r)), size = 2.4) +
  scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b", limits = c(-1, 1)) +
  theme_report() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 7),
        axis.text.y = element_text(size = 7)) +
  labs(title = "Driver-space correlations (Bulk, historical estimation sample)",
       x = NULL, y = NULL)
```

```{r corr-cap, results='asis'}
howToRead("Pairwise Pearson correlations over the Bulk estimation sample. Cells with ",
          "|r| > 0.8 between institutional candidates are forbidden from co-entering a ",
          "spec (Stage-0 Channel Screen rule). The GDP row shows why Q-centred GDP exists: ",
          "raw GDP correlates strongly with governance, Q-centred should not.")
```

# Seam-Continuity Table

```{r seam, results='asis'}
if (has_scen) {
  seam <- computeSeamDiagnostics(hist_mag, scen_mag)
  cat(sprintf("**%d of %d variable×region seams flagged** (|jump| > 0.5 historical SD).\n\n",
              sum(seam$flagged), nrow(seam)))
  flaggedShare <- seam %>%
    group_by(variable) %>%
    summarise(nFlagged = sum(flagged), n = n(),
              maxAbsNormJump = max(abs(normJump), na.rm = TRUE), .groups = "drop") %>%
    arrange(desc(nFlagged))
  interactiveTable(flaggedShare, caption = "Seam flags per variable (worst first)")
  howToRead("Per variable: how many of its regional seams (last historical vs first ",
            "scenario value) jump by more than half a historical standard deviation. ",
            "A variable with many flags has a **projection-method inconsistency** — ",
            "every variable here should be near zero flags after the GDP Q-centred fix.",
            type = "table")
  interactiveTable(head(seam, 200), caption = "Worst 200 individual seams")
  howToRead("Individual variable×region seams sorted by |normalized jump|. normJump is ",
            "in historical standard deviations; histYear/scenYear show which years were ",
            "compared. Use this to localize exactly which region a misalignment hits.",
            type = "table")
} else {
  cat("> Scenario panel not available — seam diagnostics skipped.\n")
}
```

# Coverage & Sparsity

```{r sparsity, fig.height=7}
sp <- hist_long %>%
  group_by(variable) %>%
  summarise(naShare = mean(!is.finite(value)), .groups = "drop") %>%
  arrange(desc(naShare)) %>%
  head(25)
ggplot(sp, aes(x = reorder(variable, naShare), y = naShare)) +
  geom_col(fill = "#ff7f0e") + coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  theme_report() +
  labs(title = "Missing-value share by variable (top 25, historical)",
       x = NULL, y = "Share of region-years missing")
```

```{r sparsity-cap, results='asis'}
howToRead("Share of missing region-years per variable (worst 25). Variables with high ",
          "missingness silently shrink the estimation sample through complete-case ",
          "dropping — if a key driver appears here, the model's effective N is smaller ",
          "than the panel suggests.")
```
